### 2015-03-09 README.Rmd
readme is now created by an R markdown file

---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

This is code for an ABM of a climate prediction market
in which traders adapt their belief about the climate based
on the monetary performance of their neighboring traders in
a social network. The ABM is meant to test whether climate
prediction markets foster the convergence of people’s beliefs
about the climate. The model will be used to
assess whether such convergence is impacted by factors such
as the initial belief-homophily in the social network or the
time horizon at which climate securities are realized.

Martin Van der Linden and John Nay

Vanderbilt University

Master is the main script that sources everything else.
```{r, eval=FALSE, include=TRUE}
source(master.R)
````

### 2015-03-30 main.R
Add new outcome measure and anticipate possibility of different true models.

@@ -13,7 +13,8 @@ source("assortativity_coefficient.R")
 
 main <- function(visu = TRUE, seg = 0.95, 
                  market.struct = c("CDA", "LMSR"), 
-                 market.complet  = 4, outcome = "segreg", 
+                 outcome = c("segreg","converg"),
+                 market.complet  = 4,  
                  n.traders = 100, n.edg = 150,
                  risk.tak = 0.0001, ideo = 10) {
   ### Market structure parameters:
@@ -23,6 +24,7 @@ main <- function(visu = TRUE, seg = 0.95,
   # more precise temperature intervals
   
   market.struct <- match.arg(market.struct)
+  outcome <- match.arg(outcome)
   
   library(igraph)
   if (visu){
@@ -31,7 +33,10 @@ main <- function(visu = TRUE, seg = 0.95,
   #####
   ## Set model's parameters and create corresponding network
   #####
-  net <- generate_model( ### Network parameters
+  net <- generate_model( # True model of the climate
+    t.model = 1,
+    
+    ### Network parameters
     n.traders  = n.traders,
     n.edg      = n.edg,
     seg        = seg, # determine initial segregation of the 

### 2015-03-30 populate_net.R
Adapt terminology regarding true model.
Add attributes to the graph to allow measuring util.converg, and record initial value of util.converg.

@@ -23,12 +23,32 @@ PopulateNet <- function(
   V(g)$money  <- mon.dis
   V(g)$approx <- approx.dis 
     #NOTE : in the preliminary version, approximate models are taken from Sumner and Jackson (2008) 
-            # approx = 1  ->  true model (aka true)
+            # approx = 1  ->  Full model : ACC true, both geoingeneering and policy effective (aka full)
             # approx = 2  ->  ACC is a myth (aka myth)
             # approx = 3  ->  ACC is true but all measures are ineffective (aka ineff)
             # approx = 4  ->  ACC is true and policy are effective, but not geoingeneering (aka policy)
             # approx = 5  ->  ACC is true, geoingeneering is effective, but policy is not (aka engineer)
+  
+  # Assign values for the distance between approximate models. 
+  
+  V(g)$dist <- matrix(
+                c (0,3,2,1,1,
+                   3,0,1,1,2,
+                   2,3,0,1,1,
+                   1,2,3,0,0,
+                   1,2,3,0,0),
+                nrow=5,
+                ncol=5)
+  
+  # Will be used for
+  # convergence analysis in order to compute some social utility associated
+  # with a distribution of distances to the true model emerging from the 
+  # distribution of approximate models.
+  
+  
   V(g)$behav  <- behav.dis  # Market behavior of the traders, only ZI in preliminary model
+  
+  
   
   risk.tak.dis <- sample(0:(risk.tak*10000)/10000, n.traders, replace = TRUE)
   V(g)$risk.tak <- risk.tak.dis # Tendency of a player to take risk on the market by 

### 2015-03-30 main.R
Same comments as before, just commited before saving, so there were missing elements.

@@ -192,6 +192,14 @@ main <- function(visu = TRUE, seg = 0.95,
     #See more at: http://www.babelgraph.org/wp/?p=351#sthash.yWfBpOhv.dpuf
   }
   
+  if (outcome == "converg"){
+    
+    # return difference in utility of network convergence
+    
+    converg.util(net) - net$init.converg.util
+   
+  }
+  
 }
 
 

### 2015-03-30 populate_net.R

@@ -15,10 +15,10 @@ PopulateNet <- function(
   # package requirement
   require(igraph)
   
-  # construct empty graph
+  ## construct empty graph
   g <- graph.empty(n=n.traders, directed = FALSE)
   
-  # Assign attributes to agents in graph
+  ## Assign attributes to agents in graph
 
   V(g)$money  <- mon.dis
   V(g)$approx <- approx.dis 
@@ -29,22 +29,27 @@ PopulateNet <- function(
             # approx = 4  ->  ACC is true and policy are effective, but not geoingeneering (aka policy)
             # approx = 5  ->  ACC is true, geoingeneering is effective, but policy is not (aka engineer)
   
-  # Assign values for the distance between approximate models. 
+  ## Assign values for the distance between approximate models. 
   
-  V(g)$dist <- matrix(
-                c (0,3,2,1,1,
+  g <- set.graph.attribute(g,"burn.in",burn.in)
+  
+  g$dist <- matrix (c (0,3,2,1,1,
                    3,0,1,1,2,
                    2,3,0,1,1,
                    1,2,3,0,0,
                    1,2,3,0,0),
                 nrow=5,
                 ncol=5)
+
   
   # Will be used for
   # convergence analysis in order to compute some social utility associated
   # with a distribution of distances to the true model emerging from the 
   # distribution of approximate models.
   
+  ## Record convergence utility of the initial network
+  
+  g$init.converg.util <- converg.util(g)
   
   V(g)$behav  <- behav.dis  # Market behavior of the traders, only ZI in preliminary model
   

### 2015-03-30 set_add_param.R
I've seen you do this on other functions. I am not sure why it matters, but I changed it here for consistency.

@@ -13,6 +13,6 @@ SetAddParam <- function ( g,
   
   g <- set.graph.attribute(g,"n.seq",n.seq)
   
-  return(g)
+  g
   
 }

### 2015-03-30 README.Rmd
small edits to readme. and moved where some files are being sourced(). model does not run: problem is that in the function PopulateNet() on line 34 of file populate_net.R a variable called 'burn.in' is passed into a function but this variable is not defined as an arg to PopulateNet() or inside PopulateNet().

@@
 
-Martin Van der Linden and John Nay
-
-Vanderbilt University
+[Martin Van der Linden](https://martinvdlinden.wordpress.com/) (Dept. of Economics, Vanderbilt University) and [John Nay](https://johnnay.github.io) (School of Engineering, Vanderbilt University).
 
@@
 ```{r, eval=FALSE, include=TRUE}
-source(master.R)
+source("master.R")
 ````

### 2015-03-30 generate_model.R

@@ -1,9 +1,4 @@
 generate_model <- function ( 
-  
-  # Choose true model
-  
-  t.model = 1, # by default, the true model is the full model
-  
   # Network parameters
   n.traders  = 100,
   n.edg      = 100,
@@ -33,13 +28,14 @@ generate_model <- function (
 ){
   market.struct <- match.arg(market.struct)
   ####
-  ## Sub functions
+  ## Sub functions, 
+  # these are sourced()/created in main.R, we dont need to recreate these functions everytime that generate_model() is called
   ####
-  source("populate_net.R")
-  source("shape_net.R")
-  source("set_add_param.R")
-  source("mixing_matrix.R")
-  source("assortativity_coefficient.R")
+    source("populate_net.R")
+    source("shape_net.R")
+    source("set_add_param.R")
+    source("mixing_matrix.R")
+    source("assortativity_coefficient.R")
   
   #####
   ## Set model's parameters and create corresponding network
@@ -51,11 +47,6 @@ generate_model <- function (
                       ideo = ideo
   )
   
-  ####
-  ## Set true model of the climate
-  ####
-  
-  V(net)$t.model <- t.model
   
   #####
   ## Create links in the network according to homophily in terms of approximate model

### 2015-03-30 main.R

@@ -1,6 +1,13 @@
 
 # Construct model
 source("generate_model.R")
+# Sub functions needed inside of generate_model():
+source("populate_net.R")
+source("shape_net.R")
+source("set_add_param.R")
+source("mixing_matrix.R")
+source("assortativity_coefficient.R")
+# Construct data for model
 source("generate_data.R")
 # Operate on model
 source("form_expect.R")
@@ -13,8 +20,7 @@ source("assortativity_coefficient.R")
 
 main <- function(visu = TRUE, seg = 0.95, 
                  market.struct = c("CDA", "LMSR"), 
-                 outcome = c("segreg","converg"),
-                 market.complet  = 4,  
+                 market.complet  = 4, outcome = "segreg", 
                  n.traders = 100, n.edg = 150,
                  risk.tak = 0.0001, ideo = 10) {
   ### Market structure parameters:
@@ -24,7 +30,6 @@ main <- function(visu = TRUE, seg = 0.95,
   # more precise temperature intervals
   
   market.struct <- match.arg(market.struct)
-  outcome <- match.arg(outcome)
   
   library(igraph)
   if (visu){
@@ -33,10 +38,7 @@ main <- function(visu = TRUE, seg = 0.95,
   #####
   ## Set model's parameters and create corresponding network
   #####
-  net <- generate_model( # True model of the climate
-    t.model = 1,
-    
-    ### Network parameters
+  net <- generate_model( ### Network parameters
     n.traders  = n.traders,
     n.edg      = n.edg,
     seg        = seg, # determine initial segregation of the 
@@ -190,14 +192,6 @@ main <- function(visu = TRUE, seg = 0.95,
     net$ac.init - ac.final  
     
     #See more at: http://www.babelgraph.org/wp/?p=351#sthash.yWfBpOhv.dpuf
-  }
-  
-  if (outcome == "converg"){
-    
-    # return difference in utility of network convergence
-    
-    converg.util(net) - net$init.converg.util
-   
   }
   
 }

### 2015-03-30 generate_model.R
got the model working. burn.in is now an arg to PopulateNet() func and inside generate_model() the burn in is passed in from env above. And needed to add creation of the converge util func to the main.R file. Make sure to always run rm(list=ls()) and see that sourcing main.R and then main() works without error, ensuring that the working of the model is not dependent on anything in your global workspace, if it is, then create whatever is in your global workpace (that is needed for no erorrs) in the actual code that is run by sourcing main.R.

@@ -44,7 +44,8 @@ generate_model <- function (
   net <- PopulateNet( n.traders = n.traders,
                       risk.tak = risk.tak,
                       market.complet  = market.complet,
-                      ideo = ideo
+                      ideo = ideo, 
+                      burn.in = burn.in
   )
   
   

### 2015-03-30 main.R

@@ -3,6 +3,8 @@
 source("generate_model.R")
 # Sub functions needed inside of generate_model():
 source("populate_net.R")
+# converge.util.R wasnt sourced before, but its needed in populatenet
+source("converg_util.R")
 source("shape_net.R")
 source("set_add_param.R")
 source("mixing_matrix.R")

### 2015-03-30 master.R

@@ -1,5 +1,8 @@
 # master script for conducting ABM experiments
-rm(list=ls())
+rm(list=ls()) # make sure to always run this line of code and see that the next two 
+# lines of code work without error, ensuring that the working of the model is not 
+# dependent on anything in your global workspace, if it is, then you need to create 
+# whatever is in your global workpace in the code that is sourced in the next two lines of code.
 source("main.R")
 main(market.struct = "CDA")
 

### 2015-03-30 populate_net.R

@@ -6,14 +6,15 @@ PopulateNet <- function(
                         behav.dis    = sample(1:1, n.traders, replace = TRUE),
                         risk.tak     = 1/2,
                         ideo         = 0.5,
-                        market.complet = 10              
+                        market.complet = 10,
+                        burn.in
                       ) 
   {  
 
   # Input consistency checks (to be added)
   
   # package requirement
-  require(igraph)
+  library(igraph)
   
   ## construct empty graph
   g <- graph.empty(n=n.traders, directed = FALSE)
@@ -31,7 +32,7 @@ PopulateNet <- function(
   
   ## Assign values for the distance between approximate models. 
   
-  g <- set.graph.attribute(g,"burn.in",burn.in)
+  g <- set.graph.attribute(g, "burn.in", burn.in)
   
   g$dist <- matrix (c (0,3,2,1,1,
                    3,0,1,1,2,

### 2015-03-30 README.Rmd
readme image

@@
 
-[Martin Van der Linden](https://martinvdlinden.wordpress.com/) (Dept. of Economics, Vanderbilt University) and [John Nay](https://johnnay.github.io) (School of Engineering, Vanderbilt University).
+[Martin Van der Linden](https://martinvdlinden.wordpress.com/) (Dept. of Economics, Vanderbilt University) and [John J. Nay](https://johnnay.github.io) (School of Engineering, Vanderbilt University).
 
@@
 ````
+
+Here is a snapshot of the model's network of agents at an intialization:
+
+![alt text](figures/network.png)

### 2015-03-31 main.R
main() match.arg()

@@ -22,7 +22,8 @@ source("assortativity_coefficient.R")
 
 main <- function(visu = TRUE, seg = 0.95, 
                  market.struct = c("CDA", "LMSR"), 
-                 market.complet  = 4, outcome = "segreg", 
+                 outcome = c("segreg","converg"),
+                 market.complet  = 4,
                  n.traders = 100, n.edg = 150,
                  risk.tak = 0.0001, ideo = 10) {
   ### Market structure parameters:
@@ -32,6 +33,7 @@ main <- function(visu = TRUE, seg = 0.95,
   # more precise temperature intervals
   
   market.struct <- match.arg(market.struct)
+  outcome <- match.arg(outcome)
   
   library(igraph)
   if (visu){

### 2015-03-31 main.R
recovered from overwrite.

@@ -198,6 +198,16 @@ main <- function(visu = TRUE, seg = 0.95,
     #See more at: http://www.babelgraph.org/wp/?p=351#sthash.yWfBpOhv.dpuf
   }
   
+   if (outcome == "converg"){
+    
+       # return difference in utility of network convergence
+      
+       converg.util(net) - net$init.converg.util
+    
+       }
+  
+}
+  
 }
 
 

### 2015-03-31 sobol_sa.R
improved sobol_sa plot

@@ -100,15 +100,19 @@ correct_bias <- function(x) {
 
 plot_sobol_fo <- function(x, outcome_var = "Outcome"){
   ss <- correct_bias(x$S)
-  ggplot(ss, aes(x = var, y = x_corr)) + geom_point() + xlab("Variable") + ylab("") +
+  ggplot(ss, aes(x = var, y = x_corr)) + geom_point() +
+    xlab("Variable") + ylab("Estimated Effect") +
     ggtitle(paste("First Order Effects of Variables on", outcome_var)) +
-    geom_errorbar(aes(ymax = max_ci, ymin = min_ci), width=0.25)
+    geom_errorbar(aes(ymax = max_ci, ymin = min_ci), width=0.25) +
+    ylim(c(0,1))
 }
 
 plot_sobol_total <- function(x, outcome_var = "Outcome"){
   tt <- correct_bias(x$T)
-  ggplot(tt, aes(x = var, y = x_corr)) + geom_point() + xlab("Variable") + ylab("") +
+  ggplot(tt, aes(x = var, y = x_corr)) + geom_point() + 
+    xlab("Variable") + ylab("Estimated Effect") +
     ggtitle(paste("Total Sensitivity Effects of Variables on", outcome_var)) +
-    geom_errorbar(aes(ymax = max_ci, ymin = min_ci), width=0.25)
+    geom_errorbar(aes(ymax = max_ci, ymin = min_ci), width=0.25) + 
+    ylim(c(0,1))
 }
 

### 2015-04-01 main.R
developed sa in pkg. loading it and trying to use it. still cant get the sa to work.

@@ -20,12 +20,15 @@ source("adapt.R")
 source("mixing_matrix.R")
 source("assortativity_coefficient.R")
 
-main <- function(visu = TRUE, seg = 0.95, 
-                 market.struct = c("CDA", "LMSR"), 
-                 outcome = c("segreg","converg"),
+main <- function(seg = 0.95, 
+                 risk.tak = 0.0001,
                  market.complet  = 4,
                  n.traders = 100, n.edg = 150,
-                 risk.tak = 0.0001, ideo = 10) {
+                ideo = 10,
+                 
+                 market.struct = c("CDA", "LMSR"), 
+                 out = c("segreg","converg"),
+                 visu = TRUE) {
   ### Market structure parameters:
   # market.struct, choses between CDA and LMSR
   # market.complet is number of securities which
@@ -33,7 +36,7 @@ main <- function(visu = TRUE, seg = 0.95,
   # more precise temperature intervals
   
   market.struct <- match.arg(market.struct)
-  outcome <- match.arg(outcome)
+  out <- match.arg(out)
   
   library(igraph)
   if (visu){
@@ -100,7 +103,7 @@ main <- function(visu = TRUE, seg = 0.95,
   for (t in from:to){
     #####
     ## Traders calibrate their approximate model and determine their 
-    ## expected distribution for the future outcome
+    ## expected distribution for the future out
     #####
     net <- FormExpect(net, t, D) 
     #####
@@ -181,7 +184,7 @@ main <- function(visu = TRUE, seg = 0.95,
   # and then return one index, the one that matches the arg of main() called "outcome" 
   # outcome
   
-  if (outcome == "segreg"){
+  if (out == "segreg"){
     
     # calculate the mixing matrix
     m <- mixmat(net,'approx')
@@ -193,20 +196,20 @@ main <- function(visu = TRUE, seg = 0.95,
     # less assortative network are "better", so ac.final<net$ac.init is "good"
     # and the more positive the reported difference, the more powerful the market
     # is at breaking assortativity
-    net$ac.init - ac.final  
+    result <- net$ac.init - ac.final  
     
     #See more at: http://www.babelgraph.org/wp/?p=351#sthash.yWfBpOhv.dpuf
   }
   
-   if (outcome == "converg"){
+  if (out == "converg"){
     
-       # return difference in utility of network convergence
-      
-       converg.util(net) - net$init.converg.util
+    # return difference in utility of network convergence
     
[…]

### 2015-04-01 master.R

@@ -6,33 +6,24 @@ rm(list=ls()) # make sure to always run this line of code and see that the next 
 source("main.R")
 main(market.struct = "CDA")
 
-# param is vec length 2 created by either a stochastic metaheuristic optimization method
-# or a sampling scheme for a sens analysis
-# and main() takes in param and returns a scalar quant measure of the outcome var
-# so , TODO: decide what the important 
-# results <- main(seg = param[1], ideo = param[2], outcome = "segreg")
+devtools::install_github("JohnNay/eat", 
+                         auth_token = "08d34f040cbe8c95d89477741ceb450a9cfa42c4")
+library(eat)
+?sobol_sa
 
-
-
-
-# # TODO: create a function to put in "abm" arg of sobol_sa() func 
-# # This abm func will just be a slight modification of the existing main() func.
-# # TODO: create a list to put in "input_values" arg of sobol_sa() func 
-# # that looks like this: (just change param1 to the name of the actual param)
-# 
-# input_values <- lapply(list(param1 = NA, param2 = NA), 
-#                        function(x) list(random_function = "qunif",
-#                                         ARGS = list(min = 0, max = 1)))
+input_values <- lapply(list(seg = NA, risk.tak = NA), 
+                       function(x) list(random_function = "qunif",
+                                        ARGS = list(min = 0, max = 1)))
 # # if there are any params that are binary valued give them a binom prior distribution:
 # input_values[["param2"]] <- list(random_function = "qbinom",
 #                                  ARGS = list(size = 1, prob = 0.5))
-# 
-# 
-# source("sobol_sa.R")
-# sa_results <- sobol_sa(abm = , input_values = ,
-#                        out = "segreg", 
-#                        sample_count = 100, 
-#                        sobol_nboot = 1000, 
-#                        iterations = NULL,
-#                        parallel = TRUE)
-# 
+
+sa_results <- sobol_sa(abm = main, 
+                       input_values = input_values,
+                       out = "segreg", 
+                       sample_count = 10, 
+                       sobol_nboot = 1000, 
+                       parallel = FALSE)
+plot_sobol_fo(sa_results)
+plot_sobol_total(sa_results)
+

### 2015-04-01 README.Rmd
Small addition.
grammar
I Admire your politness but told you : I am a defender of the alphabetical order :)

@@
 
-[Martin Van der Linden](https://martinvdlinden.wordpress.com/) (Dept. of Economics, Vanderbilt University) and [John J. Nay](https://johnnay.github.io) (School of Engineering, Vanderbilt University).
+[John J. Nay](https://johnnay.github.io) (School of Engineering, Vanderbilt University)
+ and [Martin Van der Linden](https://martinvdlinden.wordpress.com/) (Dept. of Economics, Vanderbilt University).
 
@@
 ![alt text](figures/network.png)
+
+The colors of the nodes correspond to different beliefs about the climate.

### 2015-04-01 main.R
got sa working!

@@ -20,11 +20,7 @@ source("adapt.R")
 source("mixing_matrix.R")
 source("assortativity_coefficient.R")
 
-main <- function(seg = 0.95, 
-                 risk.tak = 0.0001,
-                 market.complet  = 4,
-                 n.traders = 100, n.edg = 150,
-                ideo = 10,
+main <- function(parameters,
                  
                  market.struct = c("CDA", "LMSR"), 
                  out = c("segreg","converg"),
@@ -34,6 +30,14 @@ main <- function(seg = 0.95,
   # market.complet is number of securities which
   # are traded. With higher securities, traders can trade on
   # more precise temperature intervals
+  
+  seg <- parameters[1] # = 0.95, 
+  risk.tak <- parameters[2]*0.0001 # SA creates everything in 0-1 interval, so im scaling things to the interval we want inside here
+  
+  market.complet  = 4
+  n.traders = 100 
+  n.edg = 150
+  ideo = 10
   
   market.struct <- match.arg(market.struct)
   out <- match.arg(out)

### 2015-04-01 colored.R
interact.R is the filename sourced() in the code but the file was saved as "Interact.R". colored.r had to be changed to colored.R too. this is my first run of the model on the linux workstation and it could not deal with those mis-namings.

@@ -1,7 +1,7 @@
 Colored <- function(
-                  g = net,
-                  attri = "approx"
-                ){
+  g = net,
+  attri = "approx"
+){
   V(g)$color<- get.vertex.attribute(g,attri)
   
   d <- length(unique(V(g)$color))
@@ -16,4 +16,3 @@ Colored <- function(
   
   g
 }
-

### 2015-04-01 interact.R

@@ -1,4 +1,3 @@
-
 Interact <- function (
   g   
 ){
@@ -170,4 +169,4 @@ Interact <- function (
   g
   
 }
-  
+

### 2015-04-01 master.R

@@ -4,7 +4,7 @@ rm(list=ls()) # make sure to always run this line of code and see that the next 
 # dependent on anything in your global workspace, if it is, then you need to create 
 # whatever is in your global workpace in the code that is sourced in the next two lines of code.
 source("main.R")
-main(market.struct = "CDA")
+main(parameters = c(0.5, 0.5), market.struct = "CDA")
 
 devtools::install_github("JohnNay/eat", 
                          auth_token = "08d34f040cbe8c95d89477741ceb450a9cfa42c4")

### 2015-04-02 main.R
trying to get param ranges to not cause errors

@@ -21,10 +21,9 @@ source("mixing_matrix.R")
 source("assortativity_coefficient.R")
 
 main <- function(parameters,
-                 
                  market.struct = c("CDA", "LMSR"), 
                  out = c("segreg","converg"),
-                 visu = TRUE) {
+                 visu = FALSE) {
   ### Market structure parameters:
   # market.struct, choses between CDA and LMSR
   # market.complet is number of securities which
@@ -32,10 +31,10 @@ main <- function(parameters,
   # more precise temperature intervals
   
   seg <- parameters[1] # = 0.95, 
-  risk.tak <- parameters[2]*0.0001 # SA creates everything in 0-1 interval, so im scaling things to the interval we want inside here
+  risk.tak <- parameters[2] #0.0001 # SA creates everything in 0-1 interval, so im scaling things to the interval we want inside here
   
-  market.complet  = 4
-  n.traders = 100 
+  market.complet = 4
+  n.traders = 100
   n.edg = 150
   ideo = 10
   

### 2015-04-02 master.R

@@ -4,7 +4,7 @@ rm(list=ls()) # make sure to always run this line of code and see that the next 
 # dependent on anything in your global workspace, if it is, then you need to create 
 # whatever is in your global workpace in the code that is sourced in the next two lines of code.
 source("main.R")
-main(parameters = c(0.5, 0.5), market.struct = "CDA")
+main(parameters = c(0.01, 0.01), market.struct = "CDA")
 
 devtools::install_github("JohnNay/eat", 
                          auth_token = "08d34f040cbe8c95d89477741ceb450a9cfa42c4")
@@ -21,9 +21,10 @@ input_values <- lapply(list(seg = NA, risk.tak = NA),
 sa_results <- sobol_sa(abm = main, 
                        input_values = input_values,
                        out = "segreg", 
-                       sample_count = 10, 
+                       sample_count = 30, 
                        sobol_nboot = 1000, 
-                       parallel = FALSE)
+                       parallel = TRUE,
+                       cores = 30)
 plot_sobol_fo(sa_results)
 plot_sobol_total(sa_results)
 

### 2015-04-02 main.R
if (visu) then do any plotting, o.w. no plotting

@@ -131,7 +131,8 @@ main <- function(parameters,
   
   net <- Adapt(g = net)
   
-  plot.igraph(net,vertex.label=NA,layout=layout.fruchterman.reingold, vertex.size = 7)
+  if (visu)
+    plot.igraph(net,vertex.label=NA,layout=layout.fruchterman.reingold, vertex.size = 7)
   
   ########                            ############
   ######   Subsequent trading sequences   ########
@@ -178,7 +179,8 @@ main <- function(parameters,
       
       net <- Adapt( g = net)
       
-      plot.igraph(net,vertex.label=NA,layout=layout.fruchterman.reingold, vertex.size = 7)
+      if (visu)
+        plot.igraph(net,vertex.label=NA,layout=layout.fruchterman.reingold, vertex.size = 7)
     }
   }
   

### 2015-04-02 main.R
still getting this error message when trying to do sa: "Error in { : task 1 failed - "subscript out of bounds"

@@ -22,7 +22,7 @@ source("assortativity_coefficient.R")
 
 main <- function(parameters,
                  market.struct = c("CDA", "LMSR"), 
-                 out = c("segreg","converg"),
+                 out = c("segreg", "converg"),
                  visu = FALSE) {
   ### Market structure parameters:
   # market.struct, choses between CDA and LMSR
@@ -31,10 +31,10 @@ main <- function(parameters,
   # more precise temperature intervals
   
   seg <- parameters[1] # = 0.95, 
-  risk.tak <- parameters[2] #0.0001 # SA creates everything in 0-1 interval, so im scaling things to the interval we want inside here
+  n.traders <- ifelse(round(parameters[2]*100) < 1, 1, round(parameters[2]*100))
+  risk.tak <-  0.0001 # SA creates everything in 0-1 interval, so im scaling things to the interval we want inside here
   
   market.complet = 4
-  n.traders = 100
   n.edg = 150
   ideo = 10
   

### 2015-04-02 master.R

@@ -11,7 +11,7 @@ devtools::install_github("JohnNay/eat",
 library(eat)
 ?sobol_sa
 
-input_values <- lapply(list(seg = NA, risk.tak = NA), 
+input_values <- lapply(list(seg = NA, n.traders = NA), 
                        function(x) list(random_function = "qunif",
                                         ARGS = list(min = 0, max = 1)))
 # # if there are any params that are binary valued give them a binom prior distribution:
@@ -21,10 +21,10 @@ input_values <- lapply(list(seg = NA, risk.tak = NA),
 sa_results <- sobol_sa(abm = main, 
                        input_values = input_values,
                        out = "segreg", 
-                       sample_count = 30, 
+                       sample_count = 4, 
                        sobol_nboot = 1000, 
                        parallel = TRUE,
-                       cores = 30)
+                       cores = 7)
 plot_sobol_fo(sa_results)
 plot_sobol_total(sa_results)
 

### 2015-04-02 behav.R
correct comment

@@ -22,9 +22,9 @@ Behav <- function (
   ### Determine bounds for securities
   ######
   
-  min.trad <- 1.0001 * min(D[,"temp"])  # every temperature below 1.1 * min.temp gets  
+  min.trad <- 1.0001 * min(D[,"temp"])  # every temperature below 1.0001 * min.temp gets  
   # included in a single security
-  max.trad <- 0.9999 * max(D[,"temp"])  # every temperature above 0.9 * max.temp gets  
+  max.trad <- 0.9999 * max(D[,"temp"])  # every temperature above 0.9999 * max.temp gets  
   # included in a single security
   
   #########

### 2015-04-02 generate_model.R
additional comment

@@ -3,7 +3,10 @@ generate_model <- function (
   n.traders  = 100,
   n.edg      = 100,
   seg        = 0.9, # determine initial segregation of the 
-  # network. The higher seg, the higher the initial segregation
+  # network. The higher seg, the higher the initial segregation 
+  # In the currect procedure, the probability that a link be formed is
+  #            - (1-seg)/(number of possible nodes), if the two traders have different approximate models
+  #            - 1/(number of possible nodes)      , if the two traders have the same approximate model
   
   # Market structure parameters
   market.struct   = c("CDA", "LMSR"), # chose between CDA and LMSR

### 2015-04-02 main.R
correct and add comments

@@ -55,10 +55,11 @@ main <- function(parameters,
     # network. The higher seg, the higher the initial segregation
     
     ### Behavior parameters:
-    risk.tak = risk.tak, # IN PER/10000 Determine the distribution of risk
+    risk.tak = risk.tak, # IN PERCENT. Determines the distribution of risk
     # taking behavior. The higher risk.taking, the more agent
     # will try to buy (resp. sell) lower (resp. higher) than
-    # their reservation price.
+    # their reservation price. Agent i tries to buy at (resrv_i * (1 - risk.tak_i))
+    # and tries to sell at (resrv_i * (1 + risk.tak_i))
     ideo = ideo,     # IN PERCENT. Determines the degree of "ideology" embedded
     # in the approximate models. If ideo is high, traders will not
     # revise their approximate models easily, even when faced with

### 2015-04-02 populate_net.R
The security distribution did not account for the fact that two default securities were later added. This would have caused errors for market.complet < 3. Also there would have been a mismatch between market.complet and the actual number of securities (the actual number of securities used to be market.complet +2).

@@ -69,7 +69,10 @@ PopulateNet <- function(
   ideo.dis <- sample(0:((1-ideo)*100)/100, n.traders, replace = TRUE)
   V(g)$ideo <- ideo.dis
   
-  secu.dis     = list(rep(1,market.complet))
+  secu.dis     = list(rep(1,market.complet + 2))    # The "+2" accounts for the fact
+                  # that in the market implementation in behav.r, we have 2 default
+                  # securities that cover temperature below 1.0001 * min.temp (empirical min)
+                  # and above 0.9999 * max.temp (empirical max)
   V(g)$secu   <- secu.dis   # Securities held by the traders
   
     

### 2015-04-02 main.R
working on param ranges

@@ -30,13 +30,14 @@ main <- function(parameters,
   # are traded. With higher securities, traders can trade on
   # more precise temperature intervals
   
-  seg <- parameters[1] # = 0.95, 
-  n.traders <- ifelse(round(parameters[2]*100) < 1, 1, round(parameters[2]*100))
-  risk.tak <-  0.0001 # SA creates everything in 0-1 interval, so im scaling things to the interval we want inside here
+  # SA creates everything in 0-1 interval, so im scaling things to the interval we want inside here
+  seg <- parameters[1] # (0,1) # = 0.95, 
+  ideo <- parameters[2]*100 # (0,1) but in percentage so *100
+  risk.tak <- parameters[3] 
+  market.complet <- parameters[4] 
   
-  market.complet = 4
-  n.edg = 150
-  ideo = 10
+  n.edg <- 150
+  n.traders <- 100 #ifelse(round(parameters[2]*100) < 1, 1, round(parameters[2]*100))
   
   market.struct <- match.arg(market.struct)
   out <- match.arg(out)

### 2015-06-11 master.R
finished sim and plotted results

@@ -100,11 +100,18 @@ ggplot2::ggplot(d, aes(x = convergence)) + geom_histogram(binwidth = 0.04) +
   ggtitle(paste(sample_count, 
                  "Convergence of Belief Outcomes with LHS Input Parameters")) +
   theme_bw()
-ggplot2::ggplot(d, aes(x = convergence)) + geom_histogram(binwidth = 0.04) +
-  xlim(c(-0.6, 0.6)) +
-  ylab("Count") + xlab("Convergence of Beliefs") + 
+
+d$true.model <- factor(d$true.model)
+library(plyr)
+cdat <- plyr::ddply(d, "true.model", summarize, outcome_mean = mean(convergence))
+ggplot2::ggplot(d, aes(x = convergence, fill = true.model)) + 
+  geom_density(alpha = 0.3) +
+  # xlim(c(-0.6, 0.6)) +
+  ylab("Density") + xlab("Convergence of Beliefs") + 
   ggtitle(paste(sample_count, 
                 "Convergence of Belief Outcomes with LHS Input Parameters")) +
+  geom_vline(data=cdat, aes(xintercept= outcome_mean,  color = true.model),
+             linetype="dashed", size=1) + 
   theme_bw()
 
 ##############################################################################

### 2015-06-11 paper/older/prediction_market.tex
first draft/skeleton of paper. lets contact each other when we are working on this paper so we do not lead to conflicts that need to be resolved and merged. unless otherwise noted, assume that Nay is working on the paper until further notice/emailing me about it.

\documentclass{sig-alternate}

\usepackage[utf8]{inputenc}
\usepackage{natbib}

%\usepackage[utf8]{inputenc}
%\usepackage[backend = bibtex, citestyle=authoryear]{biblatex}

%\usepackage[american]{babel}
%\usepackage{csquotes}
%\usepackage[style=apa]{biblatex}
%\DeclareLanguageMapping{american}{american-apa}

%\addbibresource{library.bib}
%\usepackage{fullpage}

%Math mode

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
%\usepackage{amsthm}
\usepackage{dsfont}
%\usepackage{mathtools}
\usepackage{xfrac}
%\usepackage{thmtools}
\usepackage{bm}

% CODE

\usepackage{courier}

\usepackage{listings}
\usepackage{color} %red, green, blue, yellow, cyan, magenta, black, white
\definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
\definecolor{mylilas}{RGB}{170,55,241}
\usepackage{marginnote}
\usepackage[linktocpage, breaklinks, colorlinks = true, linkcolor = Brown, citecolor = blue]{hyperref}
\usepackage{enumerate}

%PS_tricks

\usepackage[usenames,dvipsnames]{pstricks}
\usepackage{epsfig}
\usepackage{pst-grad} % For gradients
\usepackage{pst-plot} % For axes
\usepackage{float}

\usepackage{nameref}
\usepackage{mdframed}
\usepackage{accents}

%Geometry
%\usepackage[lmargin=3.5cm,rmargin=3.5cm,tmargin=2cm,bmargin=2cm,marginpar=3cm]{geometry}

%Proof style

%\renewcommand\qedsymbol{$\blacksquare$}

% %Margin style

\renewcommand*{\marginfont}{\footnotesize}

% %Graphic path

\graphicspath{{graphics/}}

% %Theorem Environments

%\declaretheorem [name = Theorem]{theo}
%\declaretheorem [name = Corollary]{coro}
%\declaretheorem [name = Lemma]{lem}
%\declaretheorem [name = Fairness Axiom]{ax}
%\declaretheorem [name = Preference axiom]{pref}
%\declaretheorem [name = Profile axiom]{prof}
%\declaretheorem [name = Observation]{obs}
%\declaretheorem [name = Definition]{defi}
%\declaretheorem [name = Social ordering function]{sof}
%\declaretheorem [name = Proposition, unnumbered]{prop}
%\declaretheorem [name = Example, unnumbered]{expl}
%
% %In-text hyppereff

\newcommand{\nref}{\nameref}

% %Short commands for math mode

%Style

\newcommand{\mb}[1]{\mathbf{#1}}     % bold letters in mathmode
\newcommand{\bs}[1]{\boldsymbol{#1}} % bold symbols in mathmode
\newcommand{\bb}[1]{\mathbb{#1}}     % Doppelstrichbuchstaben im Mathematikmodus

\newcommand{\mc}[1]{\mathcal{#1}}    % calligraphy style

%Sets

\newcommand{\rr}{\bb{R}} % set of reals in mathmode

%Fractions

\newcommand{\pfraca}[2]{\bigl(\frac{#1}{#2}\bigr)}    % ratios with small parantheses
\newcommand{\pfracb}[2]{\Bigl(\frac{#1}{#2}\Bigr)}    % ratios with big parantheses
\newcommand{\pfracc}[2]{\biggl(\frac{#1}{#2}\biggr)}  % ratios with bigger parantheses
\newcommand{\pfracd}[2]{\Biggl(\frac{#1}{#2}\Biggr)}  % ratios with huge parantheses

%Partial derivatives

\newcommand{\di}[1]{\frac{\partial}{\partial #1}}     % shortcut for first partial derivative
\newcommand{\dii}[1]{\frac{\partial^2}{\partial #1^2}}% shortcut for second partial derivative

%other

\newcommand{\ra}{\rightarrow}

%Convergence

\newcommand{\cd}{\overset{d}{\rightarrow}}    % convergence in distribution
\newcommand{\cp}{\overset{p}{\rightarrow}}    % convergence in probability
\newcommand{\cas}{\overset{as}{\rightarrow}}  % convergence almost surely

%Text in math environment

\newcommand{\sgn}{~\text{sgn}~}
\newcommand{\f}{~\text{if}~}
\newcommand{\frall}{~\text{for all}~}
\newcommand{\ow}{~\text{otherwise}~}
\newcommand{\nd}{~\text{and}~}
\newcommand{\thn}{~\text{then}~}
\newcommand{\id}{\coloneqq}

%Variant of variables (tilde, lower bar,...)

\newcommand{\lbar}[1]{\underline{#1\mkern-4mu}\mkern4mu }

\newcommand{\x}{\bm{x}}
\newcommand{\m}{\bm{m}}
\newcommand{\z}{\bm{z}}
\newcommand{\g}{\bm{\gamma}}
\newcommand{\h}{\bm{h}}
\newcommand{\y}{\bm{y}}
\newcommand{\e}{\bm{e}}
\newcommand{\la}{\bm{\lambda}}
\newcommand{\ce}{\bm{c}}
\renewcommand{\b}{\bm{\beta}}
\newcommand{\hb}{\hat{\b}}

\clubpenalty=10000
\widowpenalty = 10000

\usepackage[hyphenbreaks]{breakurl}
%\usepackage[hyphens]{url}

\begin{document}

	\title{Prediction Markets and Congruence of Predictive Models:
  An Agent-Based Simulation of Climate Prediction Markets}
  %\substitle{An Agent-Based Simulation of Climate Prediction Markets}

	\numberofauthors{2} 
	\author{
    \alignauthor
		John J. Nay
		\alignauthor
		Martin Van der Linden
	}

	\maketitle
	\begin{abstract}
We present an agent-based model (ABM) of a climate prediction market in which traders adapt their belief about the climate based on the monetary performance of their neighboring traders in a social network. We use the ABM to test whether climate prediction markets fosters a convergence of market participants' beliefs regarding a ``true'' climate model. We provide preliminary sensitivity analysis for the impact on the convergence of factors such as the initial belief-homophily in the social network and the degree of risk-taking of traders on the market.
	\end{abstract}

	\section{Introduction}

	The previous two decades saw a strong polarization of the climate change debate. Although the scientific consensus on the anthropogenic nature of climate change strongly increased,  the average belief about anthropogenic climate change did not evolve much within the public (see \cite{Vandenbergh2013f} for all the claims in this paragraph). In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily, indicating that the question is becoming more and more politicized and potentially disconnected from scientific evidence.

	This is worrisome given the costs of a misinformed climate policy. If anthropocentric climate change is a myth but is taken to be true, a tremendous amount of public resources will be wasted on adaptation and mitigation efforts. On the other hand, if anthropocentric climate change is wrongly consider as false, the costs of inaction would be huge. Given that effective climate policies require long-lasting measures to be taken, the importance of an accurate consensus on the issue is manifest.

	Attempts to foster such consensus face many \linebreak  social and psychological challenges (see \cite[Sections II.B to II.D]{Vandenbergh2013f} for a review). Recently, people have argued that such challenges could be met by setting up climate prediction markets (see for instance \cite{Hsu2011}, and \cite{Vandenbergh2013f}). The idea of using prediction markets to efficiently aggregate information about uncertain events has been discussed for decades (see \cite{Horn2014c} for a recent review). Prediction markets have been shown to have interesting theoretical properties (e.g. \cite{Set1998}, \cite{Hanson2012}, ) and to perform well in terms of prediction accuracy in experiments (e.g \cite{Hanson2006}, \cite{Healy2010}), agent-based models (henceforth ABM) ( e.g \cite{Klingert2012c}, \cite{Jumadinova2011a}), and in the real-world (see \cite{Wolfers2006} for a review).

However, to the best of our knowledge, the idea that prediction markets can generate a consensus \emph{on the factors} affecting  uncertain events has never been quantitatively tested. ABMs of prediction markets have been studied ( e.g. \cite{Klingert2012c}, \cite{Tseng2010a}, or\cite{Jumadinova2011a}), some of which feature communication between agents. In these models, however, beliefs about the uncertain outcomes are constructed in rather abstract ways. In particular, beliefs are never based on structural models from which agents would derive causal implications. Therefore, although convergence of beliefs could be studied in previous ABMs, these models do not investigate the convergence of the underlying explanatory models agents use for prediction (more details in section \ref{related}).

From a public policy perspective, this is a crucial point. Effective climate policy not only requires an accurate consensus on future climate \emph{outcomes}. It also requires an accurate consensus on the causal \emph{mechanisms}  influencing such outcomes. If people agree that temperature will rise, but some believe it will be due to greenhouse gases, while others believe that it will be caused by increased solar activity, inconsistent and ineffective policy will may be implemented. In this project, we investigate whether prediction markets can foster such convergence in predictive models.

The main questions we want to answer are:

\begin{enumerate}
	\item Do traders' predictive models converge to the true model, or can traders converge to a ``bad" equilibrium in which most people believe in a wrong predictive model?
	\item How is convergence to the true model affected by the parameters and assumptions of the model such as:
	\begin{enumerate}
		 \item The initial homophily of the network. A strongly homophil network has traders connect preferably with other traders having the same beliefs. Given the political polarization of beliefs in climate change, this homophily is likely to be a feature of a climate change belief network. One may therefore wonder under which conditions of initial homophily could prediction market decluster the belief networks.
		 \item The timing of the market. \cite{Vandenbergh2013f} have argued that prediction market should trade securities based on long term outcomes in order to guarantee that the market provides information about climate rather than weather. However, long term trading might slow down convergence. For a trader $i$ to change her predictive model and adopt the model of another trader $j$, $i$ must be convinced  that $j$'s model is better. If $i$ relies on the market performances of $j$ to evaluate $j$'s models, $i$ will have to wait that securities be realized before she adapt their explanatory model. If the securities are based on long term outcomes, then traders will not adapt their model for before a while. Therefore, there may be a trade-off between speed and accuracy. 
		 \item The model assumed to be the true model of the climate. An important argument in favor of a climate prediction market is that it is supposed to be an apartisan proposal. Defenders of climate prediction markets argue that climate prediction markets will foster convergence to the best approximate model, whatever the actual true model is. If climate change is not anthropogenic -- so the argument goes, traders will converge to non-anthropogenic in pretty much the same way as they would converge to anthropogenic models if the climate is truly influenced by human activities. 
	\end{enumerate}
\end{enumerate}

In this report, we provide preliminary results regarding question 1 and 2.a), together with sensitivity analyses for other parameters of the model. See section \ref{results} for important caveats regarding these results. Despite these caveats, our first exploratory investigation raise a couple of interesting questions.

First, we observe that convergence to the true model is by no means inevitable. In the model, traders adapt their beliefs based on their neighbors monetary performance on the market (more in Section \ref{implem}). Therefore, one might expect that traders with the right model of the climate will perform better and  convince their (social network) neighbors to believe in the true model. We find that this is not always the case. On \url{https://www.youtube.com/watch?v=MS4YfhTU-WY}, one can see the evolution of the belief network for some particular values of the parameter. In this case, the green and red nodes correspond to traders believeing in predictive models which are qualitatively very far from the true model (see below). One sees that these models are progressively eliminated from the network. On \url{https://www.youtube.com/watch?v=aDB7UY7bDGE} however, initial homophily is increased and the network is sufficiently clustered at the outset to prevent convergence. In Section \ref{results} we provide further evidence on the possibility of non-convergence.  The reason of such non-convergence and the precise conditions on parameters which favor convergence are an open question that we plan to tackle in the near future.

% So far, we have been able to construct a minimal model of a prediction market. In  the model, traders update their explanatory model of the climate based on the model of their neighbors in a social network. If a trader believe in model $x$ and is surrounded by traders believing in model $y$ who happen to do much better then her on the market, then there is a probability that $x$ will change her explanatory model for the model of trader $x$. 
% 
% At this point the model is not complete. In particular, it features a single market mechanism known as the continuous double auction ( CDA). We plan on implementing at least a second market mechanism known as the logarithmic market scoring rule (LMSR, \cite{Hanson2012}) in order to compare the performance of both mechanism in terms of prediction accuracy and explanatory model convergence. We also need to work on relating some of the parameters with the theoretical literature on network analysis and decision making.\footnote{For instance, we have a parameter which influence the initial level of clustering in the network. We need to work on relating the value of this coefficient with commonly used clustering measures in the network literature.}

  \section{Related Work}
	\label{related}

	\cite{Tseng2010a} study an ABM of a continuous double auction model with various types of market strategies (including two variants of the zero-intelligence agents). They compare the behavior of their simulation with data from a prediction market for the outcome of political elections. They find that despite their simplicity, zero-intelligence agents capture some salient features of real market data. \cite{Klingert2012c} reach similar conclusions with respect to experimental data from  \cite{Hanson2006}. In their simulation, \cite{Klingert2012c} compare continuous double auctions with logarithmic market scoring rules and find that the latter generally outperform continuous double auctions in terms of prediction accuracy. In some variants of the two last models, agents' strategies feature learning, but these are rather basic and unrelated to a formal predictive model.

	\cite{Jumadinova2011a} conducted research more similar to our project, as they explicitly include information about uncertain events in their model. The paper studies a continuous double auction model in which agents update their beliefs about the uncertain events based on information on the event. New beliefs are modeled as a weighted sum of the last period's beliefs and market prices. The agent's information set changes every period. The better the information set, the more likely the agent is to put higher weights on last period's prices when revising their beliefs. Not much structure is imposed on the information set and the way it is used to generate the weights in the update function. Therefore, the model does not allow for investigating the convergence of predictive models. \cite{Ontanon2009} study the effect of deliberation on a simulated prediction market. They assume that agents use Case-Based Reasoning \citep{Aamodt1994} to debate about  uncertain outcomes with their neighbors in a social network.  Although the model could in theory be used to assess convergence of predictive models by looking at the arguments agents adopt after the deliberation, this is not done by the authors. Also, the way beliefs are formed and revised through argumentation is rather abstract and might not be ideal to study climate prediction markets.

	\section{Model}

	\label{model}

	In its current form, the ABM is composed of the following elements.

	\begin{description}
		\setlength\itemsep{0em}
		\item [A simple ``true" model $F$ of world temperatures.] The model is an additive autoregressive function and is taken  from \cite{Sumner2008b} (it is a simplified version of  RICE99 from \cite{Nordhaus}).
		\begin{align}
		\label{true}
		\begin{split}
				T_t ~ =~  \lambda T_{t-1} + \beta \underbrace{\big[ \kappa GHG_{t-1} - \alpha P_t^{1/2} + \delta Y_t  \big]}_{\equiv GHG_t} \\
				- \rho \underbrace{\big[ \omega G_{t-1} + \theta X_{t} \big]}_{\equiv G_t} +  \epsilon_t, 
		\end{split}
		\end{align}
		where
		\begin{itemize}
			\item $T_t$ is the temperature at time $t$,
			\item $GHG_t$ are greenhouse gas emissions at time $t$,
			\item $P_t$ are spendings in policies alleviating climate change at $t$,
			\item $Y_t$ is time $t$'s world Gross Domestic Product (GDP),
			\item $G_{t}$ is the effect in period $t$ of geo-ingeneering produced before $t$. 
			\item $X_t$ are spending in geo-ingeneering in period  $t$.
		\end{itemize}

		\item[ Beliefs.]  Traders use $5$ models to  forecast future temperatures in order to trade on the prediction market. These $5$ models are interpreted as the trader's beliefs about the true climate model.  Again, these models are taken from \cite{Sumner2008b}). They are meant to represent pervasive positions on climate change in the public debate. Four of these models are approximations of the true model in \eqref{true} described below, and one of them is the true model itself.\footnote{Thus some traders use the true model to make predictions. This does not mean that traders using the true model make perfectly accurate predictions. Although these traders believe in the correct \emph{functional form} of the model, they still need to calibrate it based on limited noisy data. Therefore, the values they use for  the \emph{parameters} of the model will typically still be off.}

		\begin{enumerate}
			\setlength\itemsep{0em}
			\item Anthropogenic climate change is a myth ($\beta = \rho = 0$),
			\begin{align}
			T^1_t = \lambda T_{t-1} +  \epsilon_t.
			\end{align}
			\item Anthropogenic climate change is real but policies and geoengineering are ineffective ($\omega = \rho = 0$),
			\begin{align}
			T^2_t & = \lambda T_{t-1} + \beta \big[ \kappa GHG_{t-1}  + \delta Y_t  \big] +  \epsilon_t.
			\end{align}
			\item Anthropogenic climate change is real, policies are effective, but not geoengineering ($\rho = 0$),
			\begin{align}
			T^3_t = \lambda T_{t-1} + \beta \big[ \kappa GHG_{t-1} - \alpha P_t^{1/2} + \delta Y_t  \big] +  \epsilon_t.
			\end{align}
			\item Anthropogenic climate change is real, policies are ineffective, but geoengineering works($\alpha = 0$),
			\begin{align}
			\begin{split}
		T^4_t = \lambda T_{t-1} + \beta \big[ \kappa GHG_{t-1} + \delta Y_t  \big]\\ - \rho \big[ \omega G_{t-1} + \theta X_{t} \big] +  \epsilon_t.
			\end{split}
			\end{align}

		\end{enumerate}

		\item[ Traders.] Traders are endowed with some initial money in an Experimental Currency Unit (ECU), and an initial approximate model. Approximate models are distributed at random among the traders with each model having the same probability of being assigned to any trader.  Traders use the distribution of temperatures obtained from their approximate models to determine their reservation price for different securities.

		Securities pay 1 ECU  if the actual temperature at some time $t^*$ falls between a certain range. Traders are assumed to be \emph{risk-neutral expected utility maximizers}. Therefore, their reservation price for a security $\tilde{s}$ is simply their assessment of the probability that the temperature at time $t^*$ falls in the range covered by $\tilde{s}$.

		  Based on their reservation price, traders behave as ``\emph{zero-intelligence}" agents \citep{Gode1993}, that is they sell at a random price above their reservation, and buy at a random price below their reservation. These strategies are simple but have been shown to provide good approximations of the behavior of prediction markets (see for instance \cite{Klingert2012c}), and financial markets more broadly.

		\item[Continuous-Double Auction (CDA) market structure.] Continuous-Double auctions ( or some variants thereof) are common \emph{market mechanisms}, i.e. procedures to match buy and sell orders. CDA are notably used on large stock markets (see for instance \cite{Tseng2010a}). Our model features a particular version of CDA (more in section \ref{implem}).

		\item[A social network.] Traders are part of a social network. Every time a security is realized, each trader $i$ looks at the performance of their neighbors in the network. If one of $i$'s neighbors, say $j$, is richer than $i$, $i$ interprets this has an indication that $j$ has a better approximate model. Then $i$ considers adopting $j$'s approximate model.\footnote{Traders start with the same initial amount of money, so differences in money among traders can only come from  traders' interactions through the market.}

		An example of such social network is depicted in Figure \ref{socnet}. This corresponds to the initial social network for some of our simulations. The color of the nodes identify the approximate model of the trader. Notice the belief-homophily of the initial network which can be seen by the high clustering of traders with the same approximate model. Although traders can change their approximate model (i.e. colors can change in the graph), the connections between traders do not change as the market unfolds (i.e. the edges are fixed).
	\end{description}

	\begin{figure}
		\begin{center}
		\includegraphics[scale = 0.5]{network.png}
		\end{center}
		\caption{An example of the initial state of a belief network \label{socnet}}
	\end{figure}

	\subsubsection*{Timing}

	The time periods $t$ are grouped in trading \emph{sequences}. In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence. For instance, the third trading sequence might start in period $t=300$ and end in period $t = 399$. In this case, a security traded in the third sequence pay 1 ECU if the temperature at $t=400$ falls into the range of temperatures covered by the security.

	   At each time $t$, traders are assumed to know the past value of the temperature $T_{0:t}$ and the vector of past explanatory variables $Z_{0:t}$. In a sequence finishing at time $t^*$,  traders also have common knowledge of $Z_{t:t^*}$, the future values of explanatory variables up to $t^*$. However, at any $t$, traders do not know the value of any future temperatures. In particular, in a given sequence, traders do not know the value of $T_{t^*}$. Traders can only predict $T_{t^*}$ using their approximate model and their knowledge of $Z_{t:t^*}$. Notice that because $Z_{t:t^*}$ is common knowledge, in each period $t$ every trader $i$ with the same approximate model forms the same expectation 
	   \begin{align*}
	  \mu_{ti} \equiv E_t( T_{t^*} ~|~ T_{0:t}, Z_{t:t^*},  \text{ approximate model of } i \text { at } t).
	   \end{align*}
	   We assume that the probability density that trader $i$ associate with $T_{t^*} = \tilde{T}$ at any $t$ is
	   \begin{align*}
	   f_{it}(T) = \mu_{ti} + N(0, \sigma_T),
	   \end{align*}

	   where $N$ stands for a normal distribution and $\sigma_T$ is a fixed number determined by the true variance of temperatures.

	    At each time $t$, traders
	\begin{itemize}
		\setlength\itemsep{0em}
		\item recalibrate their approximate model based on the new set of past data available at $t$,
		\item compute their belief-density for $T_{t^*}$ and use it to determine the expected value they attached to each security,
		\item and trade on the market based on their market strategy at $t$.
	\end{itemize}

	At $t^*$, when the sequence ends, there is only one security $s^*$ associated with a range of temperatures including the actual $T_{t^*}$. Thus,
	\begin{itemize}
			\setlength\itemsep{0em}
		\item traders receive 1 ECU per unit of $s^*$ they own, and
		\item  consider adapting their approximate model as described above.
	\end{itemize}

	\section{Implementation}
	\label{implem}
	The model is implemented in the \texttt{R} programming language,  using the \texttt{igraph} package to model the social network. Each trader in the model is a node of an \texttt{igraph} network. Because the model uses time-series data and  traders calibrate statistical models based on these data, \texttt{R}, primarily used for statistical computing, is particularly convenient.

	 Traders calibrate their approximate models based on past-data using Pooled Ordinary Least Squares. This is questionable from a statistical point of view, but provides an approximation of the way traders may form their forecast on actual prediction markets. Because the model is in \texttt{R}, it is also very easy to change the estimation method or model to virtually any statistical model, and the effect of altering models may easily be studied in the future.

	 In each period $\tilde{t}$, our version of the CDA works as follows.

	\begin{enumerate}
			\setlength\itemsep{0em}
		\item Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
		\item Every trader $i$ also chooses at random a security $s_i^S$ she will try to sell among the securities she owns a positive amount of (if any).
		\item Traders then decide of their selling $p_i^S$ and buying price $p_i^S$. They do so based on their believed-density of temperatures at the end of the sequence, and following the zero-intelligence rule described above.
		\item Traders go to the market one at the time, in a random order.
		\item When $i$ comes to the market, she place \emph{limit orders} in the order book. These orders  specify that $i$ is willing to buy $s_i^B$ at  any price below $p_i^B$, and to sell $s_i^S$ at any price above $p_i^S$.
		\item The market maker then tries to match $i$'s orders with some order which was put in the order book \emph{before} $i$ came to the market.
		\item If there are  outstanding sell offers for $s_i^B$ at price $p$ lower than $p_i^B$, then a trade is concluded. Trader $i$ buys one unit from the sellers who sells at the \emph{highest} price below $p_i^B$, and the sell and buy offers are removed from the order book.
		\item If there are  outstanding buy offers for $s_i^S$ at price $p$ higher than $p_i^S$, then a trade is concluded. Trader $i$ sells one unit to the buyer who buys at the \emph{highest} price above $p_i^S$, and the  sell and buy offers are removed from the order book.
		\item Whenever all traders have come to the market, any remaining outstanding offer is removed from the order book, and the trading period is concluded. 
	\end{enumerate}

The model depends on the following parameters. 
\begin{description}
	\setlength\itemsep{0em}
	\item[Network parameters.]~
	\begin{description}
			\setlength\itemsep{0em}
		\item [\texttt{n.traders}:] the number of traders.
		\item [\texttt{n.edg} :] the number of edges in the social network. This number is fixed throughout the experiments.   
		\item [\texttt{seg} :] determines the initial degree of homophily in the network. The higher \texttt{seg}, the higher the initial homophily.\footnote{ When constructing the \texttt{n.edg} edges of the network, the probability that a link between two traders be formed depends on whether the traders share the same approximate model in the following way
			\begin{align*}
			\begin{cases}
			\frac{(1-\texttt{seg})}{\texttt{n.edg}},  &\text{if the traders have different approximate models}\\
			\frac{1}{\texttt{n.edg}},  & \text{otherwise}
			\end{cases}
			\end{align*}}
	\end{description}

\item[Market structure parameter.]~

		\begin{description}
			\setlength\itemsep{0em}
			\item [\texttt{market.complet}.] Determines market's completeness, i.e. the number of securities which can be traded. With more securities, the interval of temperatures corresponding to each security is smaller. Traders can then trade on more precise temperature intervals. Higher values of \texttt{market.complet} may, however, reduce the number of exchanges. Because traders pick the securities they buy and sell at random, the probability that a match between sellers and buyers is found is lower for higher values of \texttt{market.complet}.
		\end{description}

\item[Behavioral parameters.]~

\begin{description}
	\setlength\itemsep{0em}
	\item [\texttt{risk.tak}.] Determines the distribution of risk taking behavior. The higher \texttt{risk.tak}, the more traders	will try to buy (resp. sell) lower (resp. higher) than their reservation price. Formally, at time $t$, trader $i$ picks her buying (resp. selling) prices randomly in the interval  $[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it}  * (1 - \texttt{risk.tak}_i)]$ (resp. $[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it} * (1  + \texttt{risk.tak}_i)]$), where $\texttt{reserv}_{it}$ is $i$'s reservation price at time $t$ for the security $i$ picked to buy (resp. sell).
	\item [\texttt{ideo}.]  Determines the degree of ``ideology" of traders. If \texttt{ideo} is high, traders will not
	 revise their approximate models easily, even when faced with
	 strong evidence that their neighbors are doing better than them.\footnote{ For each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$. The value of $d_{i}$ is the probability that $i$ adopts one of her neighbors' approximate model if this neighbor is doing better than $i$ at the end of the sequence (in monetary terms).}
\end{description}

	\item[Timing parameters.]~

	\begin{description}
			\setlength\itemsep{0em}
		\item [\texttt{burn.in}.] The number of burn-in periods in which no securities are traded (necessary to allow approximate models to be estimated in the first period).
		\item [\texttt{n.seq}.] Number of trading sequences.
		\item [\texttt{horizon}.] Number of trading periods in each trading sequence.
	\end{description}

\end{description}

	\section{Results}

	\label{results}

In this section, we present preliminary sensitivity analyses and exploratory results based on the above model.  In general, we will be interested in the fraction of traders who believe in the true model. For sensitivity analysis, the relevant outcome will be the difference between the fraction of traders who believe in the true model at the beginning of the experiment, and the fraction who believe in the true model at the end of all trading sequences.

	\begin{figure*}
		\begin{center}
			\includegraphics[scale = 0.5]{sobol_sa.png}
			\caption{ Effect of parameters on the difference between the initial and final fraction of traders who believe in the true model \label{nondirect}}
		\end{center}
	\end{figure*}

Before presenting these results, let us mention some important caveat. First, the model currently uses artificial data. The values of the explanatory variables are in not connected with the observed past values of greenhouse gas concentrations, GDP, etc. As a consequence, the results should be viewed as no more than a proof of concept, and no conclusions on actual {climate} prediction markets should be drawn from them. A second reason to be cautious about the following results is that they rely on varying only a handful of the model's parameters, namely
\begin{itemize}
	\item \texttt{ideo}, 
	\item \texttt{market.complet}, 
	\item \texttt{risk.tak}, and
	\item  \texttt{seg}.
\end{itemize}

Other parameters are fixed to the following values.

\begin{itemize}
	\item \texttt{n.traders} = 100,
	\item \texttt{n.edg} = 150,
	\item \texttt{burn.in} = 100,
	\item \texttt{n.seq} = 20, and
	\item \texttt{horizon} = 10.
\end{itemize}

We first present two sensitivity analyses for the four parameters above. Both sensitivity analyses are based on a Latin Hypercube Sampling of the four parameters, with uniform prior distributions. The support for the distributions are:

\begin{itemize}
	\item \texttt{ideo} : $(0,1)$
	\item \texttt{market.complet} : $ \{1,\dots,1000\}$,  
	\item \texttt{risk.tak}  : $(0,1)$, and
	\item  \texttt{seg} : $(0,1)$.
\end{itemize}

	The first analysis is similar to an ANOVA, and is based on  variance decomposition below \cite[Section 2.3.1]{Confalonieri2010}. Let the four varied parameters be denoted $x_i$ for $i =1,\dots,4$, and the difference between the final and the initial fraction of traders who believe in the true model be $Y$. Then we have
	\begin{align}
	\label{variance}
	V(Y) = \sum_{i=1}^4 \Big[  D_i + 
	\sum_{i \leq j \leq 4}^4 \big[ D_{ij} + \dots + 
	D_{1234} \big] \Big],
	\end{align}

	where $D_i = V(E[Y|x_i])$, $D_{ij} = V(E[Y|x_i,x_j]) - D_i - D_j$ and so on. Dividing both sides of \eqref{variance} by $V(Y)$ we obtain a normalized sum. The first terms of this normalized sum are
	\begin{align}
	S_i = \frac{V(E[Y|x_i])}{V(Y)},
	\end{align}
	which are called the first-order effects (of $i$). First-order effect $S_i$ represents the proportion of the variance which is explained by changes in $x_i$ once the effect of variations in the other $x_j$ ( $j\neq i$) has been accounted for. In Figure \ref{nondirect},  the value of $S_i$ is represented in red for the four parameters, together with their respective confidence intervals. One can see that none of the parameters' first-order effects is significant.

	\begin{figure*}
		\begin{center}
			\includegraphics[scale = 0.5]{pc_sa.png}
			\caption{Partial rank correlation coefficients and confidence interval \label{prcc}}
		\end{center}
	\end{figure*}

	 Depicted in blue in Figure \ref{nondirect} are the total effects of the parameters. For each parameter $x_i$, the total effect corresponds to the sum of the net impacts of every possible combination of $x_i$ with other parameters. Formally, for each $x_i$ the blue dot and confidence interval in Figure \ref{nondirect} corresponds to
	 \begin{align}
	 \begin{split}
ST_i =  S_i  + \sum_{j\neq i}  S_{ij} + \dots +  S_{1234} 
 \end{split},
	 \end{align}

	where  $S_{ij} = \sfrac{ V(E[Y|x_i,x_j]) - \allowbreak V(E[Y|x_i])- V(E[Y|x_j])}{V(Y)}$ and the other terms are defined similarly. Figure \ref{nondirect} indicates that most the effect of the parameters happens through interactions with other parameters.

	\begin{figure}
		\includegraphics[scale = 0.4]{Rplot01.png}
		\includegraphics[scale = 0.4]{Rplot02.png}
		\includegraphics[scale = 0.4]{Rplot03.png}
		\caption{Evolution of the fraction of traders who believe in the true model (y-axis) as a function of the number of trading sequences (x-axis) for 30 random draws of the parameters.\label{example}}				
	\end{figure}

	The variance analysis does not provide information on the sign of the first-order effects. To investigate signed-effect, we turn to a second sensitivity analysis known as partial \emph{rank} correlation coefficient \citep{Marino2008}. Partial correlation computes the linear relationship between the part of the variation of $x_i$ and $Y$ that are linearly independent of other $x_j$  ($j \neq i$). First, we compute predicted value $\hat{x}_i$ and $\hat{Y}$ based on an OLS regression of $x_i$ and $Y$ on $x_j$  ($j\neq i$)
	\begin{align}
	\begin{split}
	\hat{x}_i = c^0_{OLS} + \sum_{j \neq i} c^j_{OLS} x_j,\\
	\hat{Y} = b^0_{OLS} + \sum_{j \neq i} b^j_{OLS} x_j.
	\end{split}
	\end{align}
	Then we compute the coefficient of correlation between $(x_i - \hat{x}_i)$ and $(\hat{Y} - \hat{Y})$.

	The only difference with the partial correlation and partial \emph{rank} correlation (which we use here), is that the dependent variable $Y$ is first ranked-transformed in order to foster more uniform residuals.\footnote{That is every $Y_s$ in the data set is replace by a number $f(Y_s) \equiv |\{Y_r~|~ Y_s > Y_r\}|$, where $Y_r$ are the other observed values of $Y$ in the data set.} Such transformation preserves the sign of the correlation coefficient. However, because data are transformed, we loose the meaning of the coefficient's magnitude with respect to the original data. The partial rank correlation coefficients are reported in Figure \ref{prcc}.

	 It is surprising that the net effect of \texttt{seg} on $Y$ is positive. This suggests that a higher initial belief-homophily in the network increases (or reduces the decrease in) the adoption of the true model. Intuitively, traders endowed with the correct explanatory model should be better at predicting climate outcomes. Therefore, if they are connected to traders with different approximate models, one could expect that these other traders will eventually adopt the true model. If that were true, the more traders endowed with the true model are disseminated across the network (i.e. the smaller \texttt{seg}) the better the result should be in terms of adoptions of the true model. The results  in Figure \ref{prcc} indicates that, if at all present, this kind of dynamics are not sufficient to explain changes in the adoption of the true model.

	Exploratory investigations also cast a doubt on the ability of the modeled prediction market to foster convergence to the true model. Figure \ref{example} contains plots for the evolution through trading sequences of the fraction of traders who believe in the true model. Each line corresponds to a different random value for the four parameters. The figure show a wide variety of dynamics and indicate that under many parameter configurations, the fraction of traders who believe in the true model significantly reduces (sometimes almost completely).

	Finally, it appears that the current version of the model is too random. Figure \ref{random} represents the evolution of $Y$ for 10 different runs of the model \emph{using the same value for the parameters}, namely

	\begin{itemize}
		\item \texttt{ideo} = 0.5,
		\item \texttt{market.complet} = 100,
		\item \texttt{risk.tak} and = 0.02, and
		\item  \texttt{seg} = 0.5.
	\end{itemize}

	As can be seen from the figure, a lot of randomness is built into the model independently of changes in the parameters. For instance, notice that the initial fraction of agents who believe in the true models varies from one run of the model to the other.  The probability that a trader starts with the true model is fixed at $\sfrac{1}{5}$. But only 100 traders are assigned an initial approximate models which explains such variability. Other randomness built into the model includes the distribution of the risk and ideology parameters $\texttt{risk.tak}_i$ and $\texttt{ideo}_i$.\footnote{Currently, $\texttt{risk.tak}_i$ is drawn at random from $[0, \texttt{risk.tak}]$. Thus, increasing \texttt{risk.tak} increases the chance that traders have higher $\texttt{risk.tak}_i$, but there is a significant amount of randomness in the assignment of the $\texttt{risk.tak}_i$. The same is true for $\texttt{ideo}_i$.} This might partly explain why we do not get clearer results in the sensitivity analysis.

	\section{Conclusion}

	In this report, we described a minimal ABM of a climate prediction market in which traders adapt their belief of the climate based on the monetary performance of their neighboring traders in a social network. The market mechanism is a variant of the famous continuous double-auction, and traders are risk-neutral expected utility maximizers who behave on the market as zero-intelligence agents \citep{Gode1993}.
		\section{Related Work}
	Future developments involve :

	\begin{itemize}
		\item Reducing the randomness built in the model and simplify the model in order to obtain clearer results. One way of simplifying the model would be to reduce the number of approximate models (e.g. anthropogenic climate change is a myth vs. anthropogenic climate change is true).
		\item Implementing a second market rule (LMSR, \cite{Hanson2012}).
		\item Using actual time-series data for the temperature and the explanatory variables. At this stage, the data used in the model are randomly generated time series with no connection to actual temperatures, or to observed values of the explanatory variables.
		\item Varying more parameters. In particular, changes in timing parameters should be studied in order to tackle the issue of optimal timing in climate prediction markets.
		\item Studying the model under different assumption about the true model in order to check whether climate prediction markets are an apartisan proposal.
	\end{itemize}

	\begin{figure}
		\includegraphics[scale = 0.4]{random.png}
		\caption{Evolution of the fraction of traders who believe in the true model (y-axis) as a function of the number of trading sequences (x-axis) for 10 different run of the model with \emph{fixed} values of the parameters ( \texttt{ideo} = 0.5,
			 \texttt{market.complet} = 100,
			 \texttt{risk.tak} and = 0.02, and
			  \texttt{seg} = 0.5.). \label{random}}
	\end{figure}

	\bibliography{library}
	\bibliographystyle{apalike}

\end{document}

### 2015-06-11 prediction_market.Rnw
first draft/skeleton of paper. lets contact each other when we are working on this paper so we do not lead to conflicts that need to be resolved and merged. unless otherwise noted, assume that Nay is working on the paper until further notice/emailing me about it.

\documentclass{article}

\begin{document}

\title{Climate Prediction Markets and Convergence of Predictive Models: An Agent-based Simulation Analysis}
\author{not sure about the author order}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abstract}
We present an agent-based model of a climate prediction market in which traders adapt their belief about the climate based on the monetary performance of their neighboring traders in a social network. We use the model to test whether climate prediction markets foster a convergence of market participants' beliefs regarding a ``true'' climate model by conducting a sensitivity analysis of the impact of a variety of factors. We find that only the imposed ``true model'' itself significantly affects the convergence of beliefs. If the true model from which the temperature data is generated is auto-regressive, with no human-induced effect, there is, on average. no convergence of traders' models due to participation in the prediction market. On other hand, if anthropogenic climate change is real, traders' beliefs are more likely than not to converge toward the correct model of the world after participation in the market.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}

The previous two decades saw a strong polarization of the climate change debate. Although the scientific consensus on the anthropogenic nature of climate change strongly increased, the average belief about anthropogenic climate change did not evolve much within the public (see \cite{Vandenbergh2013f} for all the claims in this paragraph). In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily, indicating that the question is becoming increasingly politicized and potentially disconnected from scientific evidence.

This is troubling given the costs of a misinformed climate policy. If anthropocentric climate change is a myth but is taken to be true, a tremendous amount of public resources will be wasted on adaptation and mitigation efforts. On the other hand, if climate change is human induced but not recognized as so, the costs of inaction would be huge. Given that effective climate policies require long-lasting measures to be taken, the importance of an accurate consensus on the issue is manifest.

Attempts to foster such consensus face many social and psychological challenges. Recently, people have argued that such challenges could be met by setting up climate prediction markets (see for instance \cite{Hsu2011}, and \cite{Vandenbergh2013f}). The idea of using prediction markets to efficiently aggregate information about uncertain events has been discussed for decades (see \cite{Horn2014c} for a recent review). Prediction markets have been shown to have interesting theoretical properties (e.g. \cite{Set1998}, \cite{Hanson2012}) and to perform well in terms of prediction accuracy in experiments (e.g \cite{Hanson2006}, \cite{Healy2010}), agent-based models (henceforth ABM) ( e.g \cite{Klingert2012c}, \cite{Jumadinova2011a}), and in the real-world (see \cite{Wolfers2006} for a review).

% NOTE: the follwing two paragraphs is some good lit reivew, but whether it is included in the main article text depends on the potential journal we are submitting to. journals like science and nature do not want lit review, or claims to priority. So, do not delete these following two paragraphs that are commented out until we are accepted into a journal that doesnt want this.

% \cite{Tseng2010a} study an ABM of a continuous double auction model with various types of market strategies (including two variants of the zero-intelligence agents). They compare the behavior of their simulation with data from a prediction market for the outcome of political elections. They find that despite their simplicity, zero-intelligence agents capture some salient features of real market data. \cite{Klingert2012c} reach similar conclusions with respect to experimental data from  \cite{Hanson2006}. In their simulation, \cite{Klingert2012c} compare continuous double auctions with logarithmic market scoring rules and find that the latter generally outperform continuous double auctions in terms of prediction accuracy. In some variants of the two last models, agents' strategies feature learning, but these are rather basic and unrelated to a formal predictive model. 
% 	
% \cite{Jumadinova2011a} conducted research more similar to our project, as they explicitly include information about uncertain events in their model. The paper studies a continuous double auction model in which agents update their beliefs about the uncertain events based on information on the event. New beliefs are modeled as a weighted sum of the last period's beliefs and market prices. The agent's information set changes every period. The better the information set, the more likely the agent is to put higher weights on last period's prices when revising their beliefs. Not much structure is imposed on the information set and the way it is used to generate the weights in the update function. Therefore, the model does not allow for investigating the convergence of predictive models. \cite{Ontanon2009} study the effect of deliberation on a simulated prediction market. They assume that agents use Case-Based Reasoning \cite{Aamodt1994} to debate about  uncertain outcomes with their neighbors in a social network.  Although the model could in theory be used to assess convergence of predictive models by looking at the arguments agents adopt after the deliberation, this is not done by the authors. Also, the way beliefs are formed and revised through argumentation is rather abstract and might not be ideal to study climate prediction markets.

However, to the best of our knowledge, the idea that prediction markets can generate a consensus \emph{on the factors} affecting  uncertain events has never been quantitatively tested. ABMs of prediction markets have been studied ( e.g. \cite{Klingert2012c}, \cite{Tseng2010a}, or\cite{Jumadinova2011a}), some of which feature communication between agents. In these models, however, beliefs about the uncertain outcomes are constructed in rather abstract ways. In particular, beliefs are not based on structural models from which agents would derive causal implications. Therefore, although convergence of beliefs could be studied in previous ABMs, these models do not investigate the convergence of the underlying explanatory models agents use for prediction.

From a public policy perspective, this is a crucial point. Effective climate policy not only requires an accurate consensus on future climate \emph{outcomes}. It also requires an accurate consensus on the causal \emph{mechanisms}  influencing such outcomes. If people agree that temperature will rise, but some believe it will be due to greenhouse gases, while others believe that it will be caused by increased solar activity, inconsistent and ineffective policy will may be implemented. Our model is designed to investigate whether prediction markets can foster such convergence in predictive models.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Model}

TODO: Martin use this text below (it is commented out) from the original document and flesh it out for the Supplementary Material and then write a couple paragraph summary and the key equations to put in this main article right here.
% 	\section{Model}
% 	
% 	\label{model}
% 	
% 	In its current form, the ABM is composed of the following elements.
% 	
% 	\begin{description}
% 		\setlength\itemsep{0em}
% 		\item [A simple ``true" model $F$ of world temperatures.] The model is an additive autoregressive function and is taken  from \cite{Sumner2008b} (it is a simplified version of  RICE99 from \cite{Nordhaus}).
% 		\begin{align}
% 		\label{true}
% 		\begin{split}
% 				T_t ~ =~  \lambda T_{t-1} + \beta \underbrace{\big[ \kappa GHG_{t-1} - \alpha P_t^{1/2} + \delta Y_t  \big]}_{\equiv GHG_t} \\
% 				- \rho \underbrace{\big[ \omega G_{t-1} + \theta X_{t} \big]}_{\equiv G_t} +  \epsilon_t, 
% 		\end{split}
% 		\end{align}
% 		where
% 		\begin{itemize}
% 			\item $T_t$ is the temperature at time $t$,
% 			\item $GHG_t$ are greenhouse gas emissions at time $t$,
% 			\item $P_t$ are spendings in policies alleviating climate change at $t$,
% 			\item $Y_t$ is time $t$'s world Gross Domestic Product (GDP),
% 			\item $G_{t}$ is the effect in period $t$ of geo-ingeneering produced before $t$. 
% 			\item $X_t$ are spending in geo-ingeneering in period  $t$.
% 		\end{itemize}
% 		
% 		
% 		\item[ Beliefs.]  Traders use $5$ models to  forecast future temperatures in order to trade on the prediction market. These $5$ models are interpreted as the trader's beliefs about the true climate model.  Again, these models are taken from \cite{Sumner2008b}). They are meant to represent pervasive positions on climate change in the public debate. Four of these models are approximations of the true model in \eqref{true} described below, and one of them is the true model itself.\footnote{Thus some traders use the true model to make predictions. This does not mean that traders using the true model make perfectly accurate predictions. Although these traders believe in the correct \emph{functional form} of the model, they still need to calibrate it based on limited noisy data. Therefore, the values they use for  the \emph{parameters} of the model will typically still be off.} 
% 		
% 		\begin{enumerate}
% 			\setlength\itemsep{0em}
% 			\item Anthropogenic climate change is a myth ($\beta = \rho = 0$),
% 			\begin{align}
% 			T^1_t = \lambda T_{t-1} +  \epsilon_t.
% 			\end{align}
% 			\item Anthropogenic climate change is real but policies and geoengineering are ineffective ($\omega = \rho = 0$),
% 			\begin{align}
% 			T^2_t & = \lambda T_{t-1} + \beta \big[ \kappa GHG_{t-1}  + \delta Y_t  \big] +  \epsilon_t.
% 			\end{align}
% 			\item Anthropogenic climate change is real, policies are effective, but not geoengineering ($\rho = 0$),
% 			\begin{align}
% 			T^3_t = \lambda T_{t-1} + \beta \big[ \kappa GHG_{t-1} - \alpha P_t^{1/2} + \delta Y_t  \big] +  \epsilon_t.
% 			\end{align}
% 			\item Anthropogenic climate change is real, policies are ineffective, but geoengineering works($\alpha = 0$),
% 			\begin{align}
% 			\begin{split}
% 		T^4_t = \lambda T_{t-1} + \beta \big[ \kappa GHG_{t-1} + \delta Y_t  \big]\\ - \rho \big[ \omega G_{t-1} + \theta X_{t} \big] +  \epsilon_t.
% 			\end{split}
% 			\end{align}
% 			
% 		\end{enumerate}
% 		
% 		\item[ Traders.] Traders are endowed with some initial money in an Experimental Currency Unit (ECU), and an initial approximate model. Approximate models are distributed at random among the traders with each model having the same probability of being assigned to any trader.  Traders use the distribution of temperatures obtained from their approximate models to determine their reservation price for different securities. 
% 		
% 		Securities pay 1 ECU  if the actual temperature at some time $t^*$ falls between a certain range. Traders are assumed to be \emph{risk-neutral expected utility maximizers}. Therefore, their reservation price for a security $\tilde{s}$ is simply their assessment of the probability that the temperature at time $t^*$ falls in the range covered by $\tilde{s}$.
% 		
% 		  Based on their reservation price, traders behave as ``\emph{zero-intelligence}" agents \citep{Gode1993}, that is they sell at a random price above their reservation, and buy at a random price below their reservation. These strategies are simple but have been shown to provide good approximations of the behavior of prediction markets (see for instance \cite{Klingert2012c}), and financial markets more broadly. 
% 		  
% 		\item[Continuous-Double Auction (CDA) market structure.] Continuous-Double auctions ( or some variants thereof) are common \emph{market mechanisms}, i.e. procedures to match buy and sell orders. CDA are notably used on large stock markets (see for instance \cite{Tseng2010a}). Our model features a particular version of CDA (more in section \ref{implem}). 
% 		
% 		\item[A social network.] Traders are part of a social network. Every time a security is realized, each trader $i$ looks at the performance of their neighbors in the network. If one of $i$'s neighbors, say $j$, is richer than $i$, $i$ interprets this has an indication that $j$ has a better approximate model. Then $i$ considers adopting $j$'s approximate model.\footnote{Traders start with the same initial amount of money, so differences in money among traders can only come from  traders' interactions through the market.} 
% 		
% 		An example of such social network is depicted in Figure \ref{socnet}. This corresponds to the initial social network for some of our simulations. The color of the nodes identify the approximate model of the trader. Notice the belief-homophily of the initial network which can be seen by the high clustering of traders with the same approximate model. Although traders can change their approximate model (i.e. colors can change in the graph), the connections between traders do not change as the market unfolds (i.e. the edges are fixed).
% 	\end{description}
% 	
% 	\begin{figure}
% 		\begin{center}
% 		\includegraphics[scale = 0.5]{network.png}
% 		\end{center}
% 		\caption{An example of the initial state of a belief network \label{socnet}}
% 	\end{figure}
% 	
% 	\subsubsection*{Timing}
% 	
% 	The time periods $t$ are grouped in trading \emph{sequences}. In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence. For instance, the third trading sequence might start in period $t=300$ and end in period $t = 399$. In this case, a security traded in the third sequence pay 1 ECU if the temperature at $t=400$ falls into the range of temperatures covered by the security. 
% 	
% 	   At each time $t$, traders are assumed to know the past value of the temperature $T_{0:t}$ and the vector of past explanatory variables $Z_{0:t}$. In a sequence finishing at time $t^*$,  traders also have common knowledge of $Z_{t:t^*}$, the future values of explanatory variables up to $t^*$. However, at any $t$, traders do not know the value of any future temperatures. In particular, in a given sequence, traders do not know the value of $T_{t^*}$. Traders can only predict $T_{t^*}$ using their approximate model and their knowledge of $Z_{t:t^*}$. Notice that because $Z_{t:t^*}$ is common knowledge, in each period $t$ every trader $i$ with the same approximate model forms the same expectation 
% 	   \begin{align*}
% 	  \mu_{ti} \equiv E_t( T_{t^*} ~|~ T_{0:t}, Z_{t:t^*},  \text{ approximate model of } i \text { at } t).
% 	   \end{align*}
% 	   We assume that the probability density that trader $i$ associate with $T_{t^*} = \tilde{T}$ at any $t$ is
% 	   \begin{align*}
% 	   f_{it}(T) = \mu_{ti} + N(0, \sigma_T),
% 	   \end{align*}
% 	   
% 	   where $N$ stands for a normal distribution and $\sigma_T$ is a fixed number determined by the true variance of temperatures.       
% 	   
% 	    At each time $t$, traders
% 	\begin{itemize}
% 		\setlength\itemsep{0em}
% 		\item recalibrate their approximate model based on the new set of past data available at $t$,
% 		\item compute their belief-density for $T_{t^*}$ and use it to determine the expected value they attached to each security,
% 		\item and trade on the market based on their market strategy at $t$.
% 	\end{itemize}
% 	
% 	
% 	At $t^*$, when the sequence ends, there is only one security $s^*$ associated with a range of temperatures including the actual $T_{t^*}$. Thus,
% 	\begin{itemize}
% 			\setlength\itemsep{0em}
% 		\item traders receive 1 ECU per unit of $s^*$ they own, and
% 		\item  consider adapting their approximate model as described above.
% 	\end{itemize}
% 	
% 	\section{Implementation}
% 	\label{implem}
% 	The model is implemented in the \texttt{R} programming language,  using the \texttt{igraph} package to model the social network. Each trader in the model is a node of an \texttt{igraph} network. Because the model uses time-series data and  traders calibrate statistical models based on these data, \texttt{R}, primarily used for statistical computing, is particularly convenient. 
% 	
% 	 Traders calibrate their approximate models based on past-data using Pooled Ordinary Least Squares. This is questionable from a statistical point of view, but provides an approximation of the way traders may form their forecast on actual prediction markets. Because the model is in \texttt{R}, it is also very easy to change the estimation method or model to virtually any statistical model, and the effect of altering models may easily be studied in the future.
% 	
% 	 In each period $\tilde{t}$, our version of the CDA works as follows.
% 	
% 	\begin{enumerate}
% 			\setlength\itemsep{0em}
% 		\item Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
% 		\item Every trader $i$ also chooses at random a security $s_i^S$ she will try to sell among the securities she owns a positive amount of (if any).
% 		\item Traders then decide of their selling $p_i^S$ and buying price $p_i^S$. They do so based on their believed-density of temperatures at the end of the sequence, and following the zero-intelligence rule described above.
% 		\item Traders go to the market one at the time, in a random order.
% 		\item When $i$ comes to the market, she place \emph{limit orders} in the order book. These orders  specify that $i$ is willing to buy $s_i^B$ at  any price below $p_i^B$, and to sell $s_i^S$ at any price above $p_i^S$.
% 		\item The market maker then tries to match $i$'s orders with some order which was put in the order book \emph{before} $i$ came to the market.
% 		\item If there are  outstanding sell offers for $s_i^B$ at price $p$ lower than $p_i^B$, then a trade is concluded. Trader $i$ buys one unit from the sellers who sells at the \emph{highest} price below $p_i^B$, and the sell and buy offers are removed from the order book.
% 		\item If there are  outstanding buy offers for $s_i^S$ at price $p$ higher than $p_i^S$, then a trade is concluded. Trader $i$ sells one unit to the buyer who buys at the \emph{highest} price above $p_i^S$, and the  sell and buy offers are removed from the order book.
% 		\item Whenever all traders have come to the market, any remaining outstanding offer is removed from the order book, and the trading period is concluded. 
% 	\end{enumerate} 
% 	
% The model depends on the following parameters. 
% \begin{description}
% 	\setlength\itemsep{0em}
% 	\item[Network parameters.]~
% 	\begin{description}
% 			\setlength\itemsep{0em}
% 		\item [\texttt{n.traders}:] the number of traders.
% 		\item [\texttt{n.edg} :] the number of edges in the social network. This number is fixed throughout the experiments.   
% 		\item [\texttt{seg} :] determines the initial degree of homophily in the network. The higher \texttt{seg}, the higher the initial homophily.\footnote{ When constructing the \texttt{n.edg} edges of the network, the probability that a link between two traders be formed depends on whether the traders share the same approximate model in the following way
% 			\begin{align*}
% 			\begin{cases}
% 			\frac{(1-\texttt{seg})}{\texttt{n.edg}},  &\text{if the traders have different approximate models}\\
% 			\frac{1}{\texttt{n.edg}},  & \text{otherwise}
% 			\end{cases}
% 			\end{align*}}
% 	\end{description}
% 
% \item[Market structure parameter.]~
% 
% 		\begin{description}
% 			\setlength\itemsep{0em}
% 			\item [\texttt{market.complet}.] Determines market's completeness, i.e. the number of securities which can be traded. With more securities, the interval of temperatures corresponding to each security is smaller. Traders can then trade on more precise temperature intervals. Higher values of \texttt{market.complet} may, however, reduce the number of exchanges. Because traders pick the securities they buy and sell at random, the probability that a match between sellers and buyers is found is lower for higher values of \texttt{market.complet}.
% 		\end{description}
% 	 
% \item[Behavioral parameters.]~
% 
% \begin{description}
% 	\setlength\itemsep{0em}
% 	\item [\texttt{risk.tak}.] Determines the distribution of risk taking behavior. The higher \texttt{risk.tak}, the more traders	will try to buy (resp. sell) lower (resp. higher) than their reservation price. Formally, at time $t$, trader $i$ picks her buying (resp. selling) prices randomly in the interval  $[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it}  * (1 - \texttt{risk.tak}_i)]$ (resp. $[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it} * (1  + \texttt{risk.tak}_i)]$), where $\texttt{reserv}_{it}$ is $i$'s reservation price at time $t$ for the security $i$ picked to buy (resp. sell).
% 	\item [\texttt{ideo}.]  Determines the degree of ``ideology" of traders. If \texttt{ideo} is high, traders will not
% 	 revise their approximate models easily, even when faced with
% 	 strong evidence that their neighbors are doing better than them.\footnote{ For each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$. The value of $d_{i}$ is the probability that $i$ adopts one of her neighbors' approximate model if this neighbor is doing better than $i$ at the end of the sequence (in monetary terms).}
% \end{description}	 
% 	
% 	\item[Timing parameters.]~
% 	
% 	\begin{description}
% 			\setlength\itemsep{0em}
% 		\item [\texttt{burn.in}.] The number of burn-in periods in which no securities are traded (necessary to allow approximate models to be estimated in the first period).
% 		\item [\texttt{n.seq}.] Number of trading sequences.
% 		\item [\texttt{horizon}.] Number of trading periods in each trading sequence.
% 	\end{description}
% 
% \end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}

In this section, we present an analysis of the sensitivity of the parameters of our model on the difference between the fraction of traders who believe in the true model at the beginning of the experiment, and the fraction who believe in the true model at the end of all trading sequences. The sensitivity analysis is based on a Latin Hypercube Sampling of the parameters from the following distributions:

\begin{itemize}
	\item \texttt{ideo} : $Unif(0,1)$
	\item \texttt{market.complet} : $Unif(0,1000)$ (mapped into integer),  
	\item  \texttt{n.edge} : $Unif(100,200)$ (mapped into integer),
	\item  \texttt{n.traders} : $Unif(50,250)$ (mapped into integer),
	\item \texttt{risk.tak}  : $Unif(0,1)$,
	\item  \texttt{seg} : $Unif(0,1)$,
	\item  \texttt{true.model} : $Binom(0.5)$
\end{itemize}

We conduct a partial \emph{rank} correlation coefficient analysis on the resulting simulated data \cite{Marino2008}. Partial correlation computes the linear relationship between the part of the variation of $x_i$ and $Y$ that are linearly independent of other $x_j$  ($j \neq i$).  The only difference with the partial correlation and partial \emph{rank} correlation (which we use here), is that the dependent variable $Y$ is first ranked-transformed in order to foster more uniform residuals.\footnote{That is every $Y_s$ in the data set is replace by a number $f(Y_s) \equiv |\{Y_r~|~ Y_s > Y_r\}|$, where $Y_r$ are the other observed values of $Y$ in the data set.} Such transformation preserves the sign of the correlation coefficient.

Our model suggests that the strength of the convergence to the true model is dependent of what the true model actually is: if anthropogenic climate change is true, a prediction market is likely to cause convergence to the true model; while if anthropogenic climate change is false, and the model of the climate is in fact auto-regressive, convergence is unlikely.

\begin{figure}
\centering
<<pc_sa, eval=TRUE, echo=FALSE, message=FALSE, cache=TRUE>>=
load("output/pc2.Rda")
correct_bias <- function(x) {
  x_corr <- x[["original"]] - x[["bias"]]
  min_ci <- x[["min. c.i."]]
  max_ci <- x[["max. c.i."]]
  x <- cbind(x_corr = x_corr, min_ci = min_ci, max_ci = max_ci, x)
  cbind(var = row.names(x), x)
}
plot_pc <- function(x, outcome_var = "Outcome", 
                    xlab = "Variable"){
  ss <- correct_bias(x[[7]]) # $SRRC, $SRC, $PRRC, $PRC

  ggplot2::ggplot(ss, ggplot2::aes(x = var, y = x_corr)) + ggplot2::geom_point() +
    ggplot2::xlab(xlab) + ggplot2::ylab("Partial Rank Correlation Coefficient ") +
    ggplot2::ggtitle(paste("Estimated Effects on", outcome_var)) +
    ggplot2::geom_errorbar(ggplot2::aes(ymax = max_ci, ymin = min_ci), width=0.25) +
    ggplot2::geom_hline(yintercept = 0) +
    ggplot2::ylim(c(-1,1)) +
    ggplot2::theme_bw()
}

plot_pc(pc2, outcome_var = "Convergence of Beliefs", xlab="")
@
\caption{Partial rank correlation coefficient analysis of the effects of ... on convergence of beliefs; lines are bootstrapped 95\% confidence intervals.}
\label{fig:pc_sa}
\end{figure}

\begin{figure}
\centering
<<hist, eval=TRUE, echo=FALSE, message=FALSE, cache=TRUE, fig.width=9, fig.height=6>>=
sample_count <- 7000
load("output/d.Rda")
library(ggplot2)
d$true.model <- factor(d$true.model, levels = c(0,1), 
                       labels = c("Natural Change", "Human-Induced"))
library(plyr)
cdat <- plyr::ddply(d, "true.model", summarize, outcome_summary = median(convergence))
ggplot2::ggplot(d, aes(x = convergence, fill = true.model)) + 
  geom_density(alpha = 0.3) +
  # xlim(c(-0.6, 0.6)) +
  ylab("Density") + xlab("Convergence of Beliefs") + 
  scale_fill_discrete(name="True Model") + 
  ggtitle(paste(sample_count, 
                "Convergence of Belief Outcomes with LHS Input Parameters")) +
  geom_vline(data=cdat, aes(xintercept= outcome_summary,  color = true.model),
             linetype="dashed", size=1) + 
  theme_bw() + 
  theme(legend.justification=c(0,1), legend.position=c(0,1)) 
@
\caption{Convergence of belief outcomes with Latin Hypercube Sampled input parameters. Dashed lines are the medians of the densities (\Sexpr{cdat$outcome_summary[1]}, \Sexpr{cdat$outcome_summary[2]}).}
\label{fig:hist}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
An important argument in favor of a climate prediction market is that it is supposed to be an apartisan proposal. Defenders of climate prediction markets argue that climate prediction markets will foster convergence to the best approximate model, whatever the actual true model is. If climate change is not anthropogenic -- so the argument goes, traders will converge to non-anthropogenic in pretty much the same way as they would converge to anthropogenic models if the climate is truly influenced by human activities. Our findings contradict this assumption.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
We gratefully acknowledge the authors of \textsf{R} \cite{r_core}. This manuscript was prepared using knitr \cite{xie_knitr_2014} and benefited from our exposure to the literature on reproducible research \cite{stodden_reproducible_research_2014}. We would like to thank Yevgeniy Vorobeychik for feedback on this project.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{plain}
\bibliography{library}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2015-06-11 prediction_market.Rnw
much more paper work. more references. better fig captions -- those 2 figs are pretty much pub ready now.

@@
 
-\title{Climate Prediction Markets and Convergence of Predictive Models: An Agent-based Simulation Analysis}
+\title{Climate Prediction Markets and Convergence of Predictive Models}
 \author{not sure about the author order}
@@
 We present an agent-based model of a climate prediction market in which traders adapt their belief about the climate based on the monetary performance of their neighboring traders in a social network.
-We use the model to test whether climate prediction markets foster a convergence of market participants' beliefs regarding a ``true'' climate model by conducting a sensitivity analysis of the impact of a variety of factors.
+We analyze the model to determine how a variety of factors affect whether climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the earth's climate.
 We find that only the imposed ``true model'' itself significantly affects the convergence of beliefs.
-If the true model from which the temperature data is generated is auto-regressive, with no human-induced effect, there is, on average.
-no convergence of traders' models due to participation in the prediction market.
-On other hand, if anthropogenic climate change is real, traders' beliefs are more likely than not to converge toward the correct model of the world after participation in the market.
+If the true model from which the temperature data is generated is auto-regressive, with no human-induced effects, participation in the prediction market, on average, does not cause convergence of traders' beliefs.
+On other hand, if anthropogenic climate change is real, traders' beliefs are more likely than not to converge toward the correct climate model due to participation in the market.
 
@@
 The previous two decades saw a strong polarization of the climate change debate.
-Although the scientific consensus on the anthropogenic nature of climate change strongly increased, the average belief about anthropogenic climate change did not evolve much within the public (see \cite{Vandenbergh2013f} for all the claims in this paragraph).
+Although the scientific consensus on the anthropogenic nature of climate change strongly increased, the average belief about anthropogenic climate change did not evolve much within the public \cite{Vandenbergh2013f}.
 In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily, indicating that the question is becoming increasingly politicized and potentially disconnected from scientific evidence.
-
 This is troubling given the costs of a misinformed climate policy.
@@
 Attempts to foster such consensus face many social and psychological challenges.
-Recently, people have argued that such challenges could be met by setting up climate prediction markets (see for instance \cite{Hsu2011}, and \cite{Vandenbergh2013f}).
-The idea of using prediction markets to efficiently aggregate information about uncertain events has been discussed for decades (see \cite{Horn2014c} for a recent review).
-Prediction markets have been shown to have interesting theoretical properties (e.g.
-\cite{Set1998}, \cite{Hanson2012}) and to perform well in terms of prediction accuracy in experiments (e.g \cite{Hanson2006}, \cite{Healy2010}), agent-based models (henceforth ABM) ( e.g \cite{Klingert2012c}, \cite{Jumadinova2011a}), and in the real-world (see \cite{Wolfers2006} for a review).
+Recently, people have argued that such challenges could be met by setting up climate prediction markets \cite{Hsu2011,Vandenbergh2013f}.
+The idea of using prediction markets to efficiently aggregate information about uncertain events has been discussed for decades \cite{Horn2014c}.
+Prediction markets have been shown to have interesting theoretical properties \cite{Set1998,Hanson2012} and to perform well in terms of prediction accuracy in experiments \cite{Hanson2006,Healy2010}, agent-based models (ABM) \cite{Klingert2012c,Jumadinova2011a}, and in the real-world \cite{Wolfers2006}.
 
@@
 So, do not delete these following two paragraphs that are commented out until we are accepted into a journal that doesnt want this.
+For Nature CC we could put this in an SI Appendix.
 
@@
 
-However, to the best of our knowledge, the idea that prediction markets can generate a consensus \emph{on the factors} affecting  uncertain events has never been quantitatively tested.
-ABMs of prediction markets have been studied ( e.g.
-\cite{Klingert2012c}, \cite{Tseng2010a}, or\cite{Jumadinova2011a}), some of which feature communication between agents.
+However, to the best of our knowledge, the idea that prediction markets can generate consensus \emph{on the factors} affecting  uncertain events has never been quantitatively explored.
+ABMs of prediction markets have been studied \cite{Klingert2012c,Tseng2010a,Jumadinova2011a}, some of which feature communication between agents.
 In these models, however, beliefs about the uncertain outcomes are constructed in rather abstract ways.
-In particular, beliefs are not based on structural models from which agents would derive causal implications.
-Therefore, although convergence of beliefs could be studied in previous ABMs, these models do not investigate the convergence of the underlying explanatory models agents use for prediction.
+In particular, beliefs are not based on structural models from which agents could derive causal implications, and therefore these models are not suitable for investigating the convergence of the underlying explanatory models agents employ for prediction.
 
-From a public policy perspective, this is a crucial point.
+From a public policy perspective, changing explanatory models is crucial.
 Effective climate policy not only requires an accurate consensus on future climate \emph{outcomes}.
 It also requires an accurate consensus on the causal \emph{mechanisms} influencing such outcomes.
-If people agree that temperature will rise, but some believe it will be due to greenhouse gases, while others believe that it will be caused by increased solar activity, inconsistent and ineffective policy will may be implemented.
-Our model is designed to investigate whether prediction markets can foster such convergence in predictive models.
+If people agree that temperature will rise, but some believe it will be due to greenhouse gases, while others believe that it will be caused by increased solar activity, inconsistent and ineffective policies may be implemented.
+Our model is designed to investigate whether, and under what conditions, prediction markets can foster such convergence in predictive models.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Model}
+\section*{Model Design}
 
 TODO: Martin use this text below (it is commented out) from the original document and flesh it out for the Supplementary Material and then write a couple paragraph summary and the key equations to put in this main article right here.
+
+In particular, lets go over main design and then define what each of these mean in our model: 	
+\begin{itemize}
+	\item \texttt{ideo} $\sim Unif(0,1)$
+	\item \texttt{market.complet} $\sim Unif(0,1000)$ (mapped into integer)
+	\item  \texttt{n.edge} $\sim Unif(100,200)$ (mapped into integer)
+	\item  \texttt{n.traders} $\sim Unif(50,250)$ (mapped into integer)
+	\item \texttt{risk.tak}  $\sim Unif(0,1)$
+	\item  \texttt{seg} $\sim Unif(0,1)$
+	\item  \texttt{true.model} $\sim Binom(0.5)$
+\end{itemize}
+
 % 	\section{Model}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Results}
+\section*{Model Analysis}
 
-In this section, we present an analysis of the sensitivity of the parameters of our model on the difference between the fraction of traders who believe in the true model at the beginning of the experiment, and the fraction who believe in the true model at the end of all trading sequences.
-The sensitivity analysis is based on a Latin Hypercube Sampling of the parameters from the following distributions:
+We estimated the effects of the parameters of our model on the difference between the fraction of traders who believe in the true model at the beginning of the experiment and the fraction who believe in the true model at the end of all trading sequences.
+The sensitivity analysis is based on a Latin Hypercube Sampling of 10,000 parameter sets from the following distributions \cite{beachkofski_improved_2002,carnell_lhs_2012}.
 
 \begin{itemize}
-	\item \texttt{ideo} : $Unif(0,1)$
-	\item \texttt{market.complet} : $Unif(0,1000)$ (mapped into integer),  
-	\item  \texttt{n.edge} : $Unif(100,200)$ (mapped into integer),
-	\item  \texttt{n.traders} : $Unif(50,250)$ (mapped into integer),
-	\item \texttt{risk.tak}  : $Unif(0,1)$,
-	\item  \texttt{seg} : $Unif(0,1)$,
-	\item  \texttt{true.model} : $Binom(0.5)$
+	\item \texttt{ideo} $\sim Unif(0,1)$
+	\item \texttt{market.complet} $\sim Unif(0,1000)$ (mapped into integer)
+	\item  \texttt{n.edge} $\sim Unif(100,200)$ (mapped into integer)
+	\item  \texttt{n.traders} $\sim Unif(50,250)$ (mapped into integer)
+	\item \texttt{risk.tak}  $\sim Unif(0,1)$
+	\item  \texttt{seg} $\sim Unif(0,1)$
+	\item  \texttt{true.model} $\sim Binom(0.5)$
 \end{itemize}
 
-We conduct a partial \emph{rank} correlation coefficient analysis on the resulting simulated data \cite{Marino2008}.
-Partial correlation computes the linear relationship between the part of the variation of $x_i$ and $Y$ that are linearly independent of other $x_j$  ($j \neq i$).
-The only difference with the partial correlation and partial \emph{rank} correlation (which we use here), is that the dependent variable $Y$ is first ranked-transformed in order to foster more uniform residuals.\footnote{That is every $Y_s$ in the data set is replace by a number $f(Y_s) \equiv |\{Y_r~|~ Y_s > Y_r\}|$, where $Y_r$ are the other observed values of $Y$ in the data set.} Such transformation preserves the sign of the correlation coefficient.
+We use the model to simulate 10,000 outcomes based on the input paramter sets and then conduct a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, resulting simulated vector, $y$ \cite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
+Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).
+The only difference with the partial correlation and partial \emph{rank} correlation (which we use here), is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships.\footnote{That is every value, $y_s$, in the data set is replace by a number $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, where $y_r$ are the other observed values of $y$ in the data set.}
 
-Our model suggests that the strength of the convergence to the true model is dependent of what the true model actually is: if anthropogenic climate change is true, a prediction market is likely to cause convergence to the true model; while if anthropogenic climate change is false, and the model of the climate is in fact auto-regressive, convergence is unlikely.
+Our sensitivity analysis, illustrated by Fig.~\ref{fig:pc_sa}, strongly suggests that the magnitude of convergence to the true model is conditional on what the true model actually is: if anthropogenic climate change is true, a prediction market is likely to cause convergence to the true model; while if anthropogenic climate change is false, and the model of the climate is in fact auto-regressive, convergence to this true model is unlikely (Fig.~\ref{fig:hist}).
 
-\begin{figure}
+\begin{figure}[h]
 \centering
@@
 @
-\caption{Partial rank correlation coefficient analysis of the effects of ... on convergence of beliefs; lines are bootstrapped 95\% confidence intervals.}
+\caption{Partial rank correlation coefficient analysis \cite{pujol_sensitivity:_2014,saltelli_sensitivity_2009} of the effects of ideology, market completeness, number of edges in social network, number of traders in market, risk taking propensities, segregation measure of the social network, and the true climate model on the convergence of traders' beliefs.
+Lines are bootstrapped 95\% confidence intervals.}
 \label{fig:pc_sa}
@@
 @
-\caption{Convergence of belief outcomes with Latin Hypercube Sampled input parameters.
-Dashed lines are the medians of the densities (\Sexpr{cdat$outcome_summary[1]}, \Sexpr{cdat$outcome_summary[2]}).}
+\caption{Density of \Sexpr{sample_count} convergence of belief model outcomes with \Sexpr{formatC(sample_count, big.mark=',')} Latin Hypercube Sampled input parameter sets.
+\Sexpr{formatC(sum(ifelse(d$true.model=="Human-Induced", 1, 0)), big.mark=',')} of the Latin Hypercube Sampled input parameter sets have ``human-induced'' climate change as the true model and \Sexpr{formatC(sum(ifelse(d$true.model=="Natural Change", 1, 0)), big.mark=',')} have ``natural change'' climate change as the true model.
+Dashed lines are the medians of the two densities (\Sexpr{cdat$outcome_summary[1]}, \Sexpr{cdat$outcome_summary[2]}).}
 \label{fig:hist}
@@
 \section*{Discussion}
-An important argument in favor of a climate prediction market is that it is supposed to be an apartisan proposal.
-Defenders of climate prediction markets argue that climate prediction markets will foster convergence to the best approximate model, whatever the actual true model is.
-If climate change is not anthropogenic -- so the argument goes, traders will converge to non-anthropogenic in pretty much the same way as they would converge to anthropogenic models if the climate is truly influenced by human activities.
+Proponents of climate prediction markets argue that climate prediction markets will foster convergence to the best approximate model, whatever the actual true model is.
+It is framed as an apartisan proposal: if climate change is not anthropogenic, traders will converge to non-anthropogenic beliefs similar to how they would converge to anthropogenic beliefs if the climate is truly influenced by human activities.
 Our findings contradict this assumption.
+
+TODO: Jonathan, perhaps talk about the policy implications of this finding...
 

### 2015-06-11 prediction_market.Rnw
more fig work. TODO: jonathan prob knows all the relevant cites for the sentences in the intro section that are currently lacking cites.

@@
 @
-\caption{Partial rank correlation coefficient analysis \cite{pujol_sensitivity:_2014,saltelli_sensitivity_2009} of the effects of ideology, market completeness, number of edges in social network, number of traders in market, risk taking propensities, segregation measure of the social network, and the true climate model on the convergence of traders' beliefs.
+\caption{Partial rank correlation coefficient analysis based on 10,000 simulated parameter sets \cite{pujol_sensitivity:_2014,saltelli_sensitivity_2009} of the effects of ideology, market completeness, number of edges in social network, number of traders in market, risk taking propensities, segregation measure of the social network, and the true climate model on the convergence of traders' beliefs.
 Lines are bootstrapped 95\% confidence intervals.}
@@
   scale_fill_discrete(name="True Model") + 
-  ggtitle(paste(sample_count, 
+  ggtitle(paste(formatC(sample_count, big.mark=','), 
                 "Convergence of Belief Outcomes with LHS Input Parameters")) +
@@
 @
-\caption{Density of \Sexpr{sample_count} convergence of belief model outcomes with \Sexpr{formatC(sample_count, big.mark=',')} Latin Hypercube Sampled input parameter sets.
+\caption{Density of \Sexpr{formatC(sample_count, big.mark=',')} convergence of belief model outcomes with \Sexpr{formatC(sample_count, big.mark=',')} Latin Hypercube Sampled input parameter sets.
 \Sexpr{formatC(sum(ifelse(d$true.model=="Human-Induced", 1, 0)), big.mark=',')} of the Latin Hypercube Sampled input parameter sets have ``human-induced'' climate change as the true model and \Sexpr{formatC(sum(ifelse(d$true.model=="Natural Change", 1, 0)), big.mark=',')} have ``natural change'' climate change as the true model.
-Dashed lines are the medians of the two densities (\Sexpr{cdat$outcome_summary[1]}, \Sexpr{cdat$outcome_summary[2]}).}
+Dashed lines are the medians of the two densities (\Sexpr{cdat$outcome_summary[1]}, \Sexpr{cdat$outcome_summary[2]}).
+Possible values of convergence of beliefs range from -1 to 1.}
 \label{fig:hist}

### 2015-06-11 prediction_market.Rnw
added the goel et al (2010) prediction without markets idea/cite

@@
 
-From a public policy perspective, changing explanatory models is crucial.
-Effective climate policy not only requires an accurate consensus on future climate \emph{outcomes}.
+From a public policy perspective, changing the explanatory models of market participants is one of the most important roles that prediction markets might play.\footnote{and perhaps the most important potential social benefit of prediction markets given the comparable predictive power of statistical models, which are less expensive than setting up and maintaining prediction markets \cite{goel_prediction_2010}.} Effective climate policy not only requires an accurate consensus on future climate \emph{outcomes}.
 It also requires an accurate consensus on the causal \emph{mechanisms} influencing such outcomes.

### 2015-06-14 colored.R
small code changes

@@ -1,7 +1,6 @@
-Colored <- function(
-  g = net,
-  attri = "approx"
-){
+Colored <- function(g,
+  attri = "approx"){
+  
   V(g)$color<- get.vertex.attribute(g,attri)
   
   d <- length(unique(V(g)$color))

### 2015-06-14 main.R

@@ -1,4 +1,6 @@
 
+library(igraph)
+
 # Construct model:
 source("generate_model.R")
 # Sub functions needed inside of generate_model():
@@ -29,7 +31,6 @@ main <- function(parameters,
   # are traded. With higher securities, traders can trade on
   # more precise temperature intervals
   
-  
   # SA creates everything in 0-1 interval, so im scaling things to the interval we want inside here
   seg <- parameters[1] # continuous value in (0,1)
   ideo <- parameters[2] # continuous value in (0,1)
@@ -46,7 +47,6 @@ main <- function(parameters,
   market.struct <- match.arg(market.struct)
   out <- match.arg(out)
   
-  library(igraph)
   if (visu){
     source("colored.R")
   }
@@ -98,7 +98,7 @@ main <- function(parameters,
   # Visualize network (optional)
   if (visu){
     net <- Colored(net)
-    plot.igraph(net,vertex.label=NA,layout=layout.fruchterman.reingold, vertex.size = 7)
+    igraph::plot.igraph(net,vertex.label=NA,layout=layout.fruchterman.reingold, vertex.size = 7)
   }
   #############################################
   #############################################

### 2015-06-14 main2.R

@@ -1,4 +1,6 @@
 
+library(igraph)
+
 # Construct model:
 source("generate_model.R")
 # Sub functions needed inside of generate_model():
@@ -46,7 +48,6 @@ main2 <- function(parameters,
   market.struct <- match.arg(market.struct)
   out <- match.arg(out)
   
-  library(igraph)
   if (visu){
     source("colored.R")
   }
@@ -98,7 +99,7 @@ main2 <- function(parameters,
   # Visualize network (optional)
   if (visu){
     net <- Colored(net)
-    plot.igraph(net,vertex.label=NA,layout=layout.fruchterman.reingold, vertex.size = 7)
+    igraph::plot.igraph(net,vertex.label=NA,layout=layout.fruchterman.reingold, vertex.size = 7)
   }
   #############################################
   #############################################

### 2015-06-14 master.R

@@ -6,7 +6,7 @@ rm(list=ls()) # make sure to always run this line of code and see that the next 
 source("main.R")
 # test
 s <- append(runif(4, min = 0.0001, max = 0.9999), sample(0:1, 1))
-main(parameters = s, out = "converg")
+main(parameters = s, out = "converg", visu = TRUE)
 
 
 # ######

### 2015-06-15 main2.R
ready for looking at convergence over iterations and plot for it

@@ -219,11 +219,6 @@ main2 <- function(parameters,
     }
   }
   
-  # TODO:
-  # need to compute some set of outcome measures here, like prices of a particular indices
-  # and then return one index, the one that matches the arg of main() called "outcome" 
-  # outcome
-  
   if (out == "segreg" && record == FALSE){
     
     # calculate the mixing matrix

### 2015-06-15 master.R

@@ -8,25 +8,6 @@ source("main.R")
 s <- append(runif(4, min = 0.0001, max = 0.9999), sample(0:1, 1))
 main(parameters = s, out = "converg", visu = TRUE)
 
-
-# ######
-# ##   Record values of outcomes through time and plot average through iterations
-# ######
-# 
-# s <- runif(4, min = 0.0001, max = 0.9999)
-# outcome.evolution <- main(parameters = s, out = "converg", record = TRUE)
-# 
-# for (j in 1:99){
-#   s <- runif(4, min = 0.0001, max = 0.9999)
-#   outcome.evolution <- rbind(outcome.evolution, main(parameters = s, out = "converg", record = TRUE))
-# }
-# 
-# average.convergence <- colMeans(outcome.evolution)
-# 
-# plot(average.convergence, type = "b", main = "Average convergence \nas a function of the number of trading sequences",
-#                                       xlab = "Number of trading sequences",
-#                                       ylab = "Average convergence (n = 100)")
-
 #####
 ## Sensitivity analysis
 #####
@@ -119,6 +100,25 @@ ggplot2::ggplot(d, aes(x = convergence, fill = true.model)) +
 ##############################################################################
 rm(list=ls())
 source("main2.R")
+# test once
+s <- c(runif(4, min = 0.0001, max = 0.9999), sample(0:1, 1), runif(2, min = 0.0001, max = 0.9999))
+outcome.evolution <- main2(parameters = s, out = "converg", visu = TRUE, record = TRUE)
+# run 100
+doParallel::registerDoParallel(cores = parallel::detectCores() - 1)
+runs <- 100
+outcome.evolution <- foreach::`%dopar%`(foreach::foreach(i=seq(runs), .combine='rbind'), {
+  main2(parameters = c(runif(4, min = 0.0001, max = 0.9999), sample(0:1, 1), runif(2, min = 0.0001, max = 0.9999)), 
+        out = "converg", visu = FALSE, record = TRUE)
+})
+average_convergence <- data.frame(avg = colMeans(outcome.evolution), 
+                                  trading_seq = seq(length(colMeans(outcome.evolution))))
+save(average_convergence, file = "output/average_convergence.Rda")
+library(ggplot2)
+ggplot(data=average_convergence, aes(x= trading_seq, y=avg)) +
+  geom_line() + ggtitle("Average convergence \nas a function of the number of trading sequences") +
[…]

### 2015-06-15 master.R
computed and plotted avg convergence over trading sequences

@@ -103,11 +103,25 @@ source("main2.R")
 # test once
 s <- c(runif(4, min = 0.0001, max = 0.9999), sample(0:1, 1), runif(2, min = 0.0001, max = 0.9999))
 outcome.evolution <- main2(parameters = s, out = "converg", visu = TRUE, record = TRUE)
-# run 100
+# run many
+input_values <- lapply(list(seg = NA, ideo = NA, risk.tak = NA,
+                            market.complet = NA), 
+                       function(x) list(random_function = "qunif",
+                                        ARGS = list(min = 0.0001, max = 0.9999)))
+input_values$true.model <- list(random_function = "qbinom",
+                                ARGS = list(size = 1, prob = 0.5))
+input_values$n.edg <- list(random_function = "qunif",
+                           ARGS = list(min = 0.0001, max = 0.9999))
+input_values$n.traders <- list(random_function = "qunif",
+                               ARGS = list(min = 0.0001, max = 0.19999))
+input_names <- names(input_values)
+cores <- parallel::detectCores() - 1
+sample_count <- 1000
+input_set <- create_set(input_values, input_names, sample_count = sample_count,
+                        constraints = "none")
 doParallel::registerDoParallel(cores = parallel::detectCores() - 1)
-runs <- 100
-outcome.evolution <- foreach::`%dopar%`(foreach::foreach(i=seq(runs), .combine='rbind'), {
-  main2(parameters = c(runif(4, min = 0.0001, max = 0.9999), sample(0:1, 1), runif(2, min = 0.0001, max = 0.9999)), 
+outcome.evolution <- foreach::`%dopar%`(foreach::foreach(i=seq(nrow(input_set)), .combine='rbind'), {
+  main2(parameters = as.numeric(input_set[i, ]), 
         out = "converg", visu = FALSE, record = TRUE)
 })
 average_convergence <- data.frame(avg = colMeans(outcome.evolution), 
@@ -115,8 +129,9 @@ average_convergence <- data.frame(avg = colMeans(outcome.evolution),
 save(average_convergence, file = "output/average_convergence.Rda")
 library(ggplot2)
 ggplot(data=average_convergence, aes(x= trading_seq, y=avg)) +
-  geom_line() + ggtitle("Average convergence \nas a function of the number of trading sequences") +
-  xlab("Number of trading sequences") + ylab(paste0("Average convergence (n = ", runs, ")"))
[…]

### 2015-06-15 master.R
broke down convergence over iterations into those with true model as ACC and true model as natural change and see two distinct temporal trends that show that the prediction market causes opposite convergence effects for the two.

@@ -124,14 +124,47 @@ outcome.evolution <- foreach::`%dopar%`(foreach::foreach(i=seq(nrow(input_set)),
   main2(parameters = as.numeric(input_set[i, ]), 
         out = "converg", visu = FALSE, record = TRUE)
 })
+# TRUE model is even prob between Human-induced and Natural
 average_convergence <- data.frame(avg = colMeans(outcome.evolution), 
-                                  trading_seq = seq(length(colMeans(outcome.evolution))))
+                                  trading_seq = seq(length(colMeans(outcome.evolution))),
+                                  true_mod = "Both")
+# TRUE model is Human-induced
+input_values$true.model <- list(random_function = "qbinom",
+                                ARGS = list(size = 1, prob = 1))
+input_set <- create_set(input_values, input_names, sample_count = sample_count,
+                        constraints = "none")
+doParallel::registerDoParallel(cores = parallel::detectCores() - 1)
+outcome.evolution <- foreach::`%dopar%`(foreach::foreach(i=seq(nrow(input_set)), .combine='rbind'), {
+  main2(parameters = as.numeric(input_set[i, ]), 
+        out = "converg", visu = FALSE, record = TRUE)
+})
+average_convergence <- rbind(average_convergence,
+                             data.frame(avg = colMeans(outcome.evolution), 
+                                        trading_seq = seq(length(colMeans(outcome.evolution))),
+                                        true_mod = "Human-Induced"))
+# TRUE model is Natural Change
+input_values$true.model <- list(random_function = "qbinom",
+                                ARGS = list(size = 1, prob = 0))
+input_set <- create_set(input_values, input_names, sample_count = sample_count,
+                        constraints = "none")
+doParallel::registerDoParallel(cores = parallel::detectCores() - 1)
+outcome.evolution <- foreach::`%dopar%`(foreach::foreach(i=seq(nrow(input_set)), .combine='rbind'), {
+  main2(parameters = as.numeric(input_set[i, ]), 
+        out = "converg", visu = FALSE, record = TRUE)
+})
+average_convergence <- rbind(average_convergence,
+                             data.frame(avg = colMeans(outcome.evolution), 
+                                        trading_seq = seq(length(colMeans(outcome.evolution))),
+                                        true_mod = "Natural"))
[…]

### 2015-06-15 prediction_market.tex
added temporal evolution results to paper

--- 2015-06-11 prediction_market.Rnw
@@
-\documentclass{article}
+\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
 
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
+
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 \centering
-<<pc_sa, eval=TRUE, echo=FALSE, message=FALSE, cache=TRUE>>=
-load("output/pc2.Rda")
-correct_bias <- function(x) {
-  x_corr <- x[["original"]] - x[["bias"]]
-  min_ci <- x[["min. c.i."]]
-  max_ci <- x[["max. c.i."]]
-  x <- cbind(x_corr = x_corr, min_ci = min_ci, max_ci = max_ci, x)
-  cbind(var = row.names(x), x)
-}
-plot_pc <- function(x, outcome_var = "Outcome", 
-                    xlab = "Variable"){
-  ss <- correct_bias(x[[7]]) # $SRRC, $SRC, $PRRC, $PRC
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/pc_sa-1}
 
-  ggplot2::ggplot(ss, ggplot2::aes(x = var, y = x_corr)) + ggplot2::geom_point() +
-    ggplot2::xlab(xlab) + ggplot2::ylab("Partial Rank Correlation Coefficient ") +
-    ggplot2::ggtitle(paste("Estimated Effects on", outcome_var)) +
-    ggplot2::geom_errorbar(ggplot2::aes(ymax = max_ci, ymin = min_ci), width=0.25) +
-    ggplot2::geom_hline(yintercept = 0) +
-    ggplot2::ylim(c(-1,1)) +
-    ggplot2::theme_bw()
-}
-
-plot_pc(pc2, outcome_var = "Convergence of Beliefs", xlab="")
-@
+\end{knitrout}
 \caption{Partial rank correlation coefficient analysis based on 10,000 simulated parameter sets \cite{pujol_sensitivity:_2014,saltelli_sensitivity_2009} of the effects of ideology, market completeness, number of edges in social network, number of traders in market, risk taking propensities, segregation measure of the social network, and the true climate model on the convergence of traders' beliefs.
@@
 \centering
-<<hist, eval=TRUE, echo=FALSE, message=FALSE, cache=TRUE, fig.width=9, fig.height=6>>=
-sample_count <- 7000
-load("output/d.Rda")
-library(ggplot2)
-d$true.model <- factor(d$true.model, levels = c(0,1), 
-                       labels = c("Natural Change", "Human-Induced"))
-library(plyr)
-cdat <- plyr::ddply(d, "true.model", summarize, outcome_summary = median(convergence))
-ggplot2::ggplot(d, aes(x = convergence, fill = true.model)) + 
-  geom_density(alpha = 0.3) +
-  # xlim(c(-0.6, 0.6)) +
-  ylab("Density") + xlab("Convergence of Beliefs") + 
-  scale_fill_discrete(name="True Model") + 
-  ggtitle(paste(formatC(sample_count, big.mark=','), 
-                "Convergence of Belief Outcomes with LHS Input Parameters")) +
-  geom_vline(data=cdat, aes(xintercept= outcome_summary,  color = true.model),
-             linetype="dashed", size=1) + 
-  theme_bw() + 
-  theme(legend.justification=c(0,1), legend.position=c(0,1)) 
-@
-\caption{Density of \Sexpr{formatC(sample_count, big.mark=',')} convergence of belief model outcomes with \Sexpr{formatC(sample_count, big.mark=',')} Latin Hypercube Sampled input parameter sets.
-\Sexpr{formatC(sum(ifelse(d$true.model=="Human-Induced", 1, 0)), big.mark=',')} of the Latin Hypercube Sampled input parameter sets have ``human-induced'' climate change as the true model and \Sexpr{formatC(sum(ifelse(d$true.model=="Natural Change", 1, 0)), big.mark=',')} have ``natural change'' climate change as the true model.
-Dashed lines are the medians of the two densities (\Sexpr{cdat$outcome_summary[1]}, \Sexpr{cdat$outcome_summary[2]}).
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/hist-1}
+
+\end{knitrout}
+\caption{Density of 7,000 convergence of belief model outcomes with 7,000 Latin Hypercube Sampled input parameter sets.
+3,500 of the Latin Hypercube Sampled input parameter sets have ``human-induced'' climate change as the true model and 3,500 have ``natural change'' climate change as the true model.
+Dashed lines are the medians of the two densities (0, 0.11).
 Possible values of convergence of beliefs range from -1 to 1.}
 \label{fig:hist}
+\end{figure}
+
+\begin{figure}
+\centering
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}
+\includegraphics[width=\maxwidth]{figure/temporal-1}
+
+\end{knitrout}
+\caption{Temporal evolution of convergence of beliefs.}
+\label{fig:temporal}
 \end{figure}

### 2015-06-15 prediction_market.Rnw
caption for temporal evolution figure

@@
 
+\begin{figure}
+\centering
+<<temporal, eval=TRUE, echo=FALSE, message=FALSE, cache=TRUE, fig.width=10, fig.height=6>>=
+sample_count <- 1000
+load("output/average_convergence.Rda")
+dat <- average_convergence[average_convergence$true_mod != "Both", ]
+dat$true_mod <- factor(dat$true_mod, levels = c("Natural", "Human-Induced"), 
+                       labels = c("Natural Change", "Human-Induced"))
+library(ggplot2)
+ggplot(data= dat, aes(x= trading_seq, y=avg, color = true_mod)) +
+  geom_smooth(method = "loess") + 
+  ggtitle("Average Convergence Over Trading Sequences") +
+  xlab("Trading Sequences") + 
+  ylab(paste0("Average Convergence (n = ", formatC(sample_count, big.mark = ','), ")")) + 
+  theme_bw() + theme(legend.justification=c(0,1), legend.position=c(0,1)) + 
+  scale_color_discrete(name="True Model")
+@
+\caption{Temporal evolution of convergence of beliefs for the ABM with \Sexpr{formatC(sample_count, big.mark=',')} model runs with ``True Model'' parameter set to natural climate change and \Sexpr{formatC(sample_count, big.mark=',')} model runs with ``True Model'' parameter set to human-induced climate change.
+All other parameters are sampled from their full support with Latin Hypercube Sampling; therefore, the plot represents, in effect, a marginalizing out of the effects of all other parameters to show the moderating effect that the true model has on the effect of the prediction market on the convergence of traders' beliefs about what the true model is.
+The prediction market causes opposite convergence effects for the two true climate models.}
+\label{fig:temporal}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-06-15 prediction_market.tex
caption for temporal evolution figure

@@
 \end{knitrout}
-\caption{Temporal evolution of convergence of beliefs.}
+\caption{Temporal evolution of convergence of beliefs for the ABM with 1,000 model runs with ``True Model'' parameter set to natural climate change and 1,000 model runs with ``True Model'' parameter set to human-induced climate change.
+All other parameters are sampled from their full support with Latin Hypercube Sampling; therefore, the plot represents, in effect, a marginalizing out of the effects of all other parameters to show the moderating effect that the true model has on the effect of the prediction market on the convergence of traders' beliefs about what the true model is.
+The prediction market causes opposite convergence effects for the two true climate models.}
 \label{fig:temporal}

### 2015-06-15 main2.R
changed main2 so when recording temporal evoluition we subtract initial convergence each time

@@ -154,7 +154,7 @@ main2 <- function(parameters,
     
     if(out == "converg"){
       
-      result <- length(V(net)$approx[V(net)$approx == 1])/length(V(net))
+      result <- (length(V(net)$approx[V(net)$approx == 1])/length(V(net))) - net$init.converg.util
       
     }
   }
@@ -211,7 +211,7 @@ main2 <- function(parameters,
         
         if(out == "converg"){
           
-          result <- append(result,length(V(net)$approx[V(net)$approx == 1])/length(V(net)))
+          result <- append(result, (length(V(net)$approx[V(net)$approx == 1])/length(V(net))) - net$init.converg.util)
           
         }
       }

### 2015-06-15 master.R
new output with convergence measures same as the histogram plot and SA

@@ -104,6 +104,8 @@ source("main2.R")
 s <- c(runif(4, min = 0.0001, max = 0.9999), sample(0:1, 1), runif(2, min = 0.0001, max = 0.9999))
 outcome.evolution <- main2(parameters = s, out = "converg", visu = TRUE, record = TRUE)
 # run many
+sample_count <- 1000
+# SA
 input_values <- lapply(list(seg = NA, ideo = NA, risk.tak = NA,
                             market.complet = NA), 
                        function(x) list(random_function = "qunif",
@@ -115,33 +117,20 @@ input_values$n.edg <- list(random_function = "qunif",
 input_values$n.traders <- list(random_function = "qunif",
                                ARGS = list(min = 0.0001, max = 0.19999))
 input_names <- names(input_values)
-cores <- parallel::detectCores() - 1
-sample_count <- 1000
-input_set <- create_set(input_values, input_names, sample_count = sample_count,
-                        constraints = "none")
-doParallel::registerDoParallel(cores = parallel::detectCores() - 1)
-outcome.evolution <- foreach::`%dopar%`(foreach::foreach(i=seq(nrow(input_set)), .combine='rbind'), {
-  main2(parameters = as.numeric(input_set[i, ]), 
-        out = "converg", visu = FALSE, record = TRUE)
-})
-# TRUE model is even prob between Human-induced and Natural
-average_convergence <- data.frame(avg = colMeans(outcome.evolution), 
-                                  trading_seq = seq(length(colMeans(outcome.evolution))),
-                                  true_mod = "Both")
 # TRUE model is Human-induced
 input_values$true.model <- list(random_function = "qbinom",
                                 ARGS = list(size = 1, prob = 1))
-input_set <- create_set(input_values, input_names, sample_count = sample_count,
+library(eat)
+input_set <- eat::create_set(input_values, input_names, sample_count = sample_count,
                         constraints = "none")
 doParallel::registerDoParallel(cores = parallel::detectCores() - 1)
 outcome.evolution <- foreach::`%dopar%`(foreach::foreach(i=seq(nrow(input_set)), .combine='rbind'), {
   main2(parameters = as.numeric(input_set[i, ]), 
         out = "converg", visu = FALSE, record = TRUE)
 })
-average_convergence <- rbind(average_convergence,
-                             data.frame(avg = colMeans(outcome.evolution), 
-                                        trading_seq = seq(length(colMeans(outcome.evolution))),
[…]

### 2015-06-15 prediction_market.Rnw
requisite changes to the paper

@@
 load("output/average_convergence.Rda")
-dat <- average_convergence[average_convergence$true_mod != "Both", ]
+dat <- average_convergence
 dat$true_mod <- factor(dat$true_mod, levels = c("Natural", "Human-Induced"), 

### 2015-08-27 prediction_market.Rnw
Work on paper's plot showing the effect of the timing parameters.

--- 2015-06-11 prediction_market.Rnw
@@
 
-\title{Climate Prediction Markets and Convergence of Predictive Models}
-\author{not sure about the author order}
+\title{Simulated Climate Prediction Markets and Convergence of Predictive Models}
+\author{ }
 
@@
 
+\begin{figure}
+\centering
+<<temporal, eval=TRUE, echo=FALSE, message=FALSE, cache=TRUE, fig.width=10, fig.height=10>>=
+# set[[1]] <- list(burn.in = 4, n.seq = 28, horizon = 4)
+# set[[2]] <- list(burn.in = 6, n.seq = 22, horizon = 5)
+# set[[3]] <- list(burn.in = 16, n.seq = 20, horizon = 5)
+load("output/average_convergence.Rda")
+sample_count <- 1000
+library(ggplot2)
+plot_data <- average_convergence
+plot_data[ , "set"] <- factor(plot_data[ , "set"], levels = gtools::mixedsort(unique(plot_data[ , "set"])))
+ggplot(data=plot_data, aes(x= trading_seq, y=avg, color = true_mod)) +
+  geom_smooth(method = "loess") + #geom_line() +
+  ggplot2::facet_wrap(~set, ncol = 1) +
+  ggtitle("Average Convergence Over Trading Sequences") +
+  xlab("Trading Sequences") + ylab(paste0("Average Convergence (n = ", sample_count, ")")) + 
+  theme_bw() + theme(legend.justification=c(1,0), legend.position=c(1,0)) + 
+  scale_color_discrete(name="True Model")
+@
+\caption{Temporal evolution of convergence of beliefs for the ABM with \Sexpr{formatC(sample_count, big.mark=',')} model runs with ``True Model'' parameter set to natural climate change and \Sexpr{formatC(sample_count, big.mark=',')} model runs with ``True Model'' parameter set to human-induced climate change.
+All other parameters besides burn.in, n.seq, and horizon are sampled from their full support with Latin Hypercube Sampling; therefore, the plot represents, in effect, a marginalizing out of the effects of all other parameters to show the moderating effect that the true model has on the effect of the prediction market on the convergence of traders' beliefs about what the true model is.
+The prediction market causes opposite convergence effects for the two true climate models.
+The top figure used the timing settings: burn.in = 4, n.seq = 28, horizon = 4.
+Middle figure: burn.in = 6, n.seq = 22, horizon = 5.
+Bottom figure: burn.in = 16, n.seq = 20, horizon = 5.}
+\label{fig:temporal}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 We gratefully acknowledge the authors of \textsf{R} \cite{r_core}.
-This manuscript was prepared using knitr \cite{xie_knitr_2014} and benefited from our exposure to the literature on reproducible research \cite{stodden_reproducible_research_2014}.
+This manuscript was prepared using knitr \cite{xie_knitr_2014}.
 We would like to thank Yevgeniy Vorobeychik for feedback on this project.

### 2015-08-28 prediction_market.Rnw
just corrected a typo.

@@
 
-From a public policy perspective, changing the explanatory models of market participants is one of the most important roles that prediction markets might play.\footnote{and perhaps the most important potential social benefit of prediction markets given the comparable predictive power of statistical models, which are less expensive than setting up and maintaining prediction markets \cite{goel_prediction_2010}.} Effective climate policy not only requires an accurate consensus on future climate \emph{outcomes}.
+From a public policy perspective, changing the explanatory models of market participants is one of the most important roles that prediction markets might play.\footnote{And perhaps the most important potential social benefit of prediction markets given the comparable predictive power of statistical models, which are less expensive than setting up and maintaining prediction markets \cite{goel_prediction_2010}.} Effective climate policy not only requires an accurate consensus on future climate \emph{outcomes}.
 It also requires an accurate consensus on the causal \emph{mechanisms} influencing such outcomes.

### 2015-08-28 Temp Latex work/Supplemental material/Supplemental material.tex
Knitr looks neat but I like my old latex editor (Texstudio) better than working from inside Rstudio to edit latex text.
Over the years, I have personalized it with shortcuts and stuff which make me more productive, and going back really slows me down.

So I am creating this directory to store temporary latex files. Once I am done with working on asection of the paper, I'll put the section back on the .Rwd file.

If there is any better way to proceed, just let me know.

--- 2015-06-11 paper/older/prediction_market.tex
@@
-\documentclass{sig-alternate}
-
-\usepackage[utf8]{inputenc}
+\documentclass[10pt,a4paper]{article}
+\usepackage{fullpage}
+\usepackage[latin1]{inputenc}
 \usepackage{natbib}
-
-%\usepackage[utf8]{inputenc}
-%\usepackage[backend = bibtex, citestyle=authoryear]{biblatex}
-
-%\usepackage[american]{babel}
-%\usepackage{csquotes}
-%\usepackage[style=apa]{biblatex}
-%\DeclareLanguageMapping{american}{american-apa}
-
-%\addbibresource{library.bib}
-%\usepackage{fullpage}
-
-%Math mode
-
 \usepackage{amsmath}
@@
 \usepackage{amssymb}
-%\usepackage{amsthm}
-\usepackage{dsfont}
-%\usepackage{mathtools}
-\usepackage{xfrac}
-%\usepackage{thmtools}
-\usepackage{bm}
-
-% CODE
-
-\usepackage{courier}
-
-\usepackage{listings}
-\usepackage{color} %red, green, blue, yellow, cyan, magenta, black, white
-\definecolor{mygreen}{RGB}{28,172,0} % color values Red, Green, Blue
-\definecolor{mylilas}{RGB}{170,55,241}
-\usepackage{marginnote}
+\usepackage{amsthm}
+\usepackage{mathtools}
+\usepackage{animate}
 \usepackage[linktocpage, breaklinks, colorlinks = true, linkcolor = Brown, citecolor = blue]{hyperref}
 \usepackage{enumerate}
-
-%PS_tricks
-
+\usepackage[dvips]{graphicx}
 \usepackage[usenames,dvipsnames]{pstricks}
@@
 \usepackage{float}
+\usepackage{bm}
+\usepackage{nameref}
+\usepackage{xfrac}
+\usepackage{accents}
+\usepackage[stable]{footmisc}
+\usepackage[nice]{nicefrac}
+\usepackage{thmtools}
+\usepackage{blindtext}
+\usepackage{tcolorbox} 
+\tcbuselibrary{breakable}
 
-\usepackage{nameref}
-\usepackage{mdframed}
-\usepackage{accents}
+\newcounter{test}
 
-%Geometry
-%\usepackage[lmargin=3.5cm,rmargin=3.5cm,tmargin=2cm,bmargin=2cm,marginpar=3cm]{geometry}
+\newtcolorbox{framet}[2][%
+breakable,
+arc=7pt,
+outer arc=0pt,
+coltitle=black,
+fonttitle=\bfseries,
+boxrule=2pt,
+colframe=Brown,
+colback=Apricot,
+title after break={Frame~\thetest\ (Continued)}
+]{%
+before upper={
+  \stepcounter{test}\textbf{Frame~\thetest.\ }%
+},
+label={#2},
+#1}
 
-%Proof style
+\newcommand{\aligncom}[1]{%
+  \text{\phantom{(#1)}} \tag{#1}
+}
 
-%\renewcommand\qedsymbol{$\blacksquare$}
+\newcounter{frame} \setcounter{frame}{0}
 
-% %Margin style
+\renewcommand{\theframe}{\Roman{frame}}
 
-\renewcommand*{\marginfont}{\footnotesize}
-
-% %Graphic path
+\newenvironment{FFrame}[1]%
+{
+\refstepcounter{frame}
+\vspace{1ex} \flushleft\makebox[13ex][l]{\textbf{Frame} \theframe} \texttt{#1}
+}
 
@@
 
-% %Theorem Environments
+\declaretheorem [name = Theorem]{theo}
+\declaretheorem [name = Lemma]{lem}
+\declaretheorem [name = Axiom]{ax}
+\declaretheorem [name = Preference axiom]{pref}
+\declaretheorem [name = Observation]{obs}
+\declaretheorem [name = Definition]{defi}
+\declaretheorem [name = Example]{expl}
+\declaretheorem [name = Social ordering function]{sof}
+\declaretheorem [name = Problem]{prob}
+\declaretheorem [name = Assumption]{ass}
+\declaretheorem [name = Proposition]{prop}
 
-%\declaretheorem [name = Theorem]{theo}
-%\declaretheorem [name = Corollary]{coro}
-%\declaretheorem [name = Lemma]{lem}
-%\declaretheorem [name = Fairness Axiom]{ax}
-%\declaretheorem [name = Preference axiom]{pref}
-%\declaretheorem [name = Profile axiom]{prof}
-%\declaretheorem [name = Observation]{obs}
-%\declaretheorem [name = Definition]{defi}
-%\declaretheorem [name = Social ordering function]{sof}
-%\declaretheorem [name = Proposition, unnumbered]{prop}
-%\declaretheorem [name = Example, unnumbered]{expl}
-%
-% %In-text hyppereff
+\DeclareFontFamily{OT1}{pzc}{}
+\DeclareFontShape{OT1}{pzc}{m}{it}%
+              {<-> s * [0.900] pzcmi7t}{}
+\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}%
+                                 {m}{it}
+
+%\newtheorem{ax}{Axiom}
+%\newtheorem{pref}{Preference axiom}
+%\newtheorem{obs}{Observation}
+%\newtheorem{defi}{Definition}
+%\newtheorem{sof}{Social ordering function}
 
@@
 %Style
-
 \newcommand{\mb}[1]{\mathbf{#1}}     % bold letters in mathmode
@@
 \newcommand{\bb}[1]{\mathbb{#1}}     % Doppelstrichbuchstaben im Mathematikmodus
-
 \newcommand{\mc}[1]{\mathcal{#1}}    % calligraphy style
-
 %Sets
-
 \newcommand{\rr}{\bb{R}} % set of reals in mathmode
-
 %Fractions
-
 \newcommand{\pfraca}[2]{\bigl(\frac{#1}{#2}\bigr)}    % ratios with small parantheses
@@
 \newcommand{\pfracd}[2]{\Biggl(\frac{#1}{#2}\Biggr)}  % ratios with huge parantheses
-
 %Partial derivatives
-
 \newcommand{\di}[1]{\frac{\partial}{\partial #1}}     % shortcut for first partial derivative
 \newcommand{\dii}[1]{\frac{\partial^2}{\partial #1^2}}% shortcut for second partial derivative
-
 %other
-
-\newcommand{\ra}{\rightarrow}
-
+\newcommand{\ra}{\rightarrow}            % convergence
 %Convergence
-
 \newcommand{\cd}{\overset{d}{\rightarrow}}    % convergence in distribution
@@
 \newcommand{\cas}{\overset{as}{\rightarrow}}  % convergence almost surely
-
 %Text in math environment
-
 \newcommand{\sgn}{~\text{sgn}~}
@@
 \newcommand{\id}{\coloneqq}
-
 %Variant of variables (tilde, lower bar,...)
-
 \newcommand{\lbar}[1]{\underline{#1\mkern-4mu}\mkern4mu }
-
 \newcommand{\x}{\bm{x}}
@@
 
-\clubpenalty=10000
-\widowpenalty = 10000
-
-\usepackage[hyphenbreaks]{breakurl}
-%\usepackage[hyphens]{url}
+\title{}
+\author{Martin Van der Linden}
 
@@
 
-	\title{Prediction Markets and Congruence of Predictive Models:
-  An Agent-Based Simulation of Climate Prediction Markets}
-  %\substitle{An Agent-Based Simulation of Climate Prediction Markets}
+\maketitle
 
-	\numberofauthors{2} 
-	\author{
-    \alignauthor
-		John J. Nay
-		\alignauthor
-		Martin Van der Linden
-	}
-
-	\maketitle
-	\begin{abstract}
-We present an agent-based model (ABM) of a climate prediction market in which traders adapt their belief about the climate based on the monetary performance of their neighboring traders in a social network.
-We use the ABM to test whether climate prediction markets fosters a convergence of market participants' beliefs regarding a ``true'' climate model.
-We provide preliminary sensitivity analysis for the impact on the convergence of factors such as the initial belief-homophily in the social network and the degree of risk-taking of traders on the market.
-	\end{abstract}
-
-	\section{Introduction}
-
-	The previous two decades saw a strong polarization of the climate change debate.
-Although the scientific consensus on the anthropogenic nature of climate change strongly increased,  the average belief about anthropogenic climate change did not evolve much within the public (see \cite{Vandenbergh2013f} for all the claims in this paragraph).
-In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily, indicating that the question is becoming more and more politicized and potentially disconnected from scientific evidence.
-
-	This is worrisome given the costs of a misinformed climate policy.
-If anthropocentric climate change is a myth but is taken to be true, a tremendous amount of public resources will be wasted on adaptation and mitigation efforts.
-On the other hand, if anthropocentric climate change is wrongly consider as false, the costs of inaction would be huge.
-Given that effective climate policies require long-lasting measures to be taken, the importance of an accurate consensus on the issue is manifest.
-
-	Attempts to foster such consensus face many \linebreak  social and psychological challenges (see \cite[Sections II.B to II.D]{Vandenbergh2013f} for a review).
-Recently, people have argued that such challenges could be met by setting up climate prediction markets (see for instance \cite{Hsu2011}, and \cite{Vandenbergh2013f}).
-The idea of using prediction markets to efficiently aggregate information about uncertain events has been discussed for decades (see \cite{Horn2014c} for a recent review).
-Prediction markets have been shown to have interesting theoretical properties (e.g.
-\cite{Set1998}, \cite{Hanson2012}, ) and to perform well in terms of prediction accuracy in experiments (e.g \cite{Hanson2006}, \cite{Healy2010}), agent-based models (henceforth ABM) ( e.g \cite{Klingert2012c}, \cite{Jumadinova2011a}), and in the real-world (see \cite{Wolfers2006} for a review).
-
-However, to the best of our knowledge, the idea that prediction markets can generate a consensus \emph{on the factors} affecting  uncertain events has never been quantitatively tested.
-ABMs of prediction markets have been studied ( e.g.
-\cite{Klingert2012c}, \cite{Tseng2010a}, or\cite{Jumadinova2011a}), some of which feature communication between agents.
-In these models, however, beliefs about the uncertain outcomes are constructed in rather abstract ways.
-In particular, beliefs are never based on structural models from which agents would derive causal implications.
-Therefore, although convergence of beliefs could be studied in previous ABMs, these models do not investigate the convergence of the underlying explanatory models agents use for prediction (more details in section \ref{related}).
-
-From a public policy perspective, this is a crucial point.
-Effective climate policy not only requires an accurate consensus on future climate \emph{outcomes}.
-It also requires an accurate consensus on the causal \emph{mechanisms}  influencing such outcomes.
-If people agree that temperature will rise, but some believe it will be due to greenhouse gases, while others believe that it will be caused by increased solar activity, inconsistent and ineffective policy will may be implemented.
-In this project, we investigate whether prediction markets can foster such convergence in predictive models.
-
-The main questions we want to answer are:
-
-\begin{enumerate}
-	\item Do traders' predictive models converge to the true model, or can traders converge to a ``bad" equilibrium in which most people believe in a wrong predictive model?
-	\item How is convergence to the true model affected by the parameters and assumptions of the model such as:
-	\begin{enumerate}
-		 \item The initial homophily of the network.
-A strongly homophil network has traders connect preferably with other traders having the same beliefs.
-Given the political polarization of beliefs in climate change, this homophily is likely to be a feature of a climate change belief network.
-One may therefore wonder under which conditions of initial homophily could prediction market decluster the belief networks.
-		 \item The timing of the market.
-\cite{Vandenbergh2013f} have argued that prediction market should trade securities based on long term outcomes in order to guarantee that the market provides information about climate rather than weather.
-However, long term trading might slow down convergence.
-For a trader $i$ to change her predictive model and adopt the model of another trader $j$, $i$ must be convinced  that $j$'s model is better.
-If $i$ relies on the market performances of $j$ to evaluate $j$'s models, $i$ will have to wait that securities be realized before she adapt their explanatory model.
-If the securities are based on long term outcomes, then traders will not adapt their model for before a while.
-Therefore, there may be a trade-off between speed and accuracy.
-		 \item The model assumed to be the true model of the climate.
-An important argument in favor of a climate prediction market is that it is supposed to be an apartisan proposal.
-Defenders of climate prediction markets argue that climate prediction markets will foster convergence to the best approximate model, whatever the actual true model is.
-If climate change is not anthropogenic -- so the argument goes, traders will converge to non-anthropogenic in pretty much the same way as they would converge to anthropogenic models if the climate is truly influenced by human activities.
-	\end{enumerate}
-\end{enumerate}
-
-In this report, we provide preliminary results regarding question 1 and 2.a), together with sensitivity analyses for other parameters of the model.
-See section \ref{results} for important caveats regarding these results.
-Despite these caveats, our first exploratory investigation raise a couple of interesting questions.
-
-First, we observe that convergence to the true model is by no means inevitable.
-In the model, traders adapt their beliefs based on their neighbors monetary performance on the market (more in Section \ref{implem}).
-Therefore, one might expect that traders with the right model of the climate will perform better and  convince their (social network) neighbors to believe in the true model.
-We find that this is not always the case.
-On \url{https://www.youtube.com/watch?v=MS4YfhTU-WY}, one can see the evolution of the belief network for some particular values of the parameter.
-In this case, the green and red nodes correspond to traders believeing in predictive models which are qualitatively very far from the true model (see below).
-One sees that these models are progressively eliminated from the network.
-On \url{https://www.youtube.com/watch?v=aDB7UY7bDGE} however, initial homophily is increased and the network is sufficiently clustered at the outset to prevent convergence.
-In Section \ref{results} we provide further evidence on the possibility of non-convergence.
-The reason of such non-convergence and the precise conditions on parameters which favor convergence are an open question that we plan to tackle in the near future.
-
-% So far, we have been able to construct a minimal model of a prediction market.
-In  the model, traders update their explanatory model of the climate based on the model of their neighbors in a social network.
-If a trader believe in model $x$ and is surrounded by traders believing in model $y$ who happen to do much better then her on the market, then there is a probability that $x$ will change her explanatory model for the model of trader $x$.
-% 
-% At this point the model is not complete.
-In particular, it features a single market mechanism known as the continuous double auction ( CDA).
-We plan on implementing at least a second market mechanism known as the logarithmic market scoring rule (LMSR, \cite{Hanson2012}) in order to compare the performance of both mechanism in terms of prediction accuracy and explanatory model convergence.
-We also need to work on relating some of the parameters with the theoretical literature on network analysis and decision making.\footnote{For instance, we have a parameter which influence the initial level of clustering in the network.
-We need to work on relating the value of this coefficient with commonly used clustering measures in the network literature.}
-
-  \section{Related Work}
-	\label{related}
-
-	\cite{Tseng2010a} study an ABM of a continuous double auction model with various types of market strategies (including two variants of the zero-intelligence agents).
-They compare the behavior of their simulation with data from a prediction market for the outcome of political elections.
-They find that despite their simplicity, zero-intelligence agents capture some salient features of real market data.
-\cite{Klingert2012c} reach similar conclusions with respect to experimental data from  \cite{Hanson2006}.
-In their simulation, \cite{Klingert2012c} compare continuous double auctions with logarithmic market scoring rules and find that the latter generally outperform continuous double auctions in terms of prediction accuracy.
-In some variants of the two last models, agents' strategies feature learning, but these are rather basic and unrelated to a formal predictive model.
-
-	\cite{Jumadinova2011a} conducted research more similar to our project, as they explicitly include information about uncertain events in their model.
-The paper studies a continuous double auction model in which agents update their beliefs about the uncertain events based on information on the event.
-New beliefs are modeled as a weighted sum of the last period's beliefs and market prices.
-The agent's information set changes every period.
-The better the information set, the more likely the agent is to put higher weights on last period's prices when revising their beliefs.
-Not much structure is imposed on the information set and the way it is used to generate the weights in the update function.
-Therefore, the model does not allow for investigating the convergence of predictive models.
-\cite{Ontanon2009} study the effect of deliberation on a simulated prediction market.
-They assume that agents use Case-Based Reasoning \citep{Aamodt1994} to debate about  uncertain outcomes with their neighbors in a social network.
-Although the model could in theory be used to assess convergence of predictive models by looking at the arguments agents adopt after the deliberation, this is not done by the authors.
-Also, the way beliefs are formed and revised through argumentation is rather abstract and might not be ideal to study climate prediction markets.
-
-	\section{Model}
-
-	\label{model}
-
-	In its current form, the ABM is composed of the following elements.
-
-	\begin{description}
-		\setlength\itemsep{0em}
-		\item [A simple ``true" model $F$ of world temperatures.] The model is an additive autoregressive function and is taken  from \cite{Sumner2008b} (it is a simplified version of  RICE99 from \cite{Nordhaus}).
-		\begin{align}
-		\label{true}
-		\begin{split}
-				T_t ~ =~  \lambda T_{t-1} + \beta \underbrace{\big[ \kappa GHG_{t-1} - \alpha P_t^{1/2} + \delta Y_t  \big]}_{\equiv GHG_t} \\
-				- \rho \underbrace{\big[ \omega G_{t-1} + \theta X_{t} \big]}_{\equiv G_t} +  \epsilon_t, 
-		\end{split}
-		\end{align}
-		where
-		\begin{itemize}
-			\item $T_t$ is the temperature at time $t$,
-			\item $GHG_t$ are greenhouse gas emissions at time $t$,
-			\item $P_t$ are spendings in policies alleviating climate change at $t$,
-			\item $Y_t$ is time $t$'s world Gross Domestic Product (GDP),
-			\item $G_{t}$ is the effect in period $t$ of geo-ingeneering produced before $t$. 
-			\item $X_t$ are spending in geo-ingeneering in period  $t$.
-		\end{itemize}
-
-		\item[ Beliefs.]  Traders use $5$ models to  forecast future temperatures in order to trade on the prediction market.
-These $5$ models are interpreted as the trader's beliefs about the true climate model.
-Again, these models are taken from \cite{Sumner2008b}).
-They are meant to represent pervasive positions on climate change in the public debate.
-Four of these models are approximations of the true model in \eqref{true} described below, and one of them is the true model itself.\footnote{Thus some traders use the true model to make predictions.
-This does not mean that traders using the true model make perfectly accurate predictions.
-Although these traders believe in the correct \emph{functional form} of the model, they still need to calibrate it based on limited noisy data.
-Therefore, the values they use for  the \emph{parameters} of the model will typically still be off.}
-
-		\begin{enumerate}
-			\setlength\itemsep{0em}
-			\item Anthropogenic climate change is a myth ($\beta = \rho = 0$),
-			\begin{align}
-			T^1_t = \lambda T_{t-1} +  \epsilon_t.
-			\end{align}
-			\item Anthropogenic climate change is real but policies and geoengineering are ineffective ($\omega = \rho = 0$),
-			\begin{align}
-			T^2_t & = \lambda T_{t-1} + \beta \big[ \kappa GHG_{t-1}  + \delta Y_t  \big] +  \epsilon_t.
-			\end{align}
-			\item Anthropogenic climate change is real, policies are effective, but not geoengineering ($\rho = 0$),
-			\begin{align}
-			T^3_t = \lambda T_{t-1} + \beta \big[ \kappa GHG_{t-1} - \alpha P_t^{1/2} + \delta Y_t  \big] +  \epsilon_t.
-			\end{align}
-			\item Anthropogenic climate change is real, policies are ineffective, but geoengineering works($\alpha = 0$),
-			\begin{align}
-			\begin{split}
-		T^4_t = \lambda T_{t-1} + \beta \big[ \kappa GHG_{t-1} + \delta Y_t  \big]\\ - \rho \big[ \omega G_{t-1} + \theta X_{t} \big] +  \epsilon_t.
-			\end{split}
-			\end{align}
-
-		\end{enumerate}
-
-		\item[ Traders.] Traders are endowed with some initial money in an Experimental Currency Unit (ECU), and an initial approximate model.
-Approximate models are distributed at random among the traders with each model having the same probability of being assigned to any trader.
-Traders use the distribution of temperatures obtained from their approximate models to determine their reservation price for different securities.
-
-		Securities pay 1 ECU  if the actual temperature at some time $t^*$ falls between a certain range.
-Traders are assumed to be \emph{risk-neutral expected utility maximizers}.
-Therefore, their reservation price for a security $\tilde{s}$ is simply their assessment of the probability that the temperature at time $t^*$ falls in the range covered by $\tilde{s}$.
-
-		  Based on their reservation price, traders behave as ``\emph{zero-intelligence}" agents \citep{Gode1993}, that is they sell at a random price above their reservation, and buy at a random price below their reservation.
-These strategies are simple but have been shown to provide good approximations of the behavior of prediction markets (see for instance \cite{Klingert2012c}), and financial markets more broadly.
-
-		\item[Continuous-Double Auction (CDA) market structure.] Continuous-Double auctions ( or some variants thereof) are common \emph{market mechanisms}, i.e.
-procedures to match buy and sell orders.
-CDA are notably used on large stock markets (see for instance \cite{Tseng2010a}).
-Our model features a particular version of CDA (more in section \ref{implem}).
-
-		\item[A social network.] Traders are part of a social network.
-Every time a security is realized, each trader $i$ looks at the performance of their neighbors in the network.
-If one of $i$'s neighbors, say $j$, is richer than $i$, $i$ interprets this has an indication that $j$ has a better approximate model.
-Then $i$ considers adopting $j$'s approximate model.\footnote{Traders start with the same initial amount of money, so differences in money among traders can only come from  traders' interactions through the market.}
-
-		An example of such social network is depicted in Figure \ref{socnet}.
-This corresponds to the initial social network for some of our simulations.
-The color of the nodes identify the approximate model of the trader.
-Notice the belief-homophily of the initial network which can be seen by the high clustering of traders with the same approximate model.
-Although traders can change their approximate model (i.e.
-colors can change in the graph), the connections between traders do not change as the market unfolds (i.e.
-the edges are fixed).
-	\end{description}
-
-	\begin{figure}
-		\begin{center}
-		\includegraphics[scale = 0.5]{network.png}
-		\end{center}
-		\caption{An example of the initial state of a belief network \label{socnet}}
-	\end{figure}
-
-	\subsubsection*{Timing}
-
-	The time periods $t$ are grouped in trading \emph{sequences}.
-In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence.
-For instance, the third trading sequence might start in period $t=300$ and end in period $t = 399$.
-In this case, a security traded in the third sequence pay 1 ECU if the temperature at $t=400$ falls into the range of temperatures covered by the security.
-
-	   At each time $t$, traders are assumed to know the past value of the temperature $T_{0:t}$ and the vector of past explanatory variables $Z_{0:t}$.
-In a sequence finishing at time $t^*$,  traders also have common knowledge of $Z_{t:t^*}$, the future values of explanatory variables up to $t^*$.
-However, at any $t$, traders do not know the value of any future temperatures.
-In particular, in a given sequence, traders do not know the value of $T_{t^*}$.
-Traders can only predict $T_{t^*}$ using their approximate model and their knowledge of $Z_{t:t^*}$.
-Notice that because $Z_{t:t^*}$ is common knowledge, in each period $t$ every trader $i$ with the same approximate model forms the same expectation
-	   \begin{align*}
-	  \mu_{ti} \equiv E_t( T_{t^*} ~|~ T_{0:t}, Z_{t:t^*},  \text{ approximate model of } i \text { at } t).
-	   \end{align*}
-	   We assume that the probability density that trader $i$ associate with $T_{t^*} = \tilde{T}$ at any $t$ is
-	   \begin{align*}
-	   f_{it}(T) = \mu_{ti} + N(0, \sigma_T),
-	   \end{align*}
-
-	   where $N$ stands for a normal distribution and $\sigma_T$ is a fixed number determined by the true variance of temperatures.
-
-	    At each time $t$, traders
-	\begin{itemize}
-		\setlength\itemsep{0em}
-		\item recalibrate their approximate model based on the new set of past data available at $t$,
-		\item compute their belief-density for $T_{t^*}$ and use it to determine the expected value they attached to each security,
-		\item and trade on the market based on their market strategy at $t$.
-	\end{itemize}
-
-	At $t^*$, when the sequence ends, there is only one security $s^*$ associated with a range of temperatures including the actual $T_{t^*}$. Thus,
-	\begin{itemize}
-			\setlength\itemsep{0em}
-		\item traders receive 1 ECU per unit of $s^*$ they own, and
-		\item  consider adapting their approximate model as described above.
-	\end{itemize}
-
-	\section{Implementation}
-	\label{implem}
-	The model is implemented in the \texttt{R} programming language,  using the \texttt{igraph} package to model the social network.
-Each trader in the model is a node of an \texttt{igraph} network.
-Because the model uses time-series data and  traders calibrate statistical models based on these data, \texttt{R}, primarily used for statistical computing, is particularly convenient.
-
-	 Traders calibrate their approximate models based on past-data using Pooled Ordinary Least Squares.
-This is questionable from a statistical point of view, but provides an approximation of the way traders may form their forecast on actual prediction markets.
-Because the model is in \texttt{R}, it is also very easy to change the estimation method or model to virtually any statistical model, and the effect of altering models may easily be studied in the future.
-
-	 In each period $\tilde{t}$, our version of the CDA works as follows.
-
-	\begin{enumerate}
-			\setlength\itemsep{0em}
-		\item Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
-		\item Every trader $i$ also chooses at random a security $s_i^S$ she will try to sell among the securities she owns a positive amount of (if any).
-		\item Traders then decide of their selling $p_i^S$ and buying price $p_i^S$.
-They do so based on their believed-density of temperatures at the end of the sequence, and following the zero-intelligence rule described above.
-		\item Traders go to the market one at the time, in a random order.
-		\item When $i$ comes to the market, she place \emph{limit orders} in the order book.
-These orders  specify that $i$ is willing to buy $s_i^B$ at  any price below $p_i^B$, and to sell $s_i^S$ at any price above $p_i^S$.
-		\item The market maker then tries to match $i$'s orders with some order which was put in the order book \emph{before} $i$ came to the market.
-		\item If there are  outstanding sell offers for $s_i^B$ at price $p$ lower than $p_i^B$, then a trade is concluded.
-Trader $i$ buys one unit from the sellers who sells at the \emph{highest} price below $p_i^B$, and the sell and buy offers are removed from the order book.
-		\item If there are  outstanding buy offers for $s_i^S$ at price $p$ higher than $p_i^S$, then a trade is concluded.
-Trader $i$ sells one unit to the buyer who buys at the \emph{highest} price above $p_i^S$, and the  sell and buy offers are removed from the order book.
-		\item Whenever all traders have come to the market, any remaining outstanding offer is removed from the order book, and the trading period is concluded. 
-	\end{enumerate}
-
-The model depends on the following parameters. 
-\begin{description}
-	\setlength\itemsep{0em}
-	\item[Network parameters.]~
-	\begin{description}
-			\setlength\itemsep{0em}
-		\item [\texttt{n.traders}:] the number of traders.
-		\item [\texttt{n.edg} :] the number of edges in the social network. This number is fixed throughout the experiments.   
-		\item [\texttt{seg} :] determines the initial degree of homophily in the network.
-The higher \texttt{seg}, the higher the initial homophily.\footnote{ When constructing the \texttt{n.edg} edges of the network, the probability that a link between two traders be formed depends on whether the traders share the same approximate model in the following way
-			\begin{align*}
-			\begin{cases}
-			\frac{(1-\texttt{seg})}{\texttt{n.edg}},  &\text{if the traders have different approximate models}\\
-			\frac{1}{\texttt{n.edg}},  & \text{otherwise}
-			\end{cases}
-			\end{align*}}
-	\end{description}
-
-\item[Market structure parameter.]~
-
-		\begin{description}
-			\setlength\itemsep{0em}
-			\item [\texttt{market.complet}.] Determines market's completeness, i.e.
-the number of securities which can be traded.
-With more securities, the interval of temperatures corresponding to each security is smaller.
-Traders can then trade on more precise temperature intervals.
-Higher values of \texttt{market.complet} may, however, reduce the number of exchanges.
-Because traders pick the securities they buy and sell at random, the probability that a match between sellers and buyers is found is lower for higher values of \texttt{market.complet}.
-		\end{description}
-
-\item[Behavioral parameters.]~
-
-\begin{description}
-	\setlength\itemsep{0em}
-	\item [\texttt{risk.tak}.] Determines the distribution of risk taking behavior.
-The higher \texttt{risk.tak}, the more traders	will try to buy (resp.
-sell) lower (resp.
-higher) than their reservation price.
-Formally, at time $t$, trader $i$ picks her buying (resp.
-selling) prices randomly in the interval  $[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it}  * (1 - \texttt{risk.tak}_i)]$ (resp.
-$[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it} * (1  + \texttt{risk.tak}_i)]$), where $\texttt{reserv}_{it}$ is $i$'s reservation price at time $t$ for the security $i$ picked to buy (resp.
-sell).
-	\item [\texttt{ideo}.]  Determines the degree of ``ideology" of traders. If \texttt{ideo} is high, traders will not
-	 revise their approximate models easily, even when faced with
-	 strong evidence that their neighbors are doing better than them.\footnote{ For each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$.
-The value of $d_{i}$ is the probability that $i$ adopts one of her neighbors' approximate model if this neighbor is doing better than $i$ at the end of the sequence (in monetary terms).}
-\end{description}
-
-	\item[Timing parameters.]~
-
-	\begin{description}
-			\setlength\itemsep{0em}
-		\item [\texttt{burn.in}.] The number of burn-in periods in which no securities are traded (necessary to allow approximate models to be estimated in the first period).
-		\item [\texttt{n.seq}.] Number of trading sequences.
-		\item [\texttt{horizon}.] Number of trading periods in each trading sequence.
-	\end{description}
-
-\end{description}
-
-	\section{Results}
-
-	\label{results}
-
-In this section, we present preliminary sensitivity analyses and exploratory results based on the above model.
-In general, we will be interested in the fraction of traders who believe in the true model.
-For sensitivity analysis, the relevant outcome will be the difference between the fraction of traders who believe in the true model at the beginning of the experiment, and the fraction who believe in the true model at the end of all trading sequences.
-
-	\begin{figure*}
-		\begin{center}
-			\includegraphics[scale = 0.5]{sobol_sa.png}
-			\caption{ Effect of parameters on the difference between the initial and final fraction of traders who believe in the true model \label{nondirect}}
-		\end{center}
-	\end{figure*}
-
-Before presenting these results, let us mention some important caveat.
-First, the model currently uses artificial data.
-The values of the explanatory variables are in not connected with the observed past values of greenhouse gas concentrations, GDP, etc.
-As a consequence, the results should be viewed as no more than a proof of concept, and no conclusions on actual {climate} prediction markets should be drawn from them.
-A second reason to be cautious about the following results is that they rely on varying only a handful of the model's parameters, namely
-\begin{itemize}
-	\item \texttt{ideo}, 
-	\item \texttt{market.complet}, 
-	\item \texttt{risk.tak}, and
-	\item  \texttt{seg}.
-\end{itemize}
-
-Other parameters are fixed to the following values.
-
-\begin{itemize}
-	\item \texttt{n.traders} = 100,
-	\item \texttt{n.edg} = 150,
-	\item \texttt{burn.in} = 100,
-	\item \texttt{n.seq} = 20, and
-	\item \texttt{horizon} = 10.
-\end{itemize}
-
-We first present two sensitivity analyses for the four parameters above.
-Both sensitivity analyses are based on a Latin Hypercube Sampling of the four parameters, with uniform prior distributions.
-The support for the distributions are:
-
-\begin{itemize}
-	\item \texttt{ideo} : $(0,1)$
-	\item \texttt{market.complet} : $ \{1,\dots,1000\}$,  
-	\item \texttt{risk.tak}  : $(0,1)$, and
-	\item  \texttt{seg} : $(0,1)$.
-\end{itemize}
-
-	The first analysis is similar to an ANOVA, and is based on  variance decomposition below \cite[Section 2.3.1]{Confalonieri2010}.
-Let the four varied parameters be denoted $x_i$ for $i =1,\dots,4$, and the difference between the final and the initial fraction of traders who believe in the true model be $Y$.
-Then we have
-	\begin{align}
-	\label{variance}
-	V(Y) = \sum_{i=1}^4 \Big[  D_i + 
-	\sum_{i \leq j \leq 4}^4 \big[ D_{ij} + \dots + 
-	D_{1234} \big] \Big],
-	\end{align}
-
-	where $D_i = V(E[Y|x_i])$, $D_{ij} = V(E[Y|x_i,x_j]) - D_i - D_j$ and so on. Dividing both sides of \eqref{variance} by $V(Y)$ we obtain a normalized sum. The first terms of this normalized sum are
-	\begin{align}
-	S_i = \frac{V(E[Y|x_i])}{V(Y)},
-	\end{align}
-	which are called the first-order effects (of $i$).
-First-order effect $S_i$ represents the proportion of the variance which is explained by changes in $x_i$ once the effect of variations in the other $x_j$ ( $j\neq i$) has been accounted for.
-In Figure \ref{nondirect},  the value of $S_i$ is represented in red for the four parameters, together with their respective confidence intervals.
-One can see that none of the parameters' first-order effects is significant.
-
-	\begin{figure*}
-		\begin{center}
-			\includegraphics[scale = 0.5]{pc_sa.png}
-			\caption{Partial rank correlation coefficients and confidence interval \label{prcc}}
-		\end{center}
-	\end{figure*}
-
-	 Depicted in blue in Figure \ref{nondirect} are the total effects of the parameters.
-For each parameter $x_i$, the total effect corresponds to the sum of the net impacts of every possible combination of $x_i$ with other parameters.
-Formally, for each $x_i$ the blue dot and confidence interval in Figure \ref{nondirect} corresponds to
-	 \begin{align}
-	 \begin{split}
-ST_i =  S_i  + \sum_{j\neq i}  S_{ij} + \dots +  S_{1234} 
- \end{split},
-	 \end{align}
-
-	where  $S_{ij} = \sfrac{ V(E[Y|x_i,x_j]) - \allowbreak V(E[Y|x_i])- V(E[Y|x_j])}{V(Y)}$ and the other terms are defined similarly.
-Figure \ref{nondirect} indicates that most the effect of the parameters happens through interactions with other parameters.
-
-	\begin{figure}
-		\includegraphics[scale = 0.4]{Rplot01.png}
-		\includegraphics[scale = 0.4]{Rplot02.png}
-		\includegraphics[scale = 0.4]{Rplot03.png}
-		\caption{Evolution of the fraction of traders who believe in the true model (y-axis) as a function of the number of trading sequences (x-axis) for 30 random draws of the parameters.\label{example}}
-	\end{figure}
-
-	The variance analysis does not provide information on the sign of the first-order effects.
-To investigate signed-effect, we turn to a second sensitivity analysis known as partial \emph{rank} correlation coefficient \citep{Marino2008}.
-Partial correlation computes the linear relationship between the part of the variation of $x_i$ and $Y$ that are linearly independent of other $x_j$  ($j \neq i$).
-First, we compute predicted value $\hat{x}_i$ and $\hat{Y}$ based on an OLS regression of $x_i$ and $Y$ on $x_j$  ($j\neq i$)
-	\begin{align}
-	\begin{split}
-	\hat{x}_i = c^0_{OLS} + \sum_{j \neq i} c^j_{OLS} x_j,\\
-	\hat{Y} = b^0_{OLS} + \sum_{j \neq i} b^j_{OLS} x_j.
-	\end{split}
-	\end{align}
-	Then we compute the coefficient of correlation between $(x_i - \hat{x}_i)$ and $(\hat{Y} - \hat{Y})$.
-
-	The only difference with the partial correlation and partial \emph{rank} correlation (which we use here), is that the dependent variable $Y$ is first ranked-transformed in order to foster more uniform residuals.\footnote{That is every $Y_s$ in the data set is replace by a number $f(Y_s) \equiv |\{Y_r~|~ Y_s > Y_r\}|$, where $Y_r$ are the other observed values of $Y$ in the data set.} Such transformation preserves the sign of the correlation coefficient.
-However, because data are transformed, we loose the meaning of the coefficient's magnitude with respect to the original data.
-The partial rank correlation coefficients are reported in Figure \ref{prcc}.
-
-	 It is surprising that the net effect of \texttt{seg} on $Y$ is positive.
-This suggests that a higher initial belief-homophily in the network increases (or reduces the decrease in) the adoption of the true model.
-Intuitively, traders endowed with the correct explanatory model should be better at predicting climate outcomes.
-Therefore, if they are connected to traders with different approximate models, one could expect that these other traders will eventually adopt the true model.
-If that were true, the more traders endowed with the true model are disseminated across the network (i.e.
-the smaller \texttt{seg}) the better the result should be in terms of adoptions of the true model.
-The results  in Figure \ref{prcc} indicates that, if at all present, this kind of dynamics are not sufficient to explain changes in the adoption of the true model.
-
-	Exploratory investigations also cast a doubt on the ability of the modeled prediction market to foster convergence to the true model.
-Figure \ref{example} contains plots for the evolution through trading sequences of the fraction of traders who believe in the true model.
-Each line corresponds to a different random value for the four parameters.
-The figure show a wide variety of dynamics and indicate that under many parameter configurations, the fraction of traders who believe in the true model significantly reduces (sometimes almost completely).
-
-	Finally, it appears that the current version of the model is too random.
-Figure \ref{random} represents the evolution of $Y$ for 10 different runs of the model \emph{using the same value for the parameters}, namely
-
-	\begin{itemize}
-		\item \texttt{ideo} = 0.5,
-		\item \texttt{market.complet} = 100,
-		\item \texttt{risk.tak} and = 0.02, and
-		\item  \texttt{seg} = 0.5.
-	\end{itemize}
-
-	As can be seen from the figure, a lot of randomness is built into the model independently of changes in the parameters.
-For instance, notice that the initial fraction of agents who believe in the true models varies from one run of the model to the other.
-The probability that a trader starts with the true model is fixed at $\sfrac{1}{5}$.
-But only 100 traders are assigned an initial approximate models which explains such variability.
-Other randomness built into the model includes the distribution of the risk and ideology parameters $\texttt{risk.tak}_i$ and $\texttt{ideo}_i$.\footnote{Currently, $\texttt{risk.tak}_i$ is drawn at random from $[0, \texttt{risk.tak}]$.
-Thus, increasing \texttt{risk.tak} increases the chance that traders have higher $\texttt{risk.tak}_i$, but there is a significant amount of randomness in the assignment of the $\texttt{risk.tak}_i$.
-The same is true for $\texttt{ideo}_i$.} This might partly explain why we do not get clearer results in the sensitivity analysis.
-
-	\section{Conclusion}
-
-	In this report, we described a minimal ABM of a climate prediction market in which traders adapt their belief of the climate based on the monetary performance of their neighboring traders in a social network.
-The market mechanism is a variant of the famous continuous double-auction, and traders are risk-neutral expected utility maximizers who behave on the market as zero-intelligence agents \citep{Gode1993}.
-		\section{Related Work}
-	Future developments involve :
-
-	\begin{itemize}
-		\item Reducing the randomness built in the model and simplify the model in order to obtain clearer results.
-One way of simplifying the model would be to reduce the number of approximate models (e.g.
-anthropogenic climate change is a myth vs.
-anthropogenic climate change is true).
-		\item Implementing a second market rule (LMSR, \cite{Hanson2012}).
-		\item Using actual time-series data for the temperature and the explanatory variables.
-At this stage, the data used in the model are randomly generated time series with no connection to actual temperatures, or to observed values of the explanatory variables.
-		\item Varying more parameters. In particular, changes in timing parameters should be studied in order to tackle the issue of optimal timing in climate prediction markets.
-		\item Studying the model under different assumption about the true model in order to check whether climate prediction markets are an apartisan proposal.
-	\end{itemize}
-
-	\begin{figure}
-		\includegraphics[scale = 0.4]{random.png}
-		\caption{Evolution of the fraction of traders who believe in the true model (y-axis) as a function of the number of trading sequences (x-axis) for 10 different run of the model with \emph{fixed} values of the parameters ( \texttt{ideo} = 0.5,
-			 \texttt{market.complet} = 100,
-			 \texttt{risk.tak} and = 0.02, and
-			  \texttt{seg} = 0.5.). \label{random}}
-	\end{figure}
-
-	\bibliography{library}
+\bibliography{library.bib}
 \bibliographystyle{apalike}

### 2015-08-29 Temp Latex work/Supplemental material/Supplemental material.tex
First (rough) draft of detailed description of the model for supplemental material (plus small modification to visualization tools in main.R)
This is a rough draft but feel free to check it out and to give me feedbacks next time we meet.

@@
 \documentclass[10pt,a4paper]{article}
-\usepackage{fullpage}
 \usepackage[latin1]{inputenc}
@@
 
+   \subsection{Detailed description of the model}
+
+ 	\label{model}
+
+ 	\subsubsection{Elements of the model}
+
+ 	The ABM is made of the following elements.
+
+ 	\begin{description}
+ 		\setlength\itemsep{0em}
+ 		\item [A simple ``true" model of world temperatures.] 
+ 		The true model is determined by parameter  \texttt{true.model}.
+
+ 		\begin{description}
+ 			\item[If \texttt{true.model} = 0,] then anthropogenic climate change is a myth and the temperatures follow the autoregressive process 
+ 			\begin{align}
+ 			\label{eq:true0}
+ 			\begin{split}
+ 			T_t ~ =~  \lambda T_{t-1} +  \epsilon_t, 
+ 			\end{split}
+ 			\end{align}
+ 			where $T_t$ is the temperature at time $t$, and $\epsilon^T_t$ is a random error term.
+
+ 			The value of $\lambda$ is obtained by calibration on actual data.
+Specifically, $\lambda$ is the ordinary least square estimator of \eqref{eq:true0}, where the temperature time series are US yearly averages based from 1895 to 2010, based on US monthly averages from \url{http://www7.ncdc.noaa.gov/CDO/CDODivisionalSelect.jsp} ( TAVG in the data set, retrieved on May 5th 2015).
+
+ 			The distribution of $\epsilon_t$ is the empirical distribution of the residual from this linear regression.
+
+ 			\item[If \texttt{true.model} = 1,] then anthropogenic climate is true and the temperatures follow the process 
+ 			\begin{align}
+ 			\label{eq:true1}
+ 			\begin{split}
+ 			T_t ~ =~  \beta GHG_{t-1} +  \mu_t, 
+ 			\end{split}
+ 			\end{align}
+ 			where $GHG_t$ is the level of greenhouse gas emissions at time $t$, and $\epsilon^{GHG}_t$ is a random error term.
+
+ 			Again, the value of $\beta$ is obtained by calibration on actual data.
+Specifically, $\beta$ is the ordinary least square estimator of \eqref{eq:true1}, where the temperature time series are as in \texttt{true.model} = 0, and the $GHG$ time series are Global Mean CO2 Mixing Ratios (ppm) on the same time period from \url{http://data.giss.nasa.gov/modelforce/ghgases/Fig1A.ext.txt} ( MixR in the data set, retrieved on May 5th 2015).
+
+ 			The distribution of $\mu_t$ is the empirical distribution of the residual from this linear regression.
+ 		\end{description}
+
+ 		In either cases (\texttt{true.model} = 0 and \texttt{true.model} = 1), the chosen true model is used to generate and artificial temperature time series $\{\hat{T}_t\}_{t=1895}^{t=2010}$.
+
+		We emphasize the artificial time series is in general different from the temperature in the empirical data set.	
+		\begin{itemize}
+			\item When  \texttt{true.model} = 0, $\hat{T}_{1895}$ is chosen to be the  temperature in 1895 from the empirical data set, and $\{\hat{T}_t\}_{t=1896}^{t=2010}$ are computed recursively using \eqref{eq:true0}.
+			\item When  \texttt{true.model} = 1, the value of $GHG_t$ for all $t \in \{1895,\dots,2010\}$ is taken from the empirical data set and $\{\hat{T}_t\}_{t=1895}^{t=2010}$ are computed using \eqref{eq:true1}.
+		\end{itemize}
+
+ 		\item[ Beliefs.]  Traders use either \eqref{eq:true0} or \eqref{eq:true1} to  forecast future temperatures in order to trade on the prediction market.
+These $2$ models are interpreted as the trader's beliefs about the true climate model.
+They are meant to represent pervasive positions on climate change in the public debate.
+In order to approximatively match the current configuration of beliefs in climate change in the US, model \eqref{eq:true0} is randomly assigned to half of the traders, while \eqref{eq:true1} is assigned to the other half of the traders.
+
+ 		Thus, both when \texttt{true.model} = 0 and when  \texttt{true.model} = 1,  some traders use the true model to make predictions.
+However, this does not mean that traders using the true model make perfectly accurate predictions.
+Although these traders believe in the correct \emph{functional form} of the model, they still need to calibrate it based on limited noisy data.
+Therefore, the values they use as  the \emph{parameters} of the model will typically still be different from $\lambda$ and $\beta$ (see description of the model's timing below).
+
+ 		\item[ Traders.] Traders are initial endowed with one single unit Experimental Currency Unit (ECU).  
+ 		Traders use their believed model to forecast the distribution of future temperatures and determine their reservation price for different securities.
+Securities pay 1 ECU  if the actual temperature at some time $t^*$ falls between a certain range.
+Traders are assumed to be \emph{risk-neutral expected utility maximizers}.
+Therefore, their reservation price for a security $\tilde{s}$ is simply their assessment of the probability that the temperature at time $t^*$ falls in the range covered by $\tilde{s}$.
+
+ 		 Based on their reservation price, traders behave as ``\emph{zero-intelligence}" agents \citep{Gode1993}, that is they sell at a random price above their reservation, and buy at a random price below their reservation.
+The distribution from which buy and sell prices are drawn is determined by the traders risk attitude, characterized by parameter \texttt{risk.tak}$_i$ (more details below).
+These strategies are simple but have been shown to provide good approximations of the behavior on prediction markets (see for instance \cite{Klingert2012c}), and in financial markets more broadly.
+
+ 		\item[Continuous-Double Auction (CDA) market structure.] Continuous-Double auctions (or some variants thereof) are common \emph{market mechanisms}, i.e.
+procedures to match buy and sell orders.
+CDA are notably used on large stock markets (see for instance \cite{Tseng2010a}).
+Our model features a particular version of CDA through which traders exchange securities (more in section \ref{implem}).
+
+ 		\item[A social network.] Traders are part of a social network.
+Every time a security is realized, each trader $i$ looks at the performance of their neighbors in the network.
+If one of $i$'s neighbors, say $j$, is richer than $i$, $i$ interprets this has an indication that $j$ has a better approximate model.
+Then $i$ considers adopting $j$'s approximate model.\footnote{Traders start with the same initial amount of money, so differences in money among traders can only come from  traders' interactions through the market.} For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is characterized by parameter \texttt{ideo}$_i$.
+
+ 		An example of such social network is depicted in Figure \ref{socnet}.
+ 		We assume that the initial network is segregated, that is traders who believe in \eqref{eq:true0} (resp.
+\eqref{eq:true1}) are more likely to be linked with other traders who believe in \eqref{eq:true0}  (resp.
+\eqref{eq:true1}).
+The degree of segregation is determined by parameter \texttt{seg} (more details below).
+Although traders can change their approximate model as time passes, the \emph{connections} between traders do not change as the market unfolds (i.e.
+the edges are fixed).
+ 	\end{description}
+
+ 	\begin{figure}
+ 		\begin{center}
+ 		\includegraphics[width = 0.8\textwidth]{figures/segregated.png}
+ 		\end{center}
+ 		\caption{An example of the initial state of a belief network with 50 traders.
+Red nodes are traders who initially believe in \eqref{eq:true0} and black nodes are traders who initially believe in \eqref{eq:true1}.
+Notice that the  \label{socnet}}
+ 	\end{figure}
+
+ 	\subsubsection{Timing}
+
+ 	The time periods $t$ are grouped in trading \emph{sequences}.
+In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence.
+For instance, the third trading sequence might start in period $t=1964$ and end in period $t = 1970$.
+In this case, a security traded in the third sequence pay 1 ECU if the temperature at $t=1970$ falls into the range of temperatures covered by the security.
+
+ 	   At each time $t$, traders are assumed to know the past value of the temperature $\hat{T}_{0:t}$ and greenhouse gas emissions  $GHG_{0:t}$.
+In a sequence finishing at time $t^*$,  traders also have common knowledge of $GHG_{t:t^*}$, the future values of greenhouse gas emissions up to $t^*$.
+However, at any $t$, traders do not know the value of any future \emph{temperatures}.
+In particular, in a given sequence, traders do not know the value of $\hat{T}_{t^*}$.
+Traders can only predict $\hat{T}_{t^*}$ using their approximate model and their knowledge of $\hat{T}_{0:t}$ and  $GHG_{0:t^*}$.
+Notice that because $GHG_{0:t^*}$ is common knowledge, in each period $t$ every trader $i$ with the same approximate model forms the same \emph{pseudo}-expectation (see footnote \ref{foot:pseudo})
+ 	   \begin{align}
+ 	   \label{eqref:distrib}
+ 	  \mu_{ti} \equiv E_t( \hat{T}_{t^*} ~|~ \hat{T}_{0:t}, GHG_{0:t^*},   i's \text{ approximate model of } \text { at } t).
+ 	   \end{align}
+ 	   We assume that the believed probability distribution of $\hat{T}_{t^*}$ for trader $i$  at any time $t$ is
+ 	   \begin{align*}
+ 	   \mu_{ti} + \eta_t,
+ 	   \end{align*}
+
+ 	   where the distribution $\eta_t$ is obtained by trader $i$ by bootstrapping the residuals from the regression of her believed true model over the past periods.\footnote{This explains the use of the term \emph{pseudo}-expectation.
+Because $\eta_t$ is not bound to have expected value zero, a traders' effective expected value of $\hat{T}_{t^*}$ may be different from $\mu_{ti}$.
+\label{foot:pseudo}}
+
+ 	    At each time $t$, traders
+ 	\begin{itemize}
+ 		\setlength\itemsep{0em}
+ 		\item recalibrate their approximate model based on the new set of past data available at $t$ (via pooled ordinary least squares),
+ 		\item compute their belief-density for $\hat{T}_{t^*}$ and use it to determine the expected value they attached to each security,
+ 		\item and trade on the CDA market based on their zero-intelligence decision rule as follows. 
+ 		\begin{enumerate}
+ 			\setlength\itemsep{0em}
+ 			\item Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
+ 			\item Every trader $i$ also chooses at random among the securities she owns a positive amount of (if any) a security $s_i^S$ she will try to sell.
+ 			\item Traders then decide of their selling $p_i^S$ and buying price $p_i^S$.
+To do so, they compute the expected value of the securities $s_i^B$ and $s_i^S$ according to \eqref{eqref:distrib}.
+Then they set a selling and a buying price following the zero-intelligence rule described above.
+ 			\item Traders go to the market one at the time, in a random order.
+ 			\item When $i$ comes to the market, she place \emph{limit orders} in the order book.
+These orders  specify that $i$ is willing to buy $s_i^B$ at  any price below $p_i^B$, and to sell $s_i^S$ at any price above $p_i^S$.
+ 			\item The market maker then tries to match $i$'s orders with some order which was put in the order book \emph{before} $i$ came to the market.
+ 			\item If there are  outstanding sell offers for $s_i^B$ at price $p$ lower than $p_i^B$, then a trade is concluded.
+Trader $i$ buys one unit from the sellers who sells at the \emph{highest} price below $p_i^B$, and the sell and buy offers are removed from the order book.
+ 			\item If there are  outstanding buy offers for $s_i^S$ at price $p$ higher than $p_i^S$, then a trade is concluded.
+Trader $i$ sells one unit to the buyer who buys at the \emph{highest} price above $p_i^S$, and the  sell and buy offers are removed from the order book.
+ 			\item Whenever all traders have come to the market, any remaining outstanding offer is removed from the order book, and the trading period is concluded. 
+ 		\end{enumerate} 
+ 	\end{itemize}
+
+ 	At $t^*$, when the sequence ends, there is only one security $s^*$ associated with a range of temperatures including the actual $\hat{T}_{t^*}$. Thus,
+ 	\begin{itemize}
+ 			\setlength\itemsep{0em}
+ 		\item traders receive 1 ECU per unit of $s^*$ they own, and
+ 		\item  consider adapting their neighbors' approximate model as described above.
+ 	\end{itemize}
+
+ 	\subsubsection{The models' parameters and details about their effects}
+ 	\label{param}
+
+ The model depends on the following parameters. 
+ \begin{description}
+ 	\setlength\itemsep{0em}
+ 	\item[Network parameters.]~
+ 	\begin{description}
+ 			\setlength\itemsep{0em}
+ 		\item [\texttt{n.traders}:] the number of traders.
+ 		\item [\texttt{n.edg} :] the number of edges in the social network. This number is fixed throughout the experiments.   
+ 		\item [\texttt{seg} :] determines the initial degree of homophily in the network. The higher \texttt{seg}, the higher the initial homophily.
+
+ 		When constructing the \texttt{n.edg} edges of the network, the probability that a link between two traders be formed depends on whether the traders share the same approximate model in the following way
+ 			\begin{align*}
+ 			\begin{cases}
+ 			\frac{(1-\texttt{seg})}{\texttt{n.edg}},  &\text{if the traders have different approximate models}\\
+ 			\frac{1}{\texttt{n.edg}},  & \text{otherwise}
+ 			\end{cases}
+ 			\end{align*}
+ 	\end{description}
+
+ \item[Market structure parameter.]~
+
+ 		\begin{description}
+ 			\setlength\itemsep{0em}
+ 			\item [\texttt{market.complet}.] Determines market's completeness, i.e.
+the number of securities which can be traded.
+With more securities, the interval of temperatures corresponding to each security is smaller.
+Traders can then trade on more precise temperature intervals.
+Higher values of \texttt{market.complet} may, however, reduce the number of exchanges.
+Because traders pick the securities they buy and sell at random, the probability that a match between sellers and buyers is found is lower for higher values of \texttt{market.complet}.
+ 		\end{description}
+
+ \item[Behavioral parameters.]~
+
+ \begin{description}
+ 	\setlength\itemsep{0em}
+ 	\item [\texttt{risk.tak}.] Determines the distribution of risk taking behavior.
+The higher \texttt{risk.tak}, the more traders	will try to buy (resp.
+sell) lower (resp.
+higher) than their reservation price.
+Formally, at time $t$, trader $i$ picks her buying (resp.
+selling) prices randomly in the interval  $[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it}  * (1 - \texttt{risk.tak}_i)]$ (resp.
+$[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it} * (1  + \texttt{risk.tak}_i)]$), where $\texttt{reserv}_{it}$ is $i$'s reservation price at time $t$ for the security $i$ picked to buy (resp.
+sell).
+
+ 	For each trader $i$, the level of \texttt{risk.tak}$_i$ is drawn uniformly at random from $[0,$\texttt{risk.tak} $]$
+
+ 	\item [\texttt{ideo}$_i$.]  Determines the degree of ``ideology" of traders. If \texttt{ideo} is high, traders will not
+ 	 revise their approximate models easily, even when faced with
+ 	 strong evidence that their neighbors are doing better than them.
+
+ 	  For each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$.
+The value of $d_{i}$ is the probability that $i$ adopts one of her neighbors' approximate model if this neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
+
+ 	  For each trader $i$, the level of \texttt{ideo}$_i$ is drawn uniformly at random from $[0,$\texttt{ideo} $]$
+ \end{description}
+
+ 	\item[Timing parameters.]~
+
+ 	\begin{description}
+ 			\setlength\itemsep{0em}
+ 		\item [\texttt{burn.in}.] The number of burn-in periods in which no securities are traded (necessary to allow believed models to be estimated in the first period).
+ 		\item [\texttt{n.seq}.] Number of trading sequences.
+ 		\item [\texttt{horizon}.] Number of trading periods in each trading sequence.
+ 	\end{description}
+
+ \end{description}
+
 \bibliography{library.bib}

### 2015-08-29 prediction_market.Rnw
First (rough) draft of detailed description of the model for supplemental material (plus small modification to visualization tools in main.R)
This is a rough draft but feel free to check it out and to give me feedbacks next time we meet.

@@
 \documentclass{article}
+
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{mathtools}
+\usepackage[linktocpage, breaklinks, colorlinks = true, citecolor = blue]{hyperref}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Supplemental material}
+
+   \subsection{Detailed description of the model}
+
+ 	\label{model}
+
+ 	\subsubsection{Elements of the model}
+
+ 	The ABM is made of the following elements.
+
+ 	\begin{description}
+ 		\setlength\itemsep{0em}
+ 		\item [A simple ``true" model of world temperatures.] 
+ 		The true model is determined by parameter  \texttt{true.model}.
+
+ 		\begin{description}
+ 			\item[If \texttt{true.model} = 0,] then anthropogenic climate change is a myth and the temperatures follow the autoregressive process 
+ 			\begin{align}
+ 			\label{eq:true0}
+ 			\begin{split}
+ 			T_t ~ =~  \lambda T_{t-1} +  \epsilon_t, 
+ 			\end{split}
+ 			\end{align}
+ 			where $T_t$ is the temperature at time $t$, and $\epsilon^T_t$ is a random error term.
+
+ 			The value of $\lambda$ is obtained by calibration on actual data.
+Specifically, $\lambda$ is the ordinary least square estimator of \eqref{eq:true0}, where the temperature time series are US yearly averages based from 1895 to 2010, based on US monthly averages from \url{http://www7.ncdc.noaa.gov/CDO/CDODivisionalSelect.jsp} ( TAVG in the data set, retrieved on May 5th 2015).
+
+ 			The distribution of $\epsilon_t$ is the empirical distribution of the residual from this linear regression.
+
+ 			\item[If \texttt{true.model} = 1,] then anthropogenic climate is true and the temperatures follow the process 
+ 			\begin{align}
+ 			\label{eq:true1}
+ 			\begin{split}
+ 			T_t ~ =~  \beta GHG_{t-1} +  \mu_t, 
+ 			\end{split}
+ 			\end{align}
+ 			where $GHG_t$ is the level of greenhouse gas emissions at time $t$, and $\epsilon^{GHG}_t$ is a random error term.
+
+ 			Again, the value of $\beta$ is obtained by calibration on actual data.
+Specifically, $\beta$ is the ordinary least square estimator of \eqref{eq:true1}, where the temperature time series are as in \texttt{true.model} = 0, and the $GHG$ time series are Global Mean CO2 Mixing Ratios (ppm) on the same time period from \url{http://data.giss.nasa.gov/modelforce/ghgases/Fig1A.ext.txt} ( MixR in the data set, retrieved on May 5th 2015).
+
+ 			The distribution of $\mu_t$ is the empirical distribution of the residual from this linear regression.
+ 		\end{description}
+
+ 		In either cases (\texttt{true.model} = 0 and \texttt{true.model} = 1), the chosen true model is used to generate and artificial temperature time series $\{\hat{T}_t\}_{t=1895}^{t=2010}$.
+
+		We emphasize the artificial time series is in general different from the temperature in the empirical data set.	
+		\begin{itemize}
+			\item When  \texttt{true.model} = 0, $\hat{T}_{1895}$ is chosen to be the  temperature in 1895 from the empirical data set, and $\{\hat{T}_t\}_{t=1896}^{t=2010}$ are computed recursively using \eqref{eq:true0}.
+			\item When  \texttt{true.model} = 1, the value of $GHG_t$ for all $t \in \{1895,\dots,2010\}$ is taken from the empirical data set and $\{\hat{T}_t\}_{t=1895}^{t=2010}$ are computed using \eqref{eq:true1}.
+		\end{itemize}
+
+ 		\item[ Beliefs.]  Traders use either \eqref{eq:true0} or \eqref{eq:true1} to  forecast future temperatures in order to trade on the prediction market.
+These $2$ models are interpreted as the trader's beliefs about the true climate model.
+They are meant to represent pervasive positions on climate change in the public debate.
+In order to approximatively match the current configuration of beliefs in climate change in the US, model \eqref{eq:true0} is randomly assigned to half of the traders, while \eqref{eq:true1} is assigned to the other half of the traders.
+
+ 		Thus, both when \texttt{true.model} = 0 and when  \texttt{true.model} = 1,  some traders use the true model to make predictions.
+However, this does not mean that traders using the true model make perfectly accurate predictions.
+Although these traders believe in the correct \emph{functional form} of the model, they still need to calibrate it based on limited noisy data.
+Therefore, the values they use as  the \emph{parameters} of the model will typically still be different from $\lambda$ and $\beta$ (see description of the model's timing below).
+
+ 		\item[ Traders.] Traders are initial endowed with one single unit Experimental Currency Unit (ECU).  
+ 		Traders use their believed model to forecast the distribution of future temperatures and determine their reservation price for different securities.
+Securities pay 1 ECU  if the actual temperature at some time $t^*$ falls between a certain range.
+Traders are assumed to be \emph{risk-neutral expected utility maximizers}.
+Therefore, their reservation price for a security $\tilde{s}$ is simply their assessment of the probability that the temperature at time $t^*$ falls in the range covered by $\tilde{s}$.
+
+ 		 Based on their reservation price, traders behave as ``\emph{zero-intelligence}" agents (\cite{Gode1993}), that is they sell at a random price above their reservation, and buy at a random price below their reservation.
+The distribution from which buy and sell prices are drawn is determined by the traders risk attitude, characterized by parameter \texttt{risk.tak}$_i$ (more details below).
+These strategies are simple but have been shown to provide good approximations of the behavior on prediction markets (see for instance \cite{Klingert2012c}), and in financial markets more broadly.
+
+ 		\item[Continuous-Double Auction (CDA) market structure.] Continuous-Double auctions (or some variants thereof) are common \emph{market mechanisms}, i.e.
+procedures to match buy and sell orders.
+CDA are notably used on large stock markets (see for instance \cite{Tseng2010a}).
+Our model features a particular version of CDA through which traders exchange securities (more in section \ref{implem}).
+
+ 		\item[A social network.] Traders are part of a social network.
+Every time a security is realized, each trader $i$ looks at the performance of their neighbors in the network.
+If one of $i$'s neighbors, say $j$, is richer than $i$, $i$ interprets this has an indication that $j$ has a better approximate model.
+Then $i$ considers adopting $j$'s approximate model.\footnote{Traders start with the same initial amount of money, so differences in money among traders can only come from  traders' interactions through the market.} For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is characterized by parameter \texttt{ideo}$_i$.
+
+ 		An example of such social network is depicted in Figure \ref{socnet}.
+ 		We assume that the initial network is segregated, that is traders who believe in \eqref{eq:true0} (resp.
+\eqref{eq:true1}) are more likely to be linked with other traders who believe in \eqref{eq:true0}  (resp.
+\eqref{eq:true1}).
+The degree of segregation is determined by parameter \texttt{seg} (more details below).
+Although traders can change their approximate model as time passes, the \emph{connections} between traders do not change as the market unfolds (i.e.
+the edges are fixed).
+ 	\end{description}
+
+ 	\begin{figure}
+ 		\begin{center}
+ 		\includegraphics[width = 0.8\textwidth]{figures/segregated.png}
+ 		\end{center}
+ 		\caption{An example of the initial state of a belief network with 50 traders.
+Red nodes are traders who initially believe in \eqref{eq:true0} and black nodes are traders who initially believe in \eqref{eq:true1}.
+Notice that the  \label{socnet}}
+ 	\end{figure}
+
+ 	\subsubsection{Timing}
+
+ 	The time periods $t$ are grouped in trading \emph{sequences}.
+In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence.
+For instance, the third trading sequence might start in period $t=1964$ and end in period $t = 1970$.
+In this case, a security traded in the third sequence pay 1 ECU if the temperature at $t=1970$ falls into the range of temperatures covered by the security.
+
+ 	   At each time $t$, traders are assumed to know the past value of the temperature $\hat{T}_{0:t}$ and greenhouse gas emissions  $GHG_{0:t}$.
+In a sequence finishing at time $t^*$,  traders also have common knowledge of $GHG_{t:t^*}$, the future values of greenhouse gas emissions up to $t^*$.
+However, at any $t$, traders do not know the value of any future \emph{temperatures}.
+In particular, in a given sequence, traders do not know the value of $\hat{T}_{t^*}$.
+Traders can only predict $\hat{T}_{t^*}$ using their approximate model and their knowledge of $\hat{T}_{0:t}$ and  $GHG_{0:t^*}$.
+Notice that because $GHG_{0:t^*}$ is common knowledge, in each period $t$ every trader $i$ with the same approximate model forms the same \emph{pseudo}-expectation (see footnote \ref{foot:pseudo})
+ 	   \begin{align}
+ 	   \label{eqref:distrib}
+ 	  \mu_{ti} \equiv E_t( \hat{T}_{t^*} ~|~ \hat{T}_{0:t}, GHG_{0:t^*},   i's \text{ approximate model of } \text { at } t).
+ 	   \end{align}
+ 	   We assume that the believed probability distribution of $\hat{T}_{t^*}$ for trader $i$  at any time $t$ is
+ 	   \begin{align*}
+ 	   \mu_{ti} + \eta_t,
+ 	   \end{align*}
+
+ 	   where the distribution $\eta_t$ is obtained by trader $i$ by bootstrapping the residuals from the regression of her believed true model over the past periods.\footnote{This explains the use of the term \emph{pseudo}-expectation.
+Because $\eta_t$ is not bound to have expected value zero, a traders' effective expected value of $\hat{T}_{t^*}$ may be different from $\mu_{ti}$.
+\label{foot:pseudo}}
+
+ 	    At each time $t$, traders
+ 	\begin{itemize}
+ 		\setlength\itemsep{0em}
+ 		\item recalibrate their approximate model based on the new set of past data available at $t$ (via pooled ordinary least squares),
+ 		\item compute their belief-density for $\hat{T}_{t^*}$ and use it to determine the expected value they attached to each security,
+ 		\item and trade on the CDA market based on their zero-intelligence decision rule as follows. 
+ 		\begin{enumerate}
+ 			\setlength\itemsep{0em}
+ 			\item Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
+ 			\item Every trader $i$ also chooses at random among the securities she owns a positive amount of (if any) a security $s_i^S$ she will try to sell.
+ 			\item Traders then decide of their selling $p_i^S$ and buying price $p_i^S$.
+To do so, they compute the expected value of the securities $s_i^B$ and $s_i^S$ according to \eqref{eqref:distrib}.
+Then they set a selling and a buying price following the zero-intelligence rule described above.
+ 			\item Traders go to the market one at the time, in a random order.
+ 			\item When $i$ comes to the market, she place \emph{limit orders} in the order book.
+These orders  specify that $i$ is willing to buy $s_i^B$ at  any price below $p_i^B$, and to sell $s_i^S$ at any price above $p_i^S$.
+ 			\item The market maker then tries to match $i$'s orders with some order which was put in the order book \emph{before} $i$ came to the market.
+ 			\item If there are  outstanding sell offers for $s_i^B$ at price $p$ lower than $p_i^B$, then a trade is concluded.
+Trader $i$ buys one unit from the sellers who sells at the \emph{highest} price below $p_i^B$, and the sell and buy offers are removed from the order book.
+ 			\item If there are  outstanding buy offers for $s_i^S$ at price $p$ higher than $p_i^S$, then a trade is concluded.
+Trader $i$ sells one unit to the buyer who buys at the \emph{highest} price above $p_i^S$, and the  sell and buy offers are removed from the order book.
+ 			\item Whenever all traders have come to the market, any remaining outstanding offer is removed from the order book, and the trading period is concluded. 
+ 		\end{enumerate} 
+ 	\end{itemize}
+
+ 	At $t^*$, when the sequence ends, there is only one security $s^*$ associated with a range of temperatures including the actual $\hat{T}_{t^*}$. Thus,
+ 	\begin{itemize}
+ 			\setlength\itemsep{0em}
+ 		\item traders receive 1 ECU per unit of $s^*$ they own, and
+ 		\item  consider adapting their neighbors' approximate model as described above.
+ 	\end{itemize}
+
+ 	\subsubsection{The models' parameters and details about their effects}
+ 	\label{param}
+
+ The model depends on the following parameters. 
+ \begin{description}
+ 	\setlength\itemsep{0em}
+ 	\item[Network parameters.]~
+ 	\begin{description}
+ 			\setlength\itemsep{0em}
+ 		\item [\texttt{n.traders}:] the number of traders.
+ 		\item [\texttt{n.edg} :] the number of edges in the social network. This number is fixed throughout the experiments.   
+ 		\item [\texttt{seg} :] determines the initial degree of homophily in the network. The higher \texttt{seg}, the higher the initial homophily.
+
+ 		When constructing the \texttt{n.edg} edges of the network, the probability that a link between two traders be formed depends on whether the traders share the same approximate model in the following way
+ 			\begin{align*}
+ 			\begin{cases}
+ 			\frac{(1-\texttt{seg})}{\texttt{n.edg}},  &\text{if the traders have different approximate models}\\
+ 			\frac{1}{\texttt{n.edg}},  & \text{otherwise}
+ 			\end{cases}
+ 			\end{align*}
+ 	\end{description}
+
+ \item[Market structure parameter.]~
+
+ 		\begin{description}
+ 			\setlength\itemsep{0em}
+ 			\item [\texttt{market.complet}.] Determines market's completeness, i.e.
+the number of securities which can be traded.
+With more securities, the interval of temperatures corresponding to each security is smaller.
+Traders can then trade on more precise temperature intervals.
+Higher values of \texttt{market.complet} may, however, reduce the number of exchanges.
+Because traders pick the securities they buy and sell at random, the probability that a match between sellers and buyers is found is lower for higher values of \texttt{market.complet}.
+ 		\end{description}
+
+ \item[Behavioral parameters.]~
+
+ \begin{description}
+ 	\setlength\itemsep{0em}
+ 	\item [\texttt{risk.tak}.] Determines the distribution of risk taking behavior.
+The higher \texttt{risk.tak}, the more traders	will try to buy (resp.
+sell) lower (resp.
+higher) than their reservation price.
+Formally, at time $t$, trader $i$ picks her buying (resp.
+selling) prices randomly in the interval  $[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it}  * (1 - \texttt{risk.tak}_i)]$ (resp.
+$[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it} * (1  + \texttt{risk.tak}_i)]$), where $\texttt{reserv}_{it}$ is $i$'s reservation price at time $t$ for the security $i$ picked to buy (resp.
+sell).
+
+ 	For each trader $i$, the level of \texttt{risk.tak}$_i$ is drawn uniformly at random from $[0,$\texttt{risk.tak} $]$
+
+ 	\item [\texttt{ideo}$_i$.]  Determines the degree of ``ideology" of traders. If \texttt{ideo} is high, traders will not
+ 	 revise their approximate models easily, even when faced with
+ 	 strong evidence that their neighbors are doing better than them.
+
+ 	  For each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$.
+The value of $d_{i}$ is the probability that $i$ adopts one of her neighbors' approximate model if this neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
+
+ 	  For each trader $i$, the level of \texttt{ideo}$_i$ is drawn uniformly at random from $[0,$\texttt{ideo} $]$
+ \end{description}
+
+ 	\item[Timing parameters.]~
+
+ 	\begin{description}
+ 			\setlength\itemsep{0em}
+ 		\item [\texttt{burn.in}.] The number of burn-in periods in which no securities are traded (necessary to allow believed models to be estimated in the first period).
+ 		\item [\texttt{n.seq}.] Number of trading sequences.
+ 		\item [\texttt{horizon}.] Number of trading periods in each trading sequence.
+ 	\end{description}
+
+ \end{description}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Acknowledgements}

### 2015-08-29 main.R

@@ -96,10 +96,20 @@ main <- function(parameters,
   
   
   # Visualize network (optional)
+  
+  
+  
   if (visu){
+    # color the network
     net <- Colored(net)
-    igraph::plot.igraph(net,vertex.label=NA,layout=layout.fruchterman.reingold, vertex.size = 7)
+    # set weights on edges
+    E(net)$weights = as.numeric(V(net)$approx[get.edges(g, E(g))[,1]] 
+                              - V(net)$approx[get.edges(g, E(g))[,2]] == 0)
+    l<-layout.fruchterman.reingold(net,weights=E(net)$weight)
+    # plot the network 
+    igraph::plot.igraph(net,vertex.label=NA,layout=l, vertex.size = 7)
   }
+  
   #############################################
   #############################################
   #########           RUN             #########

### 2015-09-03 prediction_market.Rnw
Completing uncomplete sentence.
Added additional sens analysis to appendix to show robustness of results.

@@
 Red nodes are traders who initially believe in \eqref{eq:true0} and black nodes are traders who initially believe in \eqref{eq:true1}.
-Notice that the  \label{socnet}}
+Notice how red (resp.
+black) nodes preferentially attach with red (resp.
+black) nodes.\label{socnet}}
  	\end{figure}
@@
 
+ \subsection{Additional sensitivity analyses}
+
+We also conducted a standardized rank regression coefficient analysis to confirm that our results are robust to the analysis technique \cite{pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
+We obtain the almost identical results as the partial rank correlation coefficient analysis (Fig.~\ref{fig:src_sa}).
+
+ \begin{figure}[h]
+\centering
+<<src_sa, eval=TRUE, echo=FALSE, message=FALSE, cache=TRUE>>=
+load("output/src2.Rda")
+eat::plot(src2, outcome_var = "Convergence of Beliefs", xlab="")
+@
+\caption{Standardized rank regression coefficient analysis based on 10,000 simulated parameter sets \cite{pujol_sensitivity:_2014,saltelli_sensitivity_2009} of the effects of ideology, market completeness, number of edges in social network, number of traders in market, risk taking propensities, segregation measure of the social network, and the true climate model on the convergence of traders' beliefs.
+Lines are bootstrapped 95\% confidence intervals.}
+\label{fig:src_sa}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-09-03 prediction_market.Rnw
Small changes to srrc plot

@@
 load("output/src2.Rda")
-eat::plot(src2, outcome_var = "Convergence of Beliefs", xlab="")
+eat::plot(src2, outcome_var = "Convergence of Beliefs",
+          ylab = "Standardized Rank Regression Coefficient", xlab="")
 @

### 2015-09-17 timeseries/acftest.R
Added time series analysis.

@@ -1,0 +1,59 @@
+#
+# Adapted from 
+#
+require(ggplot2)
+require(tidyr)
+require(dplyr)
+
+acftest <- function(residue, model.arma, model.ar1, lag.max = NULL,
+                    type = c('correlation','partial'),
+                    ci = 0.96, ci.type=c('white','ma'))
+{
+  type = match.arg(type)
+  ci.type = match.arg(ci.type)
+  
+  a <- acf(residue$r, plot=FALSE, lag.max = lag.max, type = type)
+  a.df <- with(a, data.frame(acf = acf, lag = lag))
+  a11 <- ARMAacf(coef(model.arma)[1], coef(model.arma)[2], tail(a$lag, 1), 
+                 pacf = type == 'partial')
+  a10 <- ARMAacf(coef(model.ar1)[1], 0, tail(a$lag,1),
+                 pacf = type == 'partial')
+  a.df <- cbind(a.df, arma11 = a11, ar1 = a10)
+ 
+  if (ci > 0) { 
+  clim0  <- qnorm((1+ci)/2) / sqrt(a$n.used)
+  
+  if (ci.type == 'ma') {
+    clim <- clim0 * sqrt(cumsum( 2 * c(1,2*a.df$acf[-1]^2)))
+    clim[1] <- NA
+  } else {
+    clim <- rep_len(clim0, nrow(a.df))
+  }
+  a.df$clim <- clim
+  }
+  
+  ylabel <- switch(type,
+    partial = "Partial ACF",
+    correlation = "ACF"
+  )
+  
+  p <- ggplot(a.df, aes(x = lag, xend=lag, y = acf, yend=0)) +
+    geom_segment(size=1) + 
+    geom_point(aes(y=ar1, shape="AR(1)", color = "AR(1)"), size=4) +
+    geom_point(aes(y=arma11, shape="ARMA(1,1)", color = "ARMA(1,1)"), size=4)
+  
+  if (ci > 0) {
+    p <- p +
+    geom_line(aes(y = clim), color="dark blue", linetype="dashed") +
+    geom_line(aes(y = -clim), color="dark blue", linetype="dashed")
+  }
+  
+  p <- p + 
+    scale_color_brewer(type="qual", palette=2, name = "Model") + 
+    scale_shape(name = "Model") +
+    labs(x = "Lag", y = ylabel) +
+    theme_bw(base_size = 20) +
+    theme(legend.position=c(0.99,0.99),
+          legend.justification=c(1,1))
+  p
+}

### 2015-09-17 timeseries/download_giss_data.R

@@ -1,0 +1,15 @@
+## File Download and File
+## GISS monthly data import script develodped by http://LearnR.wordpress.com   
+
+giss_url <- c(land.sea = "http://data.giss.nasa.gov/gistemp/tabledata/GLB.Ts+dSST.txt",
+              land = "http://data.giss.nasa.gov/gistemp/tabledata/GLB.Ts.txt")
+giss_data.source.text <- c( land.sea = paste0("Data source: NASA, ", giss_url['land.sea']),
+                            land = paste0("Data source: NASA, ", giss_url['land']))
+
+giss_file <- c(land.sea = file.path('data',"GLB.Ts+dSST.txt"), 
+               land = file.path('data','GLB.Ts.txt'))
+
+update.data <- function() {
+  for (loc in c('land.sea','land'))
+    download.file(giss_url[loc], giss_file[loc])
+}

### 2015-09-17 timeseries/load_keeling.R

@@ -1,0 +1,27 @@
+############## RClimate Script: Mauna Loa Monthly CO2    ###################################
+## Script stored on http://chartsgraphs.wordpress.com account for Users to source()       ##
+## Download and process Monthly CO2 Data File                                             ##
+## Developed by D Kelly O'Day to demonstrate use of source() function for climate data    ##
+##                   http:chartsgraphs.wordpress.com    1/16/10                           ##
+############################################################################################
+require(dplyr, quietly = TRUE, warn.conflicts = FALSE)
+
+source('download_keeling_data.R')
+
+month.names <- c('Jan','Feb','Mar','Apr','May','Jun','Jul','Aug','Sep','Oct','Nov','Dec')
+
+get.co2 <- function(file = keeling_filename) {
+  co2_data <- read.table(file,row.names=NULL, header=T)
+  
+  df <- co2_data %>% select(year = frac.year, monthly = co2.interpolated, 
+                            annual = co2.interpolated.adj)
+  
+  monthly.cycle <- co2_data %>% group_by(month) %>% 
+    summarize(co2 = mean(co2.interpolated - co2.interpolated.adj, na.rm=T))
+  monthly.cycle <- rbind(monthly.cycle, monthly.cycle[1,])
+  monthly.cycle$month[13] <- 13
+  
+  invisible(list(keeling = df, seasonal = monthly.cycle, 
+                 up.to.date = tail(na.omit(co2_data),1) %>% select(year, month)))
+}
+

### 2015-09-17 timeseries/paleo_co2.R

@@ -1,0 +1,89 @@
+library(dplyr)
+
+epica_co2_url <- 'ftp://ftp.ncdc.noaa.gov/pub/data/paleo/icecore/antarctica/epica_domec/edc-co2-2008.txt'
+epica_temp_url <- 'ftp://ftp.ncdc.noaa.gov/pub/data/paleo/icecore/antarctica/epica_domec/edc3deuttemp2007.txt'
+law_co2_url <- 'ftp://ftp.ncdc.noaa.gov/pub/data/paleo/icecore/antarctica/law/law2006.txt'
+
+epica_co2_file <- basename(epica_co2_url)
+epica_temp_file <- basename(epica_temp_url)
+law_co2_file <- basename(law_co2_url)
+
+get_epica_co2_data <- function(data.file = epica_co2_file, data.dir = 'data',
+                           epica_url = epica_co2_url, quiet = TRUE) {
+  if (! is.null(data.dir)) {
+    data.file <- file.path(data.dir, data.file)
+  }
+  download.file(epica_url, destfile = data.file, quiet = quiet, mode = 'wb')
+}
+
+get_epica_temp_data <- function(data.file = epica_temp_file, data.dir = 'data', 
+                               epica_url = epica_temp_url, quiet = TRUE) {
+  if (! is.null(data.dir)) {
+    data.file <- file.path(data.dir, data.file)
+  }
+  download.file(epica_url, destfile = data.file, quiet = quiet, mode = 'wb')
+}
+
+get_epica_data <- function(data.dir = 'data', quiet = TRUE) {
+  get_epica_co2_data(data.dir = data.dir, quiet = quiet)
+  get_epica_temp_data(data.dir = data.dir, quiet = quiet)
+}
+
+get_law_dome_data <- function(data.file = law_co2_file, data.dir = 'data', 
+                              law_url = law_co2_url, quiet = TRUE) {
+  if (! is.null(data.dir)) {
+    data.file <- file.path(data.dir, data.file)
+  }
+  download.file(law_url, destfile = data.file, quiet = quiet, mode = 'wb')
+}
+
+load_epica_co2 <- function(data.file = epica_co2_file, data.dir = 'data') {
+  if (! is.null(data.dir)) {
+    data.file <- file.path(data.dir, data.file)
+  }
+  #
+  # EPICA-C file has three sets of records. 
+  # Skip to the composite record (EPICA + Vostok)
+  #
[…]

### 2015-09-17 timeseries/model_temp.Rmd

---
title: "Modeling Temperature"
author: "Jonathan Gilligan"
date: "September 17, 2015"
output: pdf_document
---
```{r, include=FALSE}
library(knitr)
opts_chunk$set(cache=TRUE)
```
# Modeling the temperature

We model the temperature as a time-series with ARMA noise.

First, we load the data.
```{r load_data}
library(dplyr)
source('load_giss.R')
source('load_keeling.R')
source('paleo_co2.R')

scinot <- function(x, places, uthreshold = 3, lthreshold = 1-places) {
  lx <- log10(x)
  ex <- floor(lx)
  mx <- lx %% 1
  if (ex >= uthreshold || ex <= lthreshold)
    paste(signif(10^mx,places),' \\times 10^{',ex,'}',sep='')
  else 
    signif(x,places)
}

t <- load_giss_data()$data %>% select(time, t = t.anom)
t <- na.omit(t)

co2 <- get.co2()$keeling %>% select(time = year, co2 = annual) %>% mutate(time = (floor(time * 12) + 0.5)/12)

co2.law <- load_law_dome_co2() %>% select(time = year, co2) %>% mutate(time = time + 0.5)

sunspots <- read.csv('data/SN_ms_tot_V2.0.csv', sep=';', header=F)
names(sunspots) <- c('year','month','frac.year','ss','sd','nobs','def')
sunspots <- sunspots %>% mutate(ss = ifelse(ss >= 0, ss, NA)) %>% mutate(time = year + (month - 0.5)/12) %>% select(time,ss)

tsi <- tsi <- read.table('data/TSI_TIM_Reconstruction.txt', comment.char = ';')
names(tsi) <- c('time','tsi')
tsi$xtsi <- tsi$tsi
for(i in seq_along(tsi)[-1]) tsi$xtsi[i] <- 0.05 * tsi$tsi[i] + 0.95 * tsi$xtsi[i-1]

data <- merge(t, co2, all.x = TRUE)

data.solar <- t %>% mutate(time = floor(time) + 0.5) %>% group_by(time) %>% summarize(t = mean(t)) %>% ungroup()
data.solar <- merge(data.solar, co2.law, all.x = TRUE)
data.solar <- merge(data.solar, sunspots, all.x = TRUE)
data.solar <- merge(data.solar, tsi, all.x = TRUE)
```

Now fit a linear model of $t$ vs. $\mathrm{CO}_2$.
```{r, lin_model, dependson=c("load_data"), cache.extra=c(data)}
d <- na.omit(data %>% select(time, t, co2))
co2.lin <- lm(t ~ co2, data = d)
co2.res <- data.frame(time = d$time, r = residuals(co2.lin))

p <- predict(co2.lin)
```

Plot the data vs. model to check whether things look sensible:
```{r plot_lin_model, dependson=c("lin_model"), cache.extra=c(d,p)}
library(ggplot2)

ggplot(cbind(d,predict = p), aes(x = co2, y = t)) + geom_point(size = 2, color = "dark blue", alpha=0.5) + 
  geom_line(aes(y=predict), size=1) +
  labs(x = expression(paste(CO[2]," (ppm)")), 
       y = expression(paste("Temperature anomaly ", (degree * C)))) +
  theme_bw(base_size=20)

```

Next, we look at the autocorrelation characteristics of the fit residuals:
```{r test_acf, fig.scap=c('Autocorrelation','Partial autocorrelation'), fig.cap="Compare autocorrelation of fit residuals to AR(1) and ARMA(1,1) models.", dependson=c("lin_model"), cache.extra=c(d,co2.res), warning=FALSE}
library(tseries)
library(nlme)

source('acftest.R')

co2.ar1 <- arma(co2.res$r, order=c(1,0), include.intercept = FALSE, method="BFGS")
co2.arma11 <- arma(co2.res$r, order=c(1,1), include.intercept = FALSE, method="BFGS")

co2.aic <- c(ar1 = summary(co2.ar1)$aic, arma11 = summary(co2.arma11)$aic)
use_arma_co2 <- co2.aic['ar1'] > co2.aic['arma11']

acftest(co2.res, co2.arma11, co2.ar1, ci.type='ma')
acftest(co2.res, co2.arma11, co2.ar1, type='partial', ci.type='ma')
```

According to the Akaike Information Criterion, the `r ifelse(use_arma_co2, 'ARMA(1,1)', 'AR(1)')` model is \(`r scinot(exp((max(co2.aic)-min(co2.aic))/2),2,3)`\) times more likely than the the `r ifelse(!use_arma_co2, 'ARMA(1,1)', 'AR(1)')` model. 
```{r fit_gls, dependson=c("line_model", "test_acf"), cache.extra=c(co2.arma11,d)}
if (use_arma_co2) {
  cs <- corARMA(coef(co2.arma11), p=1, q=1, form=~1)
} else {
  cs <- corAR1(coef(co2.ar1), form = ~1)
}
model.cst <- Initialize(cs, data = d)
model.gls <- gls(t ~ co2, data = d, correlation=model.cst)
p <- predict(model.gls)
d$predicted <- predict(model.gls)
```
```{r plot_gls, dependson=c('fit_gls'), cache.extra=c(model.gls, data)}
ggplot(d, aes(x = co2, y = t)) + geom_point(color = "dark blue", size=2, alpha = 0.5) +
  geom_line(aes(y = predicted), size=1) + 
  labs(x = expression(paste(CO[2]," (ppm)")), 
       y = expression(paste("Temperature anomaly ", (degree * C)))) +
  theme_bw(base_size = 20)

ggplot(d, aes(x = time, y = t)) + geom_point(color = "dark blue", size=2, alpha = 0.5) +
  geom_line(aes(y = predicted), size=1) + 
  labs(x = "Year",
       y = expression(paste("Temperature anomaly ", (degree * C)))) +
  theme_bw(base_size = 20)
```

Now fit a linear model of $t$ vs. TSI #.
```{r lin_model_tsi, dependson=c("load_data"), cache.extra=c(data)}
d.tsi <- na.omit(data.solar %>% select(time, t, tsi = xtsi))
tsi.lin <- lm(t ~ tsi, data = d.tsi)
tsi.res <- data.frame(time = d.tsi$time, r = residuals(tsi.lin))

p.tsi <- predict(tsi.lin)
```

Plot the data vs. model to check whether things look sensible:
```{r plot_lin_model_tsi, dependson=c("lin_model_tsi"), cache.extra=c(tsi.lin, d.tsi, p.tsi)}
library(ggplot2)

ggplot(cbind(d.tsi,predict = p.tsi), aes(x = tsi, y = t)) + geom_point(size = 2, color = "dark blue", alpha=0.5) + 
  geom_line(aes(y=predict), size=1) +
  labs(x = expression(paste("Total Solar Irradiance ", W / m^2)),
       y = expression(paste("Temperature anomaly ", (degree * C)))) +
  theme_bw(base_size=20)

```

Next, we look at the autocorrelation characteristics of the fit residuals:
```{r test_acf_tsi, fig.scap=c('Autocorrelation','Partial autocorrelation'), fig.cap="Compare autocorrelation of fit residuals to AR(1) and ARMA(1,1) models.", dependson=c("lin_model_tsi"), warning=FALSE}

tsi.ar1 <- arma(tsi.res$r, order=c(1,0), include.intercept = FALSE, method="BFGS")
tsi.arma11 <- arma(tsi.res$r, order=c(1,1), include.intercept = FALSE, method="BFGS")

tsi.aic <- c(ar1 = summary(tsi.ar1)$aic, arma11 = summary(tsi.arma11)$aic)
use_arma_tsi <- tsi.aic['ar1'] > tsi.aic['arma11']

acftest(tsi.res, tsi.arma11, tsi.ar1, ci.type='ma')
acftest(tsi.res, tsi.arma11, tsi.ar1, type='partial', ci.type='ma')
```

According to the Akaike Information Criterion, the `r ifelse(use_arma_tsi, 'ARMA(1,1)', 'AR(1)')` model is \(`r scinot(exp((max(tsi.aic)-min(tsi.aic))/2),2,3)`\) times more likely than the the `r ifelse(!use_arma_tsi, 'ARMA(1,1)', 'AR(1)')` model. 
Now that this is settled, let's do a generalized least-squares fit using an `r ifelse(use_arma_tsi, 'ARMA(1,1)', 'AR(1)')` noise model:
```{r fit_gls_tsi, dependson=c("line_model", "test_acf")}
if (use_arma_tsi) {
  cs.tsi <- corARMA(coef(tsi.arma11), p = 1, q = 1, form = ~1)
} else {
  cs.tsi <- corAR1(coef(tsi.ar1), form=~1)
}
model.cst.tsi <- Initialize(cs.tsi, data = d.tsi)
model.gls.tsi <- gls(t ~ tsi, data = d.tsi, correlation=model.cst.tsi)
d.tsi$predicted.tsi <- predict(model.gls.tsi)
```
```{r plot_gls_tsi, dependson=c("fit_gls_tsi")}
ggplot(d.tsi, aes(x = tsi, y = t)) + geom_point(color = "dark blue", size=2, alpha = 0.5) +
  geom_line(aes(y = predicted.tsi), size=1) + 
  labs(x = expression(paste("Total Solar Irradiance ", W / m^2)),
       y = expression(paste("Temperature anomaly ", (degree * C)))) +
  theme_bw(base_size = 20)

ggplot(d.tsi, aes(x = time, y = t)) + geom_point(color = "dark blue", size=2, alpha = 0.5) +
  geom_line(aes(y = predicted.tsi), size=1) + 
  labs(x = "Year",
       y = expression(paste("Temperature anomaly ", (degree * C)))) +
  theme_bw(base_size = 20)
```

### 2015-10-12 timeseries/model_temp.Rmd
Working on object-oriented way of handling time series.

@@
 date: "September 17, 2015"
-output: pdf_document
+output:
+  html_document:
+    keep_md: yes
+  pdf_document: default
 ---
@@
 
-t <- load_giss_data()$data %>% select(time, t = t.anom)
-t <- na.omit(t)
-
-co2 <- get.co2()$keeling %>% select(time = year, co2 = annual) %>% mutate(time = (floor(time * 12) + 0.5)/12)
-
-co2.law <- load_law_dome_co2() %>% select(time = year, co2) %>% mutate(time = time + 0.5)
+t <- load_giss_data()$data %>% group_by(year) %>% summarize(t.anom = mean(t.anom)) %>% ungroup() %>% filter(!is.na(t.anom)) %>% mutate(year = year + 0.5)
+co2 <- readRDS('data/smoothed_co2.Rds')
 
@@
 names(sunspots) <- c('year','month','frac.year','ss','sd','nobs','def')
-sunspots <- sunspots %>% mutate(ss = ifelse(ss >= 0, ss, NA)) %>% mutate(time = year + (month - 0.5)/12) %>% select(time,ss)
+sunspots <- sunspots %>% mutate(ss = ifelse(ss >= 0, ss, NA)) %>% group_by(year) %>% summarize(ss = mean(ss)) %>% ungroup() %>% filter(! is.na(ss)) %>% mutate(year = year + 0.5)
 
 tsi <- tsi <- read.table('data/TSI_TIM_Reconstruction.txt', comment.char = ';')
-names(tsi) <- c('time','tsi')
+names(tsi) <- c('year','tsi')
 tsi$xtsi <- tsi$tsi
@@
 data <- merge(t, co2, all.x = TRUE)
-
-data.solar <- t %>% mutate(time = floor(time) + 0.5) %>% group_by(time) %>% summarize(t = mean(t)) %>% ungroup()
-data.solar <- merge(data.solar, co2.law, all.x = TRUE)
-data.solar <- merge(data.solar, sunspots, all.x = TRUE)
+data.solar <- merge(data, sunspots, all.x = TRUE)
 data.solar <- merge(data.solar, tsi, all.x = TRUE)
@@
 ```{r, lin_model, dependson=c("load_data"), cache.extra=c(data)}
-d <- na.omit(data %>% select(time, t, co2))
-co2.lin <- lm(t ~ co2, data = d)
-co2.res <- data.frame(time = d$time, r = residuals(co2.lin))
+d <- na.omit(data %>% select(year, t.anom, co2))
+co2.lin <- lm(t.anom ~ co2, data = d)
+co2.res <- data.frame(year = d$year, r = residuals(co2.lin))
 
@@
 
-ggplot(cbind(d,predict = p), aes(x = co2, y = t)) + geom_point(size = 2, color = "dark blue", alpha=0.5) + 
-  geom_line(aes(y=predict), size=1) +
+ggplot(cbind(d,predict = p), aes(x = co2, y = t.anom)) + 
+  geom_point(size = 2, color = "dark blue", alpha=0.5, na.rm=T) + 
+  geom_line(aes(y=predict), size=1, na.rm=T) +
   labs(x = expression(paste(CO[2]," (ppm)")), 
@@
 model.cst <- Initialize(cs, data = d)
-model.gls <- gls(t ~ co2, data = d, correlation=model.cst)
+model.gls <- gls(t.anom ~ co2, data = d, correlation=model.cst)
 p <- predict(model.gls)
@@
 ```{r plot_gls, dependson=c('fit_gls'), cache.extra=c(model.gls, data)}
-ggplot(d, aes(x = co2, y = t)) + geom_point(color = "dark blue", size=2, alpha = 0.5) +
-  geom_line(aes(y = predicted), size=1) + 
+ggplot(d, aes(x = co2, y = t.anom)) + 
+  geom_point(color = "dark blue", size=2, alpha = 0.5, na.rm=T) +
+  geom_line(aes(y = predicted), size=1, na.rm=T) + 
   labs(x = expression(paste(CO[2]," (ppm)")), 
@@
 
-ggplot(d, aes(x = time, y = t)) + geom_point(color = "dark blue", size=2, alpha = 0.5) +
-  geom_line(aes(y = predicted), size=1) + 
+ggplot(d, aes(x = year, y = t.anom)) + 
+  geom_point(color = "dark blue", size=2, alpha = 0.5, na.rm=T) +
+  geom_line(aes(y = predicted), size=1, na.rm=T) + 
   labs(x = "Year",
@@
 ```{r lin_model_tsi, dependson=c("load_data"), cache.extra=c(data)}
-d.tsi <- na.omit(data.solar %>% select(time, t, tsi = xtsi))
-tsi.lin <- lm(t ~ tsi, data = d.tsi)
-tsi.res <- data.frame(time = d.tsi$time, r = residuals(tsi.lin))
+d.tsi <- na.omit(data.solar %>% select(year, t.anom, tsi = xtsi))
+tsi.lin <- lm(t.anom ~ tsi, data = d.tsi)
+tsi.res <- data.frame(year = d.tsi$year, r = residuals(tsi.lin))
 
@@
 
-ggplot(cbind(d.tsi,predict = p.tsi), aes(x = tsi, y = t)) + geom_point(size = 2, color = "dark blue", alpha=0.5) + 
-  geom_line(aes(y=predict), size=1) +
+ggplot(cbind(d.tsi,predict = p.tsi), aes(x = tsi, y = t.anom)) + 
+  geom_point(size = 2, color = "dark blue", alpha=0.5, na.rm=T) + 
+  geom_line(aes(y=predict), size=1, na.rm=T) +
   labs(x = expression(paste("Total Solar Irradiance ", W / m^2)),
@@
 model.cst.tsi <- Initialize(cs.tsi, data = d.tsi)
-model.gls.tsi <- gls(t ~ tsi, data = d.tsi, correlation=model.cst.tsi)
+model.gls.tsi <- gls(t.anom ~ tsi, data = d.tsi, correlation=model.cst.tsi)
 d.tsi$predicted.tsi <- predict(model.gls.tsi)
@@
 ```{r plot_gls_tsi, dependson=c("fit_gls_tsi")}
-ggplot(d.tsi, aes(x = tsi, y = t)) + geom_point(color = "dark blue", size=2, alpha = 0.5) +
-  geom_line(aes(y = predicted.tsi), size=1) + 
+ggplot(d.tsi, aes(x = tsi, y = t.anom)) + 
+  geom_point(color = "dark blue", size=2, alpha = 0.5, na.rm=T) +
+  geom_line(aes(y = predicted.tsi), size=1, na.rm=T) + 
   labs(x = expression(paste("Total Solar Irradiance ", W / m^2)),
@@
 
-ggplot(d.tsi, aes(x = time, y = t)) + geom_point(color = "dark blue", size=2, alpha = 0.5) +
-  geom_line(aes(y = predicted.tsi), size=1) + 
+ggplot(d.tsi, aes(x = year, y = t.anom)) + 
+  geom_point(color = "dark blue", size=2, alpha = 0.5, na.rm=T) +
+  geom_line(aes(y = predicted.tsi), size=1, na.rm=T) + 
   labs(x = "Year",

### 2015-10-12 timeseries/build_co2_data.R

@@ -21,5 +21,14 @@ x$source <- factor(x$source, levels = c(levels(x$source), 'smooth'))
 y <- loess(co2 ~ year, x, span = 0.2)
 x <- rbind(x, data.frame(source = 'smooth', year = x$year, co2 = y$fitted))
 
+library(ggplot2)
 ggplot(x , aes(x = year, y = co2, color = source)) + geom_line() + geom_point() + theme_bw()
+
+
+co2 <- x %>% filter(source == 'smooth') %>% select(year, co2) %>% arrange(year)
+years <- with(co2, seq(floor(min(year)), floor(max(year)))) + 0.5
+smoothed_co2 <- as.data.frame(spline(co2, xout=years)) %>% na.omit()
+names(smoothed_co2) <- c('year', 'co2')
+
+saveRDS(smoothed_co2, file = 'data/smoothed_co2.Rds')
 

### 2015-10-12 timeseries/climate_model.R

@@ -1,0 +1,174 @@
+#
+#
+#
+library(nlme)
+library(tseries)
+library(dplyr)
+
+
+setClass("climate_model", 
+         slots = list(
+           climate="data.frame",     # data frame with columns year, t.anom, and covariates
+           covariates="list",        # covariates affecting temperature
+           burn_in = "integer",      # time-steps before starting
+           future_len = "integer",   # time-steps after starting
+           label = "character",
+           p = "integer",
+           q = "integer",
+           arma="list",
+           model="list",
+           future = "data.frame"
+         ),
+         prototype = list(
+           climate=data.frame(),
+           covariates=list(),
+           burn_in = as.integer(NA),
+           future_len = as.integer(NA),
+           label=as.character(NA),
+           p = as.integer(NA),
+           q = as.integer(NA),
+           arma= as.list(NULL),
+           model=as.list(NULL),
+           future = data.frame() # true path for simulated temperature after burn-in
+         )
+)
+
+setGeneric("mdl_formula", 
+           function(x) {
+             standardGeneric("mdl_formula")
+           }
+)
+
+setMethod("mdl_formula", "list", 
+          function(x) {
+            as.formula(paste0('t.anom ~ ', paste(x, collapse=" + ")))
+          }
+)
+
+setMethod("mdl_formula", "character", 
+          function(x) {
+            as.formula(paste0('t.anom ~ ', paste(x, collapse=" + ")))
+          }
+)
+
+
+
+setMethod("mdl_formula", signature = "climate_model",
+          function(x) {
+            as.formula(paste0('t.anom ~ ', paste(x@covariates, collapse=" + ")))
+          }
+)
+
+
+setGeneric("fit_arma", function(p, q, res, covariates, data) {
+  standardGeneric("fit_arma")
+})
+
+setMethod("fit_arma", signature(res = "numeric", covariates = "missing", data = "missing"),
+          function(p, q, res, covariates, data) {
+            invisible(arma(res, order=c(p,q), include.intercept = FALSE, method="BFGS"))
+          }
+)
+
+
[…]

### 2015-10-12 timeseries/prepare_data.R

@@ -1,0 +1,22 @@
+library(dplyr)
+source('load_giss.R')
+
+t <- load_giss_data()$data %>% group_by(year) %>% summarize(t.anom = mean(t.anom)) %>% ungroup() %>% filter(!is.na(t.anom)) %>% mutate(year = year + 0.5)
+co2 <- readRDS('data/smoothed_co2.Rds')
+
+
+sunspots <- read.csv('data/SN_ms_tot_V2.0.csv', sep=';', header=F)
+names(sunspots) <- c('year','month','frac.year','ss','sd','nobs','def')
+sunspots <- sunspots %>% mutate(ss = ifelse(ss >= 0, ss, NA)) %>% group_by(year) %>% summarize(ss = mean(ss)) %>% ungroup() %>% filter(! is.na(ss)) %>% mutate(year = year + 0.5)
+
+tsi <- tsi <- read.table('data/TSI_TIM_Reconstruction.txt', comment.char = ';')
+names(tsi) <- c('year','tsi')
+tsi$xtsi <- tsi$tsi
+for(i in seq_along(tsi)[-1]) tsi$xtsi[i] <- 0.05 * tsi$tsi[i] + 0.95 * tsi$xtsi[i-1]
+
+data <- merge(t, co2, all.x = TRUE)
+data <- merge(data, sunspots, all.x = TRUE)
+data <- merge(data, tsi, all.x = TRUE)
+data <- data %>% select(year, t.anom, co2, sunspots = ss, tsi = xtsi) %>% mutate(year = floor(year))
+data <- na.omit(data)
+

### 2015-10-12 timeseries/build_co2_data.R
Self-contained script to download CO2 data from law dome and Cape Grim and synthesize a smoothed series.

@@ -2,33 +2,74 @@
 #
 #
 
-source('paleo_co2.R')
+library(dplyr)
 
-co2 <- load_law_dome_co2()
-grim <- read.csv('data/CapeGrim_CO2_data_download.csv', header=T, skip=24, stringsAsFactors = F)
-grim <- na.omit(grim)
-grim <- grim %>% group_by(YYYY) %>% summarize(co2 = mean(CO2.ppm.)) %>% ungroup()
-names(grim)[1] <- 'year'
-grim$year <- as.numeric(grim$year)
-grim$source <- 'Cape Grim'
-grim <- grim[,names(co2)]
+law_co2_url <- 'ftp://ftp.ncdc.noaa.gov/pub/data/paleo/icecore/antarctica/law/law2006.txt'
+law_co2_file <- basename(law_co2_url)
 
-x <- co2 %>% filter(year >= 1900, source == 'Ice Core')
-x <- rbind(x, grim)
+grim_url <- 'http://www.csiro.au/greenhouse-gases/GreenhouseGas/data/CapeGrim_CO2_data_download.txt'
+grim_file <- basename(grim_url)
 
-x$source <- factor(x$source, levels = c(levels(x$source), 'smooth'))
+get_law_dome_data <- function(data.file = law_co2_file, data.dir = 'data', 
+                              law_url = law_co2_url, quiet = TRUE) {
+  if (! is.null(data.dir)) {
+    data.file <- file.path(data.dir, data.file)
+  }
+  download.file(law_url, destfile = data.file, quiet = quiet, mode = 'wb')
+}
 
-y <- loess(co2 ~ year, x, span = 0.2)
-x <- rbind(x, data.frame(source = 'smooth', year = x$year, co2 = y$fitted))
+get_grim <- function(data.file = grim_file, data.dir = 'data', quiet = TRUE) {
+  download.file(grim_url, destfile = file.path(data.dir, grim_file), quiet = quiet, mode = 'wb')
+}
 
-library(ggplot2)
-ggplot(x , aes(x = year, y = co2, color = source)) + geom_line() + geom_point() + theme_bw()
+load_law_dome_co2 <- function(data.file = law_co2_file, data.dir = 'data') {
+  if (! is.null(data.dir)) {
+    data.file <- file.path(data.dir, data.file)
+  }
+  
+  co2 <- read.fwf(data.file,widths = c(13,11,10), 
[…]

### 2015-10-12 timeseries/download_keeling_data.R
Deleted unnecessary scripts.

@@ -1,19 +1,0 @@
-#
-#
-#
-
-mlo.url <- 'http://scrippsco2.ucsd.edu/sites/default/files/data/in_situ_co2/monthly_mlo.csv'
-keeling_filename = file.path('data', 'co2_mm_mlo.txt')
-
-update.co2 <- function() {
-  
-  link <-url(mlo.url)
-  co2_data <- read.csv(link, header = F, skip = 57,
-                       colClasses = rep_len("numeric", 10),
-                       na.strings = '-99.99')
-  names(co2_data) <- c("year", "month", "date.excel", "frac.year",
-                       "co2", "co2.adj", "co2.fit", "co2.fit.adj",
-                       "co2.interpolated", "co2.interpolated.adj")
-  write.table(co2_data,file=keeling_filename,row.names=F)
-}
-

### 2015-10-21 timeseries_example.Rmd
Updated timeseries package and climate model.

---
title: "Timeseries Example"
author: "Jonathan Gilligan"
date: "October 21, 2015"
output: html_document
html_document:
  mathjax: default
---
# Example of running the timeseries functions
\(\require{mhchem}\def\COO{\ce{CO2}}\)

## Initializing the model

The first thing we need to do is get the data and initialize models for which the true model (the physically correct one) is TSI and $\log\COO$:
```{r initialize, echo=T, eval=T}
source('prepare_data.R')
source('climate_model.R')

library(ggplot2)
library(dplyr)
library(tidyr)

theme_set(theme_bw(base_size=20))

data <- prepare_climate_data('rcp 4.5')

### 2015-10-21 timeseries/climate_model.R

@@ -77,7 +77,7 @@ setMethod("fit_arma", signature(res = "numeric", covariates = "missing", data = 
 )
 
 
-setMethod("fit_arma", c(res = "missing", covariates = "list", data = "data.frame"),
+setMethod("fit_arma", signature(res = "missing", covariates = "list", data = "data.frame"),
           function(p, q, res, covariates, data) {
             if (is.null(res)) {
               f <- mdl_formula(covariates)
@@ -85,7 +85,7 @@ setMethod("fit_arma", c(res = "missing", covariates = "list", data = "data.frame
               lin.model <- lm(f, data = data)
               res <- residuals(lin.model)
             }
-            fit_arma(mdl, p, q, res)
+            fit_arma(p, q, res)
           }
 )
 
@@ -144,15 +144,16 @@ setMethod("initialize", "climate_model",
           }
 )
 
-predict_future <- function(mdl, n_today, n_horizon, samples, arma_model = NULL, 
+predict_future <- function(mdl, n_today, n_horizon, covars, samples, arma_model = NULL, 
                            auto_arma = FALSE, max_p = 2, max_q = 2, 
                            trace = FALSE) {
+  covars <- as.list(covars)
   past_data <- mdl@future[1:n_today,]
   if (is.null(arma_model)) {
     if (auto_arma)
-      am <- auto_arma(past_data, mdl@covariates, max_p = max_p, max_q = max_q)
+      am <- auto_arma(past_data, covars, max_p = max_p, max_q = max_q)
     else
-      am <- arma_model(past_data, mdl@covariates, mdl@p, mdl@q)
+      am <- arma_model(past_data, covars, mdl@p, mdl@q)
     arma_model <- am$model
   } else {
     s <- summary(arma_model)
@@ -160,11 +161,11 @@ predict_future <- function(mdl, n_today, n_horizon, samples, arma_model = NULL,
   }
   
   if (trace) message("About to fit gls_model: Best arma = (", am$p, ",", am$q, "), aic = ", am$aic)
-  gls_model <- fit_model(past_data, arma_model, mdl@covariates)
+  gls_model <- fit_model(past_data, arma_model, covars)
   
   if (trace) message("About to predict future temperature")
   future.temp <- predict(gls_model, newdata = mdl@future[n_today + seq_len(n_horizon), 
-                                                         unlist(mdl@covariates), drop=FALSE])
[…]

### 2015-10-21 timeseries/test_climate_model.R

@@ -17,67 +17,131 @@ future_covars <- data$future
 
 today <- which(climate_data$year == 1999)
 
-mdl.co2 <- new('climate_model', climate = climate_data, covariates = list('log.co2'))
-mdl.co2 <- init_model(mdl.co2, n_history = today, n_future = 100, covars = list('log.co2'), future_covars = future_covars)
+mdl.co2 <- new('climate_model', climate = climate_data)
+mdl.co2 <- init_model(mdl.co2, n_history = today, n_future = 100, true_covars = list('log.co2'), future_covars = future_covars)
 
-future <- update_model(mdl.co2, n_today = today, n_horizon = 10)
+future.co2.co2 <- update_model(mdl.co2, n_today = today, n_horizon = 10, trader_covars = list('log.co2'))
 
-p1 <- ggplot(future@future, aes(x = year, y = t.anom)) + 
+p1a <- ggplot(future.co2.co2@future, aes(x = year, y = t.anom)) + 
   # simulated future temperatures
   geom_point() + geom_line() + 
   # vertical bands: predictions
-  geom_point(data = future@prediction %>% filter(sim <= 500), alpha = 0.02) + 
+  geom_point(data = future.co2.co2@prediction %>% filter(sim <= 500), alpha = 0.02) + 
   # actual temperatures
-  geom_point(data = future@climate, color = 'blue')
+  geom_point(data = future.co2.co2@climate, color = 'blue') +
+  labs(x = "Year", y = "T anomaly", title = "True = log(CO2), Trader = log(CO2)")
 
-plot(p1)
+plot(p1a)
 
-interval_prob(future, 10, c(0.5,0.7))
+interval_prob(future.co2.co2, 10, c(0.5,0.7))
 
-#' Total kludge. Because I don't have a good model yet, I just reverse the last century of TSI to 
-#' simulate the next century.
-mdl.tsi <- new('climate_model', climate = climate_data, covariates = list("slow.tsi"))
-mdl.tsi <- init_model(mdl.tsi, n_history = today, n_future = 100, covars = list('slow.tsi'), future_covars = future_covars)
+future.tsi.co2 <- update_model(mdl.co2, n_today = today, n_horizon = 10, trader_covars = list('slow.tsi'))
 
[…]

### 2015-10-21 timeseries/timeseries_example.Rmd

---
html_document:
  mathjax: default
author: "Jonathan Gilligan"
date: "October 21, 2015"
output: html_document
title: "Timeseries Example"
---
# Example of running the timeseries functions
\(\require{mhchem}\def\COO{\ce{CO2}}\)

## Initializing the model

The first thing we need to do is get the data and initialize models for which the true model (the physically correct one) is Slow TSI and $\log\COO$:
```{r setup, echo=T, eval=T, message = F}
library(ggplot2)
library(dplyr)
library(tidyr)

theme_set(theme_bw(base_size=20))

source('prepare_data.R')
source('climate_model.R')
```
```{r load_data, echo=T, eval=T, cache=T, dependson=c('setup')}
data <- prepare_climate_data('rcp 4.5')

climate_data = data$data
future_data = data$future

mdl <- new('climate_model', climate = climate_data)

max_future <- 50
history_start <- 100
```

Next, we initialize two models, one for which the true model is $\log\COO$ and the other for which the true model is Slow TSI.
We initialize each model with the first `r history_start` years of data, and in each model, we predict a true future for the next `max_future` years. This true future is based on the best linear fit to the relevant covariate for each model (Slow TSI or $\log\COO$).

```{r init_models, echo=T, eval=T, cache=T, dependson=c('load_data')}
mdl.co2.true <- init_model(mdl, n_history = history_start, n_future = max_future, true_covars = list('log.co2'), future_covars = future_data)
mdl.tsi.true <- init_model(mdl, n_history = history_start, n_future = max_future, true_covars = list('slow.tsi'), future_covars = future_data)
```

Next, we run predictions for two types of traders with each model:
```{r trader_predict, echo=T, eval=T, cache=T, dependson=c('init_models'), cache_extra = c(mdl.co2.true, mdl.tsi.true)}
today <- history_start
trader_horizon <- 10

trader.co2.true.co2 <- update_model(mdl.co2.true, n_today = today, n_horizon = trader_horizon, trader_covars = list('log.co2'))
trader.tsi.true.co2 <- update_model(mdl.co2.true, n_today = today, n_horizon = trader_horizon, trader_covars = list('slow.tsi'))

trader.co2.true.tsi <- update_model(mdl.tsi.true, n_today = today, n_horizon = trader_horizon, trader_covars = list('log.co2'))
trader.tsi.true.tsi <- update_model(mdl.tsi.true, n_today = today, n_horizon = trader_horizon, trader_covars = list('slow.tsi'))
```

Now w report the predictions:
```{r predictions, eval=T, echo=T, cache=T, dependson=c('trader_predict')}
low = 0.5
high = 0.6

predictions <- data.frame(true.model = c('log CO2', 'log CO2', 'Slow TSI', 'Slow TSI'),
                          trader.model = rep_len(c('log CO2', 'Slow TSI'), 4),
                          prob = c(
                            interval_prob(trader.co2.true.co2, n_horizon = trader_horizon, t.range = c(low, high)),
                            interval_prob(trader.tsi.true.co2, n_horizon = trader_horizon, t.range = c(low, high)),
                            interval_prob(trader.co2.true.tsi, n_horizon = trader_horizon, t.range = c(low, high)),
                            interval_prob(trader.tsi.true.tsi, n_horizon = trader_horizon, t.range = c(low, high))
                          )
)

```

| Trader's Model | True model = $\log\COO$ | True model = Slow TSI |
|:---------------|------------------------:|------------------:|
| $\log\COO$     | `r round(predictions$prob[predictions$true.model == 'log CO2' & predictions$trader.model == 'log CO2'],3)` | `r round(predictions$prob[predictions$true.model == 'Slow TSI' & predictions$trader.model == 'log CO2'],3)` |
| Slow TSI     | `r round(predictions$prob[predictions$true.model == 'log CO2' & predictions$trader.model == 'Slow TSI'],3)` | `r round(predictions$prob[predictions$true.model == 'Slow TSI' & predictions$trader.model == 'Slow TSI'],3)` |

### 2015-10-21 timeseries/old_example/build_co2_data.R
Restored old example so it can run.

@@ -7,7 +7,7 @@ library(dplyr)
 law_co2_url <- 'ftp://ftp.ncdc.noaa.gov/pub/data/paleo/icecore/antarctica/law/law2006.txt'
 law_co2_file <- basename(law_co2_url)
 
-grim_url <- 'http://www.csiro.au/greenhouse-gases/GreenhouseGas/data/CapeGrim_CO2_data_download.txt'
+grim_url <- 'http://www.csiro.au/greenhouse-gases/GreenhouseGas/data/CapeGrim_CO2_data_download.csv'
 grim_file <- basename(grim_url)
 
 get_law_dome_data <- function(data.file = law_co2_file, data.dir = 'data', 

### 2015-10-21 timeseries/old_example/download_keeling_data.R

@@ -1,0 +1,19 @@
+#
+#
+#
+
+mlo.url <- 'http://scrippsco2.ucsd.edu/sites/default/files/data/in_situ_co2/monthly_mlo.csv'
+keeling_filename = file.path('data', 'co2_mm_mlo.txt')
+
+update.co2 <- function() {
+  
+  link <-url(mlo.url)
+  co2_data <- read.csv(link, header = F, skip = 57,
+                       colClasses = rep_len("numeric", 10),
+                       na.strings = '-99.99')
+  names(co2_data) <- c("year", "month", "date.excel", "frac.year",
+                       "co2", "co2.adj", "co2.fit", "co2.fit.adj",
+                       "co2.interpolated", "co2.interpolated.adj")
+  write.table(co2_data,file=keeling_filename,row.names=F)
+}
+

### 2015-10-21 data_and_predictions.R
Clean up and wip on integration of timeseries to ABM

@@ -9,6 +9,8 @@ DataPrediction <- function(
   n_future   = 0
 ){
   
+  theme_set(theme_bw(base_size=20))
+  
   source("timeseries/prepare_data.R",chdir=T)
   source("timeseries/climate_model.R",chdir=T)
   
@@ -17,22 +19,28 @@ DataPrediction <- function(
   climate_data <- data$data
   future_covars <- data$future
   
-  if (true.model == 1){
-     covariates <- list('log.co2')
-  }
-  else {
-     covariates <- list('tsi')
-  }
+  # Load data and create model
   
-  mdl <- new("climate_model", climate = climate_data, covariates = covariates)
-
-  mdl.co2 <- init_model(mdl, n_history = n_history, n_future = n_future,
-                    covars = list('log.co2'), future_covars = future_covars)
-
-  mdl.tsi <- init_model(mdl, n_history = n_history, n_future = n_future,
-                      covars = list('tsi'), future_covars = future_covars)
-
+  mdl <- new("climate_model", climate = climate_data)
   
+  # Set timing parameters
+  
+  history_start = n_history
+  future_length = n_future
+  
+  # Initialize the true models
+
+  mdl.co2 <- init_model(mdl, n_history = history_start,
+                        n_future = future_length, true_covars = list('log.co2'),
+                        future_covars = future_covars)
+  
+  mdl.tsi <- init_model(mdl, n_history = history_start,
+                        n_future = future_length, true_covars = list('slow.tsi'),
+                        future_covars = future_covars)
+
+  mdl.co2
+  
+  mdl.tsi
   
   
 }

### 2015-10-21 main.R

@@ -20,29 +20,37 @@ source("adapt.R")
 source("mixing_matrix.R")
 source("assortativity_coefficient.R")
 
-main <- function(parameters,
+main2 <- function(parameters,
+                  iterations = 10,
+                  burn.in = 4,
+                  n.seq = 28,
+                  horizon = 4,
                  market.struct = c("CDA", "LMSR"), 
                  out = c("segreg", "converg"),
                  visu = FALSE,
                  record = FALSE) {
+  
+  stopifnot((burn.in + n.seq * horizon) == 116)
+  
   ### Market structure parameters:
   # market.struct, choses between CDA and LMSR
   # market.complet is number of securities which
   # are traded. With higher securities, traders can trade on
   # more precise temperature intervals
   
+  
   # SA creates everything in 0-1 interval, so im scaling things to the interval we want inside here
   seg <- parameters[1] # continuous value in (0,1)
   ideo <- parameters[2] # continuous value in (0,1)
   risk.tak <- parameters[3] # continuous value in (0,1)
   market.complet <- ifelse(parameters[4]*1000 < 1, 1, round(parameters[4]*1000)) # integer in (1, 1000)
-  true.model <- parameters[5] + 1 # the true model, 1 for tsi, and 2 for log.CO2
+  true.model <- parameters[5] + 1 # the true model, 1 for ACC is a myth, 2 for ACC is fction of GHG
+  n.edg <- round(parameters[6]*100) + 100 # integer in (100, 200)
+  n.traders <- round(parameters[7]*100) + 50 # integer in (50, 250)
   
-  cat(paste("seg", seg, "ideo", ideo, "risk.tak", risk.tak, "market.complet", market.complet,
-            "true.model", true.model, "\n"))
   
-  n.edg <- 150
-  n.traders <- 100 #ifelse(round(parameters[2]*100) < 1, 1, round(parameters[2]*100))
+  cat("seg", seg, "ideo", ideo, "risk.tak", risk.tak, "market.complet", market.complet,
+            "true.model", true.model, "\n")
   
   market.struct <- match.arg(market.struct)
   out <- match.arg(out)
@@ -50,216 +58,209 @@ main <- function(parameters,
   if (visu){
     source("colored.R")
   }
-  #####
-  ## Set model's parameters and create corresponding network
-  #####
-  net <- generate_model( ### Network parameters
-    n.traders  = n.traders,
-    n.edg      = n.edg,
-    seg        = seg, # determine initial segregation of the
[…]

### 2015-10-21 main2.R

@@ -1,266 +1,0 @@
-
-library(igraph)
-
-# Construct model:
-source("generate_model.R")
-# Sub functions needed inside of generate_model():
-source("populate_net.R")
-source("shape_net.R")
-source("mixing_matrix.R")
-source("assortativity_coefficient.R")
-# Construct data for model:
-library(pracma)
-source("generate_data.R")
-# Operate on model:
-source("form_expect.R")
-source("behav.R")
-source("interact.R")
-source("payoffs.R")
-source("adapt.R")
-source("mixing_matrix.R")
-source("assortativity_coefficient.R")
-
-main2 <- function(parameters,
-                  iterations = 10,
-                  burn.in = 4,
-                  n.seq = 28,
-                  horizon = 4,
-                 market.struct = c("CDA", "LMSR"), 
-                 out = c("segreg", "converg"),
-                 visu = FALSE,
-                 record = FALSE) {
-  
-  stopifnot((burn.in + n.seq * horizon) == 116)
-  
-  ### Market structure parameters:
-  # market.struct, choses between CDA and LMSR
-  # market.complet is number of securities which
-  # are traded. With higher securities, traders can trade on
-  # more precise temperature intervals
-  
-  
-  # SA creates everything in 0-1 interval, so im scaling things to the interval we want inside here
-  seg <- parameters[1] # continuous value in (0,1)
-  ideo <- parameters[2] # continuous value in (0,1)
-  risk.tak <- parameters[3] # continuous value in (0,1)
-  market.complet <- ifelse(parameters[4]*1000 < 1, 1, round(parameters[4]*1000)) # integer in (1, 1000)
-  true.model <- parameters[5] + 1 # the true model, 1 for ACC is a myth, 2 for ACC is fction of GHG
-  n.edg <- round(parameters[6]*100) + 100 # integer in (100, 200)
-  n.traders <- round(parameters[7]*100) + 50 # integer in (50, 250)
-  
-  
-  cat("seg", seg, "ideo", ideo, "risk.tak", risk.tak, "market.complet", market.complet,
-            "true.model", true.model, "\n")
-  
-  market.struct <- match.arg(market.struct)
-  out <- match.arg(out)
-  
-  if (visu){
-    source("colored.R")
-  }
-  
-  # Making this a list rather than a vector, because if record==TRUE, each result is itself a vector > length 1
-  result_final <- as.list(rep(NA, iterations))
-  
-  for(iteration in seq(iterations)){
-    #####
[…]

### 2015-10-22 timeseries/timeseries_example.Rmd
Updated interval_prob to give an option for closed or open intervals, and include NA arguments for one element of t.range, to make one-sided tests (rather than inside range).

@@
 source('climate_model.R')
+
+prt <- function(x,digits,flag='',mark='') {
+  formatC(x, digits=digits, format='f', flag=flag, big.mark=mark)
+}
 ```
-```{r load_data, echo=T, eval=T, cache=T, dependson=c('setup')}
+```{r load_data, echo=T, eval=T, cache=T}
 data <- prepare_climate_data('rcp 4.5')
@@
                           prob = c(
-                            interval_prob(trader.co2.true.co2, n_horizon = trader_horizon, t.range = c(low, high)),
-                            interval_prob(trader.tsi.true.co2, n_horizon = trader_horizon, t.range = c(low, high)),
-                            interval_prob(trader.co2.true.tsi, n_horizon = trader_horizon, t.range = c(low, high)),
-                            interval_prob(trader.tsi.true.tsi, n_horizon = trader_horizon, t.range = c(low, high))
+                            interval_prob(trader.co2.true.co2, n_horizon = trader_horizon, t.range = c(low, high), TRUE),
+                            interval_prob(trader.tsi.true.co2, n_horizon = trader_horizon, t.range = c(low, high), TRUE),
+                            interval_prob(trader.co2.true.tsi, n_horizon = trader_horizon, t.range = c(low, high), TRUE),
+                            interval_prob(trader.tsi.true.tsi, n_horizon = trader_horizon, t.range = c(low, high), TRUE)
                           )
@@
 |:---------------|------------------------:|------------------:|
-| $\log\COO$     | `r round(predictions$prob[predictions$true.model == 'log CO2' & predictions$trader.model == 'log CO2'],3)` | `r round(predictions$prob[predictions$true.model == 'Slow TSI' & predictions$trader.model == 'log CO2'],3)` |
-| Slow TSI     | `r round(predictions$prob[predictions$true.model == 'log CO2' & predictions$trader.model == 'Slow TSI'],3)` | `r round(predictions$prob[predictions$true.model == 'Slow TSI' & predictions$trader.model == 'Slow TSI'],3)` |
+| $\log\COO$     | `r prt(predictions$prob[predictions$true.model == 'log CO2' & predictions$trader.model == 'log CO2'],3)` | `r prt(predictions$prob[predictions$true.model == 'Slow TSI' & predictions$trader.model == 'log CO2'],3)` |
+| Slow TSI     | `r prt(predictions$prob[predictions$true.model == 'log CO2' & predictions$trader.model == 'Slow TSI'],3)` | `r prt(predictions$prob[predictions$true.model == 'Slow TSI' & predictions$trader.model == 'Slow TSI'],3)` |

### 2015-10-22 timeseries/climate_model.R

@@ -265,9 +265,36 @@ update_model <- function(mdl, n_today, n_horizon, trader_covars = NULL, samples 
   invisible(mdl)
 }
 
-interval_prob <- function(mdl, n_horizon, t.range) {
+range_check <- function(x, t.range, closed) {
+  if (closed) {
+    if (is.na(t.range[1])) {
+      within_lower <- TRUE
+    } else {
+      within_lower <- x >= t.range[1]
+    }
+    if (is.na(t.range[2])) {
+      within_upper <- TRUE
+    } else {
+      within_upper <- x <= t.range[2]
+    }
+  } else {
+    if (is.na(t.range[1])) {
+      within_lower <- TRUE
+    } else {
+      within_lower <- x > t.range[1]
+    }
+    if (is.na(t.range[2])) {
+      within_upper <- TRUE
+    } else {
+      within_upper <- x < t.range[2]
+    }
+  }
+  prob <- sum(within_lower & within_upper) / length(x)
+  prob
+}
+
+interval_prob <- function(mdl, n_horizon, t.range, closed) {
   if (n_horizon > mdl@horizon) stop("n_horizon(", n_horizon, ") > mdl@horizon(", mdl@horizon, ")")
   prediction <- mdl@prediction %>% filter(step == mdl@today + n_horizon)
-  prob <- sum(findInterval(prediction$t.anom, t.range, rightmost.closed=TRUE) == 1) / nrow(prediction)
-  prob
+  range_check(prediction$t.anom, t.range, closed)
 }

### 2015-12-06 poster_plot.R
Final version of figures for AGU poster.

@@ -14,8 +14,8 @@ plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0
 
 
 p <- ggplot(data=plot_data, aes(x= trading_seq * 6 + 2014, y=(convergence + 1) * 50, color = true_mod)) +
-  geom_point(position = position_jitter(w = 0.05, h = 0), alpha = 0.1) + geom_smooth(size=1, fill=NA) +
-  scale_color_brewer(palette="Set1", name="True Model", labels = c(LogCo2 = expression(CO[2]), SlowTSI = "TSI")) +
+  geom_point(position = position_jitter(w = 0.07, h = 0), alpha = 0.1) + geom_smooth(size=1, fill=NA) +
+  scale_color_brewer(palette="Set1", name="True driver of future climate", labels = c(LogCo2 = expression(CO[2]), SlowTSI = "TSI")) +
   facet_wrap(~set, ncol = 1) +
   labs(x = "Year", y = paste0("% Traders Believing True Model (n = ",
               prettyNum(length(complete.cases(plot_data$convergence)), big.mark = ','), ")"),

### 2015-12-06 poster_plot_history.R

@@ -16,8 +16,8 @@ plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0
 start <- 1880 + 51
 
 p <- ggplot(data=plot_data, aes(x= trading_seq * 6 + start, y=(convergence + 1) * 50)) +
-  geom_point(position = position_jitter(w = 0.05, h = 0), alpha = 0.1) + geom_smooth(size=1, fill=NA) +
-  scale_color_brewer(palette="Set1", name="True Model", labels = c(LogCo2 = expression(CO[2]), SlowTSI = "TSI")) +
+  geom_point(position = position_jitter(w = 0.07, h = 0), alpha = 0.2) + geom_smooth(size=1, fill=NA) +
+  scale_color_brewer(palette="Set1", labels = c(LogCo2 = expression(CO[2]), SlowTSI = "TSI"), guide = "none") +
   ylim(0,100) +
   facet_wrap(~set, ncol = 1) +
   labs(x = "Year",  

### 2015-12-06 timeseries/climate_plot.R

@@ -1,5 +1,3 @@
-rm(list = ls())
-
 library(ggplot2)
 library(dplyr)
 library(tidyr)
@@ -11,7 +9,7 @@ TRACE_CLIMATE_MODEL <- TRUE
 PARALLEL_STAN <- TRUE
 WHICH_MODEL <- 'ar1'
 
-theme_set(theme_bw(base_size=30))
+theme_set(theme_bw(base_size=20))
 
 data <- prepare_climate_data('rcp 8.5')
 climate_data <- data$data
@@ -20,16 +18,27 @@ future_data <- data$future
 today <- nrow(climate_data)
 n_future <- 2100 - max(climate_data$year)
 
-mdl.co2 <- new('climate_model', climate = climate_data)
-mdl.co2 <- init_model(mdl.co2, n_history = today, n_future = n_future, true_covar = 'log.co2', future_covars = future_data)
-mdl.tsi <- init_model(mdl.co2, n_history = today, n_future = n_future, true_covar = 'slow.tsi', future_covars = future_data)
+if (! all(exists('mdl.co2'), exists('mdl.tsi')) ) {
+  mdl.co2 <- new('climate_model', climate = climate_data)
+  mdl.co2 <- init_model(mdl.co2, n_history = today, n_future = n_future, true_covar = 'log.co2', future_covars = future_data)
+  mdl.tsi <- init_model(mdl.co2, n_history = today, n_future = n_future, true_covar = 'slow.tsi', future_covars = future_data)
+  
+  save(mdl.co2, mdl.tsi, file = "covar_models.Rda")
+}
 
-t <- rbind(cbind(mdl.co2@future, covar = 'CO2'), cbind(mdl.tsi@future, covar="TSI"))
+t <- mdl.co2@climate %>% select(year, t.anom) %>% mutate(covar = "Historical")
+change.year <- max(t$year)
+co2 <- mdl.co2@future %>% select(year, t.anom) %>% filter(year >= change.year) %>%
+  mutate(covar = "CO2")
+tsi <- mdl.tsi@future %>% select(year, t.anom) %>% filter(year >= change.year) %>%
+  mutate(covar = "TSI")
+t <- t %>% bind_rows(co2, tsi) %>% 
+  mutate(covar = ordered(covar, levels = c("CO2", "TSI", "Historical")))
 
 p1 <- ggplot(t, aes(x = year, y = t.anom, color = covar)) + 
   geom_line(size=1) + 
[…]

### 2015-12-06 README.Rmd

--- 2015-03-30 README.Rmd
@@
 
-This is code for an ABM of a climate prediction market
-in which traders adapt their belief about the climate based
-on the monetary performance of their neighboring traders in
-a social network. The ABM is meant to test whether climate
-prediction markets foster the convergence of people’s beliefs
-about the climate. The model will be used to
-assess whether such convergence is impacted by factors such
-as the initial belief-homophily in the social network or the
-time horizon at which climate securities are realized.
+This is code for an agent-based model (ABM) of a climate prediction market in which traders adapt their belief about the climate based on the monetary performance of their neighboring traders in a social network.
+The ABM is meant to test whether climate prediction markets foster the convergence of people’s beliefs about the climate.
 
-[Martin Van der Linden](https://martinvdlinden.wordpress.com/) (Dept. of Economics, Vanderbilt University) and [John Nay](https://johnnay.github.io) (School of Engineering, Vanderbilt University).
-
-Master is the main script that sources everything else.
-```{r, eval=FALSE, include=TRUE}
-source("master.R")
-````
+[John J.
+Nay](http://johnjnay.com/) (School of Engineering, Vanderbilt University), [Martin Van der Linden](https://martinvdlinden.wordpress.com/) (Economics, Vanderbilt University), and [Jonathan M.
+Gilligan](https://my.vanderbilt.edu/jonathangilligan/) (Earth and Environmental Sciences, Vanderbilt University).

### 2016-01-17 jg_master_timeseries_experiments.R
Temporary checkin

@@ -138,4 +138,4 @@ if(plot_final){
     #scale_color_discrete(name="True Model") +
     scale_color_brewer(palette="Dark2", name="True Model")
   dev.off()
-  
+}

### 2016-01-17 master_timeseries_experiments.R

@@ -10,7 +10,7 @@ library(eat)
 ## Set up parameters for climate_model to control level of tracing to screen
 ##############################################################################
 
-SHOW_CLIMATE_PLOTS <- TRUE  # Plot graph of temperature after each year?
+SHOW_CLIMATE_PLOTS <- FALSE  # Plot graph of temperature after each year?
 TRACE_CLIMATE_MODEL <- FALSE # Show diagnostic traces for stan runs?
 STAN_REFRESH <-  0           # Frequency to show stan chain progress. 0 for silent
 PARALLEL_STAN <- FALSE       # Run chains in parallel?
@@ -145,7 +145,7 @@ if(future){
   run_experiment(set = set, input_values = input_values, 
                  file_path = "output/convergence_future.Rda",
                  # Add to previous results?
-                 previous_results = "output/convergence_future.Rda",
+                 # previous_results = "output/convergence_future.Rda",
                  sample_count = sample_count,
                  burn.in = 135)
 }

### 2016-02-23 main.R
Model seems to be working well. No NA problems. Changed edge specification to be number of edges per trader.

@@ -70,12 +70,12 @@ main <- function(parameters,
   ideo <- parameters[2] # continuous value in (0,1)
   risk.tak <- parameters[3] # continuous value in (0,1)
   true.model <- parameters[4] + 1 # the true model, 1 for slow.tsi, 2 for log.co2
-  if (TRUE) {
+  if (FALSE) {
     n.edg <- round(parameters[5]*100) + 100 # integer in (100, 200)
     n.traders <- round(parameters[6]*100) + 50 # integer in (50, 250)
   } else {
     n.traders <- round(parameters[6]*200) + 50 # integer in (50, 250)
-    n.edg <- (round(parameters[5]*8) + 2) * n.traders # integer in (2 to 10 per trader)
+    n.edg <- round((parameters[5]*8 + 2) * n.traders) # integer in (2 to 10 per trader)
   }
   
   cat("seg", seg, "ideo", ideo, "risk.tak", risk.tak, "market.complet", market.complet,

### 2016-02-23 master_timeseries_experiments.R

@@ -161,7 +161,10 @@ if(plot_final){
     #plot_data$set <- factor(plot_data$set, levels = gtools::mixedsort(unique(plot_data$set)))
     # n.edg <- round(parameters[5]*100) + 100 # integer in (100, 200)
     plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"), 
-                            labels = c("n.edge = 105, seg = 0.05", "n.edge = 105, seg = 0.95", "n.edge = 195, seg = 0.05", "n.edge = 195, seg = 0.95"))
+                            labels = c(paste0(round(8 * 0.05 + 2, 1), " edges/trader, seg = 0.05"), 
+                                       paste0(round(8 * 0.05 + 2, 1), " edges/trader, seg = 0.95"), 
+                                       paste0(round(8 * 0.95 + 2, 1), " edges/trader, seg = 0.05"), 
+                                       paste0(round(8 * 0.95 + 2, 1), " edges/trader, seg = 0.95")))
     pdf("output/timeseries_past.pdf", width=8, height=18)
     ggplot(data=plot_data, aes(x= trading_seq, y=convergence, color = true_mod)) +
       geom_point(position = position_jitter(w = 0.07, h = 0)) + geom_smooth() +
@@ -184,7 +187,10 @@ if(plot_final){
     #plot_data$set <- factor(plot_data$set, levels = gtools::mixedsort(unique(plot_data$set)))
     # n.edg <- round(parameters[5]*100) + 100 # integer in (100, 200)
     plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"), 
-                            labels = c("n.edge = 105, seg = 0.05", "n.edge = 105, seg = 0.95", "n.edge = 195, seg = 0.05", "n.edge = 195, seg = 0.95"))
+                            labels = c(paste0(round(8 * 0.05 + 2, 1), " edges/trader, seg = 0.05"), 
+                                       paste0(round(8 * 0.05 + 2, 1), " edges/trader, seg = 0.95"), 
+                                       paste0(round(8 * 0.95 + 2, 1), " edges/trader, seg = 0.05"), 
+                                       paste0(round(8 * 0.95 + 2, 1), " edges/trader, seg = 0.95")))
[…]

### 2016-02-23 poster_plot.R

@@ -9,12 +9,16 @@ load("output/convergence_future.Rda")
 plot_data <- average_convergence
 # plot_data$set <- factor(plot_data$set, levels = gtools::mixedsort(unique(plot_data$set)))
 #  n.edg <- round(parameters[5]*100) + 100 # integer in (100, 200)
-plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"),
-                        labels = c("105 edges, segmentation = 0.05", "105 edges, segmentation = 0.95", "195 edges, segmentation = 0.05", "195 edges, segmentation = 0.95"))
+plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"), 
+                        labels = c(paste0(round(8 * 0.05 + 2, 1), " edges/trader, segmentation = 0.05"), 
+                                   paste0(round(8 * 0.05 + 2, 1), " edges/trader, segmentation = 0.95"), 
+                                   paste0(round(8 * 0.95 + 2, 1), " edges/trader, segmentation = 0.05"), 
+                                   paste0(round(8 * 0.95 + 2, 1), " edges/trader, segmentation = 0.95")))
 
+sequence.years <- 6
 
-p <- ggplot(data=plot_data, aes(x= trading_seq * 6 + 2014, y=(convergence + 1) * 50, color = true_mod)) +
-  geom_point(position = position_jitter(w = 0.07, h = 0), alpha = 0.1) + geom_smooth(size=1, fill=NA) +
+p <- ggplot(data=plot_data, aes(x= trading_seq * sequence.years + 2014, y=(convergence + 1) * 50, color = true_mod)) +
+  geom_point(position = position_jitter(w = 0.07 * sequence.years, h = 0), alpha = 0.1) + geom_smooth(size=1, fill=NA) +
   scale_color_brewer(palette="Set1", name="True driver of future climate", labels = c(LogCo2 = expression(CO[2]), SlowTSI = "TSI")) +
   facet_wrap(~set, ncol = 1) +
   labs(x = "Year", y = paste0("% Traders Believing True Model (n = ",

### 2016-02-23 poster_plot_history.R

@@ -9,14 +9,18 @@ load("output/jg_convergence_true_past.Rda")
 plot_data <- average_convergence
 # plot_data$set <- factor(plot_data$set, levels = gtools::mixedsort(unique(plot_data$set)))
 #  n.edg <- round(parameters[5]*100) + 100 # integer in (100, 200)
-plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"),
-                        labels = c("105 edges, segmentation = 0.05", "105 edges, segmentation = 0.95", "195 edges, segmentation = 0.05", "195 edges, segmentation = 0.95"))
+plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"), 
+                        labels = c(paste0(round(8 * 0.05 + 2, 1), " edges/trader, segmentation = 0.05"), 
+                                   paste0(round(8 * 0.05 + 2, 1), " edges/trader, segmentation = 0.95"), 
+                                   paste0(round(8 * 0.95 + 2, 1), " edges/trader, segmentation = 0.05"), 
+                                   paste0(round(8 * 0.95 + 2, 1), " edges/trader, segmentation = 0.95")))
 
 
 start <- 1880 + 51
+sequence.years <- 6
 
-p <- ggplot(data=plot_data, aes(x= trading_seq * 6 + start, y=(convergence + 1) * 50)) +
-  geom_point(position = position_jitter(w = 0.07, h = 0), alpha = 0.2) + geom_smooth(size=1, fill=NA) +
+p <- ggplot(data=plot_data, aes(x= trading_seq * sequence.years + start, y=(convergence + 1) * 50)) +
+  geom_point(position = position_jitter(w = 0.07 * sequence.years, h = 0), alpha = 0.2) + geom_smooth(size=1, fill=NA) +
   scale_color_brewer(palette="Set1", labels = c(LogCo2 = expression(CO[2]), SlowTSI = "TSI"), guide = "none") +
   ylim(0,100) +
   facet_wrap(~set, ncol = 1) +

### 2016-02-23 jg_plot_sa.R
Code seems to be working now. Now it's time to run experiments with it.

@@ -13,7 +13,7 @@ ss <- correct_bias(src_future@result[[7]])
 ss <- ss %>% mutate(var = ordered(var, levels = c('ideo','risk.tak','n.traders',
                                                   'n.edg', 'seg', 'true.model'),
                                   labels = c('Ideology', "Risk tolerance",
-                                             '# Traders', '# Edges', 'Segmentation',
+                                             '# Traders', '# Edges/Trader', 'Segmentation',
                                              "True model")))
 
 title <- bquote(atop(plain("Estimated Effects on Convergence of Beliefs"),plain("in Future Scenario ") ~

### 2016-02-23 paper.Rmd
Created skeleton of paper for WSC. Rmd is better.

---
title: "Simulated Climate Prediction Markets and Convergence of Predictive Models"
author: " "
date: "Feb 23, 2016"
output: 
  pdf_document:
    keep_tex: true
bibliography: library.bib
---

```{r fig_ref, echo=FALSE}
figRef <- local({
  tag <- numeric()
  created <- logical()
  used <- logical()
  function(label, caption, prefix="Fig.", sep=".",
      prefix.highlight="**") {
    i <- which(names(tag) == label)
    if(length(i) == 0){
      i <- length(tag) + 1
      tag <<- c(tag, i)
      names(tag)[length(tag)] <<- label
      used <<- c(used, FALSE)
      names(used)[length(used)] <<- label
      created <<- c(created, FALSE)
      names(created)[length(created)] <<- label
    }
    if(!missing(caption)){
      created[label] <<- TRUE
      paste0(prefix.highlight, prefix, " ", i, sep, prefix.highlight, " ", caption)
    } else {
      used[label] <<- TRUE
      paste(prefix, tag[label])
    }
  }
})
# adapted from a function on this page: http://galahad.well.ox.ac.uk/repro/
rinline <- function(code){
  sprintf('`r %s`', code)
}
```

```{r, caption_setting, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
library(knitr)
knit_hooks$set(htmlcap = function(before, options, envir) {
  if(!before) {
    paste('<p class="caption">',options$htmlcap,"</p>",sep="")
    }
    })
# small margins
knit_hooks$set(small_mar = function(before, options, envir) {
    if (before) par(mar = rep(0.0001, 4))  # smaller margins all around
})
# opts_chunk$set(dev="tiff", 
#                dev.args=list(compression="lzw", type="cairo"),
#                dpi=300)
```

```{r tab_ref, echo=FALSE}
tabRef <- local({
    tag <- numeric()
    created <- logical()
    used <- logical()
    function(label, caption, prefix = "Table", 
        sep = ".", prefix.highlight = "**") {
        i <- which(names(tag) == label)
        if (length(i) == 0) {
            i <- length(tag) + 1
            tag <<- c(tag, i)
            names(tag)[length(tag)] <<- label
            used <<- c(used, FALSE)
            names(used)[length(used)] <<- label
            created <<- c(created, FALSE)
            names(created)[length(created)] <<- label
        }
        if (!missing(caption)) {
            created[label] <<- TRUE
            paste0(prefix.highlight, prefix, " ", i, sep, prefix.highlight, 
                " ", caption)
        } else {
            used[label] <<- TRUE
            paste(prefix, tag[label])
        }
    }
})
```

```{r multi_plot, echo=FALSE}
multi_plot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

library(ggplot2)
library(pander)
```

# Abstract

We present an agent-based model of a climate prediction market in which traders adapt their belief about the climate based on the monetary performance of their neighboring traders in a social network. We analyze the model to determine how a variety of factors affect whether climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the earth's climate. We find that only the imposed ``true model'' itself significantly affects the convergence of beliefs. If the true model from which the temperature data is generated is auto-regressive, with no human-induced effects, participation in the prediction market, on average, does not cause convergence of traders' beliefs. On other hand, if anthropogenic climate change is real, traders' beliefs are more likely than not to converge toward the correct climate model due to participation in the market.

# 1. Introduction

The previous two decades saw a strong polarization of the climate change debate. Although the scientific consensus on the anthropogenic nature of climate change strongly increased, the average belief about anthropogenic climate change did not evolve much within the public [@Vandenbergh2013f]. In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily, indicating that the question is becoming increasingly politicized and potentially disconnected from scientific evidence. This is troubling given the costs of a misinformed climate policy. If anthropocentric climate change is a myth but is taken to be true, a tremendous amount of public resources will be wasted on adaptation and mitigation efforts. On the other hand, if climate change is human induced but not recognized as so, the costs of inaction would be huge. Given that effective climate policies require long-lasting measures to be taken, the importance of an accurate consensus on the issue is manifest.

Attempts to foster such consensus face many social and psychological challenges. Recently, people have argued that such challenges could be met by setting up climate prediction markets [@Hsu2011,@Vandenbergh2013f]. The idea of using prediction markets to efficiently aggregate information about uncertain events has been discussed for decades [@Horn2014c]. Prediction markets have been shown to have interesting theoretical properties [@Set1998,Hanson2012] and to perform well in terms of prediction accuracy in experiments [@Hanson2006,@Healy2010], agent-based models (ABM) [@Klingert2012c,@Jumadinova2011a], and in the real-world [@Wolfers2006].

[@Tseng2010a] study an ABM of a continuous double auction model with various types of market strategies (including two variants of the zero-intelligence agents). They compare the behavior of their simulation with data from a prediction market for the outcome of political elections. They find that despite their simplicity, zero-intelligence agents capture some salient features of real market data. [@Klingert2012c] reach similar conclusions with respect to experimental data from  [@Hanson2006]. In their simulation, [@Klingert2012c] compare continuous double auctions with logarithmic market scoring rules and find that the latter generally outperform continuous double auctions in terms of prediction accuracy. In some variants of the two last models, agents' strategies feature learning, but these are rather basic and unrelated to a formal predictive model.

[@Jumadinova2011a] conducted research more similar to our project, as they explicitly include information about uncertain events in their model. The paper studies a continuous double auction model in which agents update their beliefs about the uncertain events based on information on the event. New beliefs are modeled as a weighted sum of the last period's beliefs and market prices. The agent's information set changes every period. The better the information set, the more likely the agent is to put higher weights on last period's prices when revising their beliefs. Not much structure is imposed on the information set and the way it is used to generate the weights in the update function. Therefore, the model does not allow for investigating the convergence of predictive models. [@Ontanon2009] study the effect of deliberation on a simulated prediction market. They assume that agents use Case-Based Reasoning [@Aamodt1994] to debate about  uncertain outcomes with their neighbors in a social network.  Although the model could in theory be used to assess convergence of predictive models by looking at the arguments agents adopt after the deliberation, this is not done by the authors. Also, the way beliefs are formed and revised through argumentation is rather abstract and might not be ideal to study climate prediction markets.

However, to the best of our knowledge, the idea that prediction markets can generate consensus *on the factors* affecting  uncertain events has never been quantitatively explored. ABMs of prediction markets have been studied [@Klingert2012c,@Tseng2010a,@Jumadinova2011a], some of which feature communication between agents. In these models, however, beliefs about the uncertain outcomes are constructed in rather abstract ways. In particular, beliefs are not based on structural models from which agents could derive causal implications, and therefore these models are not suitable for investigating the convergence of the underlying explanatory models agents employ for prediction.

From a public policy perspective, changing the explanatory models of market participants is one of the most important roles that prediction markets might play. And perhaps the most important potential social benefit of prediction markets given the comparable predictive power of statistical models, which are less expensive than setting up and maintaining prediction markets [@goel_prediction_2010]. Effective climate policy not only requires an accurate consensus on future climate *outcomes*. It also requires an accurate consensus on the causal *mechanisms* influencing such outcomes. If people agree that temperature will rise, but some believe it will be due to greenhouse gases, while others believe that it will be caused by increased solar activity, inconsistent and ineffective policies may be implemented. Our model is designed to investigate whether, and under what conditions, prediction markets can foster such convergence in predictive models.

# 2. Model Design

# 3. Results

The sensitivity analysis is based on a Latin Hypercube Sampling of ... parameter sets from the following distributions [@beachkofski_improved_2002,@carnell_lhs_2012].

  * ideo $\sim Unif(0,1)$
	* market.complet $\sim Unif(0,1000)$ (mapped into integer)
	* n.edge $\sim Unif(100,200)$ (mapped into integer)
	* n.traders $\sim Unif(50,250)$ (mapped into integer)
	* risk.tak  $\sim Unif(0,1)$
	* seg $\sim Unif(0,1)$
	* true.model $\sim Binom(0.5)$

We use the model to simulate 10,000 outcomes based on the input paramter sets and then conduct a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, resulting simulated vector, $y$ [@Marino2008,@pujol_sensitivity:_2014,@saltelli_sensitivity_2009]. Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).  The only difference with the partial correlation and partial rank correlation (which we use here), is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships. That is every value, $y_s$, in the data set is replace by a number $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, where $y_r$ are the other observed values of $y$ in the data set.

Our sensitivity analysis, illustrated by ...,

```{r, one, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=tabRef("one", "dfd")}

```

 `r tabRef("one")`

`r figRef("data_explore")`

```{r, data_explore, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("data_explore", " "), fig.width=10, fig.height=5}

#multi_plot(p1, p2, cols = 2)
```

# 4. Discussion

# Acknowledgements

We would like to thank Yevgeniy Vorobeychik for feedback on this project.

# References

```{r, finalize, eval=TRUE, echo=FALSE}
# Print out warnings if we did not use all the figures.
if(!all(environment(figRef)$created)){
    missingFig <- which(!environment(figRef)$created)
    warning("Figure(s) ", paste(missingFig, sep=", "), " with label(s) '", 
      paste(names(environment(figRef)$created)[missingFig], sep="', '"),
      "' are referenced in the text but have never been created.")
}
if(!all(environment(figRef)$used)){
    missingRef <- which(!environment(figRef)$used)
    warning("Figure(s) ", paste(missingRef, sep=", "), " with label(s) '", 
      paste(names(environment(figRef)$used)[missingRef], sep="', '"),
      "' are present in the document but are never referred to in the text.")
}
```

### 2016-02-24 paper.Rmd
first draft of intro and related work sections

@@
 ---
-title: "Simulated Climate Prediction Markets and Convergence of Predictive Models"
+title: "Climate Prediction Markets and the Convergence of Forecasting Models"
 author: " "
@@
 
-We present an agent-based model of a climate prediction market in which traders adapt their belief about the climate based on the monetary performance of their neighboring traders in a social network.
-We analyze the model to determine how a variety of factors affect whether climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the earth's climate.
-We find that only the imposed ``true model'' itself significantly affects the convergence of beliefs.
-If the true model from which the temperature data is generated is auto-regressive, with no human-induced effects, participation in the prediction market, on average, does not cause convergence of traders' beliefs.
-On other hand, if anthropogenic climate change is real, traders' beliefs are more likely than not to converge toward the correct climate model due to participation in the market.
+We present an agent-based model of a climate prediction market.
+Trader agents adapt their beliefs about the climate based on the monetary performance of their neighboring traders in a social network.
+We analyze the model to determine how a variety of market and climate factors affect whether climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
 
@@
 In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily, indicating that the question is becoming increasingly politicized and potentially disconnected from scientific evidence.
-This is troubling given the costs of a misinformed climate policy.
-If anthropocentric climate change is a myth but is taken to be true, a tremendous amount of public resources will be wasted on adaptation and mitigation efforts.
-On the other hand, if climate change is human induced but not recognized as so, the costs of inaction would be huge.
+This is troubling given the costs of misinformed climate policy.
+If anthropogenic climate change is a myth but is taken to be true, a tremendous amount of public resources will be wasted on adaptation and mitigation efforts.
+On the other hand, if climate change is human-induced but not recognized as so, the costs of inaction would be huge.
 Given that effective climate policies require long-lasting measures to be taken, the importance of an accurate consensus on the issue is manifest.
@@
 Attempts to foster such consensus face many social and psychological challenges.
-Recently, people have argued that such challenges could be met by setting up climate prediction markets [@Hsu2011,@Vandenbergh2013f].
+Such challenges could be met by setting up climate prediction markets [@Hsu2011; @Vandenbergh2013f].
 The idea of using prediction markets to efficiently aggregate information about uncertain events has been discussed for decades [@Horn2014c].
-Prediction markets have been shown to have interesting theoretical properties [@Set1998,Hanson2012] and to perform well in terms of prediction accuracy in experiments [@Hanson2006,@Healy2010], agent-based models (ABM) [@Klingert2012c,@Jumadinova2011a], and in the real-world [@Wolfers2006].
+Prediction markets have been shown to have interesting theoretical properties [@Set1998; @Hanson2012] and to perform well in terms of prediction accuracy in experiments [@Hanson2006; @Healy2010], agent-based models (ABM) [@Klingert2012c; @Jumadinova2011a], and the real-world [@Wolfers2006].
+We simulate a prediction market for climate outcomes to explore whether, and under what conditions, prediction markets are useful for increasing convergence of climate beliefs.
 
-[@Tseng2010a] study an ABM of a continuous double auction model with various types of market strategies (including two variants of the zero-intelligence agents).
+# 2. Related Work
+
+@Tseng2010a study an ABM of a continuous double auction market with various types of market strategies (including two variants of the zero-intelligence agent).
 They compare the behavior of their simulation with data from a prediction market for the outcome of political elections.
 They find that despite their simplicity, zero-intelligence agents capture some salient features of real market data.
-[@Klingert2012c] reach similar conclusions with respect to experimental data from  [@Hanson2006].
-In their simulation, [@Klingert2012c] compare continuous double auctions with logarithmic market scoring rules and find that the latter generally outperform continuous double auctions in terms of prediction accuracy.
+@Klingert2012c reach similar conclusions with respect to experimental data from @Hanson2006.
+In their simulation, @Klingert2012c compare continuous double auctions with logarithmic market scoring rules and find that the latter generally outperform continuous double auctions in terms of prediction accuracy.
 In some variants of the two last models, agents' strategies feature learning, but these are rather basic and unrelated to a formal predictive model.
 
-[@Jumadinova2011a] conducted research more similar to our project, as they explicitly include information about uncertain events in their model.
-The paper studies a continuous double auction model in which agents update their beliefs about the uncertain events based on information on the event.
+@Jumadinova2011a conducted research more similar to ours, as they explicitly include information about uncertain events in their model.
+@Jumadinova2011a study a continuous double auction model in which agents update their beliefs about the uncertain events based on information on the event.
 New beliefs are modeled as a weighted sum of the last period's beliefs and market prices.
-The agent's information set changes every period.
+The agents' information set changes every period.
 The better the information set, the more likely the agent is to put higher weights on last period's prices when revising their beliefs.
@@
 Therefore, the model does not allow for investigating the convergence of predictive models.
-[@Ontanon2009] study the effect of deliberation on a simulated prediction market.
-They assume that agents use Case-Based Reasoning [@Aamodt1994] to debate about  uncertain outcomes with their neighbors in a social network.
+@Ontanon2009 study the effect of deliberation on a simulated prediction market.
+They assume that agents use Case-Based Reasoning [@Aamodt1994] to reason and debate about uncertain outcomes with their neighbors in a social network.
 Although the model could in theory be used to assess convergence of predictive models by looking at the arguments agents adopt after the deliberation, this is not done by the authors.
-Also, the way beliefs are formed and revised through argumentation is rather abstract and might not be ideal to study climate prediction markets.
+Also, the way beliefs are formed and revised through argumentation is abstract and not ideal for studying climate prediction markets.
 
-However, to the best of our knowledge, the idea that prediction markets can generate consensus *on the factors* affecting  uncertain events has never been quantitatively explored.
-ABMs of prediction markets have been studied [@Klingert2012c,@Tseng2010a,@Jumadinova2011a], some of which feature communication between agents.
+To the best of our knowledge, the idea that prediction markets can generate consensus *on the factors* affecting  uncertain events has never been quantitatively explored.
+ABMs of prediction markets have been studied [@Klingert2012c; @Tseng2010a; @Jumadinova2011a], some of which feature communication between agents.
 In these models, however, beliefs about the uncertain outcomes are constructed in rather abstract ways.
@@
 From a public policy perspective, changing the explanatory models of market participants is one of the most important roles that prediction markets might play.
-And perhaps the most important potential social benefit of prediction markets given the comparable predictive power of statistical models, which are less expensive than setting up and maintaining prediction markets [@goel_prediction_2010].
+This is perhaps the most important social benefit of prediction markets given the comparable predictive power of statistical models, which are less expensive than setting up and maintaining prediction markets [@goel_prediction_2010].
 Effective climate policy not only requires an accurate consensus on future climate *outcomes*.
@@
 If people agree that temperature will rise, but some believe it will be due to greenhouse gases, while others believe that it will be caused by increased solar activity, inconsistent and ineffective policies may be implemented.
-Our model is designed to investigate whether, and under what conditions, prediction markets can foster such convergence in predictive models.
+Our model is designed to investigate whether prediction markets can foster such convergence in predictive models.
 

### 2016-02-29 paper.Rmd
Paper: added the descriptions of model design and params

@@
 
+## Temperature Models
+
+JG - describe the time series models here.
+
+## Beliefs
+
+Traders use one of the two models to forecast future temperatures in order to trade on the prediction market.
+These $2$ models are interpreted as the trader's beliefs about the true climate model.
+They are meant to represent pervasive positions on climate change in the public debate.
+In order to approximatively match the current configuration of beliefs in climate change in the US, model 1 is randomly assigned to half of the traders, while 2 is assigned to the other half of the traders.
+
+ Thus, both when **true.model** = 0 and when **true.model** = 1,  some traders use the true model to make predictions.
+However, this does not mean that traders using the true model make perfectly accurate predictions.
+Although these traders believe in the correct *functional form* of the model, they still need to calibrate it based on limited noisy data.
+Therefore, the values they use as  the parameters of the model will typically still be different from the true parameters of the data-generating process.
+
+## Traders and Market
+
+Traders are initially endowed with one single unit Experimental Currency Unit (ECU).
+Traders use their believed model to forecast the distribution of future temperatures and determine their reservation price for different securities.
+Securities pay 1 ECU  if the actual temperature at some time $t^*$ falls between a certain range.
+Traders are assumed to be risk-neutral expected utility maximizers.
+Therefore, their reservation price for a security $\tilde{s}$ is simply their assessment of the probability that the temperature at time $t^*$ falls in the range covered by $\tilde{s}$.
+
+Based on their reservation price, traders behave as "zero-intelligence" agents [@Gode1993], that is, they sell at a random price above their reservation, and buy at a random price below their reservation.
+The distribution from which buy and sell prices are drawn is determined by the traders risk attitude, characterized by parameter $risk.tak_i$.
+These strategies are simple but have been shown to provide good approximations of the behavior on prediction markets (see for instance [@Klingert2012c]), and in financial markets more broadly.
+
+Continuous-double auctions (CDAs) or some variants thereof are common market mechanisms, i.e.
+procedures to match buy and sell orders.
+CDA are notably used on large stock markets (see for instance [@Tseng2010a]).
+Our model features a particular version of CDA through which traders exchange securities.
+
+## Social Network
+
+Traders are part of a social network.
+Every time a security is realized, each trader $i$ looks at the performance of their neighbors in the network.
+If one of $i$'s neighbors, say $j$, is richer than $i$, $i$ interprets this has an indication that $j$ has a better approximate model.
+Then $i$ considers adopting $j$'s approximate model.
+Traders start with the same initial amount of money, so differences in money among traders can only come from  traders' interactions through the market.} For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is characterized by parameter $ideo_i$.
+
+ An example of such social network is depicted in Figure X.
+We assume that the initial network is segregated, that is traders who believe in model 2 (1) are more likely to be linked with other traders who believe in 2 (1).
+The degree of segregation is determined by parameter **seg**.
+Although traders can change their approximate model as time passes, the connections between traders do not change as the market unfolds (i.e.
+the edges are fixed).
+
+## Time
+
+The time periods $t$ are grouped in trading *sequences*.
+In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence.
+For instance, the third trading sequence might start in period $t=1964$ and end in period $t = 1970$.
+In this case, a security traded in the third sequence pay 1 ECU if the temperature at $t=1970$ falls into the range of temperatures covered by the security.
+
+At each time $t$, traders are assumed to know the past value of the temperature $\hat{T}_{0:t}$ and greenhouse gas emissions  $GHG_{0:t}$.
+In a sequence finishing at time $t^*$,  traders also have common knowledge of $GHG_{t:t^*}$, the future values of greenhouse gas emissions up to $t^*$.
+However, at any $t$, traders do not know the value of any future *temperatures*.
+In particular, in a given sequence, traders do not know the value of $\hat{T}_{t^*}$.
+Traders can only predict $\hat{T}_{t^*}$ using their approximate model and their knowledge of $\hat{T}_{0:t}$ and  $GHG_{0:t^*}$.
+Notice that because $GHG_{0:t^*}$ is common knowledge, in each period $t$ every trader $i$ with the same approximate model forms the same pseudo-expectation
+
+ 	  $\mu_{ti} \equiv E_t( \hat{T}_{t^*} ~|~ \hat{T}_{0:t}, GHG_{0:t^*}$
+
+ 	  $i$'s approximate model at t.
+
+ 	We assume that the believed probability distribution of $\hat{T}_{t^*}$ for trader $i$  at any time $t$ is
+
+ 	   $\mu_{ti} + \eta_t$
+
+where the distribution $\eta_t$ is obtained by trader $i$ by bootstrapping the residuals from the regression of her believed true model over the past periods.
+This explains the use of the term pseudo-expectation.
+Because $\eta_t$ is not bound to have expected value zero, a traders' effective expected value of $\hat{T}_{t^*}$ may be different from $\mu_{ti}$.
+
+ At each time $t$, traders
+
+1. recalibrate their approximate model based on the new set of past data available at $t$ (via pooled ordinary least squares),
+2. compute their belief-density for $\hat{T}_{t^*}$ and use it to determine the expected value they attached to each security,
+3. and trade on the CDA market based on their zero-intelligence decision rule as follows. 
+ 	+ Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
+ 	+ Every trader $i$ also chooses at random among the securities she owns a positive amount of (if any) a security $s_i^S$ she will try to sell.
+	+ Traders then decide of their selling $p_i^S$ and buying price $p_i^S$.
+To do so, they compute the expected value of the securities $s_i^B$ and $s_i^S$.
+Then they set a selling and a buying price following the zero-intelligence rule described above.
+	+ Traders go to the market one at the time, in a random order.
+ 	+ When $i$ comes to the market, she place limit orders in the order book.
+These orders  specify that $i$ is willing to buy $s_i^B$ at  any price below $p_i^B$, and to sell $s_i^S$ at any price above $p_i^S$.
+ 	+ The market maker then tries to match $i$'s orders with some order which was put in the order book before $i$ came to the market.
+ 	+ If there are  outstanding sell offers for $s_i^B$ at price $p$ lower than $p_i^B$, then a trade is concluded.
+Trader $i$ buys one unit from the sellers who sells at the highest price below $p_i^B$, and the sell and buy offers are removed from the order book.
+ 	+ If there are  outstanding buy offers for $s_i^S$ at price $p$ higher than $p_i^S$, then a trade is concluded.
+Trader $i$ sells one unit to the buyer who buys at the highest price above $p_i^S$, and the  sell and buy offers are removed from the order book.
+ 	+ Whenever all traders have come to the market, any remaining outstanding offer is removed from the order book, and the trading period is concluded.
+
+At $t^*$, when the sequence ends, there is only one security $s^*$ associated with a range of temperatures including the actual $\hat{T}_{t^*}$. Thus,
+
+1. traders receive 1 ECU per unit of $s^*$ they own, and
+2.  consider adapting their neighbors' approximate model as described above.
+
+# Model Parameters
+
+The model depends on the following parameters.
+
+* Network parameters
+    + **n.traders**: the number of traders.
+    + **n.edg**: the number of edges in the social network.
+    + **seg**: determines the initial degree of homophily in the network. The higher **seg**, the higher the initial homophily.
+
+ * Market structure parameters
+    + **market.complet** Determines market's completeness, i.e.
+the number of securities which can be traded.
+With more securities, the interval of temperatures corresponding to each security is smaller.
+Traders can then trade on more precise temperature intervals.
+Higher values of **market.complet** may, however, reduce the number of exchanges.
+Because traders pick the securities they buy and sell at random, the probability that a match between sellers and buyers is found is lower for higher values of **market.complet**.
+
+* Behavioral parameters.
+    + **risk.tak** Determines the distribution of risk taking behavior.
+The higher risk.tak, the more traders	will try to buy (resp.
+sell) lower (resp.
+higher) than their reservation price.
+Formally, at time $t$, trader $i$ picks her buying (resp.
+selling) prices randomly in the interval  $[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it}  * (1 - \texttt{risk.tak}_i)]$ (resp.
+$[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it} * (1  + \texttt{risk.tak}_i)]$), where $\texttt{reserv}_{it}$ is $i$'s reservation price at time $t$ for the security $i$ picked to buy (resp.
+sell).
+For each trader $i$, the level of \texttt{risk.tak}$_i$ is drawn uniformly at random from $[0,$\texttt{risk.tak} $]$
+    + **ideo**  Determines the degree of "ideology" of traders.
+If ideo is high, traders will not revise their approximate models easily, even when faced with strong evidence that their neighbors are doing better than them.
+For each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$.
+The value of $d_{i}$ is the probability that $i$ adopts one of her neighbors' approximate model if this neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
+* Timing parameters
+    + **burn.in** The number of burn-in periods in which no securities are traded (necessary to allow believed models to be estimated in the first period).
+    + **n.seq** Number of trading sequences.
+    + **horizon** Number of trading periods in each trading sequence.
+
 # 3. Results

### 2016-03-01 paper.Rmd
Work on paper.

@@
 ---
-title: "Climate Prediction Markets and the Convergence of Forecasting Models"
+title: "Climate Prediction Markets and the Convergence of Forecasting Models: An Agent-Based Simulation"
 author: " "
-date: "Feb 23, 2016"
+date: "Feb 29, 2016"
 output: 
@@
 
-We present an agent-based model of a climate prediction market.
-Trader agents adapt their beliefs about the climate based on the monetary performance of their neighboring traders in a social network.
-We analyze the model to determine how a variety of market and climate factors affect whether climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
+We present an agent-based simulation model of a climate prediction market.
+Traders adapt their beliefs about the future global temperatures based on the monetary performance of their neighboring traders in a social network.
+We conduct simulation experiments with the model to determine how a variety of market factors and the climate data-generating process affect whether climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
 
@@
 
-The previous two decades saw a strong polarization of the climate change debate.
+The climate change debate has become strongly polarized over the past twenty years.
 Although the scientific consensus on the anthropogenic nature of climate change strongly increased, the average belief about anthropogenic climate change did not evolve much within the public [@Vandenbergh2013f].
-In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily, indicating that the question is becoming increasingly politicized and potentially disconnected from scientific evidence.
-This is troubling given the costs of misinformed climate policy.
-If anthropogenic climate change is a myth but is taken to be true, a tremendous amount of public resources will be wasted on adaptation and mitigation efforts.
-On the other hand, if climate change is human-induced but not recognized as so, the costs of inaction would be huge.
+In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily as the question is becoming increasingly politicized and potentially disconnected from scientific evidence.
+The costs of misinformed climate policies are high.
+If anthropogenic climate change is a myth but is taken to be true, public resources will be spent on unnecessary efforts.
+On the other hand, if climate change is human-induced but not recognized as so, the costs of inaction could be devastating.
 Given that effective climate policies require long-lasting measures to be taken, the importance of an accurate consensus on the issue is manifest.
 
-Attempts to foster such consensus face many social and psychological challenges.
-Such challenges could be met by setting up climate prediction markets [@Hsu2011; @Vandenbergh2013f].
-The idea of using prediction markets to efficiently aggregate information about uncertain events has been discussed for decades [@Horn2014c].
-Prediction markets have been shown to have interesting theoretical properties [@Set1998; @Hanson2012] and to perform well in terms of prediction accuracy in experiments [@Hanson2006; @Healy2010], agent-based models (ABM) [@Klingert2012c; @Jumadinova2011a], and the real-world [@Wolfers2006].
-We simulate a prediction market for climate outcomes to explore whether, and under what conditions, prediction markets are useful for increasing convergence of climate beliefs.
+Attempts to foster such consensus face many social and psychological challenges, some of which could be adressed by creating climate prediction markets [@Hsu2011; @Vandenbergh2013f].
+The idea of using prediction markets to efficiently aggregate information about uncertain events has been widely discussed [@Horn2014c].
+Prediction markets have been shown to have interesting theoretical properties [@Set1998; @Hanson2012] and to perform well in terms of prediction accuracy and information aggregation in experiments [@Hanson2006; @Healy2010], agent-based models (ABM) [@Klingert2012c; @Jumadinova2011a], and the real-world [@Wolfers2006].
+We simulate a prediction market where traders exchange securities related to climate outcomes to explore whether, and under what market and climate conditions, prediction markets may be useful for increasing convergence of climate beliefs.
 
@@
 
-JG - describe the time series models here.
+JG - describe the temperature time series models here.
 
-## Beliefs
+## Climate Beliefs
 
-Traders use one of the two models to forecast future temperatures in order to trade on the prediction market.
-These $2$ models are interpreted as the trader's beliefs about the true climate model.
-They are meant to represent pervasive positions on climate change in the public debate.
-In order to approximatively match the current configuration of beliefs in climate change in the US, model 1 is randomly assigned to half of the traders, while 2 is assigned to the other half of the traders.
+Traders use one of the two models to forecast future temperatures.
+These models are interpreted as the trader's beliefs about the true climate process, i.e.
+the driving factors of long-term global temperatures.
+They represent pervasive positions on climate change in the public debate.
+In order to approximatively match the current configuration of beliefs in climate change in the United States, model 1 is randomly assigned to half of the traders, while 2 is assigned to the other half of the traders.
 
- Thus, both when **true.model** = 0 and when **true.model** = 1,  some traders use the true model to make predictions.
+Thus, both when **true.model** = 0 and **true.model** = 1,  some traders use the true data-generating model to make predictions.
 However, this does not mean that traders using the true model make perfectly accurate predictions.
 Although these traders believe in the correct *functional form* of the model, they still need to calibrate it based on limited noisy data.
-Therefore, the values they use as  the parameters of the model will typically still be different from the true parameters of the data-generating process.
+Therefore, the values they use as the parameters of the model will typically still be different from the exact parameters of the data-generating process.
 
@@
 
-Traders are initially endowed with one single unit Experimental Currency Unit (ECU).
-Traders use their believed model to forecast the distribution of future temperatures and determine their reservation price for different securities.
+Traders are initially endowed with a single unit Experimental Currency Unit (ECU).
+Traders use their model to forecast the distribution of future temperatures and determine their reservation price for different securities.
 Securities pay 1 ECU  if the actual temperature at some time $t^*$ falls between a certain range.
-Traders are assumed to be risk-neutral expected utility maximizers.
+Traders are risk-neutral expected utility maximizers.
 Therefore, their reservation price for a security $\tilde{s}$ is simply their assessment of the probability that the temperature at time $t^*$ falls in the range covered by $\tilde{s}$.
@@
 The distribution from which buy and sell prices are drawn is determined by the traders risk attitude, characterized by parameter $risk.tak_i$.
-These strategies are simple but have been shown to provide good approximations of the behavior on prediction markets (see for instance [@Klingert2012c]), and in financial markets more broadly.
+These strategies are simple but have been shown to provide accurate approximations of behavior in prediction markets (see for instance @Klingert2012c), and in financial markets more broadly.
 
@@
 procedures to match buy and sell orders.
-CDA are notably used on large stock markets (see for instance [@Tseng2010a]).
+CDA are notably used on large stock markets [@Tseng2010a].
 Our model features a particular version of CDA through which traders exchange securities.
@@
 Traders are part of a social network.
+Jonathan - add in changes made to social network.
 Every time a security is realized, each trader $i$ looks at the performance of their neighbors in the network.
@@
 Then $i$ considers adopting $j$'s approximate model.
-Traders start with the same initial amount of money, so differences in money among traders can only come from  traders' interactions through the market.} For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is characterized by parameter $ideo_i$.
+Traders start with the same initial amount of money, so differences in money among traders can only come from  traders' interactions through the market.
+For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is characterized by parameter $ideo_i$.
 
 An example of such social network is depicted in Figure X.
-We assume that the initial network is segregated, that is traders who believe in model 2 (1) are more likely to be linked with other traders who believe in 2 (1).
+We assume that the initial network is segregated, that is, traders who believe in model 2 (1) are more likely to be linked with other traders who believe in 2 (1).
 The degree of segregation is determined by parameter **seg**.
-Although traders can change their approximate model as time passes, the connections between traders do not change as the market unfolds (i.e.
-the edges are fixed).
+Although traders can change their approximate model over time, the connections between traders do not change as the market unfolds -- the edges are fixed.
 
@@
 
-The time periods $t$ are grouped in trading *sequences*.
+The time periods, $t$, are grouped in trading *sequences*.
 In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence.
-For instance, the third trading sequence might start in period $t=1964$ and end in period $t = 1970$.
-In this case, a security traded in the third sequence pay 1 ECU if the temperature at $t=1970$ falls into the range of temperatures covered by the security.
+For instance, the third trading sequence might start in period $t = 1964$ and end in period $t = 1970$.
+In this case, a security traded in the third sequence pays 1 ECU if the temperature at $t = 1970$ falls into the range of temperatures covered by the security.
 
@@
 Traders can only predict $\hat{T}_{t^*}$ using their approximate model and their knowledge of $\hat{T}_{0:t}$ and  $GHG_{0:t^*}$.
-Notice that because $GHG_{0:t^*}$ is common knowledge, in each period $t$ every trader $i$ with the same approximate model forms the same pseudo-expectation
+Notice that because $GHG_{0:t^*}$ is common knowledge, in each period $t$ every trader $i$ with the same approximate model forms the same pseudo-expectation $\mu_{ti} \equiv E_t( \hat{T}_{t^*} ~|~ \hat{T}_{0:t}, GHG_{0:t^*})$, $i$'s approximate model at $t$.
 
- 	  $\mu_{ti} \equiv E_t( \hat{T}_{t^*} ~|~ \hat{T}_{0:t}, GHG_{0:t^*}$
-
- 	  $i$'s approximate model at t.
-
- 	We assume that the believed probability distribution of $\hat{T}_{t^*}$ for trader $i$  at any time $t$ is
-
- 	   $\mu_{ti} + \eta_t$
-
-where the distribution $\eta_t$ is obtained by trader $i$ by bootstrapping the residuals from the regression of her believed true model over the past periods.
+We assume that the believed probability distribution of $\hat{T}_{t^*}$ for trader $i$  at any time $t$ is $\mu_{ti} + \eta_t$, where the distribution $\eta_t$ is obtained by trader $i$ by bootstrapping the residuals from the regression of her believed true model over the past periods.
 This explains the use of the term pseudo-expectation.
@@
 
-The model depends on the following parameters.
+The model depends on the following parameters, which we vary in simulation experiments to determine their effects on the convergence of beliefs.
 
@@
 
-The sensitivity analysis is based on a Latin Hypercube Sampling of ... parameter sets from the following distributions [@beachkofski_improved_2002,@carnell_lhs_2012].
+The sensitivity analysis is based on a Latin Hypercube Sampling of parameter sets from the following distributions [@beachkofski_improved_2002,@carnell_lhs_2012].
 
@@
 
-We use the model to simulate 10,000 outcomes based on the input paramter sets and then conduct a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, resulting simulated vector, $y$ [@Marino2008,@pujol_sensitivity:_2014,@saltelli_sensitivity_2009].
+We use the model to simulate 10,000 outcomes based on the input paramter sets and then conduct a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, resulting simulated vector, $y$ [@Marino2008; @pujol_sensitivity:_2014; @saltelli_sensitivity_2009].
 Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).
@@
 
-Our sensitivity analysis, illustrated by ...,
+Our sensitivity analysis, Fig X.
 

### 2016-03-01 master.R
master small

@@ -59,8 +59,8 @@ if(estimate_replicates){
 ##############################################################################
 
 future <- TRUE
-past <- TRUE
-sample_count <- 120
+past <- FALSE
+sample_count <- 150
 
 if(past){
   main2 <- function(parameters,

### 2016-03-01 paper.Rmd
paper plot

@@
 author: " "
-date: "Feb 29, 2016"
+date: "March 1, 2016"
 output: 
@@
 
-Our sensitivity analysis, Fig X.
+Our sensitivity analysis, `r figRef("sa")`, shows that
 
-```{r, one, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=tabRef("one", "dfd")}
+```{r sa, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("sa", "Sensitivty analysis"), fig.width=5, fig.height=5}
+library(eat)
+load("output/src_future.Rda")
+plot(src_future, outcome_var = "Convergence") + theme_classic() 
+#multi_plot(p1, p2, cols = 2)
+```
+
+```{r sa2, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=tabRef("sa2", "")}
 
@@
  `r tabRef("one")`
-
-`r figRef("data_explore")`
-
-```{r, data_explore, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("data_explore", " "), fig.width=10, fig.height=5}
-
-#multi_plot(p1, p2, cols = 2)
-```
 

### 2016-03-01 paper.Rmd
paper

@@
 
-Traders use one of the two models to forecast future temperatures.
-These models are interpreted as the trader's beliefs about the true climate process, i.e.
-the driving factors of long-term global temperatures.
+Traders use one of the two models to forecast future temperature anomalies.
+These models are interpreted as the trader's beliefs about the true climate process, the driving factor of long-term global temperatures.
 They represent pervasive positions on climate change in the public debate.
@@
 
-Thus, both when **true.model** = 0 and **true.model** = 1,  some traders use the true data-generating model to make predictions.
-However, this does not mean that traders using the true model make perfectly accurate predictions.
+Thus, both when **true.model** = 0 and **true.model** = 1, approximately half of the traders use the true data-generating model to make predictions.
+However, traders using the true model do not necessarily make perfectly accurate predictions.
 Although these traders believe in the correct *functional form* of the model, they still need to calibrate it based on limited noisy data.
@@
 The distribution from which buy and sell prices are drawn is determined by the traders risk attitude, characterized by parameter $risk.tak_i$.
-These strategies are simple but have been shown to provide accurate approximations of behavior in prediction markets (see for instance @Klingert2012c), and in financial markets more broadly.
+These trading strategies are simple but have been shown to provide accurate approximations of behavior in prediction markets [@Klingert2012c], and in financial markets more broadly.
 
+Our model features a particular version of CDA through which traders exchange securities.
 Continuous-double auctions (CDAs) or some variants thereof are common market mechanisms, i.e.
@@
 CDA are notably used on large stock markets [@Tseng2010a].
-Our model features a particular version of CDA through which traders exchange securities.
 
@@
 
-Traders are part of a social network.
-Jonathan - add in changes made to social network.
+Traders are part of a social network where each agent forms two links at random, and then forms links randomly.
+This ensures that each agent is connected to at least two other agents.
 Every time a security is realized, each trader $i$ looks at the performance of their neighbors in the network.
@@
 
-# Model Parameters
+## Model Parameters
 

### 2016-03-03 master.R
more sa

@@ -60,7 +60,7 @@ if(estimate_replicates){
 
 future <- TRUE
 past <- FALSE
-sample_count <- 150
+sample_count <- 100
 
 if(past){
   main2 <- function(parameters,

### 2016-03-03 paper.Rmd

@@
 ---
-title: "Climate Prediction Markets and the Convergence of Forecasting Models: An Agent-Based Simulation"
-author: " "
-date: "March 1, 2016"
+title: "Convergence of Forecasting Models in Climate Prediction Markets: An Agent-Based Simulation"
+author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
+date: "March 2, 2016"
 output: 
@@
 library(ggplot2)
+theme_set(theme_classic())
 library(pander)
@@
 
-We present an agent-based simulation model of a climate prediction market.
+We present an agent-based simulation of a climate prediction market.
 Traders adapt their beliefs about the future global temperatures based on the monetary performance of their neighboring traders in a social network.
@@
 The climate change debate has become strongly polarized over the past twenty years.
-Although the scientific consensus on the anthropogenic nature of climate change strongly increased, the average belief about anthropogenic climate change did not evolve much within the public [@Vandenbergh2013f].
+Although the scientific consensus on the anthropogenic nature of climate change strongly increased, beliefs about anthropogenic climate change did not evolve much within the public [@Vandenbergh2013f].
 In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily as the question is becoming increasingly politicized and potentially disconnected from scientific evidence.
@@
 
-```{r sa, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("sa", "Sensitivty analysis"), fig.width=5, fig.height=5}
-library(eat)
-load("output/src_future.Rda")
-plot(src_future, outcome_var = "Convergence") + theme_classic() 
+```{r sa, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("sa", "Estimated Effects on Convergence of Beliefs in Future Scenario "), fig.width=4, fig.height=4.5}
+# library(eat)
+# load("output/src_future.Rda")
+# plot(src_future, outcome_var = "Convergence") + theme_classic() + theme(text=element_text(family='Palatino'))
+
+source("utils/sa_plot.R")
+sa_plot() + theme(text=element_text(family='Palatino'))
 #multi_plot(p1, p2, cols = 2)
+```
+
+```{r time, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("time", "Time"), fig.width=5, fig.height=10}
+source("utils/time_plot.R")
+time_plot() + theme(text=element_text(family='Palatino'))
 ```

### 2016-03-03 paper.Rmd
paper work. moved more util funcs to be sourced too.

@@
 ---
-title: "Convergence of Forecasting Models in Climate Prediction Markets: An Agent-Based Simulation"
+title: "Put Your Money Where Your Mouth Is: Prediction Markets and Climate Change"
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
-date: "March 2, 2016"
+date: "March 3, 2016"
 output: 
@@
 ---
-
 ```{r fig_ref, echo=FALSE}
-figRef <- local({
-  tag <- numeric()
-  created <- logical()
-  used <- logical()
-  function(label, caption, prefix="Fig.", sep=".",
-      prefix.highlight="**") {
-    i <- which(names(tag) == label)
-    if(length(i) == 0){
-      i <- length(tag) + 1
-      tag <<- c(tag, i)
-      names(tag)[length(tag)] <<- label
-      used <<- c(used, FALSE)
-      names(used)[length(used)] <<- label
-      created <<- c(created, FALSE)
-      names(created)[length(created)] <<- label
-    }
-    if(!missing(caption)){
-      created[label] <<- TRUE
-      paste0(prefix.highlight, prefix, " ", i, sep, prefix.highlight, " ", caption)
-    } else {
-      used[label] <<- TRUE
-      paste(prefix, tag[label])
-    }
-  }
-})
-# adapted from a function on this page: http://galahad.well.ox.ac.uk/repro/
-rinline <- function(code){
-  sprintf('`r %s`', code)
-}
+source("utils/fig_ref.R") # tabRef and figRef functions
+source("utils/multi_plot.R")
 ``` 
-
 ```{r, caption_setting, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
@@
 ```
-
-```{r tab_ref, echo=FALSE}
-tabRef <- local({
-    tag <- numeric()
-    created <- logical()
-    used <- logical()
-    function(label, caption, prefix = "Table", 
-        sep = ".", prefix.highlight = "**") {
-        i <- which(names(tag) == label)
-        if (length(i) == 0) {
-            i <- length(tag) + 1
-            tag <<- c(tag, i)
-            names(tag)[length(tag)] <<- label
-            used <<- c(used, FALSE)
-            names(used)[length(used)] <<- label
-            created <<- c(created, FALSE)
-            names(created)[length(created)] <<- label
-        }
-        if (!missing(caption)) {
-            created[label] <<- TRUE
-            paste0(prefix.highlight, prefix, " ", i, sep, prefix.highlight, 
-                " ", caption)
-        } else {
-            used[label] <<- TRUE
-            paste(prefix, tag[label])
-        }
-    }
-})
-```
-
 ```{r multi_plot, echo=FALSE}
-multi_plot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
-  library(grid)
-
-  # Make a list from the ... arguments and plotlist
-  plots <- c(list(...), plotlist)
-
-  numPlots = length(plots)
-
-  # If layout is NULL, then use 'cols' to determine layout
-  if (is.null(layout)) {
-    # Make the panel
-    # ncol: Number of columns of plots
-    # nrow: Number of rows needed, calculated from # of cols
-    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
-                    ncol = cols, nrow = ceiling(numPlots/cols))
-  }
-
- if (numPlots==1) {
-    print(plots[[1]])
-
-  } else {
-    # Set up the page
-    grid.newpage()
-    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
-
-    # Make each plot, in the correct location
-    for (i in 1:numPlots) {
-      # Get the i,j matrix positions of the regions that contain this subplot
-      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
-
-      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
-                                      layout.pos.col = matchidx$col))
-    }
-  }
-}
-
+source("utils/multi_plot.R")
 library(ggplot2)
@@
 
-We present an agent-based simulation of a climate prediction market.
+We present a simulation of a climate prediction market.
 Traders adapt their beliefs about the future global temperatures based on the monetary performance of their neighboring traders in a social network.
-We conduct simulation experiments with the model to determine how a variety of market factors and the climate data-generating process affect whether climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
+We conduct experiments with the model to determine how a variety of market factors and the climate data-generating process affect whether climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
 
@@
 
-Attempts to foster such consensus face many social and psychological challenges, some of which could be adressed by creating climate prediction markets [@Hsu2011; @Vandenbergh2013f].
+Attempts to foster such consensus face many social and psychological challenges, some of which could be adressed by creating climate prediction markets where participants can "put their money where their mouths are" [@Hsu2011; @Vandenbergh2013f].
 The idea of using prediction markets to efficiently aggregate information about uncertain events has been widely discussed [@Horn2014c].
-Prediction markets have been shown to have interesting theoretical properties [@Set1998; @Hanson2012] and to perform well in terms of prediction accuracy and information aggregation in experiments [@Hanson2006; @Healy2010], agent-based models (ABM) [@Klingert2012c; @Jumadinova2011a], and the real-world [@Wolfers2006].
+Prediction markets have been shown to have interesting theoretical properties [@Set1998; @Hanson2012] and to perform well in terms of prediction accuracy and information aggregation in experiments [@Hanson2006; @Healy2010], simulation models [@Klingert2012c; @Jumadinova2011a], and the real-world [@Wolfers2006].
 We simulate a prediction market where traders exchange securities related to climate outcomes to explore whether, and under what market and climate conditions, prediction markets may be useful for increasing convergence of climate beliefs.
@@
 
-@Tseng2010a study an ABM of a continuous double auction market with various types of market strategies (including two variants of the zero-intelligence agent).
+@Tseng2010a create an agent-based model (ABM) of a continuous double auction market with various types of market strategies, including two variants of the zero-intelligence agent.
 They compare the behavior of their simulation with data from a prediction market for the outcome of political elections.
@@
 The agents' information set changes every period.
-The better the information set, the more likely the agent is to put higher weights on last period's prices when revising their beliefs.
-Not much structure is imposed on the information set and the way it is used to generate the weights in the update function.
+The better the information set, the more likely the agent is to put higher weight on last period's prices when revising their beliefs.
+Little structure is imposed on the information set and the way it is used to generate the weights in the update function.
 Therefore, the model does not allow for investigating the convergence of predictive models.
@@
 Although the model could in theory be used to assess convergence of predictive models by looking at the arguments agents adopt after the deliberation, this is not done by the authors.
-Also, the way beliefs are formed and revised through argumentation is abstract and not ideal for studying climate prediction markets.
+Also, the way beliefs are formed and revised through argumentation is too abstract for studying climate prediction markets, where beliefs should be based on statistical time series models.
 
@@
 From a public policy perspective, changing the explanatory models of market participants is one of the most important roles that prediction markets might play.
-This is perhaps the most important social benefit of prediction markets given the comparable predictive power of statistical models, which are less expensive than setting up and maintaining prediction markets [@goel_prediction_2010].
+This is perhaps the most important social benefit of prediction markets given the comparable predictive power of statistical supervised learning models, which can provide information at a much lower cost than creating and maintaining a prediction market [@goel_prediction_2010].
 Effective climate policy not only requires an accurate consensus on future climate *outcomes*.
@@
 If people agree that temperature will rise, but some believe it will be due to greenhouse gases, while others believe that it will be caused by increased solar activity, inconsistent and ineffective policies may be implemented.
-Our model is designed to investigate whether prediction markets can foster such convergence in predictive models.
 
@@
 
-JG - describe the temperature time series models here.
+We start with the historical temperature record and project future climates under two alternative theories:
+
+Temperature proportional to log(CO2) or Total Solar Irradiance (conventional science vs. popular alternative among doubters). Add AR(1) noise (best ARMA noise model for historical record).
+
+JG - describe the temperature time series models here. Add the plots of the time series diverging thats on middle of poster. Add citations for the models to library.bib.
 
@@
 
-Traders use one of the two models to forecast future temperature anomalies.
+Traders use one of the two models (temperature depends on CO2 or TSI) to forecast future temperature anomalies.
 These models are interpreted as the trader's beliefs about the true climate process, the driving factor of long-term global temperatures.
@@
 
-## Time
+## Model Dynamics
+
+Bet on temperature six years in future.
+During six-year period, traders buy and sell futures.
+Every year: Bayesian updating of traders personal predictions for future temperature based on current year’s temperature.
+At end of six-year period, winners collect money.
+Traders revise beliefs about climate models, based on ideology and beliefs of top earners in their social network.
 
@@
 
-The sensitivity analysis is based on a Latin Hypercube Sampling of parameter sets from the following distributions [@beachkofski_improved_2002,@carnell_lhs_2012].
+## Historical Data
+
+As a validation that the market is operating correctly, we run the market with actual temperatures 1880 to 2015 and market betting from 1931 to 2014.
+As warming signal begins to show up in 1980s, traders begin to converge toward log(CO2) model.
+
+## Future Scenarios
+
+The sensitivity analysis is based on a Latin Hypercube Sampling of parameter sets from the following distributions [@beachkofski_improved_2002; @carnell_lhs_2012].
 
@@
 
-Our sensitivity analysis, `r figRef("sa")`, shows that
+Our sensitivity analysis, `r figRef("sa")`, shows that ideology, risk tolerance, the number of traders, and the segmentation of the social network significantly affect convergence.
+Segmented social networks can create an "echo-chamber" effect.
 
-```{r sa, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("sa", "Estimated Effects on Convergence of Beliefs in Future Scenario "), fig.width=4, fig.height=4.5}
+```{r sa, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("sa", "Estimated effects of model parameters on convergence of beliefs in future scenario."), fig.width=3.8, fig.height=4.5}
 # library(eat)
@@
 
-```{r time, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("time", "Time"), fig.width=5, fig.height=10}
+```{r time, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("time", "Time"), fig.width=5, fig.height=9}
 source("utils/time_plot.R")

### 2016-03-03 utils/fig_ref.R

@@ -1,0 +1,56 @@
+figRef <- local({
+  tag <- numeric()
+  created <- logical()
+  used <- logical()
+  function(label, caption, prefix="Fig.", sep=".",
+           prefix.highlight="**") {
+    i <- which(names(tag) == label)
+    if(length(i) == 0){
+      i <- length(tag) + 1
+      tag <<- c(tag, i)
+      names(tag)[length(tag)] <<- label
+      used <<- c(used, FALSE)
+      names(used)[length(used)] <<- label
+      created <<- c(created, FALSE)
+      names(created)[length(created)] <<- label
+    }
+    if(!missing(caption)){
+      created[label] <<- TRUE
+      paste0(prefix.highlight, prefix, " ", i, sep, prefix.highlight, " ", caption)
+    } else {
+      used[label] <<- TRUE
+      paste(prefix, tag[label])
+    }
+  }
+})
+# adapted from a function on this page: http://galahad.well.ox.ac.uk/repro/
+rinline <- function(code){
+  sprintf('`r %s`', code)
+}
+
+tabRef <- local({
+  tag <- numeric()
+  created <- logical()
+  used <- logical()
+  function(label, caption, prefix = "Table", 
+           sep = ".", prefix.highlight = "**") {
+    i <- which(names(tag) == label)
+    if (length(i) == 0) {
+      i <- length(tag) + 1
+      tag <<- c(tag, i)
+      names(tag)[length(tag)] <<- label
+      used <<- c(used, FALSE)
+      names(used)[length(used)] <<- label
+      created <<- c(created, FALSE)
+      names(created)[length(created)] <<- label
+    }
+    if (!missing(caption)) {
+      created[label] <<- TRUE
+      paste0(prefix.highlight, prefix, " ", i, sep, prefix.highlight, 
+             " ", caption)
+    } else {
+      used[label] <<- TRUE
+      paste(prefix, tag[label])
+    }
+  }
+})

### 2016-03-03 utils/multi_plot.R

@@ -1,0 +1,35 @@
+multi_plot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
+  library(grid)
+  
+  # Make a list from the ... arguments and plotlist
+  plots <- c(list(...), plotlist)
+  
+  numPlots = length(plots)
+  
+  # If layout is NULL, then use 'cols' to determine layout
+  if (is.null(layout)) {
+    # Make the panel
+    # ncol: Number of columns of plots
+    # nrow: Number of rows needed, calculated from # of cols
+    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
+                     ncol = cols, nrow = ceiling(numPlots/cols))
+  }
+  
+  if (numPlots==1) {
+    print(plots[[1]])
+    
+  } else {
+    # Set up the page
+    grid.newpage()
+    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
+    
+    # Make each plot, in the correct location
+    for (i in 1:numPlots) {
+      # Get the i,j matrix positions of the regions that contain this subplot
+      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
+      
+      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
+                                      layout.pos.col = matchidx$col))
+    }
+  }
+}

### 2016-03-04 master.R
more SA results. everything significant.

@@ -60,7 +60,7 @@ if(estimate_replicates){
 
 future <- TRUE
 past <- FALSE
-sample_count <- 100
+sample_count <- 80
 
 if(past){
   main2 <- function(parameters,

### 2016-03-04 paper.Rmd
paper work.

@@
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
-date: "March 3, 2016"
+date: "March 4, 2016"
 output: 
@@
 
-Bet on temperature six years in future.
-During six-year period, traders buy and sell futures.
-Every year: Bayesian updating of traders personal predictions for future temperature based on current year’s temperature.
-At end of six-year period, winners collect money.
-Traders revise beliefs about climate models, based on ideology and beliefs of top earners in their social network.
+Traders bet on temperature six years in the future.
+During the six-year period, traders buy and sell futures.
+Every year, there is Bayesian updating of traders personal predictions for future temperature based on the current year's temperature.
+At the end of each six-year period, winners collect money and traders revise beliefs about climate models, based on ideology and beliefs of top earners in their social network.
 
-The time periods, $t$, are grouped in trading *sequences*.
+Specifically, the time periods, $t$, are grouped in trading *sequences*.
 In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence.
@@
 In this case, a security traded in the third sequence pays 1 ECU if the temperature at $t = 1970$ falls into the range of temperatures covered by the security.
-
 At each time $t$, traders are assumed to know the past value of the temperature $\hat{T}_{0:t}$ and greenhouse gas emissions  $GHG_{0:t}$.
 In a sequence finishing at time $t^*$,  traders also have common knowledge of $GHG_{t:t^*}$, the future values of greenhouse gas emissions up to $t^*$.
-However, at any $t$, traders do not know the value of any future *temperatures*.
+However, at any $t$, traders do not know the value of any future temperatures.
 In particular, in a given sequence, traders do not know the value of $\hat{T}_{t^*}$.
@@
 
- At each time $t$, traders
+ At each time $t$, traders follow this process:
 
-1. recalibrate their approximate model based on the new set of past data available at $t$ (via pooled ordinary least squares),
+1. recalibrate their approximate model based on the new set of past data available at $t$,
 2. compute their belief-density for $\hat{T}_{t^*}$ and use it to determine the expected value they attached to each security,
@@
 To do so, they compute the expected value of the securities $s_i^B$ and $s_i^S$.
-Then they set a selling and a buying price following the zero-intelligence rule described above.
-	+ Traders go to the market one at the time, in a random order.
- 	+ When $i$ comes to the market, she place limit orders in the order book.
+Then they set a selling and a buying price following the zero-intelligence rule.
+	+ Traders go to the market one at the time, in an order randomly selected each $t$.
+ 	+ When $i$ comes to the market, she places limit orders in the order book.
 These orders specify that $i$ is willing to buy $s_i^B$ at any price below $p_i^B$, and to sell $s_i^S$ at any price above $p_i^S$.
@@
 
-At $t^*$, when the sequence ends, there is only one security $s^*$ associated with a range of temperatures including the actual $\hat{T}_{t^*}$. Thus,
+At $t^*$, when the sequence ends, there is only one security $s^*$ associated with a range of temperatures including the actual $\hat{T}_{t^*}$. Thus:
 
@@
 
-* Network parameters
+* Network parameters:
     + **n.traders**: the number of traders.
     + **n.edg**: the number of edges in the social network.
-    + **seg**: determines the initial degree of homophily in the network. The higher **seg**, the higher the initial homophily.
+    + **seg**: the initial degree of homophily in the network. The higher **seg**, the higher the initial homophily.
 
- * Market structure parameters
-    + **market.complet** Determines market's completeness, i.e.
-the number of securities which can be traded.
+ * Market structure parameters:
+    + **market.complet** the number of securities which can be traded.
 With more securities, the interval of temperatures corresponding to each security is smaller.
@@
 * Behavioral parameters.
-    + **risk.tak** Determines the distribution of risk taking behavior.
+    + **risk.tak** Determines the distribution of risk-taking behavior.
 The higher risk.tak, the more traders	will try to buy (resp.
@@
 The value of $d_{i}$ is the probability that $i$ adopts one of her neighbors' approximate model if this neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
+
+IM GOING TO DELETE THESE 3:
 * Timing parameters

### 2016-03-04 paper.Rmd
overview in model design

@@
 
+Traders bet on temperature six years in the future.
+During the six-year period, traders buy and sell futures.
+Every year, there is Bayesian updating of traders personal predictions for future temperature based on the current year's temperature.
+At the end of each six-year period, winners collect money and traders revise beliefs about climate models, based on ideology and beliefs of top earners in their social network.
+In this section, we describe the models used to generate future temperature data, agents beliefs about those models, the market, and the social network connecting agents.
+We conclude with details on model dynamics and an overview of the model parameters.
+
 ## Temperature Models
@@
 
-Traders bet on temperature six years in the future.
-During the six-year period, traders buy and sell futures.
-Every year, there is Bayesian updating of traders personal predictions for future temperature based on the current year's temperature.
-At the end of each six-year period, winners collect money and traders revise beliefs about climate models, based on ideology and beliefs of top earners in their social network.
-
-Specifically, the time periods, $t$, are grouped in trading *sequences*.
+The time periods, $t$, are grouped in trading *sequences*.
 In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence.
@@
 
+Jonathan plot here.
+
 ## Future Scenarios

### 2016-03-04 master.R
add figure of social network

@@ -22,7 +22,7 @@ test <- FALSE
 estimate_replicates <- FALSE
 
 if(test){
-  main(c(runif(3, min = 0.0001, max = 0.9999), 1, runif(2, min = 0.0001, max = 0.9999)))
+  main(c(runif(3, min = 0.0001, max = 0.9999), 1, runif(2, min = 0.0001, max = 0.9999)), visu =T)
   main(c(runif(3, min = 0.0001, max = 0.9999), 0, runif(2, min = 0.0001, max = 0.9999)))
 }
 

### 2016-03-07 paper.Rmd
add figure of social network
fixed history plots to color by true model (co2 and tsi)

@@
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
-date: "March 4, 2016"
+date: "March 7, 2016"
 output: 
@@
 For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is characterized by parameter $ideo_i$.
+
+![Network](output/network.pdf)
 

### 2016-03-07 poster_plot_history.R

@@ -19,7 +19,7 @@ plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0
 start <- 1880 + 51
 sequence.years <- 6
 
-p <- ggplot(data=plot_data, aes(x= trading_seq * sequence.years + start, y=(convergence + 1) * 50)) +
+p <- ggplot(data=plot_data, aes(x= trading_seq * sequence.years + start, y=(convergence + 1) * 50,  color = true_mod)) +
   geom_point(position = position_jitter(w = 0.07 * sequence.years, h = 0), alpha = 0.2) + geom_smooth(size=1, fill=NA) +
   scale_color_brewer(palette="Set1", labels = c(LogCo2 = expression(CO[2]), SlowTSI = "TSI"), guide = "none") +
   ylim(0,100) +
@@ -39,21 +39,21 @@ if (TRUE) {
 }
 
 
-p <- ggplot(data=plot_data, aes(x= trading_seq * sequence.years + start, y=(convergence + 1) * 50) +
-              geom_boxplot(aes(group = floor(trading_seq * sequence.years + start)), 
-                           position = position_dodge(width = 3), width = 2) + 
-              geom_smooth(size=1, fill=NA) +
-              scale_color_brewer(palette="Set1", labels = c(LogCo2 = expression(CO[2]), SlowTSI = "TSI"), guide = "none") +
-              facet_wrap(~set, ncol = 1) +
-              labs(x = "Year",  
-                   y = bquote(plain("% Traders Believing ") ~ CO[2] ~ Model ~ (n == 
-                                                                                 .(prettyNum(length(complete.cases(plot_data$convergence)), big.mark = ',')))),
-                   title = paste0("Convergence Over Trading Sequences\n", 
-                                  start, "-",  start - 1 + 14 * 6))
-            ylim(0,100) +
-              theme(legend.justification=c(1,0), legend.position=c(1,0),
-                    legend.key.width = grid::unit(0.05, 'npc'),
-                    legend.key = element_rect(color=NA))
+p <- ggplot(data=plot_data, aes(x= trading_seq * sequence.years + start, y=(convergence + 1) * 50,  color = true_mod)) +
+  geom_boxplot(aes(group = floor(trading_seq * sequence.years + start * (true_mod == "SlowTSI"))), 
+               position = position_dodge(width = 3), width = 2) + 
+  geom_smooth(size=1, fill=NA) +
[…]

### 2016-03-07 paper.Rmd
Plotting code for paper.

@@
 We present a simulation of a climate prediction market.
-Traders adapt their beliefs about the future global temperatures based on the monetary performance of their neighboring traders in a social network.
+Traders adapt their beliefs about future global temperatures based on the monetary performance of their neighboring traders in a social network.
 We conduct experiments with the model to determine how a variety of market factors and the climate data-generating process affect whether climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
@@
 As a validation that the market is operating correctly, we run the market with actual temperatures 1880 to 2015 and market betting from 1931 to 2014.
-As warming signal begins to show up in 1980s, traders begin to converge toward log(CO2) model.
-
-Jonathan plot here.
+As warming signal begins to show up in 1980s, traders begin to converge toward log(CO2) model `r figRef("time")`.
 
@@
 
-```{r sa, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("sa", "Estimated effects of model parameters on convergence of beliefs in future scenario."), fig.width=3.8, fig.height=4.5}
+```{r sa, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("sa", "Estimated effects of model parameters on convergence of beliefs in future scenario."), fig.width=3, fig.height=3.5}
 # library(eat)
@@
 
-```{r time, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("time", "Time"), fig.width=5, fig.height=9}
+```{r time, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("time", "Convergence over trading sequences."), fig.width=10.5, fig.height=8.5}
 source("utils/time_plot.R")
-time_plot() + theme(text=element_text(family='Palatino'))
+p1 <- history_box() + theme(text=element_text(family='Palatino'))
+p2 <- future_box() + theme(text=element_text(family='Palatino'))
+multi_plot(p1, p2, cols=2)
 ```
@@
 # 4. Discussion
+
+Our code for the model is available at [johnjnay.com/predictMarket](http://johnjnay.com/predictMarket).
 

### 2016-03-07 poster_plot_history.R

@@ -50,13 +50,13 @@ p <- ggplot(data=plot_data, aes(x= trading_seq * sequence.years + start, y=(conv
                                                                      .(prettyNum(length(complete.cases(plot_data$convergence)), big.mark = ',')))),
        title = paste0("Convergence Over Trading Sequences\n", 
                       start, "-",  start - 1 + 14 * 6)) + 
-ylim(0,100) +
+  ylim(0,100) +
   theme(legend.justification=c(1,0), legend.position=c(1,0),
         legend.key.width = grid::unit(0.05, 'npc'),
         legend.key = element_rect(color=NA))
-            
-            print(p)
-            
+
+print(p)
+
             if (TRUE) {
               pdf("output/jg_timeseries_true_past_box.pdf", width=8, height=18)
               print(p)

### 2016-03-07 utils/sa_plot.R

@@ -11,12 +11,12 @@ sa_plot <- function(){
                                     labels = c('Ideology', "Risk tolerance",
                                                '# Traders', '# Edges/Trader', 'Segmentation',
                                                "True model")))
-  title <- bquote(R^2 == .(round(src_future@r_squared, 2)) * plain(', ') ~
-                            n == .(length(src_future@sims)))
+  title <- "" #bquote(R^2 == .(round(src_future@r_squared, 2)) * plain(', ') ~
+               #             n == .(length(src_future@sims)))
   
   ggplot(ss, aes(x = var, y = x_corr)) + 
     geom_hline(yintercept = 0, size = 1, color = 'dark gray') +
-    geom_point(size = 4) +
+    geom_point(size = 2) +
     geom_errorbar(aes(ymax = max_ci, ymin = min_ci), width=0.25, size = 1) +
     labs(x = NULL, y = "Standardized Rank Regression Coefficient",
          title = title) +

### 2016-03-07 utils/time_plot.R

@@ -2,7 +2,7 @@ library(dplyr)
 library(tidyr)
 library(ggplot2)
 
-time_plot <- function(){
+future_line <- function(){
   
   load("output/convergence_future.Rda")
   
@@ -26,6 +26,67 @@ time_plot <- function(){
          title = paste0("Convergence Over Trading Sequences\n2015-",  2014 + 14 * 6)) +
     ylim(0,100) +
     theme(legend.justification=c(1,0), legend.position=c(1,0), 
+          legend.key.width = grid::unit(0.05, 'npc'),
+          legend.key = element_rect(color=NA))
+}
+
+future_box <- function(){
+  load("output/convergence_future.Rda")
+  
+  plot_data <- average_convergence
+  # plot_data$set <- factor(plot_data$set, levels = gtools::mixedsort(unique(plot_data$set)))
+  #  n.edg <- round(parameters[5]*100) + 100 # integer in (100, 200)
+  plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"), 
+                          labels = c(paste0(round(8 * 0.05 + 2, 1), " edges/trader, segmentation = 0.05"), 
+                                     paste0(round(8 * 0.05 + 2, 1), " edges/trader, segmentation = 0.95"), 
+                                     paste0(round(8 * 0.95 + 2, 1), " edges/trader, segmentation = 0.05"), 
+                                     paste0(round(8 * 0.95 + 2, 1), " edges/trader, segmentation = 0.95")))
+  
+  sequence.years <- 6
+  
+  ggplot(data=plot_data, aes(x= trading_seq * sequence.years + 2014, y=(convergence + 1) * 50, color = true_mod)) +
+    geom_boxplot(aes(group = floor(trading_seq * sequence.years + 2014 - 1.25 +  2.5 * (true_mod == "SlowTSI")), 
+                     color = true_mod), position = position_dodge(width = 3), width = 2) + 
+    geom_smooth(size=1, fill=NA) +
+    scale_color_brewer(palette="Set1", name="True driver of future climate", labels = c(LogCo2 = expression(CO[2]), SlowTSI = "TSI")) +
+    facet_wrap(~set, ncol = 1) +
+    labs(x = "Year", y="",
+#          y = paste0("% Traders Believing True Model (n = ",
[…]

### 2016-03-07 paper.Rmd
small paper edits

@@
 
-Traders bet on temperature six years in the future.
+Traders bet on global temperature anomalies six years in the future.
 During the six-year period, traders buy and sell futures.
-Every year, there is Bayesian updating of traders personal predictions for future temperature based on the current year's temperature.
-At the end of each six-year period, winners collect money and traders revise beliefs about climate models, based on ideology and beliefs of top earners in their social network.
-In this section, we describe the models used to generate future temperature data, agents beliefs about those models, the market, and the social network connecting agents.
+Every year, traders update their models and forecast future temperature based on the current year's temperature.
+At the end of each six-year period, winners collect money and traders revise beliefs about climate models, based on their ideology and the beliefs of top earners in their social network.
+In this section, we describe the models used to generate future temperature data, agents beliefs about those models, the market procedures, and the social network connecting agents.
 We conclude with details on model dynamics and an overview of the model parameters.
@@
 
-IM GOING TO DELETE THESE 3:
+IM GOING TO DELETE THESE 3 bc we set them:
 * Timing parameters

### 2016-03-10 jg_plot_sa.R
Made timeseries plots for future and true past.

@@ -1,8 +1,9 @@
 library(ggplot2)
 library(dplyr)
 library(tidyr)
+library(methods)
 
-source('utilities/pc_sa.R', chdir = T)
+source('utils/pc_sa.R', chdir = T)
 
 theme_set(theme_bw(base_size = 20))
 

### 2016-03-10 master.R

@@ -14,7 +14,7 @@ if (length(args) < 2) {
 message(paste("Number of cores", cores))
 
 source("main.R")
-# devtools::install_github("JohnNay/eat", 
+#devtools::install_github("JohnNay/eat", 
 #                          auth_token = "08d34f040cbe8c95d89477741ceb450a9cfa42c4")
 library(eat)
 

### 2016-03-10 master_timeseries_experiments.R

@@ -67,12 +67,12 @@ run_experiment <- function(set, input_values,
              trueHistory = true_history),
         error = function(e) rep(NA, n.seq))
     })
-    cat(paste("\n", n.seq, "outcome evolutions with parameters", paste(as.character(j), collapse = ", "), "\n", outcome.evolution))
+    cat(paste("\n", n.seq, "outcome evolutions with LogCO2 and parameters", paste(as.character(j), collapse = ", "), "\n", outcome.evolution))
     average_convergence <- rbind(average_convergence,
                                  data.frame(convergence = outcome.evolution,
                                             trading_seq = rep(seq(n.seq), nrow(input_set)),
-                                            true_mod = rep(rep("LogCo2", length(seq(n.seq)), nrow(input_set))),
-                                            set = rep(rep(paste(as.character(j), collapse = ", "),  length(seq(n.seq))), nrow(input_set)),
+                                            true_mod = "LogCo2",
+                                            set = paste(as.character(j), collapse = ", "),
                                             stringsAsFactors = FALSE)
     )
     
@@ -97,12 +97,12 @@ run_experiment <- function(set, input_values,
              trueHistory = true_history),
         error = function(e) rep(NA, n.seq))
     })
-    cat(paste("\n", n.seq, "outcome evolutions with parameters", paste(as.character(j), collapse = ", "), "\n", outcome.evolution))
+    cat(paste("\n", n.seq, "outcome evolutions with SlowTSI and parameters", paste(as.character(j), collapse = ", "), "\n", outcome.evolution))
     average_convergence <- rbind(average_convergence,
                                  data.frame(convergence = outcome.evolution, 
                                             trading_seq = rep(seq(n.seq), nrow(input_set)),
-                                            true_mod = rep(rep("SlowTSI", length(seq(n.seq)), nrow(input_set))),
-                                            set = rep(rep(paste(as.character(j), collapse = ", "),  length(seq(n.seq))), nrow(input_set)),
+                                            true_mod = "SlowTSI",
+                                            set = paste(as.character(j), collapse = ", "),
[…]

### 2016-03-10 poster_plot_history.R

@@ -6,7 +6,7 @@ theme_set(theme_bw(base_size = 20))
 
 load("output/jg_convergence_true_past.Rda")
 
-plot_data <- average_convergence
+plot_data <- average_convergence %>% filter(true_mod == 'LogCo2')
 # plot_data$set <- factor(plot_data$set, levels = gtools::mixedsort(unique(plot_data$set)))
 #  n.edg <- round(parameters[5]*100) + 100 # integer in (100, 200)
 plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"), 

### 2016-03-10 paper.Rmd
paper added prediction market websites and citation to comparison.
Updated plot code for paper based on JGs plot code.

@@
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
-date: "March 7, 2016"
+date: "March 10, 2016"
 output: 
@@
 
+Existing prediction markets (e.g.
+hypermind.com, betfair.com, and predictit.org) focus on nearer term events (months in advance), but have been shown to have performance superior to statistical models for some event types [@pathak_comparison_2015].
+
 # Acknowledgements

### 2016-03-10 utils/time_plot.R

@@ -23,7 +23,7 @@ future_line <- function(){
     facet_wrap(~set, ncol = 1) +
     labs(x = "Year", y = paste0("% Traders Believing True Model (n = ",
                                 prettyNum(length(complete.cases(plot_data$convergence)), big.mark = ','), ")"),
-         title = paste0("Convergence Over Trading Sequences\n2015-",  2014 + 14 * 6)) +
+         title = paste0("2015-",  2014 + 14 * 6)) +
     ylim(0,100) +
     theme(legend.justification=c(1,0), legend.position=c(1,0), 
           legend.key.width = grid::unit(0.05, 'npc'),
@@ -50,9 +50,8 @@ future_box <- function(){
     geom_smooth(size=1, fill=NA) +
     scale_color_brewer(palette="Set1", name="True driver of future climate", labels = c(LogCo2 = expression(CO[2]), SlowTSI = "TSI")) +
     facet_wrap(~set, ncol = 1) +
-    labs(x = "Year", y="",
-#          y = paste0("% Traders Believing True Model (n = ",
-#                                 prettyNum(length(complete.cases(plot_data$convergence)), big.mark = ','), ")"),
+    labs(x = "Year", y = paste0("% Traders Believing True Model (n = ",
+                                prettyNum(length(complete.cases(plot_data$convergence)), big.mark = ','), ")"),
          title = paste0("2015-",  2014 + 14 * 6)) +
     ylim(0,100) +
     theme(legend.justification=c(1,0), legend.position=c(1,0),
@@ -63,7 +62,7 @@ future_box <- function(){
 history_box <- function(){
   load("output/jg_convergence_true_past.Rda")
   
-  plot_data <- average_convergence
+  plot_data <- average_convergence %>% filter(true_mod == 'LogCo2')
   # plot_data$set <- factor(plot_data$set, levels = gtools::mixedsort(unique(plot_data$set)))
   #  n.edg <- round(parameters[5]*100) + 100 # integer in (100, 200)
   plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"), 
@@ -82,8 +81,8 @@ history_box <- function(){
     scale_color_brewer(palette="Set1", labels = c(LogCo2 = expression(CO[2]), SlowTSI = "TSI"), guide = "none") +
     facet_wrap(~set, ncol = 1) +
     labs(x = "Year",  
-         y = paste0("% Traders Believing True Model (n = ",
[…]

### 2016-03-11 paper.Rmd
Paper work

@@
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
-date: "March 10, 2016"
+date: "March 11, 2016"
 output: 
@@
 
-We present a simulation of a climate prediction market.
+We present a simulation of a climate prediction market as a computational test-bed for market and policy design.
 Traders adapt their beliefs about future global temperatures based on the monetary performance of their neighboring traders in a social network.
-We conduct experiments with the model to determine how a variety of market factors and the climate data-generating process affect whether climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
+We conduct experiments with the model to determine how a variety of market factors and the climate data-generating process affect the extent to which climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
+We simulate two alternative climate futures: one where the natural log of CO2 is the primary driver of global temperature and one where sun spots are the primary driver.
+Market participation causes traders to converge toward believing the "true" climate model; however, when the true model is CO2-driven there is more convergence.
 
@@
 As a validation that the market is operating correctly, we run the market with actual temperatures 1880 to 2015 and market betting from 1931 to 2014.
-As warming signal begins to show up in 1980s, traders begin to converge toward log(CO2) model `r figRef("time")`.
+As warming signal begins to show up in 1980s, traders begin to converge toward the log(CO2) model (`r figRef("time")`).
 
@@
 
-Our sensitivity analysis, `r figRef("sa")`, shows that ideology, risk tolerance, the number of traders, and the segmentation of the social network significantly affect convergence.
+Our sensitivity analysis, `r figRef("sa")`, shows that ideology, risk tolerance, the number of traders, the number of edges per trader, the segmentation of the social network, and the true model all significantly affect convergence.
 Segmented social networks can create an "echo-chamber" effect.
+Market participation causes traders to converge toward believing the "true" climate model; however, when the true model is CO2-driven there is more convergence due to less noise in the relationship between predictor variable (log of CO2 versus TSI) and temperature.
 
@@
 ```
+
+The sensitivity analysis averages over time (and model replicates) and thus masks time trends.
+We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and set conducted an experiment crossing 95% and 5% percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model.
+We then collected the time series of convergence across these eight designs (`r figRef("time")`).
 
@@
 
-Our code for the model is available at [johnjnay.com/predictMarket](http://johnjnay.com/predictMarket).
+Our code for the model is available at [johnjnay.com/predictMarket](http://johnjnay.com/predictMarket). This project is a computational test-bed for public policy design.
 

### 2016-03-12 paper.Rmd
small paper work

@@
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
-date: "March 11, 2016"
+date: "March 12, 2016"
 output: 
@@
 
-We present a simulation of a climate prediction market as a computational test-bed for market and policy design.
+We present a simulation model as a computational test-bed for a climate prediction market.
 Traders adapt their beliefs about future global temperatures based on the monetary performance of their neighboring traders in a social network.
-We conduct experiments with the model to determine how a variety of market factors and the climate data-generating process affect the extent to which climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
-We simulate two alternative climate futures: one where the natural log of CO2 is the primary driver of global temperature and one where sun spots are the primary driver.
-Market participation causes traders to converge toward believing the "true" climate model; however, when the true model is CO2-driven there is more convergence.
+We conduct experiments with the model to determine how a variety of market factors and the climate data-generating process may affect the extent to which climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
+We simulate two alternative climate futures: one where the natural log of CO$_2$ is the primary driver of global temperature and one where sun spots are the primary driver.
+These represent the two most plausible competing views in the public discourse.
+Market participation causes traders to converge toward believing the "true" climate model.
 
@@
 
-The climate change debate has become strongly polarized over the past twenty years.
-Although the scientific consensus on the anthropogenic nature of climate change strongly increased, beliefs about anthropogenic climate change did not evolve much within the public [@Vandenbergh2013f].
+The climate change debate has become strongly polarized over the past two decades.
+Although the scientific consensus on the anthropogenic nature of climate change strongly increased, beliefs about climate change did not evolve much within the public [@Vandenbergh2013f].
 In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily as the question is becoming increasingly politicized and potentially disconnected from scientific evidence.
@@
 We simulate a prediction market where traders exchange securities related to climate outcomes to explore whether, and under what market and climate conditions, prediction markets may be useful for increasing convergence of climate beliefs.
+The ultimate goal is to initiate a computational design process for effective climate prediction markets.
 
@@
 
-Temperature proportional to log(CO2) or Total Solar Irradiance (conventional science vs. popular alternative among doubters). Add AR(1) noise (best ARMA noise model for historical record).
+Temperature proportional to log(CO$_2$) or Total Solar Irradiance (conventional science vs. popular alternative among doubters). Add AR(1) noise (best ARMA noise model for historical record).
 
@@
 
-Traders use one of the two models (temperature depends on CO2 or TSI) to forecast future temperature anomalies.
+Traders use one of the two models (temperature depends on CO$_2$ or TSI) to forecast future temperature anomalies.
 These models are interpreted as the trader's beliefs about the true climate process, the driving factor of long-term global temperatures.
@@
 As a validation that the market is operating correctly, we run the market with actual temperatures 1880 to 2015 and market betting from 1931 to 2014.
-As warming signal begins to show up in 1980s, traders begin to converge toward the log(CO2) model (`r figRef("time")`).
+As warming signal begins to show up in 1980s, traders begin to converge toward the log(CO$_2$) model (`r figRef("time")`).
 
@@
 Segmented social networks can create an "echo-chamber" effect.
-Market participation causes traders to converge toward believing the "true" climate model; however, when the true model is CO2-driven there is more convergence due to less noise in the relationship between predictor variable (log of CO2 versus TSI) and temperature.
+Market participation causes traders to converge toward believing the "true" climate model; however, when the true model is CO$_2$-driven there is more convergence due to less noise in the relationship between predictor variable (log of CO$_2$ versus TSI) and temperature.
 

### 2016-03-17 paper.Rmd
Work on model design

@@
 ---
-title: "Put Your Money Where Your Mouth Is: Prediction Markets and Climate Change"
+title: 'Put Your Money Where Your Mouth Is: Prediction Markets and Climate Change'
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
@@
   pdf_document:
-    keep_tex: true
+    keep_tex: yes
+  html_document: default
 bibliography: library.bib
@@
 During the six-year period, traders buy and sell futures. 
-Every year, traders update their models and forecast future temperature based on the current year's temperature.
-At the end of each six-year period, winners collect money and traders revise beliefs about climate models, based on their ideology and the beliefs of top earners in their social network.
+Every year, traders update their models and forecast of future temperatures based on newly available data. 
+At the end of each six-year period, winners collect gains and traders revise beliefs about climate models, based on their ideology and the beliefs of top earners in their social network. 
 In this section, we describe the models used to generate future temperature data, agents beliefs about those models, the market procedures, and the social network connecting agents. 
@@
 
-Traders use one of the two models (temperature depends on CO$_2$ or TSI) to forecast future temperature anomalies.
+Traders use one of the two models (temperature depends on CO$_2$ or TSI) to forecast future temperature. 
 These models are interpreted as the trader's beliefs about the true climate process, the driving factor of long-term global temperatures. 
@@
 Thus, both when **true.model** = 0 and **true.model** = 1, approximately half of the traders use the true data-generating model to make predictions. 
-However, traders using the true model do not necessarily make perfectly accurate predictions.
-Although these traders believe in the correct *functional form* of the model, they still need to calibrate it based on limited noisy data.
-Therefore, the values they use as the parameters of the model will typically still be different from the exact parameters of the data-generating process.
+Traders using the true model do not necessarily make perfectly accurate predictions however. 
+Although these traders believe in the correct *functional form* of the model, they still need to calibrate their model based on limited noisy data. 
+Therefore, the values these traders assign to the parameters of the model will typically be different from the exact parameters in the data-generating process.
 
@@
 
-Traders are initially endowed with a single unit Experimental Currency Unit (ECU).
-Traders use their model to forecast the distribution of future temperatures and determine their reservation price for different securities.
-Securities pay 1 ECU  if the actual temperature at some time $t^*$ falls between a certain range.
+Traders are initially endowed with a single experimental currency unit (ECU). 
+Traders use their model to forecast the distribution of future temperature and determine their reservation price for different securities. 
+Each security pays 1 ECU at the end of the trading sequence if the temperature at the end of the sequence falls into a certain range. 
 Traders are risk-neutral expected utility maximizers. 
-Therefore, their reservation price for a security $\tilde{s}$ is simply their assessment of the probability that the temperature at time $t^*$ falls in the range covered by $\tilde{s}$.
+Therefore, their reservation price for a security is simply their assessment of the probability that the temperature will fall in the range covered by the security at the end of the sequence.
 
-Based on their reservation price, traders behave as "zero-intelligence" agents [@Gode1993], that is, they sell at a random price above their reservation, and buy at a random price below their reservation.
-The distribution from which buy and sell prices are drawn is determined by the traders risk attitude, characterized by parameter $risk.tak_i$.
-These trading strategies are simple but have been shown to provide accurate approximations of behavior in prediction markets [@Klingert2012c], and in financial markets more broadly.
+Based on their reservation price, traders behave as "zero-intelligence" agents [@Gode1993].
+They attempt to sell securities at a random price above their reservation, and to buy securities at a random price below their reservation.
+These trading strategies are simple but provide accurate approximations of behavior in prediction markets [@Klingert2012c], and in financial markets more broadly.
 
-Our model features a particular version of CDA through which traders exchange securities.
-Continuous-double auctions (CDAs) or some variants thereof are common market mechanisms, i.e.
-procedures to match buy and sell orders.
+Based on traders' sell and buy orders, traders exchange securities following a continuous-double auctions (CDA) procedure (see Model Dynamics below for more details). 
+CDAs or some variants thereof are common procedures to match buy and sell orders. 
 CDA are notably used on large stock markets [@Tseng2010a].
@@
 This ensures that each agent is connected to at least two other agents. 
-Every time a security is realized, each trader $i$ looks at the performance of their neighbors in the network.
-If one of $i$'s neighbors, say $j$, is richer than $i$, $i$ interprets this has an indication that $j$ has a better approximate model.
-Then $i$ considers adopting $j$'s approximate model.
-Traders start with the same initial amount of money, so differences in money among traders can only come from  traders' interactions through the market.
-For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is characterized by parameter $ideo_i$.
+Every time securities are realized, each trader looks at the performance of her richest neighbor in the network. 
+Traders start with the same initial amount of ECU and differences in ECU can only come from market interactions.
+Therefore, if some trader is poorer that her richest neighbor, the trader interprets this has an signal that her richest neighbor may have a better model of the climate. 
+Then, the trader considers adopting the model of her richest neighbor. 
+For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is.
 
@@
 An example of such social network is depicted in Figure X. 
-We assume that the initial network is segregated, that is, traders who believe in model 2 (1) are more likely to be linked with other traders who believe in 2 (1).
-The degree of segregation is determined by parameter **seg**.
-Although traders can change their approximate model over time, the connections between traders do not change as the market unfolds -- the edges are fixed.
+We assume that the initial network is segregated.
+Traders who initially believe in model 1 (resp. 2) are more likely to be linked with other traders who also believe in 1 (resp. 2). 
+Although traders can change their approximate model over time, the connections between traders do not change as the market unfolds (i.e the edges are fixed).
 
@@
 
-The time periods, $t$, are grouped in trading *sequences*.
+The time periods $t$ are grouped in trading *sequences*. 
 In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence. 
@@
 In this case, a security traded in the third sequence pays 1 ECU if the temperature at $t = 1970$ falls into the range of temperatures covered by the security.
-At each time $t$, traders are assumed to know the past value of the temperature $\hat{T}_{0:t}$ and greenhouse gas emissions  $GHG_{0:t}$.
-In a sequence finishing at time $t^*$,  traders also have common knowledge of $GHG_{t:t^*}$, the future values of greenhouse gas emissions up to $t^*$.
+
+At each time $t$, traders are assumed to know the past value of the temperature ${T}_{0:t}$, carbon dioxide $CO2_{0:t}$, and total solar irradiance $TSI_{0:t}$. 
+In a sequence finishing at time $t^*$,  traders also have common knowledge of $CO2_{t:t^*}$ and $TSI_{t:t^*}$, the future values of carbon dioxide and total solar irradiance up to $t^*$.
 However, at any $t$, traders do not know the value of any future temperatures. 
-In particular, in a given sequence, traders do not know the value of $\hat{T}_{t^*}$.
-Traders can only predict $\hat{T}_{t^*}$ using their approximate model and their knowledge of $\hat{T}_{0:t}$ and  $GHG_{0:t^*}$.
-Notice that because $GHG_{0:t^*}$ is common knowledge, in each period $t$ every trader $i$ with the same approximate model forms the same pseudo-expectation $\mu_{ti} \equiv E_t( \hat{T}_{t^*} ~|~ \hat{T}_{0:t}, GHG_{0:t^*})$, $i$'s approximate model at $t$.
+In particular, in a given sequence, traders do not know the value of ${T}_{t^*}$. 
+Traders can only predict ${T}_{t^*}$ using their approximate model and their knowledge of ${T}_{0:t}$, $CO2_{0:t^*}$, and $TSI_{0:t^*}$.
+Notice that because ${T}_{0:t}$, $CO2_{0:t^*}$, and $TSI_{0:t^*}$ are common knowledge, in each period $t$, any two trader with the same approximate model $m$ forms the same stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, CO2_{0:t^*}, TSI_{0:t^*})$.
 
-We assume that the believed probability distribution of $\hat{T}_{t^*}$ for trader $i$  at any time $t$ is $\mu_{ti} + \eta_t$, where the distribution $\eta_t$ is obtained by trader $i$ by bootstrapping the residuals from the regression of her believed true model over the past periods.
-This explains the use of the term pseudo-expectation.
-Because $\eta_t$ is not bound to have expected value zero, a traders' effective expected value of $\hat{T}_{t^*}$ may be different from $\mu_{ti}$.
+ At each time $t$, traders :
 
- At each time $t$, traders follow this process:
+1. recalibrate their approximate model based on the new set of temperature data available at $t$,
+2. compute their stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, CO2_{0:t^*}, TSI_{0:t^*})$ and use it to determine the expected value they attached to each security, and
+3. trade on the CDA market based on their zero-intelligence decision rule as follows: 
+ 	+ Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
+ 	+ Every trader $i$ also chooses at random a security $s_i^S$ she will try to sell among the securities she owns a positive amount of (if any).
+	+ Traders then decide of their selling price $p_i^S$ and buying price $p_i^B$. 
+  To do so, traders first compute their expected values $E(s_i^B)$ and $E(s_i^S)$ for securities $s_i^B$ and $s_i^S$. 
+  Then traders set $p_i^S$ at random above $E(s_i^S)$ and $p_i^B$ at random below $E(s_i^B)$ (see Model Parameters below for more details).
 
-1. recalibrate their approximate model based on the new set of past data available at $t$,
-2. compute their belief-density for $\hat{T}_{t^*}$ and use it to determine the expected value they attached to each security,
-3. and trade on the CDA market based on their zero-intelligence decision rule as follows. 
- 	+ Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
- 	+ Every trader $i$ also chooses at random among the securities she owns a positive amount of (if any) a security $s_i^S$ she will try to sell.
-	+ Traders then decide of their selling $p_i^S$ and buying price $p_i^S$.
-To do so, they compute the expected value of the securities $s_i^B$ and $s_i^S$.
-Then they set a selling and a buying price following the zero-intelligence rule.
-	+ Traders go to the market one at the time, in an order randomly selected each $t$.
- 	+ When $i$ comes to the market, she places limit orders in the order book.
+	+ Traders go to the market one at the time, in an order drawn randomly for each $t$.
+ 	+ When trader $i$ comes to the market, she places limit orders in the order book. 
    These orders specify that $i$ is willing to buy $s_i^B$ at any price below $p_i^B$, and to sell $s_i^S$ at any price above $p_i^S$.
- 	+ The market maker then tries to match $i$'s orders with some order which was put in the order book before $i$ came to the market.
- 	+ If there are  outstanding sell offers for $s_i^B$ at price $p$ lower than $p_i^B$, then a trade is concluded.
-Trader $i$ buys one unit from the sellers who sells at the highest price below $p_i^B$, and the sell and buy offers are removed from the order book.
- 	+ If there are  outstanding buy offers for $s_i^S$ at price $p$ higher than $p_i^S$, then a trade is concluded.
+ 	+ The market maker then tries to match $i$'s orders with some order which was put in the book before $i$ came to the market.
+ 	+ If there are  outstanding sell offers for $s_i^B$ at a price below $p_i^B$, a trade is concluded. 
+   Trader $i$ buys one unit of $s_i^B$ from the sellers who sells at the highest price below $p_i^B$, and the sell and buy offers are removed from the order book.
+ 	+ If there are outstanding buy offers for $s_i^S$ at price above $p_i^S$, a trade is concluded. 
    Trader $i$ sells one unit to the buyer who buys at the highest price above $p_i^S$, and the sell and buy offers are removed from the order book.
@@
 
-At $t^*$, when the sequence ends, there is only one security $s^*$ associated with a range of temperatures including the actual $\hat{T}_{t^*}$. Thus:
+At $t^*$, when the sequence ends, there is only one security $s^*$ associated with a range of temperatures including the actual temperature $T_{t^*}$. 
+At $t^*$, traders:
 
-1. traders receive 1 ECU per unit of $s^*$ they own, and
+1.  receive 1 ECU per unit of $s^*$ they own, and
 2.  consider adapting their neighbors' approximate model as described above.
@@
 
- * Market structure parameters:
-    + **market.complet** the number of securities which can be traded.
-With more securities, the interval of temperatures corresponding to each security is smaller.
-Traders can then trade on more precise temperature intervals.
-Higher values of **market.complet** may, however, reduce the number of exchanges.
-Because traders pick the securities they buy and sell at random, the probability that a match between sellers and buyers is found is lower for higher values of **market.complet**.
+* Behavioral parameters.
+    + **risk.tak** : determines the distribution of risk-taking behavior. 
+    For each trader $i$, the level of \texttt{risk.tak}$_i$ is drawn uniformly at random from $[0,$\texttt{risk.tak} $]$.
+    The higher \texttt{risk.tak}$_i$, the higher (resp. lower) the price $i$ will set for the securities $i$ tries to sell (resp. buy).
+    Formally, let $\texttt{reserv}_{it}$ be the reservation price of trader $i$ at time $t$ for some security $s$.
+    Then trader $i$ picks her buying (resp.
+selling) price for $s$ uniformly at random in the interval  $[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it}  * (1 - \texttt{risk.tak}_i)]$ (resp.
+$[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it} * (1  + \texttt{risk.tak}_i)]$).
 
-* Behavioral parameters.
-    + **risk.tak** Determines the distribution of risk-taking behavior.
-The higher risk.tak, the more traders	will try to buy (resp.
-sell) lower (resp.
-higher) than their reservation price.
-Formally, at time $t$, trader $i$ picks her buying (resp.
-selling) prices randomly in the interval  $[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it}  * (1 - \texttt{risk.tak}_i)]$ (resp.
-$[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it} * (1  + \texttt{risk.tak}_i)]$), where $\texttt{reserv}_{it}$ is $i$'s reservation price at time $t$ for the security $i$ picked to buy (resp.
-sell).
-For each trader $i$, the level of \texttt{risk.tak}$_i$ is drawn uniformly at random from $[0,$\texttt{risk.tak} $]$
-    + **ideo**  Determines the degree of "ideology" of traders.
-If ideo is high, traders will not revise their approximate models easily, even when faced with strong evidence that their neighbors are doing better than them.
-For each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$.
+    + **ideo** :  determines the degree of "ideology" of traders.
+    For each trader $i$, the level of \texttt{ideo}$_i$ is drawn uniformly at random from $[0,$\texttt{ideo} $]$.
+    If ideo is high, traders will not revise their approximate models easily, even when faced with evidence that their richest neighbor is doing better than them. 
+    Formally, for each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$. 
     The value of $d_{i}$ is the probability that $i$ adopts one of her neighbors' approximate model if this neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
-
-IM GOING TO DELETE THESE 3 bc we set them:
-* Timing parameters
-    + **burn.in** The number of burn-in periods in which no securities are traded (necessary to allow believed models to be estimated in the first period).
-    + **n.seq** Number of trading sequences.
-    + **horizon** Number of trading periods in each trading sequence.
 
@@
 
-```{r time, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("time", "Convergence over trading sequences."), fig.width=10.5, fig.height=8.5}
+```{r time, eval=FALSE, echo=FALSE, message=FALSE, htmlcap=figRef("time", "Convergence over trading sequences."), fig.width=10.5, fig.height=8.5}
 source("utils/time_plot.R")

### 2016-03-17 paper.Rmd
paper: abstract and other edits throughout.

@@
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
-date: "March 12, 2016"
+date: "March 17, 2016"
 output:
@@
 Traders adapt their beliefs about future global temperatures based on the monetary performance of their neighboring traders in a social network.
-We conduct experiments with the model to determine how a variety of market factors and the climate data-generating process may affect the extent to which climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
+We conduct experiments with the model to determine how a variety of market factors and the temperature data-generating process may affect the extent to which climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
 We simulate two alternative climate futures: one where the natural log of CO$_2$ is the primary driver of global temperature and one where sun spots are the primary driver.
 These represent the two most plausible competing views in the public discourse.
-Market participation causes traders to converge toward believing the "true" climate model.
+Market participation causes traders to converge toward believing the "true" climate model under a variety of model parameterizations in a relatively small number of years, suggesting that a climate prediction market could be useful regardless of which climate change beliefs ultimately turn out to be the best model of temperature change.
 
@@
 In this section, we describe the models used to generate future temperature data, agents beliefs about those models, the market procedures, and the social network connecting agents. 
-We conclude with details on model dynamics and an overview of the model parameters.
+We conclude with details on model dynamics and an overview of the model parameters we experimentally vary.
 
@@
 Traders who initially believe in model 1 (resp. 2) are more likely to be linked with other traders who also believe in 1 (resp. 2). 
-Although traders can change their approximate model over time, the connections between traders do not change as the market unfolds (i.e the edges are fixed).
+This reflects reality where cultural and social groups are correlated with climate change beliefs. 
+Although traders can change their approximate model over time, the connections between traders do not change as the market unfolds, i.e the edges are fixed.
 
@@
 
- At each time $t$, traders :
+ At each time $t$, traders:
 
@@
 The model depends on the following parameters, which we vary in simulation experiments to determine their effects on the convergence of beliefs.
+We group the parameters into network and individual behavioral factors.
 
@@
 As a validation that the market is operating correctly, we run the market with actual temperatures 1880 to 2015 and market betting from 1931 to 2014.
-As warming signal begins to show up in 1980s, traders begin to converge toward the log(CO$_2$) model (`r figRef("time")`).
+As a warming signal begins to show up in 1980s, traders begin to converge toward the log(CO$_2$) model (`r figRef("time")`).
 
@@
 
+Our primary focus is on simulating the market in future scenarios.
+This is more relevant for policy design and more interesting theoretically due to the increasing divergence of global temperature values under the two future scenarios.
 The sensitivity analysis is based on a Latin Hypercube Sampling of parameter sets from the following distributions [@beachkofski_improved_2002; @carnell_lhs_2012].
@@
 
-We use the model to simulate 10,000 outcomes based on the input paramter sets and then conduct a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, resulting simulated vector, $y$ [@Marino2008; @pujol_sensitivity:_2014; @saltelli_sensitivity_2009].
+We use the model to simulate 10,000 outcomes based on the input paramter sets and then conduct a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of belief convergence scores, $y$ [@Marino2008; @pujol_sensitivity:_2014; @saltelli_sensitivity_2009].
 Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).
-The only difference with the partial correlation and partial rank correlation (which we use here), is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships.
+The only difference with the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships.
 That is every value, $y_s$, in the data set is replace by a number $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, where $y_r$ are the other observed values of $y$ in the data set.
 
-Our sensitivity analysis, `r figRef("sa")`, shows that ideology, risk tolerance, the number of traders, the number of edges per trader, the segmentation of the social network, and the true model all significantly affect convergence.
+Our sensitivity analysis, `r figRef("sa")`, shows that ideology, risk tolerance, the number of traders, the number of edges per trader, the segmentation of the social network, and the true model all statistically significantly affect convergence.
 Segmented social networks can create an "echo-chamber" effect.
@@
 
-```{r time, eval=FALSE, echo=FALSE, message=FALSE, htmlcap=figRef("time", "Convergence over trading sequences."), fig.width=10.5, fig.height=8.5}
+```{r time, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("time", "Convergence over trading sequences."), fig.width=10.5, fig.height=8.5}
 source("utils/time_plot.R")

### 2016-03-19 paper.Rmd
small paper work

@@
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
-date: "March 17, 2016"
+date: "March 19, 2016"
 output:
@@
 The costs of misinformed climate policies are high.
-If anthropogenic climate change is a myth but is taken to be true, public resources will be spent on unnecessary efforts.
+If climate change is not human-induced but is believed to be so, public resources will be spent on unnecessary efforts.
 On the other hand, if climate change is human-induced but not recognized as so, the costs of inaction could be devastating.
@@
 Attempts to foster such consensus face many social and psychological challenges, some of which could be adressed by creating climate prediction markets where participants can "put their money where their mouths are" [@Hsu2011; @Vandenbergh2013f].
-The idea of using prediction markets to efficiently aggregate information about uncertain events has been widely discussed [@Horn2014c].
-Prediction markets have been shown to have interesting theoretical properties [@Set1998; @Hanson2012] and to perform well in terms of prediction accuracy and information aggregation in experiments [@Hanson2006; @Healy2010], simulation models [@Klingert2012c; @Jumadinova2011a], and the real-world [@Wolfers2006].
+The idea of using prediction markets to efficiently aggregate information about uncertain event outcomes has been widely discussed [@Horn2014c].
+Prediction markets have interesting theoretical properties [@Set1998; @Hanson2012], and perform well in terms of prediction accuracy and information aggregation in experiments [@Hanson2006; @Healy2010] simulation models [@Klingert2012c; @Jumadinova2011a] and the real-world [@Wolfers2006].
 We simulate a prediction market where traders exchange securities related to climate outcomes to explore whether, and under what market and climate conditions, prediction markets may be useful for increasing convergence of climate beliefs.
-The ultimate goal is to initiate a computational design process for effective climate prediction markets.
+Our work can be extended as part of a computational design process for effective climate prediction markets.
 
@@
 
-@Tseng2010a create an agent-based model (ABM) of a continuous double auction market with various types of market strategies, including two variants of the zero-intelligence agent.
+@Tseng2010a create an agent-based model (ABM) of a continuous double auction market with multiple market strategies, including two variants of the zero-intelligence agent.
 They compare the behavior of their simulation with data from a prediction market for the outcome of political elections.
@@
 
-@Jumadinova2011a conducted research more similar to ours, as they explicitly include information about uncertain events in their model.
-@Jumadinova2011a study a continuous double auction model in which agents update their beliefs about the uncertain events based on information on the event.
+@Jumadinova2011a conducted research similar to ours, as they explicitly include information about uncertain events in their model.
+@Jumadinova2011a study a continuous double auction model in which agents update their beliefs about uncertain events based on information on the event.
 New beliefs are modeled as a weighted sum of the last period's beliefs and market prices.
@@
 @Ontanon2009 study the effect of deliberation on a simulated prediction market.
-They assume that agents use Case-Based Reasoning [@Aamodt1994] to reason and debate about uncertain outcomes with their neighbors in a social network.
+They assume that agents use Case-Based Reasoning [@Aamodt1994] to debate uncertain outcomes with their neighbors in a social network.
 Although the model could in theory be used to assess convergence of predictive models by looking at the arguments agents adopt after the deliberation, this is not done by the authors.
@@
 From a public policy perspective, changing the explanatory models of market participants is one of the most important roles that prediction markets might play.
-This is perhaps the most important social benefit of prediction markets given the comparable predictive power of statistical supervised learning models, which can provide information at a much lower cost than creating and maintaining a prediction market [@goel_prediction_2010].
+This is perhaps the most important social benefit of prediction markets given the comparable predictive power of statistical supervised learning models, which can provide information at a much lower cost than creating and maintaining a market [@goel_prediction_2010].
 Effective climate policy not only requires an accurate consensus on future climate *outcomes*.
@@
 We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and set conducted an experiment crossing 95% and 5% percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model.
-We then collected the time series of convergence across these eight designs (`r figRef("time")`).
+We then collected the time series of convergence across these eight designs for visualization of the distributions of convergence over time (`r figRef("time")`).
 
@@
 
-Our code for the model is available at [johnjnay.com/predictMarket](http://johnjnay.com/predictMarket). This project is a computational test-bed for public policy design.
+All code for the model is available at [johnjnay.com/predictMarket](http://johnjnay.com/predictMarket). This project is a computational test-bed for public policy design.
 

### 2016-03-20 paper.Rmd
Description of time-series and climate models.

@@
 ---
-title: 'Put Your Money Where Your Mouth Is: Prediction Markets and Climate Change'
+title: 'Betting and Belief: Prediction Markets and Attribution of Climate Change'
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
@@
     keep_tex: yes
+    includes:
+      in_header: header.tex
   html_document: default
@@
 Prediction markets have interesting theoretical properties [@Set1998; @Hanson2012], and perform well in terms of prediction accuracy and information aggregation in experiments [@Hanson2006; @Healy2010] simulation models [@Klingert2012c; @Jumadinova2011a] and the real-world [@Wolfers2006].
+Prediction markets have been shown to be effective at identifying which scientific publications are likely to be successfully replicated [@Dreber_2015].
+Bloch, Annan and Bowles [-@Bloch_2010] have proposed using derivatives markets to reduce the scientific uncertainty in estimating the impact of future climate change.
 We simulate a prediction market where traders exchange securities related to climate outcomes to explore whether, and under what market and climate conditions, prediction markets may be useful for increasing convergence of climate beliefs.
@@
 
-We start with the historical temperature record and project future climates under two alternative theories:
+For climate time-series, we use the annual anomaly of global mean temperature.
+For years from 1880--2014 we use the GISTEMP global mean land-sea annual temperature anomalies [@GISTEMP_2016; @Hansen_2010] and for years from 2015 onward, we project future climates under two alternative theories: In both theories, changes in global temperature are proportional to changes in a deterministic forcing plus a stochastic noise term.
+For simplicity, we choose two alternative expressions for the deterministic climate forcing: one, which corresponds to conventional climate science, takes the natural logarithm of the atmospheric carbon dioxide concentration in parts per million [@Archer_2012, p.
+37], and the other, which corresponds to an alternative theory advocated by many who doubt or reject conventional climate science, takes the total solar irradiance (the brightness of sunlight, in Watts per square meter, at the top of the atmosphere) averaged over the 11-year sunspot cycle [@Soon_2005].
+Actual scientific models of the earth's climate include many forcings, including carbon dioxide, other greenhouse gases, aerosols, total solar irradiance, and more.
+In these models, the forcing due to changing $\ce{CO2}$ concentration is larger than any of the other forcings by a wide margin [@IPCC_WG1_AR5_2013, p.
+14], so for simplicity we take this as the model.
+Choosing only one forcing term for each of the competing models simplifies comparison because each has the same number of adjustable parameters.
 
-Temperature proportional to log(CO$_2$) or Total Solar Irradiance (conventional science vs. popular alternative among doubters). Add AR(1) noise (best ARMA noise model for historical record).
+For $\ce{CO2}$ we used historical emissions through 2005, harmonized with RCP~8.5 representative concentration pathway from 2005 onward [@rcp_database_2009; @Riahi_2011] and for TSI we used the harmonized historical values, with a projection through 2100 from Velasco-Herrera *et al.* [-@Velasco_Herrera_2015].
 
-JG - describe the temperature time series models here. Add the plots of the time series diverging thats on middle of poster. Add citations for the models to library.bib.
+Warming coefficients for each model were determined by linear regression of historical temperatures from 1880--2014 against the historical values for each model's forcing term.
+The noise model was determined by fitting an $\text{ARMA}(p,q)$ model to the residuals from the regression for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal model [Vehtari_2015].
+The optimal model in both cases was AR(1).
+
+Future climates were generated by multiplying the warming coefficients for each model for warming coefficients adding stochastic AR(1) noise to deterministic warming using parameters fit to the historical data (1880--2014).
+One realization of future climates for both the $\ln(\ce{CO2})$ and TSI models is shown in Fig.~\ref{fig:climate_time_series}
+
+![fig:climate_time_series](timeseries/jg_climate.pdf)
 

### 2016-03-21 paper.Rmd
changed the climate time series figure to be created in the rmd file
is shown in `r figRef("climateplot")`.

```{r climateplot, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("climateplot", "Temperature futures in two alternate realities."), fig.width=5, fig.height=5}
source("utils/climate_plot.R")
climate_plot() + theme(text=element_text(family='Palatino'))
```

@@
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
-date: "March 19, 2016"
+date: "March 21, 2016"
 output:
@@
 We simulate two alternative climate futures: one where the natural log of CO$_2$ is the primary driver of global temperature and one where sun spots are the primary driver.
-These represent the two most plausible competing views in the public discourse.
+These represent the two most plausible competing views in the public discourse and our analysis is agnostic about which is "true".
 Market participation causes traders to converge toward believing the "true" climate model under a variety of model parameterizations in a relatively small number of years, suggesting that a climate prediction market could be useful regardless of which climate change beliefs ultimately turn out to be the best model of temperature change.
@@
 Future climates were generated by multiplying the warming coefficients for each model for warming coefficients adding stochastic AR(1) noise to deterministic warming using parameters fit to the historical data (1880--2014).
-One realization of future climates for both the $\ln(\ce{CO2})$ and TSI models is shown in Fig.~\ref{fig:climate_time_series}
+One realization of future climates for both the $\ln(\ce{CO2})$ and TSI models is shown in `r figRef("climateplot")`.
 
-![fig:climate_time_series](timeseries/jg_climate.pdf)
+```{r climateplot, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("climateplot", "Temperature futures in two alternate realities."), fig.width=5, fig.height=5}
+source("utils/climate_plot.R")
+climate_plot() + theme(text=element_text(family='Palatino'))
+```
 
@@
 
-The sensitivity analysis averages over time (and model replicates) and thus masks time trends.
+The sensitivity analysis averages over time and thus masks time trends.
 We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and set conducted an experiment crossing 95% and 5% percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model.

### 2016-03-21 timeseries/climate_plot.R

@@ -5,7 +5,7 @@ library(tidyr)
 source('prepare_data.R')
 source('climate_model.R')
 
-TRACE_CLIMATE_MODEL <- TRUE
+TRACE_CLIMATE_MODEL <- FALSE
 PARALLEL_STAN <- TRUE
 WHICH_MODEL <- 'ar1'
 

### 2016-03-21 utils/climate_plot.R

@@ -1,0 +1,23 @@
+climate_plot <- function(){
+  library(ggplot2)
+  library(dplyr)
+  library(tidyr)
+  # to generate this data: source("timeseries/climate_plot.R", chdir = TRUE)
+  load("timeseries/covar_models.Rda")
+  t <- mdl.co2@climate %>% select(year, t.anom) %>% mutate(covar = "Historical")
+  change.year <- max(t$year)
+  co2 <- mdl.co2@future %>% select(year, t.anom) %>% filter(year >= change.year) %>%
+    mutate(covar = "CO2")
+  tsi <- mdl.tsi@future %>% select(year, t.anom) %>% filter(year >= change.year) %>%
+    mutate(covar = "TSI")
+  t %>% bind_rows(co2, tsi) %>% 
+    mutate(covar = ordered(covar, levels = c("CO2", "TSI", "Historical"))) %>%
+    ggplot(aes(x = year, y = t.anom, color = covar)) + 
+    geom_line(size=1) + 
+    scale_color_brewer(palette = "Set1", labels = c(CO2 = expression(CO[2]), TSI = "TSI"), 
+                       name = "True driver of future climate") + 
+    labs(x = "Year", y = expression(paste("Temperature Anomaly  ", (degree * C)))) +
+    theme(legend.position = c(0.15,0.95), legend.justification = c(0,1), 
+          legend.key.width = grid::unit(0.05,'npc'), 
+          legend.key = element_rect(color = NA))
+}

### 2016-03-22 paper.Rmd
A few minor edits to the time series section and the bibliography to make the citations format correctly.

@@
 Actual scientific models of the earth's climate include many forcings, including carbon dioxide, other greenhouse gases, aerosols, total solar irradiance, and more.
-In these models, the forcing due to changing $\ce{CO2}$ concentration is larger than any of the other forcings by a wide margin [@IPCC_WG1_AR5_2013, p.
+In these models, the changing $\ce{CO2}$ concentration is, by a large margin, the strongest forcing [@IPCC_WG1_AR5_2013, p.
 14], so for simplicity we take this as the model.
@@
 Warming coefficients for each model were determined by linear regression of historical temperatures from 1880--2014 against the historical values for each model's forcing term.
-The noise model was determined by fitting an $\text{ARMA}(p,q)$ model to the residuals from the regression for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal model [Vehtari_2015].
+The noise model was determined by fitting an $\text{ARMA}(p,q)$ model to the residuals from the regression for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal model [@Vehtari_2015].
 The optimal model in both cases was AR(1).

### 2016-03-22 paper.Rmd
Adjusted size and caption for climate plot.

@@
 
-```{r climateplot, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("climateplot", "Temperature futures in two alternate realities."), fig.width=5, fig.height=5}
+```{r climateplot, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("climateplot", "Historical measurements of temperature and temperature futures in two alternate realities."), fig.width=4, fig.height=3}
 source("utils/climate_plot.R")

### 2016-03-22 utils/climate_plot.R

@@ -17,7 +17,7 @@ climate_plot <- function(){
     scale_color_brewer(palette = "Set1", labels = c(CO2 = expression(CO[2]), TSI = "TSI"), 
                        name = "True driver of future climate") + 
     labs(x = "Year", y = expression(paste("Temperature Anomaly  ", (degree * C)))) +
-    theme(legend.position = c(0.15,0.95), legend.justification = c(0,1), 
+    theme(legend.position = c(0.04,0.95), legend.justification = c(0,1), 
           legend.key.width = grid::unit(0.05,'npc'), 
           legend.key = element_rect(color = NA))
 }

### 2016-03-23 paper.Rmd
paper: decent amount of work on the first 2/3rds of paper.

@@
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
-date: "March 21, 2016"
+date: "March 23, 2016"
 output:
@@
 
-We present a simulation model as a computational test-bed for a climate prediction market.
-Traders adapt their beliefs about future global temperatures based on the monetary performance of their neighboring traders in a social network.
-We conduct experiments with the model to determine how a variety of market factors and the temperature data-generating process may affect the extent to which climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
+We present a simulation model as a computational test-bed for climate prediction markets.
+Traders adapt their beliefs about future temperatures based on the monetary performance of their neighboring traders in a social network.
+We conduct experiments with the model to determine how a variety of social factors and the temperature data-generating process may affect the extent to which climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
 We simulate two alternative climate futures: one where the natural log of CO$_2$ is the primary driver of global temperature and one where sun spots are the primary driver.
-These represent the two most plausible competing views in the public discourse and our analysis is agnostic about which is "true".
-Market participation causes traders to converge toward believing the "true" climate model under a variety of model parameterizations in a relatively small number of years, suggesting that a climate prediction market could be useful regardless of which climate change beliefs ultimately turn out to be the best model of temperature change.
+These represent the two most plausible competing views.
+Market participation causes most traders to converge toward believing the "true" climate model in a relatively small number of years, suggesting that a climate market could be useful.
+(150 word limit)
 
@@
 Bloch, Annan and Bowles [-@Bloch_2010] have proposed using derivatives markets to reduce the scientific uncertainty in estimating the impact of future climate change.
-We simulate a prediction market where traders exchange securities related to climate outcomes to explore whether, and under what market and climate conditions, prediction markets may be useful for increasing convergence of climate beliefs.
+We simulate a prediction market where traders exchange securities related to climate outcomes to explore whether, and under what social and climate conditions, prediction markets may be useful for increasing convergence of climate beliefs.
 Our work can be extended as part of a computational design process for effective climate prediction markets.
@@
 37], and the other, which corresponds to an alternative theory advocated by many who doubt or reject conventional climate science, takes the total solar irradiance (the brightness of sunlight, in Watts per square meter, at the top of the atmosphere) averaged over the 11-year sunspot cycle [@Soon_2005].
-Actual scientific models of the earth's climate include many forcings, including carbon dioxide, other greenhouse gases, aerosols, total solar irradiance, and more.
+Most scientific models of the earth's climate include many forcings, including carbon dioxide, other greenhouse gases, aerosols, total solar irradiance, and more.
 In these models, the changing $\ce{CO2}$ concentration is, by a large margin, the strongest forcing [@IPCC_WG1_AR5_2013, p.
@@
 
-For $\ce{CO2}$ we used historical emissions through 2005, harmonized with RCP~8.5 representative concentration pathway from 2005 onward [@rcp_database_2009; @Riahi_2011] and for TSI we used the harmonized historical values, with a projection through 2100 from Velasco-Herrera *et al.* [-@Velasco_Herrera_2015].
+For $\ce{CO2}$ we used historical emissions through 2005, harmonized with RCP 8.5 representative concentration pathway from 2005 onward [@rcp_database_2009; @Riahi_2011] and for TSI we used the harmonized historical values, with a projection through 2100 from Velasco-Herrera *et al.* [-@Velasco_Herrera_2015], which seems to be the best TSI model.
 
@@
 They represent pervasive positions on climate change in the public debate.
-In order to approximatively match the current configuration of beliefs in climate change in the United States, model 1 is randomly assigned to half of the traders, while 2 is assigned to the other half of the traders.
+In order to approximatively match the current configuration of beliefs in climate change in the United States, the CO$_2$ model is randomly assigned to half of the traders, while the TSI model is assigned to the other half of the traders.
 
-Thus, both when **true.model** = 0 and **true.model** = 1, approximately half of the traders use the true data-generating model to make predictions. 
+Thus, both when the true data generating process is CO$_2$ and TSI, approximately half of the traders use the true data-generating model to make predictions. 
 Traders using the true model do not necessarily make perfectly accurate predictions however. 
@@
 Therefore, the values these traders assign to the parameters of the model will typically be different from the exact parameters in the data-generating process.
+
+Jonathan, can you describe the model estimation process here?
 
@@
 
-Based on their reservation price, traders behave as "zero-intelligence" agents [@Gode1993].
+Based on their reservation price, agents behave as "zero-intelligence" traders [@Gode1993].
 They attempt to sell securities at a random price above their reservation, and to buy securities at a random price below their reservation.
@@
 Based on traders' sell and buy orders, traders exchange securities following a continuous-double auctions (CDA) procedure (see Model Dynamics below for more details). 
-CDAs or some variants thereof are common procedures to match buy and sell orders. 
+CDAs or some close variants are common procedures to match buy and sell orders. 
 CDA are notably used on large stock markets [@Tseng2010a].
@@
 2. compute their stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, CO2_{0:t^*}, TSI_{0:t^*})$ and use it to determine the expected value they attached to each security, and
-3. trade on the CDA market based on their zero-intelligence decision rule as follows: 
+3. trade on the CDA market as follows: 
  	+ Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
@@
 The model depends on the following parameters, which we vary in simulation experiments to determine their effects on the convergence of beliefs.
-We group the parameters into network and individual behavioral factors.
+We group the parameters into climate, network and individual behavioral factors.
+
+* Climate parameters:
+    + **true.model**: temperature data-generating process.
 
@@
 
-* Behavioral parameters.
+* Behavioral parameters:
     + **risk.tak** : determines the distribution of risk-taking behavior. 
@@
 
+We simulate two alternative climate futures: one where the natural log of CO$_2$ is the primary driver of global temperature and one where sun spots are the primary driver.
+These represent the two most plausible competing views in the public discourse and our analysis is agnostic about which is "true".
+Market participation causes traders to converge toward believing the "true" climate model under a variety of model parameterizations in a relatively small number of years, suggesting that a climate prediction market could be useful regardless of which climate change beliefs ultimately turn out to be the best model of temperature change.
+
 All code for the model is available at [johnjnay.com/predictMarket](http://johnjnay.com/predictMarket). This project is a computational test-bed for public policy design.

### 2016-03-24 paper.Rmd
paper work, more on backend

@@
 
-Traders are part of a social network where each agent forms two links at random, and then forms links randomly. 
-This ensures that each agent is connected to at least two other agents. 
+Traders are part of a social network where each agent forms two links at random, and then forms links randomly, ensuring that each agent is connected to at least two other agents. 
 Every time securities are realized, each trader looks at the performance of her richest neighbor in the network. 
 Traders start with the same initial amount of ECU and differences in ECU can only come from market interactions.
-Therefore, if some trader is poorer that her richest neighbor, the trader interprets this has an signal that her richest neighbor may have a better model of the climate. 
+Therefore, if some trader is poorer that her richest neighbor, the trader interprets this as an signal that her richest neighbor may have a better model of the climate.
 Then, the trader considers adopting the model of her richest neighbor. 
-For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is.
+For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is a parameter we vary in our experiments.
 
@@
 
-An example of such social network is depicted in Figure X. 
-We assume that the initial network is segregated.
-Traders who initially believe in model 1 (resp. 2) are more likely to be linked with other traders who also believe in 1 (resp. 2). 
-This reflects reality where cultural and social groups are correlated with climate change beliefs. 
+An example of a snapshot of a social network is depicted in Figure X. 
+We assume that the initial network is segregated: traders who initially believe in CO$_2$ forcing (resp. TSI) are more likely to be linked with other traders who also believe in CO$_2$ (resp. TSI). 
+This reflects reality where cultural and social groups are highly correlated with climate change beliefs. 
 Although traders can change their approximate model over time, the connections between traders do not change as the market unfolds, i.e the edges are fixed.
@@
 
-The time periods $t$ are grouped in trading *sequences*. 
+The time periods $t$ are grouped into trading *sequences*. 
 In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence. 
@@
 * Climate parameters:
-    + **true.model**: temperature data-generating process.
+    + **true.model**: temperature data-generating process, CO$_2$ or TSI.
 
@@
 As a validation that the market is operating correctly, we run the market with actual temperatures 1880 to 2015 and market betting from 1931 to 2014.
-As a warming signal begins to show up in 1980s, traders begin to converge toward the log(CO$_2$) model (`r figRef("time")`).
+Before 1980 there is no statistically significant warming trend so we would not expect the market to change beliefs before then.
+As a global warming signal begins to show up in 1980s, traders begin to converge toward the log(CO$_2$) model (`r figRef("time")`).
 
@@
 The sensitivity analysis averages over time and thus masks time trends.
-We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and set conducted an experiment crossing 95% and 5% percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model.
+We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and conducted an experiment crossing 95% and 5% percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model.
 We then collected the time series of convergence across these eight designs for visualization of the distributions of convergence over time (`r figRef("time")`).
@@
 
-All code for the model is available at [johnjnay.com/predictMarket](http://johnjnay.com/predictMarket). This project is a computational test-bed for public policy design.
+All code for the model is available at [johnjnay.com/predictMarket](http://johnjnay.com/predictMarket).
+This project is a computational test-bed for public policy design: our code can be extended to test the effects of various trading strategies, belief models, future climate scenarios, and market designs.
 
 Existing prediction markets (e.g.
-hypermind.com, betfair.com, and predictit.org) focus on nearer term events (months in advance), but have been shown to have performance superior to statistical models for some event types [@pathak_comparison_2015].
+hypermind.com, betfair.com, and predictit.org) focus on near term events (months in advance), but have been shown to have performance superior to statistical models for some event types [@pathak_comparison_2015].
 

### 2016-03-24 paper.Rmd
First Draft Paper

@@
 The idea of using prediction markets to efficiently aggregate information about uncertain event outcomes has been widely discussed [@Horn2014c].
-Prediction markets have interesting theoretical properties [@Set1998; @Hanson2012], and perform well in terms of prediction accuracy and information aggregation in experiments [@Hanson2006; @Healy2010] simulation models [@Klingert2012c; @Jumadinova2011a] and the real-world [@Wolfers2006].
+Prediction markets have interesting theoretical properties [@Set1998; @Hanson2012], and perform well in terms of prediction accuracy and information aggregation in experiments [@Hanson2006; @Healy2010] simulation models [@Klingert2012c; @Jumadinova2011a] and the real-world [@Wolfers2006; @pathak_comparison_2015].
 Prediction markets have been shown to be effective at identifying which scientific publications are likely to be successfully replicated [@Dreber_2015].
+
 Bloch, Annan and Bowles [-@Bloch_2010] have proposed using derivatives markets to reduce the scientific uncertainty in estimating the impact of future climate change.
+Existing prediction markets (e.g.
+hypermind.com, betfair.com, and predictit.org) focus on near term events such as elections months away, so it is difficult to extrapolate empirical findings to the climate case.
+Furthermore, we are interested in investigating the unobservable beliefs of traders.
+Therefore, we turn to simulation modeling informed by climate and economic theory.
 We simulate a prediction market where traders exchange securities related to climate outcomes to explore whether, and under what social and climate conditions, prediction markets may be useful for increasing convergence of climate beliefs.
@@
 We then collected the time series of convergence across these eight designs for visualization of the distributions of convergence over time (`r figRef("time")`).
+Under most parameterizations, within 15-20 years, the median percent of traders believing in the true model is over 75%.
 
@@
 
-```{r sa2, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=tabRef("sa2", "")}
-
-```
-
- `r tabRef("one")`
-
 # 4. Discussion
@@
 All code for the model is available at [johnjnay.com/predictMarket](http://johnjnay.com/predictMarket).
-This project is a computational test-bed for public policy design: our code can be extended to test the effects of various trading strategies, belief models, future climate scenarios, and market designs.
-
-Existing prediction markets (e.g.
-hypermind.com, betfair.com, and predictit.org) focus on near term events (months in advance), but have been shown to have performance superior to statistical models for some event types [@pathak_comparison_2015].
+This project is a computational test-bed for public policy design: our code can be extended to test the effects that trading strategies, cognitive models, future climate scenarios, and market designs have on the evolution of trader beliefs.
 

### 2016-03-25 paper.Rmd
Small typos and corrections + comments on pdf

@@
 Traders start with the same initial amount of ECU and differences in ECU can only come from market interactions.
-Therefore, if some trader is poorer that her richest neighbor, the trader interprets this as an signal that her richest neighbor may have a better model of the climate.
+Therefore, if some trader is poorer than her richest neighbor, the trader interprets it as a signal that her richest neighbor may have a better model of the climate.
 Then, the trader considers adopting the model of her richest neighbor. 
@@
 1. recalibrate their approximate model based on the new set of temperature data available at $t$,
-2. compute their stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, CO2_{0:t^*}, TSI_{0:t^*})$ and use it to determine the expected value they attached to each security, and
+2. compute their stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, CO2_{0:t^*}, TSI_{0:t^*})$ and use it to determine the expected value they attach to each security, and
 3. trade on the CDA market as follows: 
@@
  	+ If there are  outstanding sell offers for $s_i^B$ at a price below $p_i^B$, a trade is concluded. 
-   Trader $i$ buys one unit of $s_i^B$ from the sellers who sells at the highest price below $p_i^B$, and the sell and buy offers are removed from the order book.
+   Trader $i$ buys one unit of $s_i^B$ from the sellers who sells at the lowest price below $p_i^B$, and the sell and buy offers are removed from the order book.
  	+ If there are outstanding buy offers for $s_i^S$ at price above $p_i^S$, a trade is concluded. 
@@
 1.  receive 1 ECU per unit of $s^*$ they own, and
-2.  consider adapting their neighbors' approximate model as described above.
+2.  consider adapting their neighbors' approximate model as described in Social Network.
 
@@
     Formally, for each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$. 
-    The value of $d_{i}$ is the probability that $i$ adopts one of her neighbors' approximate model if this neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
+    The value of $d_{i}$ is the probability that $i$ adopts the approximate model of her richest neighbor if that neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
 

### 2016-03-26 paper.Rmd
WSC MS Word.

@@
 output:
+  word_document: default
   pdf_document:
-    keep_tex: yes
     includes:
       in_header: header.tex
+    keep_tex: yes
   html_document: default
@@
 
-# 1. Introduction
+# Introduction
 
@@
 
-# 2. Related Work
+# Related Work
 
@@
 
-# 2. Model Design
+# Model Design
 
@@
 
-# 3. Results
+# Results
 
@@
 
-# 4. Discussion
+# Discussion
 

### 2016-03-26 prediction_market_wsc.Rnw
Initial RNW version

--- 2016-03-25 paper.Rmd
@@
----
-title: 'Betting and Belief: Prediction Markets and Attribution of Climate Change'
-author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
-date: "March 23, 2016"
-output:
-  pdf_document:
-    keep_tex: yes
-    includes:
-      in_header: header.tex
-  html_document: default
-bibliography: library.bib
----
-```{r fig_ref, echo=FALSE}
-source("utils/fig_ref.R") # tabRef and figRef functions
+%**************************************************************************
+%*
+%*  Paper: ``INSTRUCTIONS FOR AUTHORS OF LATEX DOCUMENTS''
+%*
+%*  Publication: 2016 Winter Simulation Conference Author Kit
+%*
+%*  Filename: wsc16paper.tex
+%*
+%*  Date: January 31, 2001   Time:  9:45 PM
+%*      BASE of current version: Feb 01, 2010 (primary WSC'10 LaTeX file)
+%*
+%*  Word Processing System: TeXnicCenter and MiKTeX
+%*
+%*
+%*  All files need the following
+\input{wsc16style.tex}     % download from author kit.  Style files for wsc formatting. Don't remove this line - required for generating the final paper!
+
+\documentclass{wscpaperproc}
+
+<<setup, echo=F, results='hide'>>=
+library(ggplot2)
+library(pander)
 source("utils/multi_plot.R")
-``` 
-```{r, caption_setting, echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
-library(knitr)
-knit_hooks$set(htmlcap = function(before, options, envir) {
-  if(!before) {
-    paste('<p class="caption">',options$htmlcap,"</p>",sep="")
+theme_set(theme_classic())
+@
+
+%% The next versions of the hyperref command are used if you adopt the
+%% outdated latex-dvips-ps2pdf route in generating your pdf file. In
+%% this case you can use ps or eps files for graphics, but not pdf, jpeg, png etc.
+%% However, the final pdf file should embed all fonts required which means that you have to use file
+%% formats which can embed fonts. Please note that the final PDF file will not be generated on your computer!
+%% If you are using WinEdt or PCTeX, then use the following. If you are using
+%% Y&Y TeX then replace "dvips" with "dvipsone"
+
+%%\usepackage[dvips,colorlinks=true,urlcolor=blue,citecolor=black,%
+%% anchorcolor=black,linkcolor=black]{hyperref}
+%****************************************************************************
+
+  \usepackage{latexsym}
+  \usepackage{graphicx}
+  \usepackage{mathptmx}
+  \usepackage[pdftex,colorlinks=true,urlcolor=blue,citecolor=black,anchorcolor=black,linkcolor=black]{hyperref}
+  \usepackage{mhchem}
+
+\bibliographystyle{wsc}
+
+%#########################################################
+%*
+%*  The Document.
+%*
+\begin{document}
+
+%***************************************************************************
+% AUTHOR: AUTHOR NAMES GO HERE
+% FORMAT AUTHORS NAMES Like: Author1, Author2 and Author3 (last names)
+%
+%  You need to change the author listing below!
+%               Please list ALL authors using last name only, separate by a comma except
+%               for the last author, separate with "and"
+%
+\WSCpagesetup{Nay, Van der Linden, and Gilligan}
+
+% AUTHOR: Enter the title, all letters in upper case
+\title{Betting and Belief: Prediction Markets and Attribution of Climate Change}
+
+% AUTHOR: Enter the authors of the article, see end of the example document for further examples
+\author{John J. Nay\\ [12pt]
+School of Engineering\\
+Vanderbilt University \\
+PMB 351826 \\
+2301 Vanderbilt Place\\
+Nashville, TN 37235-1826, USA
+% Multiple authors are entered as follows.
+% You may also need to adjust the titlevbox size in the preamble - search for titlevboxsize
+\and
+Martin Van der Linden\\ [12pt]
+Department of Economics\\
+Vanderbilt University\\
+Nashville, TN 37235, USA\\
+\and
+Jonathan M. Gilligan\\ [12pt]
+Department of Earth \& Environmental Sciences\\
+Vanderbilt University\\
+PMB 351805\\
+2301 Vanderbilt PL\\
+Nashville, TN 37235-1805, USA
 }
-    })
-# small margins
-knit_hooks$set(small_mar = function(before, options, envir) {
-    if (before) par(mar = rep(0.0001, 4))  # smaller margins all around
-})
-# opts_chunk$set(dev="tiff", 
-#                dev.args=list(compression="lzw", type="cairo"),
-#                dpi=300)
-```
-```{r multi_plot, echo=FALSE}
-source("utils/multi_plot.R")
-library(ggplot2)
-theme_set(theme_classic())
-library(pander)
-```
 
-# Abstract
+\maketitle
+\section*{ABSTRACT}
 
@@
 We conduct experiments with the model to determine how a variety of social factors and the temperature data-generating process may affect the extent to which climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
-We simulate two alternative climate futures: one where the natural log of CO$_2$ is the primary driver of global temperature and one where sun spots are the primary driver.
+We simulate two alternative climate futures: one where the natural log of \ce{CO2} is the primary driver of global temperature and one where sun spots are the primary driver.
 These represent the two most plausible competing views.
-Market participation causes most traders to converge toward believing the "true" climate model in a relatively small number of years, suggesting that a climate market could be useful.
+Market participation causes most traders to converge toward believing the ``true'' climate model in a relatively small number of years, suggesting that a climate market could be useful.
 (150 word limit)
 
-# 1. Introduction
+\section{INTRODUCTION}
 
 The climate change debate has become strongly polarized over the past two decades.
-Although the scientific consensus on the anthropogenic nature of climate change strongly increased, beliefs about climate change did not evolve much within the public [@Vandenbergh2013f].
+Although the scientific consensus on the anthropogenic nature of climate change strongly increased, beliefs about climate change did not evolve much within the public \shortcite{Vandenbergh2013f}.
 In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily as the question is becoming increasingly politicized and potentially disconnected from scientific evidence.
@@
 
-Attempts to foster such consensus face many social and psychological challenges, some of which could be adressed by creating climate prediction markets where participants can "put their money where their mouths are" [@Hsu2011; @Vandenbergh2013f].
-The idea of using prediction markets to efficiently aggregate information about uncertain event outcomes has been widely discussed [@Horn2014c].
-Prediction markets have interesting theoretical properties [@Set1998; @Hanson2012], and perform well in terms of prediction accuracy and information aggregation in experiments [@Hanson2006; @Healy2010] simulation models [@Klingert2012c; @Jumadinova2011a] and the real-world [@Wolfers2006; @pathak_comparison_2015].
-Prediction markets have been shown to be effective at identifying which scientific publications are likely to be successfully replicated [@Dreber_2015].
+Attempts to foster such consensus face many social and psychological challenges, some of which could be adressed by creating climate prediction markets where participants can ``put their money where their mouths are'' \shortcite{Hsu2011,Vandenbergh2013f}.
+The idea of using prediction markets to efficiently aggregate information about uncertain event outcomes has been widely discussed \shortcite{Horn2014c}.
+Prediction markets have interesting theoretical properties \shortcite{Set1998,Hanson2012}, and perform well in terms of prediction accuracy and information aggregation in experiments \shortcite{Hanson2006,Healy2010} simulation models \shortcite{Klingert2012c,Jumadinova2011a} and the real-world \shortcite{Wolfers2006,pathak_comparison_2015}.
+Prediction markets have been shown to be effective at identifying which scientific publications are likely to be successfully replicated \shortcite{Dreber_2015}.
 
-Bloch, Annan and Bowles [-@Bloch_2010] have proposed using derivatives markets to reduce the scientific uncertainty in estimating the impact of future climate change.
+Bloch, Annan and Bowles \shortciteN{Bloch_2010} have proposed using derivatives markets to reduce the scientific uncertainty in estimating the impact of future climate change.
 Existing prediction markets (e.g.
@@
 
-# 2. Related Work
+\section{RELATED WORK}
 
-@Tseng2010a create an agent-based model (ABM) of a continuous double auction market with multiple market strategies, including two variants of the zero-intelligence agent.
+\shortciteN{Tseng2010a} create an agent-based model (ABM) of a continuous double auction market with multiple market strategies, including two variants of the zero-intelligence agent.
 They compare the behavior of their simulation with data from a prediction market for the outcome of political elections.
 They find that despite their simplicity, zero-intelligence agents capture some salient features of real market data.
-@Klingert2012c reach similar conclusions with respect to experimental data from @Hanson2006.
-In their simulation, @Klingert2012c compare continuous double auctions with logarithmic market scoring rules and find that the latter generally outperform continuous double auctions in terms of prediction accuracy.
+\shortciteN{Klingert2012c} reach similar conclusions with respect to experimental data from \shortciteN{Hanson2006}.
+In their simulation, \shortciteN{Klingert2012c} compare continuous double auctions with logarithmic market scoring rules and find that the latter generally outperform continuous double auctions in terms of prediction accuracy.
 In some variants of the two last models, agents' strategies feature learning, but these are rather basic and unrelated to a formal predictive model.
 
-@Jumadinova2011a conducted research similar to ours, as they explicitly include information about uncertain events in their model.
-@Jumadinova2011a study a continuous double auction model in which agents update their beliefs about uncertain events based on information on the event.
+\shortciteN{Jumadinova2011a} conducted research similar to ours, as they explicitly include information about uncertain events in their model.
+\shortciteN{Jumadinova2011a} study a continuous double auction model in which agents update their beliefs about uncertain events based on information on the event.
 New beliefs are modeled as a weighted sum of the last period's beliefs and market prices.
@@
 Therefore, the model does not allow for investigating the convergence of predictive models.
-@Ontanon2009 study the effect of deliberation on a simulated prediction market.
-They assume that agents use Case-Based Reasoning [@Aamodt1994] to debate uncertain outcomes with their neighbors in a social network.
+\shortciteN{Ontanon2009} study the effect of deliberation on a simulated prediction market.
+They assume that agents use Case-Based Reasoning \shortcite{Aamodt1994} to debate uncertain outcomes with their neighbors in a social network.
 Although the model could in theory be used to assess convergence of predictive models by looking at the arguments agents adopt after the deliberation, this is not done by the authors.
@@
 
-To the best of our knowledge, the idea that prediction markets can generate consensus *on the factors* affecting uncertain events has never been quantitatively explored.
-ABMs of prediction markets have been studied [@Klingert2012c; @Tseng2010a; @Jumadinova2011a], some of which feature communication between agents.
+To the best of our knowledge, the idea that prediction markets can generate consensus \emph{on the factors\/} affecting uncertain events has never been quantitatively explored.
+ABMs of prediction markets have been studied \shortcite{Klingert2012c,Tseng2010a,Jumadinova2011a}, some of which feature communication between agents.
 In these models, however, beliefs about the uncertain outcomes are constructed in rather abstract ways.
@@
 From a public policy perspective, changing the explanatory models of market participants is one of the most important roles that prediction markets might play.
-This is perhaps the most important social benefit of prediction markets given the comparable predictive power of statistical supervised learning models, which can provide information at a much lower cost than creating and maintaining a market [@goel_prediction_2010].
-Effective climate policy not only requires an accurate consensus on future climate *outcomes*.
-It also requires an accurate consensus on the causal *mechanisms* influencing such outcomes.
+This is perhaps the most important social benefit of prediction markets given the comparable predictive power of statistical supervised learning models, which can provide information at a much lower cost than creating and maintaining a market \shortcite{goel_prediction_2010}.
+Effective climate policy not only requires an accurate consensus on future climate \emph{outcomes\/}.
+It also requires an accurate consensus on the causal \emph{mechanisms\/} influencing such outcomes.
 If people agree that temperature will rise, but some believe it will be due to greenhouse gases, while others believe that it will be caused by increased solar activity, inconsistent and ineffective policies may be implemented.
 
-# 2. Model Design
+\section{MODEL DESIGN}
 
@@
 
-## Temperature Models
+\subsection{TEMPERATURE MODELS}
 
 For climate time-series, we use the annual anomaly of global mean temperature.
-For years from 1880--2014 we use the GISTEMP global mean land-sea annual temperature anomalies [@GISTEMP_2016; @Hansen_2010] and for years from 2015 onward, we project future climates under two alternative theories: In both theories, changes in global temperature are proportional to changes in a deterministic forcing plus a stochastic noise term.
-For simplicity, we choose two alternative expressions for the deterministic climate forcing: one, which corresponds to conventional climate science, takes the natural logarithm of the atmospheric carbon dioxide concentration in parts per million [@Archer_2012, p.
-37], and the other, which corresponds to an alternative theory advocated by many who doubt or reject conventional climate science, takes the total solar irradiance (the brightness of sunlight, in Watts per square meter, at the top of the atmosphere) averaged over the 11-year sunspot cycle [@Soon_2005].
+For years from 1880--2014 we use the GISTEMP global mean land-sea annual temperature anomalies \shortcite{GISTEMP_2016,Hansen_2010} and for years from 2015 onward, we project future climates under two alternative theories: In both theories, changes in global temperature are proportional to changes in a deterministic forcing plus a stochastic noise term.
+For simplicity, we choose two alternative expressions for the deterministic climate forcing: one, which corresponds to conventional climate science, takes the natural logarithm of the atmospheric carbon dioxide concentration in parts per million \shortcite{Archer_2012}[p.~37], and the other, which corresponds to an alternative theory advocated by many who doubt or reject conventional climate science, takes the total solar irradiance (the brightness of sunlight, in Watts per square meter, at the top of the atmosphere) averaged over the 11-year sunspot cycle \shortcite{Soon_2005}.
 Most scientific models of the earth's climate include many forcings, including carbon dioxide, other greenhouse gases, aerosols, total solar irradiance, and more.
-In these models, the changing $\ce{CO2}$ concentration is, by a large margin, the strongest forcing [@IPCC_WG1_AR5_2013, p.
-14], so for simplicity we take this as the model.
+In these models, the changing $\ce{CO2}$ concentration is, by a large margin, the strongest forcing \shortcite{IPCC_WG1_AR5_2013}[p.~14], so for simplicity we take this as the model.
 Choosing only one forcing term for each of the competing models simplifies comparison because each has the same number of adjustable parameters.
 
-For $\ce{CO2}$ we used historical emissions through 2005, harmonized with RCP 8.5 representative concentration pathway from 2005 onward [@rcp_database_2009; @Riahi_2011] and for TSI we used the harmonized historical values, with a projection through 2100 from Velasco-Herrera *et al.* [-@Velasco_Herrera_2015], which seems to be the best TSI model.
+For $\ce{CO2}$ we used historical emissions through 2005, harmonized with RCP 8.5 representative concentration pathway from 2005 onward \shortcite{rcp_database_2009,Riahi_2011} and for TSI we used the harmonized historical values, with a projection through 2100 from \shortciteN{Velasco_Herrera_2015}, which seems to be the best TSI model.
 
 Warming coefficients for each model were determined by linear regression of historical temperatures from 1880--2014 against the historical values for each model's forcing term.
-The noise model was determined by fitting an $\text{ARMA}(p,q)$ model to the residuals from the regression for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal model [@Vehtari_2015].
+The noise model was determined by fitting an $\text{ARMA}(p,q)$ model to the residuals from the regression for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal model \shortcite{Vehtari_2015}.
 The optimal model in both cases was AR(1).
@@
 Future climates were generated by multiplying the warming coefficients for each model for warming coefficients adding stochastic AR(1) noise to deterministic warming using parameters fit to the historical data (1880--2014).
-One realization of future climates for both the $\ln(\ce{CO2})$ and TSI models is shown in `r figRef("climateplot")`.
+One realization of future climates for both the $\ln(\ce{CO2})$ and TSI models is shown in Fig.~\ref{fig:climateplot}.
 
-```{r climateplot, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("climateplot", "Historical measurements of temperature and temperature futures in two alternate realities."), fig.width=4, fig.height=3}
+<<r climateplot, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Historical measurements of temperature and temperature futures in two alternate realities.", fig.width=4, fig.height=3>>=
 source("utils/climate_plot.R")
 climate_plot() + theme(text=element_text(family='Palatino'))
-```
+@
 
-## Climate Beliefs
+\subsection{CLIMATE BELIEFS}
 
-Traders use one of the two models (temperature depends on CO$_2$ or TSI) to forecast future temperature. 
+Traders use one of the two models (temperature depends on \ce{CO2} or TSI) to forecast future temperature. 
 These models are interpreted as the trader's beliefs about the true climate process, the driving factor of long-term global temperatures. 
 They represent pervasive positions on climate change in the public debate.
-In order to approximatively match the current configuration of beliefs in climate change in the United States, the CO$_2$ model is randomly assigned to half of the traders, while the TSI model is assigned to the other half of the traders.
+In order to approximatively match the current configuration of beliefs in climate change in the United States, the \ce{CO2} model is randomly assigned to half of the traders, while the TSI model is assigned to the other half of the traders.
 
-Thus, both when the true data generating process is CO$_2$ and TSI, approximately half of the traders use the true data-generating model to make predictions. 
+Thus, both when the true data generating process is \ce{CO2} and TSI, approximately half of the traders use the true data-generating model to make predictions. 
 Traders using the true model do not necessarily make perfectly accurate predictions however. 
-Although these traders believe in the correct *functional form* of the model, they still need to calibrate their model based on limited noisy data. 
+Although these traders believe in the correct \emph{functional form\/} of the model, they still need to calibrate their model based on limited noisy data. 
 Therefore, the values these traders assign to the parameters of the model will typically be different from the exact parameters in the data-generating process.
@@
 
-## Traders and Market
+\subsection{TRADERS AND MARKETS}
 
@@
 
-Based on their reservation price, agents behave as "zero-intelligence" traders [@Gode1993].
+Based on their reservation price, agents behave as ``zero-intelligence'' traders \shortcite{Gode1993}.
 They attempt to sell securities at a random price above their reservation, and to buy securities at a random price below their reservation.
-These trading strategies are simple but provide accurate approximations of behavior in prediction markets [@Klingert2012c], and in financial markets more broadly.
+These trading strategies are simple but provide accurate approximations of behavior in prediction markets \shortcite{Klingert2012c}, and in financial markets more broadly.
 
@@
 CDAs or some close variants are common procedures to match buy and sell orders. 
-CDA are notably used on large stock markets [@Tseng2010a].
+CDA are notably used on large stock markets \shortcite{Tseng2010a}.
 
-## Social Network
+\subsection{SOCIAL NETWORK}
 
@@
 For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is a parameter we vary in our experiments.
-
-![Network](output/network.pdf)
+\begin{figure}[t]
+\begin{center}
+\includegraphics[width=0.80\textwidth]{output/network.pdf}
+\caption{Trader network.}
+\end{center}
+\end{figure}
 
 An example of a snapshot of a social network is depicted in Figure X. 
-We assume that the initial network is segregated: traders who initially believe in CO$_2$ forcing (resp. TSI) are more likely to be linked with other traders who also believe in CO$_2$ (resp. TSI). 
+We assume that the initial network is segregated: traders who initially believe in \ce{CO2} forcing (resp.
+TSI) are more likely to be linked with other traders who also believe in \ce{CO2} (resp.
+TSI).
 This reflects reality where cultural and social groups are highly correlated with climate change beliefs. 
@@
 
-## Model Dynamics
+\subsection{MODEL DYNAMICS}
 
-The time periods $t$ are grouped into trading *sequences*. 
+The time periods $t$ are grouped into trading \emph{sequences\/}. 
 In a given sequence, the potential payments associated with traded securities are all based on the temperature at the end of the sequence. 
@@
 
-1. recalibrate their approximate model based on the new set of temperature data available at $t$,
-2. compute their stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, CO2_{0:t^*}, TSI_{0:t^*})$ and use it to determine the expected value they attach to each security, and
-3. trade on the CDA market as follows: 
- 	+ Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
- 	+ Every trader $i$ also chooses at random a security $s_i^S$ she will try to sell among the securities she owns a positive amount of (if any).
-	+ Traders then decide of their selling price $p_i^S$ and buying price $p_i^B$. 
+\begin{enumerate}
+\item  recalibrate their approximate model based on the new set of temperature data available at $t$,
+\item  compute their stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, CO2_{0:t^*}, TSI_{0:t^*})$ and use it to determine the expected value they attach to each security, and
+\item  trade on the CDA market as follows: 
+  \begin{itemize}
+    \item Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
+    \item Every trader $i$ also chooses at random a security $s_i^S$ she will try to sell among the securities she owns a positive amount of (if any).
+    \item Traders then decide of their selling price $p_i^S$ and buying price $p_i^B$. 
   To do so, traders first compute their expected values $E(s_i^B)$ and $E(s_i^S)$ for securities $s_i^B$ and $s_i^S$. 
@@
 
-	+ Traders go to the market one at the time, in an order drawn randomly for each $t$.
- 	+ When trader $i$ comes to the market, she places limit orders in the order book. 
+    \item Traders go to the market one at the time, in an order drawn randomly for each $t$.
+    \item When trader $i$ comes to the market, she places limit orders in the order book. 
    These orders specify that $i$ is willing to buy $s_i^B$ at any price below $p_i^B$, and to sell $s_i^S$ at any price above $p_i^S$.
- 	+ The market maker then tries to match $i$'s orders with some order which was put in the book before $i$ came to the market.
- 	+ If there are  outstanding sell offers for $s_i^B$ at a price below $p_i^B$, a trade is concluded. 
+    \item The market maker then tries to match $i$'s orders with some order which was put in the book before $i$ came to the market.
+    \item If there are  outstanding sell offers for $s_i^B$ at a price below $p_i^B$, a trade is concluded. 
    Trader $i$ buys one unit of $s_i^B$ from the sellers who sells at the lowest price below $p_i^B$, and the sell and buy offers are removed from the order book.
- 	+ If there are outstanding buy offers for $s_i^S$ at price above $p_i^S$, a trade is concluded. 
+    \item If there are outstanding buy offers for $s_i^S$ at price above $p_i^S$, a trade is concluded. 
    Trader $i$ sells one unit to the buyer who buys at the highest price above $p_i^S$, and the sell and buy offers are removed from the order book.
- 	+ Whenever all traders have come to the market, any remaining outstanding offer is removed from the order book, and the trading period is concluded.
+    \item Whenever all traders have come to the market, any remaining outstanding offer is removed from the order book, and the trading period is concluded. 
+  \end{itemize}
+\end{enumerate}
 
@@
 
-1.  receive 1 ECU per unit of $s^*$ they own, and
-2.  consider adapting their neighbors' approximate model as described in Social Network.
+\begin{enumerate}
+  \item  receive 1 ECU per unit of $s^*$ they own, and
+  \item consider adapting their neighbors' approximate model as described in Social Network.
+\end{enumerate}
 
-## Model Parameters
+\subsection{MODEL PARAMETERS}
 
@@
 We group the parameters into climate, network and individual behavioral factors.
+\begin{itemize}
+  \item Climate parameters:
+  \begin{itemize}
+    \item  extbf{true.model}: temperature data-generating process, \ce{CO2} or TSI.
+  \end{itemize}
+  \item Network parameters:
+  \begin{itemize}
+    \item  extbf{n.traders}: the number of traders.
+    \item  extbf{n.edg}: the number of edges in the social network.
+    \item  extbf{seg}: the initial degree of homophily in the network. The higher  extbf{seg}, the higher the initial homophily. 
+  \end{itemize}
 
-* Climate parameters:
-    + **true.model**: temperature data-generating process, CO$_2$ or TSI.
-
-* Network parameters:
-    + **n.traders**: the number of traders.
-    + **n.edg**: the number of edges in the social network.
-    + **seg**: the initial degree of homophily in the network. The higher **seg**, the higher the initial homophily.
-
-* Behavioral parameters:
-    + **risk.tak** : determines the distribution of risk-taking behavior. 
+  \item Behavioral parameters:
+  \begin{itemize}
+    \item  extbf{risk.tak} : determines the distribution of risk-taking behavior. 
     For each trader $i$, the level of \texttt{risk.tak}$_i$ is drawn uniformly at random from $[0,$\texttt{risk.tak} $]$.
@@
 
-    + **ideo** :  determines the degree of "ideology" of traders.
+    \item  extbf{ideo} :  determines the degree of ``ideology'' of traders.
     For each trader $i$, the level of \texttt{ideo}$_i$ is drawn uniformly at random from $[0,$\texttt{ideo} $]$.
@@
     The value of $d_{i}$ is the probability that $i$ adopts the approximate model of her richest neighbor if that neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
+  \end{itemize}
+\end{itemize}
 
-# 3. Results
+\section{RESULTS}
 
-## Historical Data
+\subsection{HISTORICAL DATA}
 
@@
 Before 1980 there is no statistically significant warming trend so we would not expect the market to change beliefs before then.
-As a global warming signal begins to show up in 1980s, traders begin to converge toward the log(CO$_2$) model (`r figRef("time")`).
+As a global warming signal begins to show up in 1980s, traders begin to converge toward the log(\ce{CO2}) model (Fig.~\ref{fig:time}).
 
-## Future Scenarios
+\subsection{FUTURE SCENARIOS}
 
@@
 This is more relevant for policy design and more interesting theoretically due to the increasing divergence of global temperature values under the two future scenarios.
-The sensitivity analysis is based on a Latin Hypercube Sampling of parameter sets from the following distributions [@beachkofski_improved_2002; @carnell_lhs_2012].
+The sensitivity analysis is based on a Latin Hypercube Sampling of parameter sets from the following distributions \shortcite{beachkofski_improved_2002,carnell_lhs_2012}.
 
-* ideo $\sim Unif(0,1)$
-* market.complet $\sim Unif(0,1000)$ (mapped into integer)
-* n.edge $\sim Unif(100,200)$ (mapped into integer)
-* n.traders $\sim Unif(50,250)$ (mapped into integer)
-* risk.tak  $\sim Unif(0,1)$
-* seg $\sim Unif(0,1)$
-* true.model $\sim Binom(0.5)$
+\begin{itemize}
+  \item ideo $\sim Unif(0,1)$
+  \item market.complet $\sim Unif(0,1000)$ (mapped into integer)
+  \item n.edge $\sim Unif(100,200)$ (mapped into integer)
+  \item n.traders $\sim Unif(50,250)$ (mapped into integer)
+  \item risk.tak  $\sim Unif(0,1)$
+  \item seg $\sim Unif(0,1)$
+  \item true.model $\sim Binom(0.5)$
+\end{itemize}
 
-We use the model to simulate 10,000 outcomes based on the input paramter sets and then conduct a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of belief convergence scores, $y$ [@Marino2008; @pujol_sensitivity:_2014; @saltelli_sensitivity_2009].
+We use the model to simulate 10,000 outcomes based on the input paramter sets and then conduct a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
 Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).
@@
 
-Our sensitivity analysis, `r figRef("sa")`, shows that ideology, risk tolerance, the number of traders, the number of edges per trader, the segmentation of the social network, and the true model all statistically significantly affect convergence.
-Segmented social networks can create an "echo-chamber" effect.
-Market participation causes traders to converge toward believing the "true" climate model; however, when the true model is CO$_2$-driven there is more convergence due to less noise in the relationship between predictor variable (log of CO$_2$ versus TSI) and temperature.
+Our sensitivity analysis, Fig.~\ref{fig:sa}, shows that ideology, risk tolerance, the number of traders, the number of edges per trader, the segmentation of the social network, and the true model all statistically significantly affect convergence.
+Segmented social networks can create an ``echo-chamber'' effect.
+Market participation causes traders to converge toward believing the ``true'' climate model; however, when the true model is \ce{CO2}-driven there is more convergence due to less noise in the relationship between predictor variable (log of \ce{CO2} versus TSI) and temperature.
 
-```{r sa, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("sa", "Estimated effects of model parameters on convergence of beliefs in future scenario."), fig.width=3, fig.height=3.5}
+<<r sa, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Estimated effects of model parameters on convergence of beliefs in future scenario.", fig.width=3, fig.height=3.5>>=
 # library(eat)
@@
 #multi_plot(p1, p2, cols = 2)
-```
+@
 
@@
 We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and conducted an experiment crossing 95% and 5% percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model.
-We then collected the time series of convergence across these eight designs for visualization of the distributions of convergence over time (`r figRef("time")`).
+We then collected the time series of convergence across these eight designs for visualization of the distributions of convergence over time (Fig.~\ref{fig:time}).
 Under most parameterizations, within 15-20 years, the median percent of traders believing in the true model is over 75%.
 
-```{r time, eval=TRUE, echo=FALSE, message=FALSE, htmlcap=figRef("time", "Convergence over trading sequences."), fig.width=10.5, fig.height=8.5}
+<<r time, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Convergence over trading sequences.", fig.width=10.5, fig.height=8.5>>=
 source("utils/time_plot.R")
@@
 multi_plot(p1, p2, cols=2)
-```
+@
 
-# 4. Discussion
+\section{DISCUSSION}
 
-We simulate two alternative climate futures: one where the natural log of CO$_2$ is the primary driver of global temperature and one where sun spots are the primary driver.
-These represent the two most plausible competing views in the public discourse and our analysis is agnostic about which is "true".
-Market participation causes traders to converge toward believing the "true" climate model under a variety of model parameterizations in a relatively small number of years, suggesting that a climate prediction market could be useful regardless of which climate change beliefs ultimately turn out to be the best model of temperature change.
+We simulate two alternative climate futures: one where the natural log of \ce{CO2} is the primary driver of global temperature and one where sun spots are the primary driver.
+These represent the two most plausible competing views in the public discourse and our analysis is agnostic about which is ``true''.
+Market participation causes traders to converge toward believing the ``true'' climate model under a variety of model parameterizations in a relatively small number of years, suggesting that a climate prediction market could be useful regardless of which climate change beliefs ultimately turn out to be the best model of temperature change.
 
-All code for the model is available at [johnjnay.com/predictMarket](http://johnjnay.com/predictMarket).
+All code for the model is available at \href{http://johnjnay.com/predictMarket}{\nolinkurl{johnjnay.com/predictMarket}}.
 This project is a computational test-bed for public policy design: our code can be extended to test the effects that trading strategies, cognitive models, future climate scenarios, and market designs have on the evolution of trader beliefs.
 
-# Acknowledgements
+\section*{ACKNOWLEDGEMENTS}
 
@@
 
-# References
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\bibliography{prediction_market_wsc}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-```{r, finalize, eval=TRUE, echo=FALSE}
-# Print out warnings if we did not use all the figures.
-if(!all(environment(figRef)$created)){
-    missingFig <- which(!environment(figRef)$created)
-    warning("Figure(s) ", paste(missingFig, sep=", "), " with label(s) '", 
-      paste(names(environment(figRef)$created)[missingFig], sep="', '"),
-      "' are referenced in the text but have never been created.")
-}
-if(!all(environment(figRef)$used)){
-    missingRef <- which(!environment(figRef)$used)
-    warning("Figure(s) ", paste(missingRef, sep=", "), " with label(s) '", 
-      paste(names(environment(figRef)$used)[missingRef], sep="', '"),
-      "' are present in the document but are never referred to in the text.")
-}
-```
+\section*{AUTHOR BIOGRAPHIES}
+
+\noindent {\bf JOHN J.
+NAY} is a Ph.D.
+Candidate at Vanderbilt University.
+He develops and applies machine learning and simulation methods to better predict behavior in complex systems.
+His research involves collaboration with social scientists, environmental scientists, computer scientists and engineers.
+His email address is \email{john.j.nay@vanderbilt.edu}, and website: \href{https://johnnay.github.io/}{https://johnnay.github.io/}.\\
+
+\noindent {\bf JONATHAN M.
+GILLIGAN} is Associate Professor of Earth and Environmental Sciences at Vanderbilt University and Associate Director for Research at the Vanderbilt Climate Change Research Network.
+His research and teaching interests center on integrating natural science, social science, behavioral science, and engineering to study coupled human-natural systems.
+He holds a  Ph.D.
+in Physics from Yale University.
+His email address is \email{jonathan.gilligan@vanderbilt.edu}, and website: \href{https://my.vanderbilt.edu/jonathangilligan/}{https://my.vanderbilt.edu/jonathangilligan/}.
+
+\end{document}

### 2016-03-26 prediction_market_wsc.Rnw
Checking in before refactoring bibliography

@@
 
-<<setup, echo=F, results='hide'>>=
+<<knitr_setup, echo=F, results='hide', cache=F>>=
+library(knitr)
+opts_chunk$set(fig.align = "center")
+@
+
+<<setup, echo=F, results='hide', cache=F>>=
 library(ggplot2)
@@
 \bibliographystyle{wsc}
+
+\newcommand{Rstat}{\textsc{R}}
 
@@
 Vanderbilt University\\
-Nashville, TN 37235, USA\\
+PMB 351819\\
+2301 Vanderbilt PL\\
+Nashville, TN 37235-1819, USA\\
 \and
@@
 Market participation causes most traders to converge toward believing the ``true'' climate model in a relatively small number of years, suggesting that a climate market could be useful.
-(150 word limit)
 
@@
 The idea of using prediction markets to efficiently aggregate information about uncertain event outcomes has been widely discussed \shortcite{Horn2014c}.
-Prediction markets have interesting theoretical properties \shortcite{Set1998,Hanson2012}, and perform well in terms of prediction accuracy and information aggregation in experiments \shortcite{Hanson2006,Healy2010} simulation models \shortcite{Klingert2012c,Jumadinova2011a} and the real-world \shortcite{Wolfers2006,pathak_comparison_2015}.
-Prediction markets have been shown to be effective at identifying which scientific publications are likely to be successfully replicated \shortcite{Dreber_2015}.
+Prediction markets have interesting theoretical properties \shortcite{Set1998,Hanson2012}, and perform well in terms of prediction accuracy and information aggregation in experiments \shortcite{Hanson2006,Healy2010} simulation models \shortcite{Klingert2012c,Jumadinova2011a} and the real-world \shortcite{Wolfers2006,pathak_comparison_2015,Dreber_2015}.
 
@@
 
-\section{RELATED WORK}
-
-\shortciteN{Tseng2010a} create an agent-based model (ABM) of a continuous double auction market with multiple market strategies, including two variants of the zero-intelligence agent.
-They compare the behavior of their simulation with data from a prediction market for the outcome of political elections.
-They find that despite their simplicity, zero-intelligence agents capture some salient features of real market data.
-\shortciteN{Klingert2012c} reach similar conclusions with respect to experimental data from \shortciteN{Hanson2006}.
-In their simulation, \shortciteN{Klingert2012c} compare continuous double auctions with logarithmic market scoring rules and find that the latter generally outperform continuous double auctions in terms of prediction accuracy.
-In some variants of the two last models, agents' strategies feature learning, but these are rather basic and unrelated to a formal predictive model.
-
-\shortciteN{Jumadinova2011a} conducted research similar to ours, as they explicitly include information about uncertain events in their model.
-\shortciteN{Jumadinova2011a} study a continuous double auction model in which agents update their beliefs about uncertain events based on information on the event.
-New beliefs are modeled as a weighted sum of the last period's beliefs and market prices.
-The agents' information set changes every period.
-The better the information set, the more likely the agent is to put higher weight on last period's prices when revising their beliefs.
-Little structure is imposed on the information set and the way it is used to generate the weights in the update function.
-Therefore, the model does not allow for investigating the convergence of predictive models.
-\shortciteN{Ontanon2009} study the effect of deliberation on a simulated prediction market.
-They assume that agents use Case-Based Reasoning \shortcite{Aamodt1994} to debate uncertain outcomes with their neighbors in a social network.
-Although the model could in theory be used to assess convergence of predictive models by looking at the arguments agents adopt after the deliberation, this is not done by the authors.
-Also, the way beliefs are formed and revised through argumentation is too abstract for studying climate prediction markets, where beliefs should be based on statistical time series models.
-
-To the best of our knowledge, the idea that prediction markets can generate consensus \emph{on the factors\/} affecting uncertain events has never been quantitatively explored.
-ABMs of prediction markets have been studied \shortcite{Klingert2012c,Tseng2010a,Jumadinova2011a}, some of which feature communication between agents.
-In these models, however, beliefs about the uncertain outcomes are constructed in rather abstract ways.
-In particular, beliefs are not based on structural models from which agents could derive causal implications, and therefore these models are not suitable for investigating the convergence of the underlying explanatory models agents employ for prediction.
-
 From a public policy perspective, changing the explanatory models of market participants is one of the most important roles that prediction markets might play.
@@
 If people agree that temperature will rise, but some believe it will be due to greenhouse gases, while others believe that it will be caused by increased solar activity, inconsistent and ineffective policies may be implemented.
+
+\section{RELATED WORK}
+
+To the best of our knowledge, the idea that prediction markets can generate consensus \emph{on the factors\/} affecting uncertain events has never been explored quantitatively.
+Agent-based simulations prediction markets have been studied \shortcite{Klingert2012c,Tseng2010a,Jumadinova2011a}, including some that feature communication between agents.
+In these models, however, beliefs about the uncertain outcomes are constructed in rather abstract ways.
+In particular, beliefs are not based on structural models from which agents could derive causal implications, so these models are not suitable for investigating the convergence of the underlying explanatory models agents employ for prediction.
+
+\shortciteN{Tseng2010a} created an agent-based model (ABM) of a continuous double auction market with multiple market strategies, including two variants of a zero-intelligence agent.
+They compared the behavior of their simulation with data from a prediction market for the outcome of political elections.
+They found that despite their simplicity, zero-intelligence agents capture some salient features of real market data.
+\shortciteN{Klingert2012c} compared the predictive accuracy different kinds of simulated markets (condinuous double auctions and logarithmic market scoring rules) and reached similar conclusions to experimental work by \shortciteN{Hanson2006}.
+None of these models featured agents learning and updating their beliefs.
+
+\shortciteN{Jumadinova2011a} created a continuous double auction model in which agents update their beliefs about uncertain events based on newly-acquired information.
+The better the information set, the more likely the agent was to put higher weight on last period's prices when revising their beliefs.
+However, little structure was imposed on the information set and the way it was used to generate the weights for updating beliefs, so that model does not permit studying the convergence of trader beliefs about predictive models.
+\shortciteN{Ontanon2009} studied the effect of deliberation on a simulated prediction market, in which agents used case-based reasoning \shortcite{Aamodt1994} to debate uncertain outcomes with their neighbors in a social network.
+This model could in principle be used to assess convergence of predictive models, but this was not done and the process of forming and revising beliefs through argumentation is not well suited to studying stochastic time series, such as those relevant to climate prediction markets.
 
@@
 The noise model was determined by fitting an $\text{ARMA}(p,q)$ model to the residuals from the regression for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal model \shortcite{Vehtari_2015}.
+This analysis was performed in \Rstat{} using the Stan probabilistic modeling language and the \texttt{rstan} package \shortcite{Stan2016,Rstan2016}.
 The optimal model in both cases was AR(1).
@@
 
-<<r climateplot, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Historical measurements of temperature and temperature futures in two alternate realities.", fig.width=4, fig.height=3>>=
+<<climateplot, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Historical measurements of temperature and temperature futures in two alternate realities.", fig.width=4, fig.height=3>>=
 source("utils/climate_plot.R")
@@
 
+At each time-step (year), the agents re-estimate the coefficients of each model using a Bayesian 
 Jonathan, can you describe the model estimation process here?
@@
 \begin{center}
-\includegraphics[width=0.80\textwidth]{output/network.pdf}
-\caption{Trader network.}
+\includegraphics[width=0.50\textwidth]{output/network.pdf}
+\caption{An example of the social network among traders.
+Yellow circles indicate traders who believe that $\log\ce{CO2}$ is responsible for climate change and blue circles indicate traders who believe that TSI is responsible.
+Lines indicate social network connections.}\label{fig:social_network}%
 \end{center}
@@
 
-An example of a snapshot of a social network is depicted in Figure X. 
+An example of a snapshot of a social network is depicted in Fig.~\ref{fig:social_network}. 
 We assume that the initial network is segregated: traders who initially believe in \ce{CO2} forcing (resp.
@@
 
-At each time $t$, traders are assumed to know the past value of the temperature ${T}_{0:t}$, carbon dioxide $CO2_{0:t}$, and total solar irradiance $TSI_{0:t}$. 
-In a sequence finishing at time $t^*$, traders also have common knowledge of $CO2_{t:t^*}$ and $TSI_{t:t^*}$, the future values of carbon dioxide and total solar irradiance up to $t^*$.
+At each time $t$, traders are assumed to know the past value of the temperature ${T}_{0:t}$, carbon dioxide $\text{CO2}_{0:t}$, and total solar irradiance $\text{TSI}_{0:t}$. 
+In a sequence finishing at time $t^*$, traders also have common knowledge of $\text{CO2}_{t:t^*}$ and $\text{TSI}_{t:t^*}$, the future values of carbon dioxide and total solar irradiance up to $t^*$.
 However, at any $t$, traders do not know the value of any future temperatures. 
 In particular, in a given sequence, traders do not know the value of ${T}_{t^*}$. 
-Traders can only predict ${T}_{t^*}$ using their approximate model and their knowledge of ${T}_{0:t}$, $CO2_{0:t^*}$, and $TSI_{0:t^*}$.
-Notice that because ${T}_{0:t}$, $CO2_{0:t^*}$, and $TSI_{0:t^*}$ are common knowledge, in each period $t$, any two trader with the same approximate model $m$ forms the same stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, CO2_{0:t^*}, TSI_{0:t^*})$.
+Traders can only predict ${T}_{t^*}$ using their approximate model and their knowledge of ${T}_{0:t}$, $\text{CO2}_{0:t^*}$, and $\text{TSI}_{0:t^*}$.
+Notice that because ${T}_{0:t}$, $\text{CO2}_{0:t^*}$, and $\text{TSI}_{0:t^*}$ are common knowledge, in each period $t$, any two trader with the same approximate model $m$ forms the same stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, \text{CO2}_{0:t^*}, \text{TSI}_{0:t^*})$.
 
@@
 \item  recalibrate their approximate model based on the new set of temperature data available at $t$,
-\item  compute their stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, CO2_{0:t^*}, TSI_{0:t^*})$ and use it to determine the expected value they attach to each security, and
+\item  compute their stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, \text{CO2}_{0:t^*}, \text{TSI}_{0:t^*})$ and use it to determine the expected value they attach to each security, and
 \item  trade on the CDA market as follows: 
@@
 
-<<r sa, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Estimated effects of model parameters on convergence of beliefs in future scenario.", fig.width=3, fig.height=3.5>>=
+<<sa, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Estimated effects of model parameters on convergence of beliefs in future scenario.", fig.width=3, fig.height=3.5>>=
 # library(eat)
@@
 
-<<r time, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Convergence over trading sequences.", fig.width=10.5, fig.height=8.5>>=
+<<time, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Convergence over trading sequences.", fig.width=10.5, fig.height=8.5, fig.pos='t'>>=
 source("utils/time_plot.R")

### 2016-03-26 utils/pc_sa.R

@@ -96,13 +96,13 @@ correct_bias <- function(x) {
 
 setMethod("plot", "pcSA",
           function(x, outcome_var = "Outcome", 
-                   xlab = "Variable", ylab = "Partial Rank Correlation Coefficient"){
+                   xlab = "Variable", ylab = "Partial rank correlation coefficient"){
             x <- x@result
             ss <- correct_bias(x[[7]]) # $SRRC, $SRC, $PRCC, $PCC
             
             ggplot2::ggplot(ss, ggplot2::aes(x = var, y = x_corr)) + ggplot2::geom_point() +
               ggplot2::xlab(xlab) + ggplot2::ylab(ylab) +
-              ggplot2::ggtitle(paste("Estimated Effects on", outcome_var)) +
+              ggplot2::ggtitle(paste("Estimated effects on", outcome_var)) +
               ggplot2::geom_errorbar(ggplot2::aes(ymax = max_ci, ymin = min_ci), width=0.25) +
               ggplot2::geom_hline(yintercept = 0) +
               ggplot2::ylim(c(-1,1)) +

### 2016-03-26 utils/time_plot.R

@@ -25,14 +25,16 @@ future_line <- function(){
                                 prettyNum(length(complete.cases(plot_data$convergence)), big.mark = ','), ")"),
          title = paste0("2015-",  2014 + 14 * 6)) +
     ylim(0,100) +
-    theme(legend.justification=c(1,0), legend.position=c(1,0), 
+    theme(plot.title = element_text(hjust=0.5),
+          legend.justification=c(1,0), legend.position=c(1,0), 
           legend.key.width = grid::unit(0.05, 'npc'),
           legend.key = element_rect(color=NA))
 }
 
 future_box <- function(){
   load("output/convergence_future.Rda")
-  
+  sequence.years <- 6
+
   plot_data <- average_convergence
   # plot_data$set <- factor(plot_data$set, levels = gtools::mixedsort(unique(plot_data$set)))
   #  n.edg <- round(parameters[5]*100) + 100 # integer in (100, 200)
@@ -41,26 +43,37 @@ future_box <- function(){
                                      paste0(round(8 * 0.05 + 2, 1), " edges/trader, segmentation = 0.95"), 
                                      paste0(round(8 * 0.95 + 2, 1), " edges/trader, segmentation = 0.05"), 
                                      paste0(round(8 * 0.95 + 2, 1), " edges/trader, segmentation = 0.95")))
+  initial_convergence <- plot_data %>% filter(trading_seq == 1) %>% 
+    mutate(trading_seq = 1/sequence.years, convergence = 0)
+  complete_data <- plot_data %>% bind_rows(initial_convergence)
   
-  sequence.years <- 6
-  
-  ggplot(data=plot_data, aes(x= trading_seq * sequence.years + 2014, y=(convergence + 1) * 50, color = true_mod)) +
-    geom_boxplot(aes(group = floor(trading_seq * sequence.years + 2014 - 1.25 +  2.5 * (true_mod == "SlowTSI")), 
-                     color = true_mod), position = position_dodge(width = 3), width = 2) + 
+  ggplot(data=complete_data, aes(x= trading_seq * sequence.years + 2014, y=(convergence + 1) * 50, color = true_mod)) +
+    geom_point(data = initial_convergence, size = 2) +
+    geom_boxplot(data = plot_data, 
+                 aes(group = floor(trading_seq * sequence.years + 2014 - 1.25 +  2.5 * (true_mod == "SlowTSI")), 
[…]

### 2016-03-27 prediction_market_wsc.tex
Lots of edits.
Moved paper fully to .Rnw format and set it up for the full WSC 2016 format.

Edited introduction and related previously published work section following Martin's comments.

Wrote details of model/belief updating in Climate Beliefs and Traders & Markets.

Rewrote discussion/conclusions to move discussion from results and separate discussion from conclusions.

Edited much of the process and parameterization
sections for clarity.

Abstract is currently 149 words.

TODO: Martin to fill in biography. John to clarify error bars in SA plot.

--- 2016-03-26 prediction_market_wsc.Rnw
@@
 
-\documentclass{wscpaperproc}
+\documentclass{wscpaperproc}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
 
-<<knitr_setup, echo=F, results='hide', cache=F>>=
-library(knitr)
-opts_chunk$set(fig.align = "center")
-@
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
 
-<<setup, echo=F, results='hide', cache=F>>=
-library(ggplot2)
-library(pander)
-source("utils/multi_plot.R")
-theme_set(theme_classic())
-@
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 
-\newcommand{Rstat}{\textsc{R}}
+\newcommand{\Rstat}{\textsf{R}}
+\newcommand{\ARMA}{\text{ARMA}}
+\newcommand{\AR}{\text{AR}}
 
@@
 %*
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
+Despite overwhelming scientific evidence, a large fraction of the American public doubts that greenhouse gases are causing global warming.
 We present a simulation model as a computational test-bed for climate prediction markets.
-Traders adapt their beliefs about future temperatures based on the monetary performance of their neighboring traders in a social network.
-We conduct experiments with the model to determine how a variety of social factors and the temperature data-generating process may affect the extent to which climate prediction markets foster a convergence of market participants' beliefs regarding a true model of the global climate.
-We simulate two alternative climate futures: one where the natural log of \ce{CO2} is the primary driver of global temperature and one where sun spots are the primary driver.
-These represent the two most plausible competing views.
-Market participation causes most traders to converge toward believing the ``true'' climate model in a relatively small number of years, suggesting that a climate market could be useful.
+Traders adapt their beliefs about future temperatures based on the profits of other traders in their social network.
+We simulate two alternative climate futures, in which global temperatures are primarily driven either by carbon dioxide or by solar irradiance.
+These represent, respectively, the scientific consensus and a hypothesis advanced by prominent skeptics.
+We conduct sensitivity analyses to determine how a variety of factors describing both the market and the physical climate may affect traders' beliefs about the cause of global climate change.
+Market participation causes most traders to converge quickly toward believing the ``true'' climate model, suggesting that a climate market could be useful at building public consensus.
 
@@
 
-\subsection{TEMPERATURE MODELS}
+\subsection{Temperature Models}%
+\label{sec:temp.models}
 
@@
 Warming coefficients for each model were determined by linear regression of historical temperatures from 1880--2014 against the historical values for each model's forcing term.
-The noise model was determined by fitting an $\text{ARMA}(p,q)$ model to the residuals from the regression for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal model \shortcite{Vehtari_2015}.
-This analysis was performed in \Rstat{} using the Stan probabilistic modeling language and the \texttt{rstan} package \shortcite{Stan2016,Rstan2016}.
-The optimal model in both cases was AR(1).
+The noise model was determined by fitting an $\ARMA(p,q)$ model to the residuals from the regression, using the \href{http://mc-stan.org}{Stan} probabilistic modeling language and the \href{https://cran.r-project.org/web/packages/rstan/}{\texttt{rstan}} package \shortcite{Stan2016,RStan2016}.
+We identified the optimal model for the autocorrelated noise term by perfroming the regression analysis for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal noise model \shortcite{Vehtari_2015}.
+In both cases (TSI and $\log\ce{CO2}$), the optimal noise model was $\AR(1)$.
 
-Future climates were generated by multiplying the warming coefficients for each model for warming coefficients adding stochastic AR(1) noise to deterministic warming using parameters fit to the historical data (1880--2014).
-One realization of future climates for both the $\ln(\ce{CO2})$ and TSI models is shown in Fig.~\ref{fig:climateplot}.
+Future climates were generated by applying the warming coefficients to Eq.~\ref{eq:climate}:
+\begin{equation}
+T_{\text{model}}(t) = \beta_{\text{model}} F_{\text{model}}(t) + \varepsilon,
+\label{eq:climate}
+\end{equation}
+where $T_{\text{model}}(t)$ is the temperature at time $t$, under a given model of what causes warming,
+$F_{\text{model}}(t)$ is the forcing (TSI or $\log\ce{CO2}$) at time $t$,
+$\beta_{\text{model}}$ is the regression coefficient, and
+$\varepsilon$ is an $\ARMA(p,q)$ noise term.
+The coefficient $\beta$ and the parameters of the ARMA noise model were fit to the historical data (1880--2014), as described above. 
+An example realization of future climates for both the $\ln(\ce{CO2})$ and TSI models is shown in Fig.~\ref{fig:climateplot}.
 
-<<climateplot, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Historical measurements of temperature and temperature futures in two alternate realities.", fig.width=4, fig.height=3>>=
-source("utils/climate_plot.R")
-climate_plot() + theme(text=element_text(family='Palatino'))
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
 
-\subsection{CLIMATE BELIEFS}
+{\centering \includegraphics[width=\maxwidth]{figure/climateplot-1}
+
+}
+
+\caption[Historical measurements of temperature and a realization of possible future temperatures under two alternate models of climate physics]{Historical measurements of temperature and a realization of possible future temperatures under two alternate models of climate physics.}\label{fig:climateplot}
+\end{figure}
+
+\end{knitrout}
+
+\subsection{Climate Beliefs}
 
@@
 
-At each time-step (year), the agents re-estimate the coefficients of each model using a Bayesian 
-Jonathan, can you describe the model estimation process here?
-
-\subsection{TRADERS AND MARKETS}
+\subsection{Traders and Markets}
 
@@
 Each security pays 1 ECU at the end of the trading sequence if the temperature at the end of the sequence falls into a certain range.
+
 Traders are risk-neutral expected utility maximizers. 
 Therefore, their reservation price for a security is simply their assessment of the probability that the temperature will fall in the range covered by the security at the end of the sequence.
+At each time-step (year), the agents use the new year's temperature data to re-estimate the coefficients for the model they believe explains climate change, using using Bayesian linear regression with an $\AR(1)$ noise model, and use the joint posterior probability distribution of regression and noise coefficients to estimate the probability distribution for the  temperature at the end of the current trading sequence.
+Stan was sufficiently fast that we could perform a full Bayesian analysis at each time step.
+Traders then use this posterior probability distribution to assign reservation prices for buying and selling securities.
 
@@
 
-\subsection{SOCIAL NETWORK}
+\subsection{Social Network}
 
@@
 \caption{An example of the social network among traders.
-Yellow circles indicate traders who believe that $\log\ce{CO2}$ is responsible for climate change and blue circles indicate traders who believe that TSI is responsible.
-Lines indicate social network connections.}\label{fig:social_network}%
+Yellow and blue circles represent traders who believe in the two different models of climate change.
+Lines indicate social network connections.
+In a highly segmented network (\texttt{seg} close to 1), most links connect like-minded traders and few connect traders with opposing beliefs.}\label{fig:social_network}%
 \end{center}
@@
 
-\subsection{MODEL DYNAMICS}
+\subsection{Model Dynamics}
 
@@
 However, at any $t$, traders do not know the value of any future temperatures. 
-In particular, in a given sequence, traders do not know the value of ${T}_{t^*}$. 
+Thus, the traders know the forcing terms for the rest of the trading sequence, but do not know the value of ${T}_{t^*}$. 
 Traders can only predict ${T}_{t^*}$ using their approximate model and their knowledge of ${T}_{0:t}$, $\text{CO2}_{0:t^*}$, and $\text{TSI}_{0:t^*}$.
-Notice that because ${T}_{0:t}$, $\text{CO2}_{0:t^*}$, and $\text{TSI}_{0:t^*}$ are common knowledge, in each period $t$, any two trader with the same approximate model $m$ forms the same stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, \text{CO2}_{0:t^*}, \text{TSI}_{0:t^*})$.
+Notice that because ${T}_{0:t}$, $\text{CO2}_{0:t^*}$, and $\text{TSI}_{0:t^*}$ are common knowledge, in each period $t$, any two trader with the same approximate model $m$ forms the same stochastic beliefs about future temperatures $P_{t,m}({T}_{t^*} ~|~ {T}_{0:t}, \text{CO2}_{0:t^*}, \text{TSI}_{0:t^*})$.
+The probability distribution $P_{t,m}$ incorporates both epistemic uncertainty (the trader does not know the true values of the coefficients of the climate model) and aleatory uncertainty (in addition to deterministic warming, the temperature exhibits stochastic noise).
 
  At each time $t$, traders:
-
 \begin{enumerate}
 \item  recalibrate their approximate model based on the new set of temperature data available at $t$,
-\item  compute their stochastic beliefs about future temperatures $F_{tm}({T}_{t^*} ~|~ {T}_{0:t}, \text{CO2}_{0:t^*}, \text{TSI}_{0:t^*})$ and use it to determine the expected value they attach to each security, and
+\label{step:calibrate}
+\item  use the posterior probability distribution from (\ref{step:calibrate}) to 
+assign beliefs about the probability distribution of future temperatures at time $t^*$: 
+$P_{t,m}({T}_{t^*} ~|~ {T}_{0:t}, \text{CO2}_{0:t^*}, \text{TSI}_{0:t^*})$ 
+and use it to determine the expected value they attach to each security, and
 \item  trade on the CDA market as follows: 
@@
   Then traders set $p_i^S$ at random above $E(s_i^S)$ and $p_i^B$ at random below $E(s_i^B)$ (see Model Parameters below for more details).
-
     \item Traders go to the market one at the time, in an order drawn randomly for each $t$.
@@
 
-\subsection{MODEL PARAMETERS}
+\subsection{Model Parameters}
 
@@
   \begin{itemize}
-    \item  extbf{true.model}: temperature data-generating process, \ce{CO2} or TSI.
+    \item  \texttt{true.model}: temperature data-generating process (\ce{CO2} or TSI).
   \end{itemize}
@@
   \begin{itemize}
-    \item  extbf{n.traders}: the number of traders.
-    \item  extbf{n.edg}: the number of edges in the social network.
-    \item  extbf{seg}: the initial degree of homophily in the network. The higher  extbf{seg}, the higher the initial homophily. 
+    \item  \texttt{n.traders}: the number of traders.
+    \item  \texttt{n.edg}: the number of edges in the social network.
+    \item  \texttt{seg}: the initial degree of homophily in the network. 
+    The higher  \texttt{seg}, the higher the initial homophily. 
   \end{itemize}
@@
   \begin{itemize}
-    \item  extbf{risk.tak} : determines the distribution of risk-taking behavior. 
-    For each trader $i$, the level of \texttt{risk.tak}$_i$ is drawn uniformly at random from $[0,$\texttt{risk.tak} $]$.
-    The higher \texttt{risk.tak}$_i$, the higher (resp. lower) the price $i$ will set for the securities $i$ tries to sell (resp. buy).
-    Formally, let $\texttt{reserv}_{it}$ be the reservation price of trader $i$ at time $t$ for some security $s$.
-    Then trader $i$ picks her buying (resp.
-selling) price for $s$ uniformly at random in the interval  $[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it}  * (1 - \texttt{risk.tak}_i)]$ (resp.
-$[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it} * (1  + \texttt{risk.tak}_i)]$).
+    \item  \texttt{risk.tak} : determines the distribution of risk tolerance with respect to successfully buying or selling securities.
+Higher risk tolerance corresponds to demanding more aggressive prices for buying and selling, and hence, a higher risk of not consummating a trade that would be mutually advantageous to the buyer and seller.
 
-    \item  extbf{ideo} :  determines the degree of ``ideology'' of traders.
-    For each trader $i$, the level of \texttt{ideo}$_i$ is drawn uniformly at random from $[0,$\texttt{ideo} $]$.
+    For each trader $i$, the level of $\texttt{risk.tak}_i$ is drawn uniformly at random from $[0,\texttt{risk.tak}]$.
+    The higher $\texttt{risk.tak}_i$, the higher the price $i$ will demand for selling securities, and the lower the price $i$ will offer for buying.
+
+    Formally, let $\texttt{reserv}_{i,t}$ be the reservation price of trader $i$ at time $t$ for some security $s$.
+    Then trader $i$ picks her buying or selling price for $s$ uniformly at random in the interval  
+    $[(1 - \texttt{risk.tak}_i) \texttt{reserv}_{i,t}, \allowbreak  \texttt{reserv}_{i,t}]$ for buying and 
+    $[\texttt{reserv}_{i,t}, \allowbreak (1  + \texttt{risk.tak}_i) \texttt{reserv}_{i,t}]$ for selling. 
+  \item  \texttt{ideo} :  determines the degree of ``ideology'' of traders.
+    For each trader $i$, the level of $\texttt{ideo}_i$ is drawn uniformly at random from $[0,\texttt{ideo}]$.
     If ideo is high, traders will not revise their approximate models easily, even when faced with evidence that their richest neighbor is doing better than them. 
@@
 
-\subsection{HISTORICAL DATA}
+\subsection{Historical Climate}
 
@@
 
-\subsection{FUTURE SCENARIOS}
+\subsection{Future Scenarios}
 
@@
 This is more relevant for policy design and more interesting theoretically due to the increasing divergence of global temperature values under the two future scenarios.
-The sensitivity analysis is based on a Latin Hypercube Sampling of parameter sets from the following distributions \shortcite{beachkofski_improved_2002,carnell_lhs_2012}.
+The sensitivity analysis is based on a Latin hypercube sampling of parameter sets from the following distributions \shortcite{beachkofski_improved_2002,carnell_lhs_2012}.
 
 \begin{itemize}
-  \item ideo $\sim Unif(0,1)$
-  \item market.complet $\sim Unif(0,1000)$ (mapped into integer)
-  \item n.edge $\sim Unif(100,200)$ (mapped into integer)
-  \item n.traders $\sim Unif(50,250)$ (mapped into integer)
-  \item risk.tak  $\sim Unif(0,1)$
-  \item seg $\sim Unif(0,1)$
-  \item true.model $\sim Binom(0.5)$
+  \item \texttt{ideo} $\sim \text{Uniform}(0,1)$
+  \item \texttt{market.complet} $\sim \text{Uniform}(0,1000)$ (mapped into integer)
+  \item \texttt{n.edge} $\sim \text{Uniform}(100,200)$ (mapped into integer)
+  \item \texttt{n.traders} $\sim \text{Uniform}(50,250)$ (mapped into integer)
+  \item \texttt{risk.tak}  $\sim \text{Uniform}(0,1)$
+  \item \texttt{seg} $\sim \text{Uniform}(0,1)$
+  \item \texttt{true.model} $\sim \text{Bernoulli}(0.5)$
 \end{itemize}
 
-We use the model to simulate 10,000 outcomes based on the input paramter sets and then conduct a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
+The future \ce{CO2} forcing is taken from the RCP~8.5 scenario \shortcite{Riahi_2011} and the future TSI forcing is taken from a prediction of 21\textsuperscript{st} Total Solar Irradiance by \shortciteN{Velasco_Herrera_2015}.
+Because of the stochastic noise term, the temperature time series were different in each simulation.
+
+We use the model to perform 10,000 simulations for each input paramter set, and then conducted a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
 Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).
-The only difference with the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships.
-That is every value, $y_s$, in the data set is replace by a number $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, where $y_r$ are the other observed values of $y$ in the data set.
+The only difference between the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships: every value, $y_s$, in the data set is replaced by a number
+$f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, 
+where $y_r$ are the observed values of $y$ in the data set.
 
-Our sensitivity analysis, Fig.~\ref{fig:sa}, shows that ideology, risk tolerance, the number of traders, the number of edges per trader, the segmentation of the social network, and the true model all statistically significantly affect convergence.
-Segmented social networks can create an ``echo-chamber'' effect.
-Market participation causes traders to converge toward believing the ``true'' climate model; however, when the true model is \ce{CO2}-driven there is more convergence due to less noise in the relationship between predictor variable (log of \ce{CO2} versus TSI) and temperature.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
 
-<<sa, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Estimated effects of model parameters on convergence of beliefs in future scenario.", fig.width=3, fig.height=3.5>>=
-# library(eat)
-# load("output/src_future.Rda")
-# plot(src_future, outcome_var = "Convergence") + theme_classic() + theme(text=element_text(family='Palatino'))
+{\centering \includegraphics[width=\maxwidth]{figure/sa-1}
 
-source("utils/sa_plot.R")
-sa_plot() + theme(text=element_text(family='Palatino'))
-#multi_plot(p1, p2, cols = 2)
-@
+}
+
+\caption{Estimated effects of model parameters on convergence of beliefs in future scenario.
+Positive numbers accelerate convergence and negative numbers retard it.
+True model refers to the true climate model being \ce{CO2}, so traders converge to the true model more quickly when carbon dioxide drives climate change than when total solar irradiance does.
+\textbf{TODO: John: explain the error bars: one sigma, two sigma, 95\% CI, \dots}}\label{fig:sa}
+\end{figure}
+
+\end{knitrout}
 
 The sensitivity analysis averages over time and thus masks time trends.
-We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and conducted an experiment crossing 95% and 5% percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model.
+We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and conducted an experiment crossing 95\textsuperscript{th} and 5\textsuperscript{th} percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model.
 We then collected the time series of convergence across these eight designs for visualization of the distributions of convergence over time (Fig.~\ref{fig:time}).
-Under most parameterizations, within 15-20 years, the median percent of traders believing in the true model is over 75%.
+Under most parameterizations, within 15--20 years, the median fraction of traders believing in the true model is over 75\%.
 
-<<time, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Convergence over trading sequences.", fig.width=10.5, fig.height=8.5, fig.pos='t'>>=
-source("utils/time_plot.R")
-p1 <- history_box() + theme(text=element_text(family='Palatino'))
-p2 <- future_box() + theme(text=element_text(family='Palatino'))
-multi_plot(p1, p2, cols=2)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[t]
+
+{\centering \includegraphics[width=\maxwidth]{figure/time-1}
+
+}
+
+\caption[Convergence over trading sequences for different degrees of social-network connection and segmentation]{Convergence over trading sequences for different degrees of social-network connection and segmentation.}\label{fig:time}
+\end{figure}
+
+\end{knitrout}
 
@@
 
+Our sensitivity analysis, Fig.~\ref{fig:sa}, shows that ideology, risk tolerance, the number of traders, the number of edges per trader, the segmentation of the social network, and the true model all statistically significantly affect convergence.
+Increasing the number of edges per trader increases the flow of information through the market, which causes traders to converge toward believing the ``true'' climate model.
+
+Segmented social networks reduce convergence by creating an ``echo-chamber'' effect, in which lack of interaction between traders with different views reduces traders' access to information that could persuade them to change their beliefs.
+This is apparent in Fig.~\ref{fig:time}, where the rate of convergence in highly segmented markets ($\text{\texttt{seg}} = 0.95$) is considerably slower, expecially in the first three trading sequences (18 years), than for markets with low segmentation ($\text{\texttt{seg}} = 0.05$).
+
+When the true model is \ce{CO2}-driven there is more convergence toward the true model.
+We believe this is because when models are fit to the historical data, the residuals from the \ce{CO2}-driven model are considerably smaller than those from the TSI-driven model.
+Thus, projections of future climate change using the TSI model will exhibit significantly larger stochastic noise than the \ce{CO2} models, and this noise makes it more difficult for traders to identify the true model when the true model is TSI.
+
+The higher the risk tolerance, the greater the range of prices at which a trader will buy and sell securities.
+This creates greater opportunities for traders to profit if their model is correct and to lose money if it is incorrect.
+This increased volatility puts more information in play, and we believe this is why greater risk taking enables traders to identify the correct model more quickly.
+We have observed a similar phenomenon in a very different context, where adding stochastic noise to player decisions in iterated games improves the accuracy with which machine-learning algorithms can identify the players' strategies \shortcite{Nay2015}.
+
+What is not clear to us is why increasing the number of traders slows convergence, even if the number of edges per trader remains fixed. This will be a topic for future research.
+
+The historical sequence in Fig.~\ref{fig:time} shows convergence to the TSI model until the early 1970s, after which traders converge toward the \ce{CO2} model.
+This is consistent with the fact that greenhouse gas concentrations did not become high enough to begin dominating natural climate variations until the 1970s or so, and it took until the 1990s for the temperature record to show clear signs of anthropogenic interference \shortcite{IPCC_WG1_AR5_2013}.
+
+\section{CONCLUSIONS}
+
 We simulate two alternative climate futures: one where the natural log of \ce{CO2} is the primary driver of global temperature and one where sun spots are the primary driver.
 These represent the two most plausible competing views in the public discourse and our analysis is agnostic about which is ``true''.
-Market participation causes traders to converge toward believing the ``true'' climate model under a variety of model parameterizations in a relatively small number of years, suggesting that a climate prediction market could be useful regardless of which climate change beliefs ultimately turn out to be the best model of temperature change.
+
+Market participation causes traders to converge toward believing the ``true'' climate model under a variety of model parameterizations in a relatively short time: In markets with low segmentation, the number of traders believing the true model rises from 50\% to 75\% in roughly 12 years, and even in highly segmented markets, 75\% convergence is achieved in 18--24 years.
+This suggests that a climate prediction market could be useful for producing broad agreement about the causes of climate change, and could have persuasive power for people who are not persuaded by the overwhelming consensus among climate researchers that greenhouse gases (and especially \ce{CO2}) are responsible for the majority of observed climate change \shortcite{IPCC_WG1_AR5_2013}.
+However, market segmentation has a large effect on the speed of convergence, so transparency and effective communication about the performance of traders with different beliefs will be important.
+
+The fact that rapid convergence to the true model occurs regardless of which model is actually ``true'' may persuade those who doubt the scientific consensus that the market is ideologically neutral, and that the deck is not stacked to produce a pre-determined result.
 
@@
 
-We would like to thank Yevgeniy Vorobeychik for feedback on this project.
+We thank Yevgeniy Vorobeychik for feedback on this project. 
+We thank Manuel Naumovich Velasco for sharing the data from his Total Solar Irradiance prediction.
+We also thank the authors of \href{http://www.r-project.org}{\textsf{R}} \shortcite{r_core}
+and 
+\href{http://mc-stan.org}{Stan} \shortcite{Stan2016}.
+This manuscript was prepared using 
+%RStudio \shortcite{rstudio} and 
+knitr \shortcite{xie_knitr_2014}.
 
@@
 His research involves collaboration with social scientists, environmental scientists, computer scientists and engineers.
-His email address is \email{john.j.nay@vanderbilt.edu}, and website: \href{https://johnnay.github.io/}{https://johnnay.github.io/}.\\
+His email address is \email{john.j.nay@vanderbilt.edu}, and website: \href{https://johnjnay.com/}{http://johnjnay.com/}.\\
+
+\noindent {\bf MARTIN VAN DER LINDEN} is a Ph.D. Candidate at Vanderbilt University. \\
+\textbf{TODO: Martin, please fill this in.}\\
+His email address is \email{martin.van.der.linden@vanderbilt.edu}, and website: \href{https://martinvdlinden.wordpress.com/}{https://martinvdlinden.wordpress.com/}.\\
 

### 2016-03-27 utils/time_plot.R

@@ -10,10 +10,10 @@ future_line <- function(){
   # plot_data$set <- factor(plot_data$set, levels = gtools::mixedsort(unique(plot_data$set)))
   #  n.edg <- round(parameters[5]*100) + 100 # integer in (100, 200)
   plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"), 
-                          labels = c(paste0(round(8 * 0.05 + 2, 1), " edges/trader, segmentation = 0.05"), 
-                                     paste0(round(8 * 0.05 + 2, 1), " edges/trader, segmentation = 0.95"), 
-                                     paste0(round(8 * 0.95 + 2, 1), " edges/trader, segmentation = 0.05"), 
-                                     paste0(round(8 * 0.95 + 2, 1), " edges/trader, segmentation = 0.95")))
+                          labels = c(paste0(round(8 * 0.05 + 2, 1), " edges per trader, segmentation = 0.05"), 
+                                     paste0(round(8 * 0.05 + 2, 1), " edges per trader, segmentation = 0.95"), 
+                                     paste0(round(8 * 0.95 + 2, 1), " edges per trader, segmentation = 0.05"), 
+                                     paste0(round(8 * 0.95 + 2, 1), " edges per trader, segmentation = 0.95")))
   
   sequence.years <- 6
   
@@ -39,10 +39,10 @@ future_box <- function(){
   # plot_data$set <- factor(plot_data$set, levels = gtools::mixedsort(unique(plot_data$set)))
   #  n.edg <- round(parameters[5]*100) + 100 # integer in (100, 200)
   plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"), 
-                          labels = c(paste0(round(8 * 0.05 + 2, 1), " edges/trader, segmentation = 0.05"), 
-                                     paste0(round(8 * 0.05 + 2, 1), " edges/trader, segmentation = 0.95"), 
-                                     paste0(round(8 * 0.95 + 2, 1), " edges/trader, segmentation = 0.05"), 
-                                     paste0(round(8 * 0.95 + 2, 1), " edges/trader, segmentation = 0.95")))
+                          labels = c(paste0(round(8 * 0.05 + 2, 1), " edges per trader, segmentation = 0.05"), 
[…]

### 2016-03-27 prediction_market_wsc.Rnw
Fix a few typos.

@@
 Warming coefficients for each model were determined by linear regression of historical temperatures from 1880--2014 against the historical values for each model's forcing term.
-The noise model was determined by fitting an $\text{ARMA}(p,q)$ model to the residuals from the regression for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal model \shortcite{Vehtari_2015}.
-This analysis was performed in \Rstat{} using the Stan probabilistic modeling language and the \texttt{rstan} package \shortcite{Stan2016,Rstan2016}.
-The optimal model in both cases was AR(1).
+The noise model was determined by fitting an $\ARMA(p,q)$ model to the residuals from the regression, using the \href{http://mc-stan.org}{Stan} probabilistic modeling language and the \href{https://cran.r-project.org/web/packages/rstan/}{\texttt{rstan}} package \shortcite{Stan2016,RStan2016}.
+We identified the optimal model for the autocorrelated noise term by perfroming the regression analysis for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal noise model \shortcite{Vehtari_2015}.
+In both cases (TSI and $\log\ce{CO2}$), the optimal noise model was $\AR(1)$.
 
-Future climates were generated by multiplying the warming coefficients for each model for warming coefficients adding stochastic AR(1) noise to deterministic warming using parameters fit to the historical data (1880--2014).
-One realization of future climates for both the $\ln(\ce{CO2})$ and TSI models is shown in Fig.~\ref{fig:climateplot}.
+Future climates were generated by applying the warming coefficients to Eq.~\ref{eq:climate}:
+\begin{equation}
+T_{\text{model}}(t) = \beta_{\text{model}} F_{\text{model}}(t) + \varepsilon,
+\label{eq:climate}
+\end{equation}
+where $T_{\text{model}}(t)$ is the temperature at time $t$, under a given model of what causes warming,
+$F_{\text{model}}(t)$ is the forcing (TSI or $\log\ce{CO2}$) at time $t$,
+$\beta_{\text{model}}$ is the regression coefficient, and
+$\varepsilon$ is an $\ARMA(p,q)$ noise term.
+The coefficient $\beta$ and the parameters of the ARMA noise model were fit to the historical data (1880--2014), as described above in Social Network.
+An example realization of future climates for both the $\ln(\ce{CO2})$ and TSI models is shown in Fig.~\ref{fig:climateplot}.
 
-<<climateplot, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Historical measurements of temperature and temperature futures in two alternate realities.", fig.width=4, fig.height=3>>=
+<<climateplot, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Historical measurements of temperature and a realization of possible future temperatures under two alternate models of climate physics.", fig.width=4, fig.height=3>>=
 source("utils/climate_plot.R")
@@
 
-\subsection{CLIMATE BELIEFS}
+\subsection{Climate Beliefs}
 
@@
 Traders start with the same initial amount of ECU and differences in ECU can only come from market interactions.
-Therefore, if some trader is poorer than her richest neighbor, the trader interprets it as a signal that her richest neighbor may have a better model of the climate.
+Therefore, if some trader is poorer than her richest neighbor, the trader interprets it as a signal than her richest neighbor may have a better model of the climate.
 Then, the trader considers adopting the model of her richest neighbor. 
@@
     \item If there are  outstanding sell offers for $s_i^B$ at a price below $p_i^B$, a trade is concluded. 
-   Trader $i$ buys one unit of $s_i^B$ from the sellers who sells at the lowest price below $p_i^B$, and the sell and buy offers are removed from the order book.
+   Trader $i$ buys one unit of $s_i^B$ from the seller who sells at the lowest price below $p_i^B$, and the sell and buy offers are removed from the order book.
     \item If there are outstanding buy offers for $s_i^S$ at price above $p_i^S$, a trade is concluded. 
@@
   \begin{itemize}
-    \item  extbf{risk.tak} : determines the distribution of risk-taking behavior. 
-    For each trader $i$, the level of \texttt{risk.tak}$_i$ is drawn uniformly at random from $[0,$\texttt{risk.tak} $]$.
-    The higher \texttt{risk.tak}$_i$, the higher (resp. lower) the price $i$ will set for the securities $i$ tries to sell (resp. buy).
-    Formally, let $\texttt{reserv}_{it}$ be the reservation price of trader $i$ at time $t$ for some security $s$.
-    Then trader $i$ picks her buying (resp.
-selling) price for $s$ uniformly at random in the interval  $[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it}  * (1 - \texttt{risk.tak}_i)]$ (resp.
-$[\texttt{reserv}_{it}, \allowbreak \texttt{reserv}_{it} * (1  + \texttt{risk.tak}_i)]$).
+    \item  \texttt{risk.tak} : determines the distribution of risk tolerance with respect to successfully buying or selling securities.
+Higher risk tolerance corresponds to demanding more aggressive prices for buying and selling, and hence, a higher risk of not consummating a trade that would be mutually advantageous to the buyer and seller.
 
-    \item  extbf{ideo} :  determines the degree of ``ideology'' of traders.
-    For each trader $i$, the level of \texttt{ideo}$_i$ is drawn uniformly at random from $[0,$\texttt{ideo} $]$.
-    If ideo is high, traders will not revise their approximate models easily, even when faced with evidence that their richest neighbor is doing better than them. 
+    For each trader $i$, the level of $\texttt{risk.tak}_i$ is drawn uniformly at random from $[0,\texttt{risk.tak}]$.
+    The higher $\texttt{risk.tak}_i$, the higher the price $i$ will demand for selling securities, and the lower the price $i$ will offer for buying.
+
+    Formally, let $\texttt{reserv}_{i,t}$ be the reservation price of trader $i$ at time $t$ for some security $s$.
+    Then trader $i$ picks her buying or selling price for $s$ uniformly at random in the interval  
+    $[(1 - \texttt{risk.tak}_i) \texttt{reserv}_{i,t}, \allowbreak  \texttt{reserv}_{i,t}]$ for buying and 
+    $[\texttt{reserv}_{i,t}, \allowbreak (1  + \texttt{risk.tak}_i) \texttt{reserv}_{i,t}]$ for selling. 
+  \item  \texttt{ideo} :  determines the degree of ``ideology'' of traders.
+    For each trader $i$, the level of $\texttt{ideo}_i$ is drawn uniformly at random from $[0,\texttt{ideo}]$.
+    If \texttt{ideo} is high, traders will not revise their approximate models easily, even when faced with evidence that their richest neighbor is doing better than them. 
     Formally, for each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$. 
@@
 This is more relevant for policy design and more interesting theoretically due to the increasing divergence of global temperature values under the two future scenarios.
-The sensitivity analysis is based on a Latin Hypercube Sampling of parameter sets from the following distributions \shortcite{beachkofski_improved_2002,carnell_lhs_2012}.
+The sensitivity analysis is based on a Latin hypercube sampling of parameter sets from the following distributions \shortcite{beachkofski_improved_2002,carnell_lhs_2012}.
 
 \begin{itemize}
-  \item ideo $\sim Unif(0,1)$
-  \item market.complet $\sim Unif(0,1000)$ (mapped into integer)
-  \item n.edge $\sim Unif(100,200)$ (mapped into integer)
-  \item n.traders $\sim Unif(50,250)$ (mapped into integer)
-  \item risk.tak  $\sim Unif(0,1)$
-  \item seg $\sim Unif(0,1)$
-  \item true.model $\sim Binom(0.5)$
+  \item \texttt{ideo} $\sim \text{Uniform}(0,1)$
+  \item \texttt{market.complet} $\sim \text{Uniform}(0,1000)$ (mapped into integer)
+  \item \texttt{n.edge} $\sim \text{Uniform}(100,200)$ (mapped into integer)
+  \item \texttt{n.traders} $\sim \text{Uniform}(50,250)$ (mapped into integer)
+  \item \texttt{risk.tak}  $\sim \text{Uniform}(0,1)$
+  \item \texttt{seg} $\sim \text{Uniform}(0,1)$
+  \item \texttt{true.model} $\sim \text{Bernoulli}(0.5)$
 \end{itemize}
 
-We use the model to simulate 10,000 outcomes based on the input paramter sets and then conduct a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
+The future \ce{CO2} forcing is taken from the RCP~8.5 scenario \shortcite{Riahi_2011} and the future TSI forcing is taken from a prediction of 21\textsuperscript{st} Total Solar Irradiance by \shortciteN{Velasco_Herrera_2015}.
+Because of the stochastic noise term, the temperature time series were different in each simulation.
+
+We use the model to perform 10,000 simulations for each input paramter set, and then conducted a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
 Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).
-The only difference with the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships.
-That is every value, $y_s$, in the data set is replace by a number $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, where $y_r$ are the other observed values of $y$ in the data set.
+The only difference between the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships: every value, $y_s$, in the data set is replaced by a number
+$f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, 
+where $y_r$ are the observed values of $y$ in the data set.
 
-Our sensitivity analysis, Fig.~\ref{fig:sa}, shows that ideology, risk tolerance, the number of traders, the number of edges per trader, the segmentation of the social network, and the true model all statistically significantly affect convergence.
-Segmented social networks can create an ``echo-chamber'' effect.
-Market participation causes traders to converge toward believing the ``true'' climate model; however, when the true model is \ce{CO2}-driven there is more convergence due to less noise in the relationship between predictor variable (log of \ce{CO2} versus TSI) and temperature.
-
-<<sa, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Estimated effects of model parameters on convergence of beliefs in future scenario.", fig.width=3, fig.height=3.5>>=
+<<sa, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Estimated effects of model parameters on convergence of beliefs in future scenario.
+Positive numbers accelerate convergence and negative numbers retard it.
+True model refers to the true climate model being \\ce{CO2}, so traders converge to the true model more quickly when carbon dioxide drives climate change than when total solar irradiance does.
+\\textbf{TODO: John: explain the error bars: one sigma, two sigma, 95\\% CI, \\dots}", fig.width=3, fig.height=3.5>>=
 # library(eat)
@@
 The sensitivity analysis averages over time and thus masks time trends.
-We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and conducted an experiment crossing 95% and 5% percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model.
+We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and conducted an experiment crossing 95\textsuperscript{th} and 5\textsuperscript{th} percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model.
 We then collected the time series of convergence across these eight designs for visualization of the distributions of convergence over time (Fig.~\ref{fig:time}).
-Under most parameterizations, within 15-20 years, the median percent of traders believing in the true model is over 75%.
+Under most parameterizations, within 15--20 years, the median fraction of traders believing in the true model is over 75\%.
 
-<<time, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Convergence over trading sequences.", fig.width=10.5, fig.height=8.5, fig.pos='t'>>=
+<<time, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Convergence over trading sequences for different degrees of social-network connection and segmentation.", fig.width=10.5, fig.height=8.5, fig.pos='t'>>=
 source("utils/time_plot.R")

### 2016-03-27 prediction_market_wsc.tex
Fix a few typos.

@@
 $\varepsilon$ is an $\ARMA(p,q)$ noise term.
-The coefficient $\beta$ and the parameters of the ARMA noise model were fit to the historical data (1880--2014), as described above. 
+The coefficient $\beta$ and the parameters of the ARMA noise model were fit to the historical data (1880--2014), as described above in Social Network.
 An example realization of future climates for both the $\ln(\ce{CO2})$ and TSI models is shown in Fig.~\ref{fig:climateplot}.
@@
     For each trader $i$, the level of $\texttt{ideo}_i$ is drawn uniformly at random from $[0,\texttt{ideo}]$.
-    If ideo is high, traders will not revise their approximate models easily, even when faced with evidence that their richest neighbor is doing better than them. 
+    If \texttt{ideo} is high, traders will not revise their approximate models easily, even when faced with evidence that their richest neighbor is doing better than them. 
     Formally, for each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$. 

### 2016-03-27 prediction_market_wsc.Rnw
Review

@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 
@@
 Prediction markets have interesting theoretical properties \shortcite{Set1998,Hanson2012}, and perform well in terms of prediction accuracy and information aggregation in experiments \shortcite{Hanson2006,Healy2010} simulation models \shortcite{Klingert2012c,Jumadinova2011a} and the real-world \shortcite{Wolfers2006,pathak_comparison_2015,Dreber_2015}.
+However, to the best of our knowledge, the idea that prediction markets can generate consensus \emph{on the factors\/} affecting uncertain events has never been explored quantitatively.
 
@@
 
-To the best of our knowledge, the idea that prediction markets can generate consensus \emph{on the factors\/} affecting uncertain events has never been explored quantitatively.
-Agent-based simulations prediction markets have been studied \shortcite{Klingert2012c,Tseng2010a,Jumadinova2011a}, including some that feature communication between agents.
+Agent-based simulations of prediction markets have been studied \shortcite{Klingert2012c,Tseng2010a,Jumadinova2011a}, including some that feature communication between agents.
 In these models, however, beliefs about the uncertain outcomes are constructed in rather abstract ways.
@@
 They found that despite their simplicity, zero-intelligence agents capture some salient features of real market data.
-\shortciteN{Klingert2012c} compared the predictive accuracy different kinds of simulated markets (condinuous double auctions and logarithmic market scoring rules) and reached similar conclusions to experimental work by \shortciteN{Hanson2006}.
+\shortciteN{Klingert2012c} compared the predictive accuracy of different kinds of simulated markets (continuous double auctions and logarithmic market scoring rules) and reached similar conclusions to experimental work by \shortciteN{Hanson2006}.
 None of these models featured agents learning and updating their beliefs.
@@
 \shortciteN{Jumadinova2011a} created a continuous double auction model in which agents update their beliefs about uncertain events based on newly-acquired information.
-The better the information set, the more likely the agent was to put higher weight on last period's prices when revising their beliefs.
+The better the information set, the more likely the agents were to put higher weight on last period's prices when revising their beliefs.
 However, little structure was imposed on the information set and the way it was used to generate the weights for updating beliefs, so that model does not permit studying the convergence of trader beliefs about predictive models.
@@
 Therefore, their reservation price for a security is simply their assessment of the probability that the temperature will fall in the range covered by the security at the end of the sequence.
-At each time-step (year), the agents use the new year's temperature data to re-estimate the coefficients for the model they believe explains climate change, using using Bayesian linear regression with an $\AR(1)$ noise model, and use the joint posterior probability distribution of regression and noise coefficients to estimate the probability distribution for the  temperature at the end of the current trading sequence.
+At each time-step (year), the agents use the new year's temperature data to re-estimate the coefficients for the model they believe explains climate change, using Bayesian linear regression with an $\AR(1)$ noise model.
+Traders use the joint posterior probability distribution of regression and noise coefficients to estimate the probability distribution for the  temperature at the end of the current trading sequence.
 Stan was sufficiently fast that we could perform a full Bayesian analysis at each time step.
+
 Traders then use this posterior probability distribution to assign reservation prices for buying and selling securities. 
-
 Based on their reservation price, agents behave as ``zero-intelligence'' traders \shortcite{Gode1993}.
@@
 Traders start with the same initial amount of ECU and differences in ECU can only come from market interactions.
-Therefore, if some trader is poorer than her richest neighbor, the trader interprets it as a signal than her richest neighbor may have a better model of the climate.
+Therefore, if some trader is poorer than her richest neighbor, the trader interprets it as a signal that her richest neighbor may have a better model of the climate.
 Then, the trader considers adopting the model of her richest neighbor. 
@@
 \begin{enumerate}
-\item  recalibrate their approximate model based on the new set of temperature data available at $t$,
+\item  recalibrate $m$, their approximate model at time $t$, based on the new set of temperature data available at $t$,
 \label{step:calibrate}
@@
     \item Traders then decide of their selling price $p_i^S$ and buying price $p_i^B$. 
-  To do so, traders first compute their expected values $E(s_i^B)$ and $E(s_i^S)$ for securities $s_i^B$ and $s_i^S$. 
+  To do so, traders first compute their expected values $E(s_i^B)$ and $E(s_i^S)$ for securities $s_i^B$ and $s_i^S$ (where expected values are with respect $i$'s approximate model at time $t$). 
   Then traders set $p_i^S$ at random above $E(s_i^S)$ and $p_i^B$ at random below $E(s_i^B)$ (see Model Parameters below for more details).
@@
   \item  receive 1 ECU per unit of $s^*$ they own, and
-  \item consider adapting their neighbors' approximate model as described in Social Network.
+  \item consider adapting their neighbors' approximate model as described in Social Network above.
 \end{enumerate}
@@
 
-    Formally, let $\texttt{reserv}_{i,t}$ be the reservation price of trader $i$ at time $t$ for some security $s$.
-    Then trader $i$ picks her buying or selling price for $s$ uniformly at random in the interval  
-    $[(1 - \texttt{risk.tak}_i) \texttt{reserv}_{i,t}, \allowbreak  \texttt{reserv}_{i,t}]$ for buying and 
-    $[\texttt{reserv}_{i,t}, \allowbreak (1  + \texttt{risk.tak}_i) \texttt{reserv}_{i,t}]$ for selling. 
+    Formally, in each period, trader $i$ picks her buying or selling price for $s$ uniformly at random in the interval  
+    $[(1 - \texttt{risk.tak}_i) \texttt{reserv}_{i}, \allowbreak  \texttt{reserv}_{i,t}]$ for buying and 
+    $[\texttt{reserv}_{i}, \allowbreak (1  + \texttt{risk.tak}_i) \texttt{reserv}_{i,t}]$ for selling. 
   \item  \texttt{ideo} :  determines the degree of ``ideology'' of traders.
@@
     If \texttt{ideo} is high, traders will not revise their approximate models easily, even when faced with evidence that their richest neighbor is doing better than them. 
-    Formally, for each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$. 
-    The value of $d_{i}$ is the probability that $i$ adopts the approximate model of her richest neighbor if that neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
+    Formally, for each trader $i$ and each sequence, \texttt{ideo}_i is the probability that $i$ adopts the approximate model of her richest neighbor if that neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
   \end{itemize}
@@
 Segmented social networks reduce convergence by creating an ``echo-chamber'' effect, in which lack of interaction between traders with different views reduces traders' access to information that could persuade them to change their beliefs.
-This is apparent in Fig.~\ref{fig:time}, where the rate of convergence in highly segmented markets ($\text{\texttt{seg}} = 0.95$) is considerably slower, expecially in the first three trading sequences (18 years), than for markets with low segmentation ($\text{\texttt{seg}} = 0.05$).
+This is apparent in Fig.~\ref{fig:time}, where the rate of convergence in highly segmented markets ($\text{\texttt{seg}} = 0.95$) is considerably slower, especially in the first three trading sequences (18 years), than for markets with low segmentation ($\text{\texttt{seg}} = 0.05$).
 

### 2016-03-27 prediction_market_wsc.Rnw
the reason it wouldn’t compile from tex to pdf was the "\texttt{idea}_i” you added which should be “$\texttt{idea}_i$”, and mainly because you added this line "\SweaveOpts{concordance=TRUE}”.

@@
 \begin{document}
-\SweaveOpts{concordance=TRUE}
 
@@
     If \texttt{ideo} is high, traders will not revise their approximate models easily, even when faced with evidence that their richest neighbor is doing better than them. 
-    Formally, for each trader $i$ and each sequence, \texttt{ideo}_i is the probability that $i$ adopts the approximate model of her richest neighbor if that neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
+    Formally, for each trader $i$ and each sequence, $\texttt{ideo}_i$ is the probability that $i$ adopts the approximate model of her richest neighbor if that neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
   \end{itemize}

### 2016-03-27 prediction_market_wsc.Rnw
Paper work. Addressed Martins two comments. Editing text too. Bib - fixed the pathak et al reference.

@@
 % AUTHOR: Enter the title, all letters in upper case
-\title{Betting and Belief: Prediction Markets and Attribution of Climate Change}
+\title{BETTING AND BELIEF: PREDICTION MARKETS AND ATTRIBUTION OF CLIMATE CHANGE}
+%Betting and Belief: Prediction Markets and Attribution of Climate Change
 
@@
 
-Despite overwhelming scientific evidence, a large fraction of the American public doubts that greenhouse gases are causing global warming.
+Despite much scientific evidence, a large fraction of the American public doubts that greenhouse gases are causing global warming.
 We present a simulation model as a computational test-bed for climate prediction markets.
@@
 We conduct sensitivity analyses to determine how a variety of factors describing both the market and the physical climate may affect traders' beliefs about the cause of global climate change.
-Market participation causes most traders to converge quickly toward believing the ``true'' climate model, suggesting that a climate market could be useful at building public consensus.
+Market participation causes most traders to converge quickly toward believing the ``true'' climate model, suggesting that a climate market could be useful for building public consensus.
 
@@
 On the other hand, if climate change is human-induced but not recognized as so, the costs of inaction could be devastating.
-Given that effective climate policies require long-lasting measures to be taken, the importance of an accurate consensus on the issue is manifest.
+Given that effective climate policies require long-lasting measures to be taken, the importance of a timely and accurate consensus on the issue is manifest.
 
-Attempts to foster such consensus face many social and psychological challenges, some of which could be adressed by creating climate prediction markets where participants can ``put their money where their mouths are'' \shortcite{Hsu2011,Vandenbergh2013f}.
+Attempts to foster such consensus face many social and psychological challenges, some of which could be addressed by creating climate prediction markets where participants can ``put their money where their mouths are'' \shortcite{Hsu2011,Vandenbergh2013f}.
 The idea of using prediction markets to efficiently aggregate information about uncertain event outcomes has been widely discussed \shortcite{Horn2014c}.
@@
 
-\subsection{Temperature Models}%
+\subsection{Temperature Models}
 \label{sec:temp.models}
@@
 The noise model was determined by fitting an $\ARMA(p,q)$ model to the residuals from the regression, using the \href{http://mc-stan.org}{Stan} probabilistic modeling language and the \href{https://cran.r-project.org/web/packages/rstan/}{\texttt{rstan}} package \shortcite{Stan2016,RStan2016}.
-We identified the optimal model for the autocorrelated noise term by perfroming the regression analysis for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal noise model \shortcite{Vehtari_2015}.
+We identified the optimal model for the auto-correlated noise term by performing the regression analysis for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal noise model \shortcite{Vehtari_2015}.
 In both cases (TSI and $\log\ce{CO2}$), the optimal noise model was $\AR(1)$.
@@
 $\varepsilon$ is an $\ARMA(p,q)$ noise term.
-The coefficient $\beta$ and the parameters of the ARMA noise model were fit to the historical data (1880--2014), as described above in Social Network.
+The coefficient $\beta$ and the parameters of the ARMA noise model were fit to the historical data (1880--2014).
 An example realization of future climates for both the $\ln(\ce{CO2})$ and TSI models is shown in Fig.~\ref{fig:climateplot}.
@@
 They represent pervasive positions on climate change in the public debate.
-In order to approximatively match the current configuration of beliefs in climate change in the United States, the \ce{CO2} model is randomly assigned to half of the traders, while the TSI model is assigned to the other half of the traders.
+In order to approximately match the current configuration of beliefs in climate change in the United States, during model initialization, the \ce{CO2} model is randomly assigned to half of the traders, while the TSI model is assigned to the other half of the traders.
+These random model assignments are made \emph{before} the market initially opens.
 
-Thus, both when the true data generating process is \ce{CO2} and TSI, approximately half of the traders use the true data-generating model to make predictions. 
+Both when the true data generating process is \ce{CO2} and TSI, at model initialization approximately half of the traders use the true data-generating model to make predictions. 
 Traders using the true model do not necessarily make perfectly accurate predictions however. 
@@
 \begin{enumerate}
-\item  recalibrate $m$, their approximate model at time $t$, based on the new set of temperature data available at $t$,
+\item  re-calibrate $m$, their approximate model at time $t$, based on the new set of temperature data available at $t$,
 \label{step:calibrate}
@@
   \begin{itemize}
-    \item Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
+    \item Every trader $i$ chooses at random a security $s_i^B$ she will try to buy.
     \item Every trader $i$ also chooses at random a security $s_i^S$ she will try to sell among the securities she owns a positive amount of (if any).
@@
 
+<<sa_n, eval=TRUE, echo=FALSE, message=FALSE>>=
+load("output/src_future.Rda")
+n_sims <- nrow(src_future@input_set)
+@
+
 Our primary focus is on simulating the market in future scenarios.
 This is more relevant for policy design and more interesting theoretically due to the increasing divergence of global temperature values under the two future scenarios.
-The sensitivity analysis is based on a Latin hypercube sampling of parameter sets from the following distributions \shortcite{beachkofski_improved_2002,carnell_lhs_2012}.
+The sensitivity analysis is based on a Latin hypercube sampling of \Sexpr{n_sims} parameter sets from the following distributions \shortcite{beachkofski_improved_2002,carnell_lhs_2012}.
 
@@
 
-We use the model to perform 10,000 simulations for each input paramter set, and then conducted a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
+We use the model to perform 10 full simulations for each of the \Sexpr{n_sims} input parameter sets and average the 10 convergence scores.
+We conducted multiple simulations for the same parameter set because there is stochasticity.
+We then conducted a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of mean belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
 Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).
@@
 \\textbf{TODO: John: explain the error bars: one sigma, two sigma, 95\\% CI, \\dots}", fig.width=3, fig.height=3.5>>=
-# library(eat)
-# load("output/src_future.Rda")
-# plot(src_future, outcome_var = "Convergence") + theme_classic() + theme(text=element_text(family='Palatino'))
-
 source("utils/sa_plot.R")
 sa_plot() + theme(text=element_text(family='Palatino'))
-#multi_plot(p1, p2, cols = 2)
 @
@@
 Candidate at Vanderbilt University.
-He develops and applies machine learning and simulation methods to better predict behavior in complex systems.
-His research involves collaboration with social scientists, environmental scientists, computer scientists and engineers.
-His email address is \email{john.j.nay@vanderbilt.edu}, and website: \href{https://johnjnay.com/}{http://johnjnay.com/}.\\
+He develops and applies machine learning and simulation methods to better predict behavior in complex systems and improve policy.
+His research involves collaboration with social scientists, environmental scientists, and engineers.
+His email address is \email{john.j.nay@vanderbilt.edu}, and website: \href{http://johnjnay.com/}{http://johnjnay.com/}.\\
 

### 2016-03-28 prediction_market_wsc.Rnw
typo

@@
     \item Traders then decide of their selling price $p_i^S$ and buying price $p_i^B$. 
-  To do so, traders first compute their expected values $E(s_i^B)$ and $E(s_i^S)$ for securities $s_i^B$ and $s_i^S$ (where expected values are with respect $i$'s approximate model at time $t$). 
+  To do so, traders first compute their expected values $E(s_i^B)$ and $E(s_i^S)$ for securities $s_i^B$ and $s_i^S$ (where expected values are with respect to $i$'s approximate model at time $t$). 
   Then traders set $p_i^S$ at random above $E(s_i^S)$ and $p_i^B$ at random below $E(s_i^B)$ (see Model Parameters below for more details).

### 2016-03-28 master.R
Update master.R

@@ -14,8 +14,7 @@ if (length(args) < 2) {
 message(paste("Number of cores", cores))
 
 source("main.R")
-#devtools::install_github("JohnNay/eat", 
-#                          auth_token = "08d34f040cbe8c95d89477741ceb450a9cfa42c4")
+#devtools::install_github("JohnNay/eat")
 library(eat)
 
 test <- FALSE

### 2016-03-28 README.Rmd
pre-AGU work on time series.

@@
 
-This is code for an agent-based model (ABM) of a climate prediction market in which traders adapt their belief about the climate based on the monetary performance of their neighboring traders in a social network.
-The ABM is meant to test whether climate prediction markets foster the convergence of people’s beliefs about the climate.
+# Betting and Belief: Prediction Markets and Attribution of Climate Change
 
@@
 Gilligan](https://my.vanderbilt.edu/jonathangilligan/) (Earth and Environmental Sciences, Vanderbilt University).
+
+Despite much scientific evidence, a large fraction of the American public doubts that greenhouse gases are causing global warming.
+We present a simulation model as a computational test-bed for climate prediction markets.
+Traders adapt their beliefs about future temperatures based on the profits of other traders in their social network.
+We simulate two alternative climate futures, in which global temperatures are primarily driven either by carbon dioxide or by solar irradiance.
+These represent, respectively, the scientific consensus and a hypothesis advanced by prominent skeptics.
+We conduct sensitivity analyses to determine how a variety of factors describing both the market and the physical climate may affect traders' beliefs about the cause of global climate change.
+Market participation causes most traders to converge quickly toward believing the "true" climate model, suggesting that a climate market could be useful for building public consensus.
+
+This repository contains all the code and data for this simulation model.
+
+Under most parameterizations, within 15-20 years, the median fraction of traders believing in the true model is over 75%:
+
+```{r time, eval=TRUE, echo=FALSE, message=FALSE, fig.width=10.5, fig.height=8.5}
+source("utils/multi_plot.R")
+source("utils/time_plot.R")
+theme_set(theme_classic())
+p1 <- history_box() + theme(text=element_text(family='Palatino'))
+p2 <- future_box() + theme(text=element_text(family='Palatino'))
+multi_plot(p1, p2, cols=2)
+```

### 2016-03-28 prediction_market_wsc.tex

@@
 On the other hand, if climate change is human-induced but not recognized as so, the costs of inaction could be devastating.
-Given that effective climate policies require long-lasting measures to be taken, the importance of an accurate consensus on the issue is manifest.
+Given that effective climate policies require long-lasting measures to be taken, the importance of a timely and accurate consensus on the issue is manifest.
 
-Attempts to foster such consensus face many social and psychological challenges, some of which could be adressed by creating climate prediction markets where participants can ``put their money where their mouths are'' \shortcite{Hsu2011,Vandenbergh2013f}.
+Attempts to foster such consensus face many social and psychological challenges, some of which could be addressed by creating climate prediction markets where participants can ``put their money where their mouths are'' \shortcite{Hsu2011,Vandenbergh2013f}.
 The idea of using prediction markets to efficiently aggregate information about uncertain event outcomes has been widely discussed \shortcite{Horn2014c}.
 Prediction markets have interesting theoretical properties \shortcite{Set1998,Hanson2012}, and perform well in terms of prediction accuracy and information aggregation in experiments \shortcite{Hanson2006,Healy2010} simulation models \shortcite{Klingert2012c,Jumadinova2011a} and the real-world \shortcite{Wolfers2006,pathak_comparison_2015,Dreber_2015}.
+However, to the best of our knowledge, the idea that prediction markets can generate consensus \emph{on the factors\/} affecting uncertain events has never been explored quantitatively.
 
@@
 \begin{enumerate}
-\item  recalibrate their approximate model based on the new set of temperature data available at $t$,
+\item  re-calibrate $m$, their approximate model at time $t$, based on the new set of temperature data available at $t$,
 \label{step:calibrate}
@@
   \begin{itemize}
-    \item Every trader $i$ choses at random a security $s_i^B$ she will try to buy.
+    \item Every trader $i$ chooses at random a security $s_i^B$ she will try to buy.
     \item Every trader $i$ also chooses at random a security $s_i^S$ she will try to sell among the securities she owns a positive amount of (if any).
     \item Traders then decide of their selling price $p_i^S$ and buying price $p_i^B$. 
-  To do so, traders first compute their expected values $E(s_i^B)$ and $E(s_i^S)$ for securities $s_i^B$ and $s_i^S$. 
+  To do so, traders first compute their expected values $E(s_i^B)$ and $E(s_i^S)$ for securities $s_i^B$ and $s_i^S$ (where expected values are with respect $i$'s approximate model at time $t$). 
   Then traders set $p_i^S$ at random above $E(s_i^S)$ and $p_i^B$ at random below $E(s_i^B)$ (see Model Parameters below for more details).
@@
     If \texttt{ideo} is high, traders will not revise their approximate models easily, even when faced with evidence that their richest neighbor is doing better than them. 
-    Formally, for each trader $i$ and each sequence, a parameter $d_{i}$ is drawn from $[0, \texttt{ideo}_i]$. 
-    The value of $d_{i}$ is the probability that $i$ adopts the approximate model of her richest neighbor if that neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
+    Formally, for each trader $i$ and each sequence, $\texttt{ideo}_i$ is the probability that $i$ adopts the approximate model of her richest neighbor if that neighbor is doing better than $i$ at the end of the sequence (in monetary terms).
   \end{itemize}
@@
 This is more relevant for policy design and more interesting theoretically due to the increasing divergence of global temperature values under the two future scenarios.
-The sensitivity analysis is based on a Latin hypercube sampling of parameter sets from the following distributions \shortcite{beachkofski_improved_2002,carnell_lhs_2012}.
+The sensitivity analysis is based on a Latin hypercube sampling of 500 parameter sets from the following distributions \shortcite{beachkofski_improved_2002,carnell_lhs_2012}.
 
@@
 
-We use the model to perform 10,000 simulations for each input paramter set, and then conducted a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
+We use the model to perform 10 full simulations for each of the 500 input parameter sets and average the 10 convergence scores.
+We conducted multiple simulations for the same parameter set because there is stochasticity.
+We then conducted a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of mean belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
 Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).

### 2016-03-28 prediction_market_wsc.Rnw
updated site
Paper work.

@@
 For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is a parameter we vary in our experiments.
+
 \begin{figure}[t]
@@
 Although traders can change their approximate model over time, the connections between traders do not change as the market unfolds, i.e the edges are fixed.
+We vary the degree of initial network segregation in our experiments.
 
@@
 Traders can only predict ${T}_{t^*}$ using their approximate model and their knowledge of ${T}_{0:t}$, $\text{CO2}_{0:t^*}$, and $\text{TSI}_{0:t^*}$.
-Notice that because ${T}_{0:t}$, $\text{CO2}_{0:t^*}$, and $\text{TSI}_{0:t^*}$ are common knowledge, in each period $t$, any two trader with the same approximate model $m$ forms the same stochastic beliefs about future temperatures $P_{t,m}({T}_{t^*} ~|~ {T}_{0:t}, \text{CO2}_{0:t^*}, \text{TSI}_{0:t^*})$.
-The probability distribution $P_{t,m}$ incorporates both epistemic uncertainty (the trader does not know the true values of the coefficients of the climate model) and aleatory uncertainty (in addition to deterministic warming, the temperature exhibits stochastic noise).
+Notice that because ${T}_{0:t}$, $\text{CO2}_{0:t^*}$, and $\text{TSI}_{0:t^*}$ are common knowledge, in each period $t$, any two traders with the same approximate model $m$ form the same stochastic beliefs about future temperatures $P_{t,m}({T}_{t^*} ~|~ {T}_{0:t}, \text{CO2}_{0:t^*}, \text{TSI}_{0:t^*})$.
+The probability distribution $P_{t,m}$ incorporates both epistemic uncertainty (the trader does not know the true values of the coefficients of the climate model) and aleatory uncertainty (in addition to deterministic warming, the temperature exhibits stochastic noise that is directly modeled).
 
@@
    These orders specify that $i$ is willing to buy $s_i^B$ at any price below $p_i^B$, and to sell $s_i^S$ at any price above $p_i^S$.
-    \item The market maker then tries to match $i$'s orders with some order which was put in the book before $i$ came to the market.
+    \item The market maker attempts to match $i$'s orders with some order which was put in the book before $i$ came to the market.
     \item If there are outstanding sell offers for $s_i^B$ at a price below $p_i^B$, a trade is concluded. 
@@
    Trader $i$ sells one unit to the buyer who buys at the highest price above $p_i^S$, and the sell and buy offers are removed from the order book.
-    \item Whenever all traders have come to the market, any remaining outstanding offer is removed from the order book, and the trading period is concluded. 
+    \item When all traders have come to the market, any remaining outstanding offer is removed from the order book, and the trading period is concluded. 
   \end{itemize}
@@
   \item  receive 1 ECU per unit of $s^*$ they own, and
-  \item consider adapting their neighbors' approximate model as described in Social Network above.
+  \item consider adapting their neighbors' approximate model as described in the behavioral parameters sub-section below.
 \end{enumerate}
@@
   \end{itemize}
+
   \item Network parameters:
@@
     $[\texttt{reserv}_{i}, \allowbreak (1  + \texttt{risk.tak}_i) \texttt{reserv}_{i,t}]$ for selling.
+
   \item  \texttt{ideo} :  determines the degree of ``ideology'' of traders.
@@
 We use the model to perform 10 full simulations for each of the \Sexpr{n_sims} input parameter sets and average the 10 convergence scores.
-We conducted multiple simulations for the same parameter set because there is stochasticity.
+We conducted multiple simulations for the same parameter set because there is stochasticity in the tempertature time series, the social network structure, and the agent decision models.
 We then conducted a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of mean belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
 Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).
-The only difference between the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships: every value, $y_s$, in the data set is replaced by a number
-$f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, 
-where $y_r$ are the observed values of $y$ in the data set.
+The difference between the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships: every value, $y_s$, in the data set is replaced by $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$,
+where $y_r$ are the observed values of $y$ in the data set. We conduct 1,000 bootstrapped estimations of the partial rank correlation coefficients to obtain 95\% confidence intervals.
 
@@
 True model refers to the true climate model being \\ce{CO2}, so traders converge to the true model more quickly when carbon dioxide drives climate change than when total solar irradiance does.
-\\textbf{TODO: John: explain the error bars: one sigma, two sigma, 95\\% CI, \\dots}", fig.width=3, fig.height=3.5>>=
+Bars represent 95\\% confidence intervals.", fig.pos='t', fig.width=3, fig.height=3.5>>=
 source("utils/sa_plot.R")
@@
 
-What is not clear to us is why increasing the number of traders slows convergence, even if the number of edges per trader remains fixed. This will be a topic for future research.
+What is not clear to us is why increasing the number of traders slows convergence, even if the number of edges per trader remains fixed.
+This will be a topic for future research where we will be increasing the number of traders into the thousands.
 
@@
 
-All code for the model is available at \href{http://johnjnay.com/predictMarket}{\nolinkurl{johnjnay.com/predictMarket}}.
+All code for the model is available at \href{http://johnjnay.com/predMarket/}{johnjnay.com/predMarket/}.
 This project is a computational test-bed for public policy design: our code can be extended to test the effects that trading strategies, cognitive models, future climate scenarios, and market designs have on the evolution of trader beliefs.
@@
 Candidate at Vanderbilt University.
-He develops and applies machine learning and simulation methods to better predict behavior in complex systems and improve policy.
+He develops machine learning and computer simulation approaches to better predict behavior in complex systems and improve policy.
 His research involves collaboration with social scientists, environmental scientists, and engineers.

### 2016-03-28 prediction_market_wsc.tex
Paper work.

@@
 We use the model to perform 10 full simulations for each of the 500 input parameter sets and average the 10 convergence scores.
-We conducted multiple simulations for the same parameter set because there is stochasticity.
+We conducted multiple simulations for the same parameter set because there is stochasticity in the tempertature time series, the social network structure, and the agent decision models.
 We then conducted a partial rank correlation coefficient analysis on the relationship between the input matrix, $X$, and the resulting simulated outcome vector of mean belief convergence scores, $y$ \shortcite{Marino2008,pujol_sensitivity:_2014,saltelli_sensitivity_2009}.
 Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).
-The only difference between the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships: every value, $y_s$, in the data set is replaced by a number
-$f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$, 
-where $y_r$ are the observed values of $y$ in the data set.
+The difference between the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships: every value, $y_s$, in the data set is replaced by $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$,
+where $y_r$ are the observed values of $y$ in the data set. We conduct 1,000 bootstrapped estimations of the partial rank correlation coefficients to obtain 95\% confidence intervals.
 
 \begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[t]
 
@@
 
-\caption{Estimated effects of model parameters on convergence of beliefs in future scenario.
+\caption[Estimated effects of model parameters on convergence of beliefs in future scenario]{Estimated effects of model parameters on convergence of beliefs in future scenario.
 Positive numbers accelerate convergence and negative numbers retard it.
 True model refers to the true climate model being \ce{CO2}, so traders converge to the true model more quickly when carbon dioxide drives climate change than when total solar irradiance does.
-\textbf{TODO: John: explain the error bars: one sigma, two sigma, 95\% CI, \dots}}\label{fig:sa}
+Bars represent 95\% confidence intervals.}\label{fig:sa}
 \end{figure}

### 2016-03-28 prediction_market_wsc.Rnw
small edit
Bio Martin

@@
 
-What is not clear to us is why increasing the number of traders slows convergence, even if the number of edges per trader remains fixed.
-This will be a topic for future research where we will be increasing the number of traders into the thousands.
+What is not clear to us is why increasing the number of traders slows convergence, even if the number of edges per trader remains fixed. This is a topic for future research.
 
@@
 
-\noindent {\bf MARTIN VAN DER LINDEN} is a Ph.D. Candidate at Vanderbilt University. \\
-\textbf{TODO: Martin, please fill this in.}\\
+\noindent {\bf MARTIN VAN DER LINDEN} is a Ph.D. Candidate at Vanderbilt University. He studies the design and performances of market and institutions using tools from economic theory and game theory.
 His email address is \email{martin.van.der.linden@vanderbilt.edu}, and website: \href{https://martinvdlinden.wordpress.com/}{https://martinvdlinden.wordpress.com/}.\\

### 2016-03-28 utils/sa_plot.R
Updated SA analysis for PRCC and changed plot ylab to match.

@@ -5,7 +5,12 @@ library(tidyr)
 sa_plot <- function(){
   source('utils/pc_sa.R', chdir = T)
   load("output/src_future.Rda")
-  ss <- correct_bias(src_future@result[[7]])
+  
+  load("output/src_future.Rda")
+  result <- sensitivity::pcc(X = src_future@input_set, y = src_future@sims, nboot = 1000, rank = TRUE)
+  ss <- correct_bias(result[[7]])
+  
+  #ss <- correct_bias(src_future@result[[7]])
   ss <- ss %>% mutate(var = ordered(var, levels = c('ideo','risk.tak','n.traders',
                                                     'n.edg', 'seg', 'true.model'),
                                     labels = c('Ideology', "Risk tolerance",
@@ -18,7 +23,7 @@ sa_plot <- function(){
     geom_hline(yintercept = 0, size = 1, color = 'dark gray') +
     geom_point(size = 2) +
     geom_errorbar(aes(ymax = max_ci, ymin = min_ci), width=0.25, size = 1) +
-    labs(x = NULL, y = "Standardized rank regression coefficient",
+    labs(x = NULL, y = "Partial Rank Correlation Coefficient",
          title = title) +
     ylim(c(-1,1)) +
     theme(panel.grid.major.x = element_blank(), 

### 2016-03-29 prediction_market_wsc.Rnw
Updated SA analysis for PRCC and changed plot ylab to match.
Fixed some typos and did a little wordsmithing.
Moved redundant discussion of historical model runs from DISCUSSION to Historical Climate.

@@
 Although the scientific consensus on the anthropogenic nature of climate change strongly increased, beliefs about climate change did not evolve much within the public \shortcite{Vandenbergh2013f}.
-In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily as the question is becoming increasingly politicized and potentially disconnected from scientific evidence.
+In addition, the divide on anthropogenic climate change between liberals and conservatives has grown steadily as the question is becoming increasingly politicized and potentially disconnected from scientific evidence \shortcite{Kahan_2011}.
 The costs of misinformed climate policies are high.
 If climate change is not human-induced but is believed to be so, public resources will be spent on unnecessary efforts.
-On the other hand, if climate change is human-induced but not recognized as so, the costs of inaction could be devastating.
-Given that effective climate policies require long-lasting measures to be taken, the importance of a timely and accurate consensus on the issue is manifest.
+On the other hand, if climate change is human-induced but recognized to be so, the costs of inaction could be devastating.
+Effective climate policies require acting quickly, so it would be valuable to bring the public to a prompt and accurate consensus on the issue.
 
@@
 
-Bloch, Annan and Bowles \shortciteN{Bloch_2010} have proposed using derivatives markets to reduce the scientific uncertainty in estimating the impact of future climate change.
+\citeN{Bloch_2010} have proposed using derivatives markets to reduce the scientific uncertainty in estimating the impact of future climate change.
 Existing prediction markets (e.g.
-hypermind.com, betfair.com, and predictit.org) focus on near term events such as elections months away, so it is difficult to extrapolate empirical findings to the climate case.
+\href{https://hypermind.com/hypermind/app.html}{hypermind}, \href{https://www.betfair.com/exchange/}{betfair}, and \href{https://www.predictit.org/}{PredictIt}) focus on near term events such as elections months away, so it is difficult to extrapolate empirical findings to the climate case.
 Furthermore, we are interested in investigating the unobservable beliefs of traders.
@@
 From a public policy perspective, changing the explanatory models of market participants is one of the most important roles that prediction markets might play.
-This is perhaps the most important social benefit of prediction markets given the comparable predictive power of statistical supervised learning models, which can provide information at a much lower cost than creating and maintaining a market \shortcite{goel_prediction_2010}.
+This is perhaps the most important social benefit of prediction markets given the predictive power of statistical supervised learning models, which could possibly provide information at a much lower cost than creating and maintaining a market \shortcite{goel_prediction_2010}.
 Effective climate policy not only requires an accurate consensus on future climate \emph{outcomes\/}.
@@
 
-Traders bet on global temperature anomalies six years in the future. 
+In our model, traders bet on global temperature anomalies six years in the future. 
 During the six-year period, traders buy and sell futures. 
@@
 For years from 1880--2014 we use the GISTEMP global mean land-sea annual temperature anomalies \shortcite{GISTEMP_2016,Hansen_2010} and for years from 2015 onward, we project future climates under two alternative theories: In both theories, changes in global temperature are proportional to changes in a deterministic forcing plus a stochastic noise term.
-For simplicity, we choose two alternative expressions for the deterministic climate forcing: one, which corresponds to conventional climate science, takes the natural logarithm of the atmospheric carbon dioxide concentration in parts per million \shortcite{Archer_2012}[p.~37], and the other, which corresponds to an alternative theory advocated by many who doubt or reject conventional climate science, takes the total solar irradiance (the brightness of sunlight, in Watts per square meter, at the top of the atmosphere) averaged over the 11-year sunspot cycle \shortcite{Soon_2005}.
+For simplicity, we choose two alternative expressions for the deterministic climate forcing: one, which corresponds to conventional climate science, takes the natural logarithm of the atmospheric carbon dioxide concentration in parts per million \shortcite[p.~37]{Archer_2012}, and the other, which corresponds to an alternative theory advocated by many who doubt or reject conventional climate science, takes the total solar irradiance (the brightness of sunlight, in Watts per square meter, at the top of the atmosphere) averaged over the 11-year sunspot cycle \shortcite{Soon_2005}.
 Most scientific models of the earth's climate include many forcings, including carbon dioxide, other greenhouse gases, aerosols, total solar irradiance, and more.
-In these models, the changing $\ce{CO2}$ concentration is, by a large margin, the strongest forcing \shortcite{IPCC_WG1_AR5_2013}[p.~14], so for simplicity we take this as the model.
-Choosing only one forcing term for each of the competing models simplifies comparison because each has the same number of adjustable parameters.
+In these models, the changing $\ce{CO2}$ concentration is, by a large margin, the strongest single forcing \shortcite[p.~14]{IPCC_WG1_AR5_2013}, so for simplicity we consider only \ce{CO2}.
+Choosing only one forcing term for each of the competing models simplifies comparison because each model has the same number of adjustable parameters.
 
-For $\ce{CO2}$ we used historical emissions through 2005, harmonized with RCP 8.5 representative concentration pathway from 2005 onward \shortcite{rcp_database_2009,Riahi_2011} and for TSI we used the harmonized historical values, with a projection through 2100 from \shortciteN{Velasco_Herrera_2015}, which seems to be the best TSI model.
+For $\ce{CO2}$ we used historical emissions through 2005, harmonized with RCP 8.5 representative concentration pathway from 2005 onward \shortcite{rcp_database_2009,Riahi_2011} and for TSI we used the harmonized historical values, with a projection through 2100 from \shortciteN{Velasco_Herrera_2015}, which is, to our knowledge, the only prediction of TSI for the entire 21\textsuperscript{st} century.
 
@@
 The noise model was determined by fitting an $\ARMA(p,q)$ model to the residuals from the regression, using the \href{http://mc-stan.org}{Stan} probabilistic modeling language and the \href{https://cran.r-project.org/web/packages/rstan/}{\texttt{rstan}} package \shortcite{Stan2016,RStan2016}.
+Stan proved more numerically stable than the \Rstat\ \href{http://CRAN.R-project.org/package=nlme}{\texttt{nlme}} package \shortcite{nlme_2016} for fitting ARMA noise models.
 We identified the optimal model for the auto-correlated noise term by performing the regression analysis for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal noise model \shortcite{Vehtari_2015}.
@@
 
-Future climates were generated by applying the warming coefficients to Eq.~\ref{eq:climate}:
+Future climates were generated by applying the future climate forcings to Eq.~\ref{eq:climate}:
 \begin{equation}
@@
 $\beta_{\text{model}}$ is the regression coefficient, and
-$\varepsilon$ is an $\ARMA(p,q)$ noise term.
+$\varepsilon$ is a noise term.
 The coefficient $\beta$ and the parameters of the ARMA noise model were fit to the historical data (1880--2014).
@@
 
-Based on traders' sell and buy orders, traders exchange securities following a continuous-double auctions (CDA) procedure (see Model Dynamics below for more details). 
+Based on traders' sell and buy orders, traders exchange securities following a continuous-double auction (CDA) procedure (see Model Dynamics below for more details). 
 CDAs or some close variants are common procedures to match buy and sell orders. 
@@
 Then, the trader considers adopting the model of her richest neighbor. 
-For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is, which is a parameter we vary in our experiments.
+For each trader, the willingness to revise her belief is determined by how ideologically loaded her belief is \shortcite{Kahan_2011}, which is a parameter we vary in our experiments.
 
@@
 An example of a snapshot of a social network is depicted in Fig.~\ref{fig:social_network}. 
-We assume that the initial network is segregated: traders who initially believe in \ce{CO2} forcing (resp.
-TSI) are more likely to be linked with other traders who also believe in \ce{CO2} (resp.
-TSI).
-This reflects reality where cultural and social groups are highly correlated with climate change beliefs. 
-Although traders can change their approximate model over time, the connections between traders do not change as the market unfolds, i.e the edges are fixed.
-We vary the degree of initial network segregation in our experiments.
+A segmentation parameter controls the homophily of the network: the extent to which traders are preferentially linked with other traders who share their initial belief in the cause of climate change (\ce{CO2} or TSI), as opposed to traders with the opposite intial belief.
+This parameter varies from 0 (no preference for like-minded traders) to 1 (traders are only connected to like-minded traders).
+%This reflects reality where cultural and social groups are highly correlated with climate change beliefs. 
+Although traders can change their beliefs over time, the connections between traders do not change as the market unfolds, i.e the edges are fixed. 
+We vary the segmentation parameter in our experiments.
 
@@
     \item  \texttt{n.edg}: the number of edges in the social network.
-    \item  \texttt{seg}: the initial degree of homophily in the network. 
-    The higher  \texttt{seg}, the higher the initial homophily. 
+    \item  \texttt{seg}: the segmentation parameter for the social network. 
   \end{itemize}
@@
 
-As a validation that the market is operating correctly, we run the market with actual temperatures 1880 to 2015 and market betting from 1931 to 2014.
-Before 1980 there is no statistically significant warming trend so we would not expect the market to change beliefs before then.
-As a global warming signal begins to show up in 1980s, traders begin to converge toward the log(\ce{CO2}) model (Fig.~\ref{fig:time}).
+As a validation that the market is operating correctly, we ran the market using actual historical temperatures from 1880 to 2014, with market betting from 1931 to 2014.
+Greenhouse gas concentrations did not become high enough to begin dominating natural climate variations until the 1970s or so, and it took until the 1990s for the temperature record to show clear signs of anthropogenic interference \shortcite{IPCC_WG1_AR5_2013}.
+Consistent with this, our simulated historical trading sequence (Fig.~\ref{fig:time}) shows convergence to belief in the TSI model until the early 1970s, after which traders converge toward believing the \ce{CO2} model.
+<<time, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Convergence over trading sequences for different degrees of social-network connection and segmentation.", fig.width=10.5, fig.height=8.5, fig.pos='t'>>=
+source("utils/time_plot.R")
+p1 <- history_box() + theme(text=element_text(family='Palatino'))
+p2 <- future_box() + theme(text=element_text(family='Palatino'))
+multi_plot(p1, p2, cols=2)
+@
 
@@
 
-We use the model to perform 10 full simulations for each of the \Sexpr{n_sims} input parameter sets and average the 10 convergence scores.
+We used the model to perform 10 full simulations for each of the \Sexpr{n_sims} input parameter sets and average the 10 convergence scores.
 We conducted multiple simulations for the same parameter set because there is stochasticity in the tempertature time series, the social network structure, and the agent decision models.
@@
 Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).
-The difference between the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first ranked-transformed in order to allow us capture potentially non-linear relationships: every value, $y_s$, in the data set is replaced by $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$,
+The difference between the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first rank-transformed in order to allow us capture potentially non-linear relationships: every value, $y_s$, in the data set is replaced by $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$,
 where $y_r$ are the observed values of $y$ in the data set. We conduct 1,000 bootstrapped estimations of the partial rank correlation coefficients to obtain 95\% confidence intervals.
@@
 We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and conducted an experiment crossing 95\textsuperscript{th} and 5\textsuperscript{th} percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model.
-We then collected the time series of convergence across these eight designs for visualization of the distributions of convergence over time (Fig.~\ref{fig:time}).
-Under most parameterizations, within 15--20 years, the median fraction of traders believing in the true model is over 75\%.
-
-<<time, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Convergence over trading sequences for different degrees of social-network connection and segmentation.", fig.width=10.5, fig.height=8.5, fig.pos='t'>>=
-source("utils/time_plot.R")
-p1 <- history_box() + theme(text=element_text(family='Palatino'))
-p2 <- future_box() + theme(text=element_text(family='Palatino'))
-multi_plot(p1, p2, cols=2)
-@
+We collected the time series of belief-convergence across these eight designs to visualize the distributions of convergence over time (Fig.~\ref{fig:time}).
+Under most parameterizations, the median fraction of traders believing in the true model reaches 75\% in 10--20 years.
 
@@
 
-The historical sequence in Fig.~\ref{fig:time} shows convergence to the TSI model until the early 1970s, after which traders converge toward the \ce{CO2} model.
-This is consistent with the fact that greenhouse gas concentrations did not become high enough to begin dominating natural climate variations until the 1970s or so, and it took until the 1990s for the temperature record to show clear signs of anthropogenic interference \shortcite{IPCC_WG1_AR5_2013}.
-
 \section{CONCLUSIONS}
 
-We simulate two alternative climate futures: one where the natural log of \ce{CO2} is the primary driver of global temperature and one where sun spots are the primary driver.
+We simulate two alternative climate futures: one where \ce{CO2} is the primary driver of global temperature and one where variations in solar intensity are the primary driver.
 These represent the two most plausible competing views in the public discourse and our analysis is agnostic about which is ``true''.
@@
 
-All code for the model is available at \href{http://johnjnay.com/predMarket/}{johnjnay.com/predMarket/}.
+All code and data for the model is available at \href{http://johnjnay.com/predMarket/}{johnjnay.com/predMarket/}.
 This project is a computational test-bed for public policy design: our code can be extended to test the effects that trading strategies, cognitive models, future climate scenarios, and market designs have on the evolution of trader beliefs.
@@
 We thank Manuel Naumovich Velasco for sharing the data from his Total Solar Irradiance prediction.
-We also thank the authors of \href{http://www.r-project.org}{\textsf{R}} \shortcite{r_core}
-and 
-\href{http://mc-stan.org}{Stan} \shortcite{Stan2016}.
-This manuscript was prepared using 
+%We also thank the authors of \href{http://www.r-project.org}{\textsf{R}} \shortcite{r_core}
+%and 
+%\href{http://mc-stan.org}{Stan} \shortcite{Stan2016}.
+%This manuscript was prepared using 
 %RStudio \shortcite{rstudio} and 
-knitr \shortcite{xie_knitr_2014}.
+%knitr \shortcite{xie_knitr_2014}.
 

### 2016-03-29 arxiv.Rnw
started arxiv rnw but errors for tex for the citations, e.g. shortcite{

--- 2016-03-29 prediction_market_wsc.Rnw
@@
-%**************************************************************************
-%*
-%*  Paper: ``INSTRUCTIONS FOR AUTHORS OF LATEX DOCUMENTS''
-%*
-%*  Publication: 2016 Winter Simulation Conference Author Kit
-%*
-%*  Filename: wsc16paper.tex
-%*
-%*  Date: January 31, 2001   Time:  9:45 PM
-%*      BASE of current version: Feb 01, 2010 (primary WSC'10 LaTeX file)
-%*
-%*  Word Processing System: TeXnicCenter and MiKTeX
-%*
-%*
-%*  All files need the following
-\input{wsc16style.tex}     % download from author kit.  Style files for wsc formatting. Don't remove this line - required for generating the final paper!
+%\input{wsc16style.tex}     % download from author kit.  Style files for wsc formatting. Don't remove this line - required for generating the final paper!
 
-\documentclass{wscpaperproc}
+\documentclass{article}
 
@@
 
-%% The next versions of the hyperref command are used if you adopt the
-%% outdated latex-dvips-ps2pdf route in generating your pdf file. In
-%% this case you can use ps or eps files for graphics, but not pdf, jpeg, png etc.
-%% However, the final pdf file should embed all fonts required which means that you have to use file
-%% formats which can embed fonts. Please note that the final PDF file will not be generated on your computer!
-%% If you are using WinEdt or PCTeX, then use the following. If you are using
-%% Y&Y TeX then replace "dvips" with "dvipsone"
-
-%%\usepackage[dvips,colorlinks=true,urlcolor=blue,citecolor=black,%
-%% anchorcolor=black,linkcolor=black]{hyperref}
-%****************************************************************************
-
   \usepackage{latexsym}
@@
 
-%#########################################################
-%*
-%*  The Document.
-%*
 \begin{document}
 
-%***************************************************************************
-% AUTHOR: AUTHOR NAMES GO HERE
-% FORMAT AUTHORS NAMES Like: Author1, Author2 and Author3 (last names)
-%
-%  You need to change the author listing below!
-%               Please list ALL authors using last name only, separate by a comma except
-%               for the last author, separate with "and"
-%
-\WSCpagesetup{Nay, Van der Linden, and Gilligan}
-
-% AUTHOR: Enter the title, all letters in upper case
-\title{BETTING AND BELIEF: PREDICTION MARKETS AND ATTRIBUTION OF CLIMATE CHANGE}
-%Betting and Belief: Prediction Markets and Attribution of Climate Change
+\title{Betting and Belief: Prediction Markets and Attribution of Climate Change}
 
 % AUTHOR: Enter the authors of the article, see end of the example document for further examples
-\author{John J. Nay\\ [12pt]
-School of Engineering\\
-Vanderbilt University \\
-PMB 351826 \\
-2301 Vanderbilt Place\\
-Nashville, TN 37235-1826, USA
-% Multiple authors are entered as follows.
-% You may also need to adjust the titlevbox size in the preamble - search for titlevboxsize
-\and
-Martin Van der Linden\\ [12pt]
-Department of Economics\\
-Vanderbilt University\\
-PMB 351819\\
-2301 Vanderbilt PL\\
-Nashville, TN 37235-1819, USA\\
-\and
-Jonathan M. Gilligan\\ [12pt]
-Department of Earth \& Environmental Sciences\\
-Vanderbilt University\\
-PMB 351805\\
-2301 Vanderbilt PL\\
-Nashville, TN 37235-1805, USA
+\author{John J. Nay, Martin Van der Linden,
+Jonathan M. Gilligan
 }
@@
 
-\section*{AUTHOR BIOGRAPHIES}
-
-\noindent {\bf JOHN J.
-NAY} is a Ph.D.
-Candidate at Vanderbilt University.
-He develops machine learning and computer simulation approaches to better predict behavior in complex systems and improve policy.
-His research involves collaboration with social scientists, environmental scientists, and engineers.
-His email address is \email{john.j.nay@vanderbilt.edu}, and website: \href{http://johnjnay.com/}{http://johnjnay.com/}.\\
-
-\noindent {\bf MARTIN VAN DER LINDEN} is a Ph.D. Candidate at Vanderbilt University. He studies the design and performances of market and institutions using tools from economic theory and game theory.
-His email address is \email{martin.van.der.linden@vanderbilt.edu}, and website: \href{https://martinvdlinden.wordpress.com/}{https://martinvdlinden.wordpress.com/}.\\
-
-\noindent {\bf JONATHAN M.
-GILLIGAN} is Associate Professor of Earth and Environmental Sciences at Vanderbilt University and Associate Director for Research at the Vanderbilt Climate Change Research Network.
-His research and teaching interests center on integrating natural science, social science, behavioral science, and engineering to study coupled human-natural systems.
-He holds a  Ph.D.
-in Physics from Yale University.
-His email address is \email{jonathan.gilligan@vanderbilt.edu}, and website: \href{https://my.vanderbilt.edu/jonathangilligan/}{https://my.vanderbilt.edu/jonathangilligan/}.
-
 \end{document}

### 2016-03-29 prediction_market_wsc.tex

@@
 
-As a validation that the market is operating correctly, we run the market with actual temperatures 1880 to 2015 and market betting from 1931 to 2014.
-Before 1980 there is no statistically significant warming trend so we would not expect the market to change beliefs before then.
-As a global warming signal begins to show up in 1980s, traders begin to converge toward the log(\ce{CO2}) model (Fig.~\ref{fig:time}).
+As a validation that the market is operating correctly, we ran the market using actual historical temperatures from 1880 to 2014, with market betting from 1931 to 2014.
+Greenhouse gas concentrations did not become high enough to begin dominating natural climate variations until the 1970s or so, and it took until the 1990s for the temperature record to show clear signs of anthropogenic interference \shortcite{IPCC_WG1_AR5_2013}.
+Consistent with this, our simulated historical trading sequence (Fig.~\ref{fig:time}) shows convergence to belief in the TSI model until the early 1970s, after which traders converge toward believing the \ce{CO2} model.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[t]
+
+{\centering \includegraphics[width=\maxwidth]{figure/time-1}
+
+}
+
+\caption[Convergence over trading sequences for different degrees of social-network connection and segmentation]{Convergence over trading sequences for different degrees of social-network connection and segmentation.}\label{fig:time}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-We use the model to perform 10 full simulations for each of the 500 input parameter sets and average the 10 convergence scores.
+We used the model to perform 10 full simulations for each of the 500 input parameter sets and average the 10 convergence scores.
 We conducted multiple simulations for the same parameter set because there is stochasticity in the tempertature time series, the social network structure, and the agent decision models.
@@
 We randomly drew from the above distributions for ideology, risk tolerance, and the number of traders, and conducted an experiment crossing 95\textsuperscript{th} and 5\textsuperscript{th} percentile values for the number of edges per trader and the segmentation of the social network and both values for the true model.
-We then collected the time series of convergence across these eight designs for visualization of the distributions of convergence over time (Fig.~\ref{fig:time}).
-Under most parameterizations, within 15--20 years, the median fraction of traders believing in the true model is over 75\%.
-
-\begin{knitrout}
-\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[t]
-
-{\centering \includegraphics[width=\maxwidth]{figure/time-1}
-
-}
-
-\caption[Convergence over trading sequences for different degrees of social-network connection and segmentation]{Convergence over trading sequences for different degrees of social-network connection and segmentation.}\label{fig:time}
-\end{figure}
-
-\end{knitrout}
+We collected the time series of belief-convergence across these eight designs to visualize the distributions of convergence over time (Fig.~\ref{fig:time}).
+Under most parameterizations, the median fraction of traders believing in the true model reaches 75\% in 10--20 years.
 
@@
 
-What is not clear to us is why increasing the number of traders slows convergence, even if the number of edges per trader remains fixed.
-This will be a topic for future research where we will be increasing the number of traders into the thousands.
-
-The historical sequence in Fig.~\ref{fig:time} shows convergence to the TSI model until the early 1970s, after which traders converge toward the \ce{CO2} model.
-This is consistent with the fact that greenhouse gas concentrations did not become high enough to begin dominating natural climate variations until the 1970s or so, and it took until the 1990s for the temperature record to show clear signs of anthropogenic interference \shortcite{IPCC_WG1_AR5_2013}.
+What is not clear to us is why increasing the number of traders slows convergence, even if the number of edges per trader remains fixed. This is a topic for future research.
 
@@
 
-We simulate two alternative climate futures: one where the natural log of \ce{CO2} is the primary driver of global temperature and one where sun spots are the primary driver.
+We simulate two alternative climate futures: one where \ce{CO2} is the primary driver of global temperature and one where variations in solar intensity are the primary driver.
 These represent the two most plausible competing views in the public discourse and our analysis is agnostic about which is ``true''.

### 2016-03-29 arXiv/arxiv.tex
Get version ready to send to arXiv.

--- 2016-03-29 prediction_market_wsc.tex
@@
-%**************************************************************************
-%*
-%*  Paper: ``INSTRUCTIONS FOR AUTHORS OF LATEX DOCUMENTS''
-%*
-%*  Publication: 2016 Winter Simulation Conference Author Kit
-%*
-%*  Filename: wsc16paper.tex
-%*
-%*  Date: January 31, 2001   Time:  9:45 PM
-%*      BASE of current version: Feb 01, 2010 (primary WSC'10 LaTeX file)
-%*
-%*  Word Processing System: TeXnicCenter and MiKTeX
-%*
-%*
-%*  All files need the following
-\input{wsc16style.tex}     % download from author kit.  Style files for wsc formatting. Don't remove this line - required for generating the final paper!
+%\input{wsc16style.tex}     % download from author kit.  Style files for wsc formatting. Don't remove this line - required for generating the final paper!
 
-\documentclass{wscpaperproc}\usepackage[]{graphicx}\usepackage[]{color}
+\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
 %% maxwidth is the original width if it is less than linewidth
@@
 
-%% The next versions of the hyperref command are used if you adopt the
-%% outdated latex-dvips-ps2pdf route in generating your pdf file. In
-%% this case you can use ps or eps files for graphics, but not pdf, jpeg, png etc.
-%% However, the final pdf file should embed all fonts required which means that you have to use file
-%% formats which can embed fonts. Please note that the final PDF file will not be generated on your computer!
-%% If you are using WinEdt or PCTeX, then use the following. If you are using
-%% Y&Y TeX then replace "dvips" with "dvipsone"
-
-%%\usepackage[dvips,colorlinks=true,urlcolor=blue,citecolor=black,%
-%% anchorcolor=black,linkcolor=black]{hyperref}
-%****************************************************************************
-
+  \usepackage[margin=1in]{geometry}
   \usepackage{latexsym}
@@
 
+\input{wscbib}
 \bibliographystyle{wsc}
@@
 \newcommand{\AR}{\text{AR}}
-
-%#########################################################
-%*
-%*  The Document.
-%*
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
@@
 
-%***************************************************************************
-% AUTHOR: AUTHOR NAMES GO HERE
-% FORMAT AUTHORS NAMES Like: Author1, Author2 and Author3 (last names)
-%
-%  You need to change the author listing below!
-%               Please list ALL authors using last name only, separate by a comma except
-%               for the last author, separate with "and"
-%
-\WSCpagesetup{Nay, Van der Linden, and Gilligan}
-
-% AUTHOR: Enter the title, all letters in upper case
-\title{BETTING AND BELIEF: PREDICTION MARKETS AND ATTRIBUTION OF CLIMATE CHANGE}
-%Betting and Belief: Prediction Markets and Attribution of Climate Change
+\title{Betting and Belief: Prediction Markets and Attribution of Climate Change}
 
 % AUTHOR: Enter the authors of the article, see end of the example document for further examples
-\author{John J. Nay\\ [12pt]
-School of Engineering\\
-Vanderbilt University \\
-PMB 351826 \\
-2301 Vanderbilt Place\\
-Nashville, TN 37235-1826, USA
-% Multiple authors are entered as follows.
-% You may also need to adjust the titlevbox size in the preamble - search for titlevboxsize
-\and
-Martin Van der Linden\\ [12pt]
-Department of Economics\\
-Vanderbilt University\\
-PMB 351819\\
-2301 Vanderbilt PL\\
-Nashville, TN 37235-1819, USA\\
-\and
-Jonathan M. Gilligan\\ [12pt]
-Department of Earth \& Environmental Sciences\\
-Vanderbilt University\\
-PMB 351805\\
-2301 Vanderbilt PL\\
-Nashville, TN 37235-1805, USA
+\author{John J. Nay, Martin Van der Linden,
+Jonathan M. Gilligan
 }
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/climateplot-1}
+{\centering \includegraphics[width=\maxwidth]{climateplot-1}
 
@@
 \begin{center}
-\includegraphics[width=0.50\textwidth]{output/network.pdf}
+\includegraphics[width=0.50\textwidth]{network.pdf}
 \caption{An example of the social network among traders.
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/time-1}
+{\centering \includegraphics[width=\maxwidth]{time-1}
 
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/sa-1}
+{\centering \includegraphics[width=\maxwidth]{sa-1}
 

### 2016-03-31 README.Rmd
update readme with arxiv citation

@@
 
+If you use these scripts, cite this paper:
+
+Nay, John J., Van der Linden, Martin, Gilligan, Jonathan. (2016) "Betting and Belief: Prediction Markets and Attribution of Climate Change." eprint arXiv:1603.08961
+
+The paper can be downloaded here: [http://arxiv.org/abs/1603.08961](http://arxiv.org/abs/1603.08961)
+
 Under most parameterizations, within 15-20 years, the median fraction of traders believing in the true model is over 75%:

### 2016-03-31 prediction_market_wsc.Rnw
One last typo
fixed type in intro of wsc rnw

@@
 If climate change is not human-induced but is believed to be so, public resources will be spent on unnecessary efforts.
-On the other hand, if climate change is human-induced but recognized to be so, the costs of inaction could be devastating.
+On the other hand, if climate change is human-induced but not recognized to be so, the costs of inaction could be devastating.
 Effective climate policies require acting quickly, so it would be valuable to bring the public to a prompt and accurate consensus on the issue.
@@
 
-\noindent {\bf MARTIN VAN DER LINDEN} is a Ph.D. Candidate at Vanderbilt University. He studies the design and performances of market and institutions using tools from economic theory and game theory.
+\noindent {\bf MARTIN VAN DER LINDEN} is a Ph.D. Candidate at Vanderbilt University. He studies the design and performance of markets and institutions using tools from economic theory and game theory.
 His email address is \email{martin.van.der.linden@vanderbilt.edu}, and website: \href{https://martinvdlinden.wordpress.com/}{https://martinvdlinden.wordpress.com/}.\\

### 2016-04-04 presentation/present.Rmd
prelim presentation. small paper typo.

---
title: 'Betting and Belief: Prediction Markets and Attribution of Climate Change'
author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
date: "April 5, 2016"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r prep}
library(ggplot2)
theme_set(theme_classic())
library(pander)
```

## Prediction Markets

```{r}
knitr::include_graphics("markfigs/predict.png")
```

## Climate Change and Prediction Markets

- Scientific consensus on anthropogenic nature of climate change strongly increased.
- BUT, beliefs about climate change did not evolve much within public. Increasingly politicized.
- Prediction markets: participants can "put their money where their mouths are."
- High prediction accuracy and information aggregation.
- Do they also change beliefs?

## Prediction Market Simulation

- Research question: *whether, and under what social and climate conditions, might prediction markets be useful for increasing convergence of climate beliefs?*
- Prediction markets (hypermind.com, betfair.com, and predictit.org) focus on near term events such as elections months away, so difficult to extrapolate to climate case.
- Investigating unobservable beliefs of traders. 
- So, simulation modeling informed by climate and economic theory.

## Two Alternate Realities

```{r}
knitr::include_graphics("markfigs/climateplot-1.png")
```

## Social Network

```{r}
knitr::include_graphics("markfigs/network.png")
```

## Model

1. Trader agents estimate forecasting model with data.
2. Make predictions for climate for 6 yrs from now to assign expected value to temp securities.
3. Trade on market.
4. Payoffs for owners of security that was correct.
5. Decide whether to adopt neighbors model based on cumulative earnings.

## Parameters

- $ideo \sim \text{Uniform}(0,1)$
- $n.edge \sim \text{Uniform}(100,200)$ (mapped into integer)
- $n.traders \sim \text{Uniform}(50,250)$ (mapped into integer)
- $risk.tak  \sim \text{Uniform}(0,1)$
- $seg \sim \text{Uniform}(0,1)$
- $true.model \sim \text{Bernoulli}(0.5)$

## Convergence Over Time

```{r}
knitr::include_graphics("markfigs/time-1.png")
```

## Sensitivity Analysis Design

- Stochasticity in tempertature data, social network structure, and agent decision models.
- 10 full simulations for each of 500 input parameter sets and average.
- Partial rank correlation coefficient analysis on relationship between input matrix, $X$, and resulting simulated outcome vector of mean belief convergence scores, $y$. 
- Partial correlation: linear relationship between part of variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$). 
- To allow potentially non-linear relationships $y$ is first rank-transformed.
- 1,000 bootstrapped estimations of PRCC to obtain 95\% confidence intervals.

## Sensitivity Analysis Results

```{r}
knitr::include_graphics("markfigs/sa-1.png")
```

## References

### 2016-04-18 presentation/present.Rmd
presentation, now creating pdf not html

@@
 author: "John J. Nay, Martin Van der Linden, Jonathan Gilligan"
-date: "April 5, 2016"
-output: slidy_presentation
+date: "April 20, 2016"
+output: beamer_presentation
 ---
@@
 ```
+
+## Trader Beliefs
+
+- Both when true data generating process is CO2 and TSI, at model initialization approximately half of traders use true data-generating model to make predictions. 
+- Traders using true model do not necessarily make perfectly accurate predictions. 
+- Although these traders believe in correct *functional form* of model, they still need to calibrate their model based on limited noisy data.
 

### 2016-04-19 presentation/present.Rmd
presentation with diagrams for model dynamics explanation

--- 2016-04-04 presentation/present.Rmd
@@
 
-- Scientific consensus on anthropogenic nature of climate change strongly increased.
-- BUT, beliefs about climate change did not evolve much within public. Increasingly politicized.
+- Scientific consensus on anthropogenic climate change increased.
+- BUT, beliefs about climate change did not evolve much within public.
 - Prediction markets: participants can "put their money where their mouths are."
-- High prediction accuracy and information aggregation.
+- High prediction accuracy, efficient information aggregation.
 - Do they also change beliefs?
@@
 
-- Research question: *whether, and under what social and climate conditions, might prediction markets be useful for increasing convergence of climate beliefs?*
-- Prediction markets (hypermind.com, betfair.com, and predictit.org) focus on near term events such as elections months away, so difficult to extrapolate to climate case.
+- *Whether, and under what social and climate conditions, might prediction markets be useful for increasing convergence of climate beliefs?*
+- Prediction markets focus on near term events such as elections months away, so difficult to extrapolate to climate case.
 - Investigating unobservable beliefs of traders. 
@@
 
+## Trader Beliefs
+
+- At initialization approximately half of traders use true data-generating model to make predictions. 
+- Traders using true model not necessarily perfectly accurate. 
+- Although correct *functional form* of model, still need to calibrate on limited noisy data.
+
 ## Social Network
@@
 
-1. Trader agents estimate forecasting model with data.
+1. Trader agents estimate forecasting model.
 2. Make predictions for climate for 6 yrs from now to assign expected value to temp securities.
-3. Trade on market.
+3. Trade.
 4. Payoffs for owners of security that was correct.

### 2016-06-10 prediction_market_wsc.Rnw
fixed adapting type
prelim presentation. small paper typo.
Paper edits based on WSC reviewers comments

@@
   \item  receive 1 ECU per unit of $s^*$ they own, and
-  \item consider adapting their neighbors' approximate model as described in the behavioral parameters sub-section below.
+  \item consider adopting their neighbors' approximate model as described in the behavioral parameters sub-section below.
 \end{enumerate}
@@
   \item \texttt{ideo} $\sim \text{Uniform}(0,1)$
-  \item \texttt{market.complet} $\sim \text{Uniform}(0,1000)$ (mapped into integer)
   \item \texttt{n.edge} $\sim \text{Uniform}(100,200)$ (mapped into integer)
@@
 Market participation causes traders to converge toward believing the ``true'' climate model under a variety of model parameterizations in a relatively short time: In markets with low segmentation, the number of traders believing the true model rises from 50\% to 75\% in roughly 12 years, and even in highly segmented markets, 75\% convergence is achieved in 18--24 years.
-This suggests that a climate prediction market could be useful for producing broad agreement about the causes of climate change, and could have persuasive power for people who are not persuaded by the overwhelming consensus among climate researchers that greenhouse gases (and especially \ce{CO2}) are responsible for the majority of observed climate change \shortcite{IPCC_WG1_AR5_2013}.
-However, market segmentation has a large effect on the speed of convergence, so transparency and effective communication about the performance of traders with different beliefs will be important.
+Ideally, we would like to compare belief convergence with and without a prediction market but because the only source of climate belief in our current model is market interactions we cannot make this comparison.
+However, we do use actual temperatures for our model under historical 20th century conditions and observe convergence to the \ce{CO2} model, while in the real world there is no convergence of beliefs.
+
+Both the historical and future simulation results suggest that a climate prediction market could be useful for producing broad agreement about the causes of climate change, and could have persuasive power for people who are not persuaded by the overwhelming consensus among climate researchers that greenhouse gases (and especially \ce{CO2}) are responsible for the majority of observed climate change \shortcite{IPCC_WG1_AR5_2013}.
+We also find that market segmentation has a large effect on the speed of convergence, so transparency and effective communication about the performance of traders with different beliefs will be important.
 
@@
 
-All code and data for the model is available at \href{http://johnjnay.com/predMarket/}{johnjnay.com/predMarket/}.
+All code and data for the model is available at \href{https://github.com/JohnNay/predMarket}{github.com/JohnNay/predMarket}.
 This project is a computational test-bed for public policy design: our code can be extended to test the effects that trading strategies, cognitive models, future climate scenarios, and market designs have on the evolution of trader beliefs.
@@
 
-We thank Yevgeniy Vorobeychik for feedback on this project. 
-We thank Manuel Naumovich Velasco for sharing the data from his Total Solar Irradiance prediction.
+We thank Yevgeniy Vorobeychik for feedback on this project and Manuel Naumovich Velasco for sharing the data from his Total Solar Irradiance prediction. 
+U.S.
+National Science Foundation grants EAR-1416964, EAR-1204685, and 1526860 partially supported this research.
+The funder had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+All views expressed are the authors' and not the funder's.
 %We also thank the authors of \href{http://www.r-project.org}{\textsf{R}} \shortcite{r_core}
@@
 Candidate at Vanderbilt University.
-He develops machine learning and computer simulation approaches to better predict behavior in complex systems and improve policy.
-His research involves collaboration with social scientists, environmental scientists, and engineers.
-His email address is \email{john.j.nay@vanderbilt.edu}, and website: \href{http://johnjnay.com/}{http://johnjnay.com/}.\\
+He develops machine learning and computer simulation approaches to forecast dynamic systems and improve law and policy.
+His research involves collaboration with lawyers, social scientists, environmental scientists, and engineers.
+His email address is \email{john.j.nay@gmail.com}, and website: \href{http://johnjnay.com/}{http://johnjnay.com/}.\\
 

### 2016-06-28 arxiv.Rnw
Updated files for arXiv.
Finished major work on ODD. Ready for proofreading and then I think it's done.

@@
 <<setup, echo=F, results='hide', cache=F>>=
-library(ggplot2)
+library(jgplot2) # devtools::install_github('jonathan-g/jgplot2', ref='jgplot2')
 library(pander)
@@
 
+  \usepackage[margin=1in]{geometry}
   \usepackage{latexsym}
@@
   \usepackage{mathptmx}
-  \usepackage[pdftex,colorlinks=true,urlcolor=blue,citecolor=black,anchorcolor=black,linkcolor=black]{hyperref}
+  \usepackage{hyperref}
   \usepackage{mhchem}
 
+\hypersetup{colorlinks=true,urlcolor=blue,citecolor=black,anchorcolor=black,linkcolor=black}
+
+\input{wscbib}
 \bibliographystyle{wsc}

### 2016-06-28 paper.Rmd
Finished major work on ODD. Ready for proofreading and then I think it's done.

@@
 source("utils/multi_plot.R")
-library(ggplot2)
+# replace ggplot2 with jgplot2 that has my enhancement for outlier.alpha on geom_boxplot.
+# library(ggplot2)
+library(jgplot2) # devtools::install_github('jonathan-g/jgplot2', ref='jgplot2')
 theme_set(theme_classic())

### 2016-06-28 prediction_market_wsc.Rnw
Finished major work on ODD. Ready for proofreading and then I think it's done.

@@
 <<setup, echo=F, results='hide', cache=F>>=
-library(ggplot2)
+# replace ggplot2 with jgplot2 that has my enhancement for outlier.alpha on geom_boxplot.
+# library(ggplot2)
+library(jgplot2) # devtools::install_github('jonathan-g/jgplot2', ref='jgplot2')
 library(pander)

### 2016-06-28 utils/time_plot.R

@@ -1,22 +1,24 @@
 library(dplyr)
 library(tidyr)
-library(ggplot2)
+# replace ggplot2 with jgplot2 that has my enhancement for outlier.alpha on geom_boxplot.
+# library(ggplot2)
+library(jgplot2) # devtools::install_github('jonathan-g/jgplot2', ref = 'jgplot2')
 
 future_line <- function(){
-  
+
   load("output/convergence_future.Rda")
-  
+
   plot_data <- average_convergence
   # plot_data$set <- factor(plot_data$set, levels = gtools::mixedsort(unique(plot_data$set)))
   #  n.edg <- round(parameters[5]*100) + 100 # integer in (100, 200)
-  plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"), 
-                          labels = c(paste0(round(8 * 0.05 + 2, 1), " edges per trader, segmentation = 0.05"), 
-                                     paste0(round(8 * 0.05 + 2, 1), " edges per trader, segmentation = 0.95"), 
-                                     paste0(round(8 * 0.95 + 2, 1), " edges per trader, segmentation = 0.05"), 
+  plot_data$set <- factor(plot_data$set, levels = c("0.05, 0.05", "0.05, 0.95", "0.95, 0.05", "0.95, 0.95"),
+                          labels = c(paste0(round(8 * 0.05 + 2, 1), " edges per trader, segmentation = 0.05"),
+                                     paste0(round(8 * 0.05 + 2, 1), " edges per trader, segmentation = 0.95"),
+                                     paste0(round(8 * 0.95 + 2, 1), " edges per trader, segmentation = 0.05"),
                                      paste0(round(8 * 0.95 + 2, 1), " edges per trader, segmentation = 0.95")))
-  
+
   sequence.years <- 6
-  
+
   ggplot(data=plot_data, aes(x= trading_seq * sequence.years + 2014, y=(convergence + 1) * 50, color = true_mod)) +
     geom_point(position = position_jitter(w = 0.07 * sequence.years, h = 0), alpha = 0.1) + geom_smooth(size=1, fill=NA) +
     scale_color_brewer(palette="Set1", name="True driver of future climate", labels = c(LogCo2 = expression(CO[2]), SlowTSI = "TSI")) +
@@ -26,7 +28,7 @@ future_line <- function(){
          title = paste0("2015-",  2014 + 14 * 6)) +
     ylim(0,100) +
[…]

### 2016-06-28 arxiv.Rnw
Switched from ggplot2 to jgplot2 (my fork of Hadley's ggplot2 which implements outlier.alpha argument to geom_boxplot). If necessary, install from devtools::install_github('jonathan-g/jgplot2', ref='jgplot2') Updated citations to WAIC.

@@
 Warming coefficients for each model were determined by linear regression of historical temperatures from 1880--2014 against the historical values for each model's forcing term.
-The noise model was determined by fitting an $\ARMA(p,q)$ model to the residuals from the regression, using the \href{http://mc-stan.org}{Stan} probabilistic modeling language and the \href{https://cran.r-project.org/web/packages/rstan/}{\texttt{rstan}} package \shortcite{Stan2016,RStan2016}.
-Stan proved more numerically stable than the \Rstat\ \href{http://CRAN.R-project.org/package=nlme}{\texttt{nlme}} package \shortcite{nlme_2016} for fitting ARMA noise models.
-We identified the optimal model for the auto-correlated noise term by performing the regression analysis for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal noise model \shortcite{Vehtari_2015}.
+The noise model was determined by fitting an $\ARMA(p,q)$ model to the residuals from the regression, using the \href{http://mc-stan.org}{Stan} probabilistic modeling language and the \href{https://cran.r-project.org/web/packages/rstan/}{\texttt{rstan}} package
+\shortcite{Stan2016,RStan2016}.
+Stan proved more numerically stable than the \Rstat\ \href{http://CRAN.R-project.org/package=nlme}{\texttt{nlme}} package
+\shortcite{nlme_2016}
+for fitting ARMA noise models.
+We identified the optimal model for the auto-correlated noise term by performing the regression analysis for all combinations of $p, q \in [0,1,2]$ and using the Widely Applicable Information Criterion to select the optimal noise model \shortcite{Watanabe_2013,Gelman_2014}.
 In both cases (TSI and $\log\ce{CO2}$), the optimal noise model was $\AR(1)$.

### 2016-06-28 prediction_market_wsc.Rnw
Switched from ggplot2 to jgplot2 (my fork of Hadley's ggplot2 which implements outlier.alpha argument to geom_boxplot). If necessary, install from devtools::install_github('jonathan-g/jgplot2', ref='jgplot2') Updated citations to WAIC.

@@
 Warming coefficients for each model were determined by linear regression of historical temperatures from 1880--2014 against the historical values for each model's forcing term.
-The noise model was determined by fitting an $\ARMA(p,q)$ model to the residuals from the regression, using the \href{http://mc-stan.org}{Stan} probabilistic modeling language and the \href{https://cran.r-project.org/web/packages/rstan/}{\texttt{rstan}} package \shortcite{Stan2016,RStan2016}.
-Stan proved more numerically stable than the \Rstat\ \href{http://CRAN.R-project.org/package=nlme}{\texttt{nlme}} package \shortcite{nlme_2016} for fitting ARMA noise models.
-We identified the optimal model for the auto-correlated noise term by performing the regression analysis for all combinations of $p, q \in [0,1,2]$ and using the Watanabe-Akaike Information Criterion to select the optimal noise model \shortcite{Vehtari_2015}.
+The noise model was determined by fitting an $\ARMA(p,q)$ model to the residuals from the regression, using the \href{http://mc-stan.org}{Stan} probabilistic modeling language and the \href{https://cran.r-project.org/web/packages/rstan/}{\texttt{rstan}} package \shortcite{Stan2016}.
+Stan proved more numerically stable than the \Rstat\ \href{http://CRAN.R-project.org/package=nlme}{\texttt{nlme}} package
+%\shortcite{nlme_2016}
+for fitting ARMA noise models.
+We identified the optimal model for the auto-correlated noise term by performing the regression analysis for all combinations of $p, q \in \lbrace 0,1,2 \rbrace$ and using the Widely Applicable Information Criterion to select the optimal noise model \shortcite{Watanabe_2013,Gelman_2014}.
 In both cases (TSI and $\log\ce{CO2}$), the optimal noise model was $\AR(1)$.
@@
 We thank Yevgeniy Vorobeychik for feedback on this project and Manuel Naumovich Velasco for sharing the data from his Total Solar Irradiance prediction.
-U.S.
-National Science Foundation grants EAR-1416964, EAR-1204685, and 1526860 partially supported this research.
-The funder had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+U.S. National Science Foundation grants EAR-1416964, EAR-1204685, and 1526860 partially supported this research.
+% The funder had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
 All views expressed are the authors' and not the funder's.

### 2016-06-28 prediction_market_wsc.Rnw
updated utils scripts to not load ggplot2. edited paper.

@@
 We conclude with details on model dynamics and an overview of the model parameters we experimentally vary.
+A full Overview, Design concepts and Details (ODD) report on our model can be found on \href{https://github.com/JohnNay/predMarket}{the project website}.
 
@@
 source("utils/climate_plot.R")
-climate_plot() + theme(text=element_text(family='Palatino'))
+climate_plot()
 @
@@
 source("utils/time_plot.R")
-p1 <- history_box() + theme(text=element_text(family='Palatino'))
-p2 <- future_box() + theme(text=element_text(family='Palatino'))
+p1 <- history_box()
+p2 <- future_box()
 multi_plot(p1, p2, cols=2)
@@
 Partial correlation computes the linear relationship between the part of the variation of $X_i$ and $y$ that are linearly independent of other $X_j$  ($j \neq i$).
-The difference between the partial correlation and the partial rank correlation that we use here is that the outcome variable $y$ is first rank-transformed in order to allow us capture potentially non-linear relationships: every value, $y_s$, in the data set is replaced by $f(y_s) \equiv |\{y_r~|~ y_s > y_r\}|$,
-where $y_r$ are the observed values of $y$ in the data set. We conduct 1,000 bootstrapped estimations of the partial rank correlation coefficients to obtain 95\% confidence intervals.
+The difference between the partial correlation and the partial rank correlation that we use here is that we first rank-transform that data in order to capture potentially non-linear relationships.
+We conduct 1,000 bootstrapped estimations of the partial rank correlation coefficients to obtain 95\% confidence intervals.
 
@@
 source("utils/sa_plot.R")
-sa_plot() + theme(text=element_text(family='Palatino'))
+sa_plot()
 @
@@
 NAY} is a Ph.D.
-Candidate at Vanderbilt University.
+Candidate at the Vanderbilt University School of Engineering and a Research Fellow at Vanderbilt Law School's Program on Law \& Innovation.
 He develops machine learning and computer simulation approaches to forecast dynamic systems and improve law and policy.

### 2016-06-28 utils/climate_plot.R

@@ -1,5 +1,4 @@
 climate_plot <- function(){
-  library(ggplot2)
   library(dplyr)
   library(tidyr)
   # to generate this data: source("timeseries/climate_plot.R", chdir = TRUE)

### 2016-06-28 utils/compute_iters.R

@@ -1,148 +1,0 @@
-
-
-setClass(Class = "computeIters",
-         slots = c(call = "language",
-                   measure = "character",
-                   result = "integer",
-                   plot_data = "data.frame",
-                   timing = "numeric",
-                   session = "ANY")
-)
-
-setMethod("summary", "computeIters",
-          function(object, digits = 3) {
-            cat(paste0("\n This process took ", object@timing, " seconds (", 
-                       round(object@timing/60/60, 2), " hours)."))
-            cat("\n Average number of iterations found to be sufficient is", object@result)
-            invisible(object)
-          }
-)
-
-setMethod("print", "computeIters",
-          function(x, ...) str(x)
-)
-
-
-setMethod("show", "computeIters",
-          function(object) {
-            cat("An object of class \"computeIters\"\n")
-            cat("\nCall:\n", deparse(object@call), "\n\n",sep="")
-            cat("Available slots:\n")
-            print(slotNames(object))
-          }
-)
-
-
-setMethod("plot", "computeIters",
-          function(x, outcome_var = "Outcome", 
-                   xlab = "Iterations", ylab = NULL){
-            
-            measure <- x@measure
-            
-            if (missing(ylab)) {
-              ylab <- measure
-            } else {
-              measure <- ylab
-            }
-            
-            x <- x@plot_data
-            
-            ggplot2::ggplot(x, 
-                            ggplot2::aes(x = iters, y = measured)) + 
-              ggplot2::geom_point() +
-              ggplot2::geom_smooth(method = "loess") +
-              ggplot2::xlab(xlab) + ggplot2::ylab(ylab) +
-              ggplot2::ggtitle(paste("Iterations and", measure, "for", outcome_var)) +
-              ggplot2::theme_bw()
-          }
-)
-
-
-# Coefficient of variation
-# @param na.rm Logical vector length one indicating whether 
-# NA values should be stripped before the computation proceeds
-coef_var <- function(x, na.rm = FALSE){
-  sd(x, na.rm = na.rm)/
-    mean(x, na.rm = na.rm)
-}
-
-compute_iters <- function(abm, 
-                          input_values,
-                          out, 
-                          sample_count = 20,
-                          repeats = 30,
-                          thresh = 0.05,
-                          initial_iters = 5,
-                          max_iters = 100,
[…]

### 2016-06-28 utils/pc_sa.R

@@ -100,13 +100,13 @@ setMethod("plot", "pcSA",
             x <- x@result
             ss <- correct_bias(x[[7]]) # $SRRC, $SRC, $PRCC, $PCC
             
-            ggplot2::ggplot(ss, ggplot2::aes(x = var, y = x_corr)) + ggplot2::geom_point() +
-              ggplot2::xlab(xlab) + ggplot2::ylab(ylab) +
-              ggplot2::ggtitle(paste("Estimated effects on", outcome_var)) +
-              ggplot2::geom_errorbar(ggplot2::aes(ymax = max_ci, ymin = min_ci), width=0.25) +
-              ggplot2::geom_hline(yintercept = 0) +
-              ggplot2::ylim(c(-1,1)) +
-              ggplot2::theme_bw()
+            ggplot(ss, aes(x = var, y = x_corr)) + geom_point() +
+              xlab(xlab) + ylab(ylab) +
+              ggtitle(paste("Estimated effects on", outcome_var)) +
+              geom_errorbar(aes(ymax = max_ci, ymin = min_ci), width=0.25) +
+              geom_hline(yintercept = 0) +
+              ylim(c(-1,1)) +
+              theme_bw()
           }
 )
 

### 2016-06-28 utils/sa_plot.R

@@ -1,4 +1,3 @@
-library(ggplot2)
 library(dplyr)
 library(tidyr)
 

### 2016-06-28 README.Rmd
Made a few wordsmithing edits:
Added a direct link to the ODD at the top of the MODEL DESIGN section, and added a mention that the model code is also available at the project web site.

Fixed an erroneous description of risk.tak in the DISCUSSION section.

Minor wordsmithing to save a line here and there, where there was only one word in the last line of a paragraph, to keep the length to 12 pages.

@@
 
-The paper can be downloaded here: [http://arxiv.org/abs/1603.08961](http://arxiv.org/abs/1603.08961)
+The paper can be downloaded here: [http://arxiv.org/abs/1603.08961](http://arxiv.org/abs/1603.08961) and the ODD specification for the agent-based prediction-market model can be found here: [https://github.com/JohnNay/predMarket/blob/master/odd.pdf](https://github.com/JohnNay/predMarket/blob/master/odd.pdf)
 

### 2016-06-28 prediction_market_wsc.Rnw
Made a few wordsmithing edits:
Added a direct link to the ODD at the top of the MODEL DESIGN section, and added a mention that the model code is also available at the project web site.

Fixed an erroneous description of risk.tak in the DISCUSSION section.

Minor wordsmithing to save a line here and there, where there was only one word in the last line of a paragraph, to keep the length to 12 pages.

@@
 library(knitr)
-opts_chunk$set(fig.align = "center")
+opts_chunk$set(fig.align = "center", cache = FALSE)
 @
@@
 We conclude with details on model dynamics and an overview of the model parameters we experimentally vary.
-A full Overview, Design concepts and Details (ODD) report on our model can be found on \href{https://github.com/JohnNay/predMarket}{the project website}.
+The model code and a full \href{https://github.com/JohnNay/predMarket/blob/master/odd.pdf}{Overview, Design concepts and Details (ODD) specification} can be found on \href{https://github.com/JohnNay/predMarket}{the project website}.
 
@@
 
-The higher the risk tolerance, the greater the range of prices at which a trader will buy and sell securities.
-This creates greater opportunities for traders to profit if their model is correct and to lose money if it is incorrect.
-This increased volatility puts more information in play, and we believe this is why greater risk taking enables traders to identify the correct model more quickly.
+Traders with higher risk tolerance will price buy and sell orders more aggressively, so they will earn more (or lose less) on completed trades, but will have a greater risk of failing to complete trades.
+Among traders with the correct model of climate, those with higher risk tolerance  will earn more profit on completed trades (and their counterparties will lose more).
+A trader's wealth is an important source of information, and we believe this is why greater risk taking enables traders to identify the correct model more quickly.
+Additionally, the risk of failing to complete trades adds volatility, which also puts information in play.
 We have observed a similar phenomenon in a very different context, where adding stochastic noise to player decisions in iterated games improves the accuracy with which machine-learning algorithms can identify the players' strategies \shortcite{Nay2015}.
@@
 All code and data for the model is available at \href{https://github.com/JohnNay/predMarket}{github.com/JohnNay/predMarket}.
-This project is a computational test-bed for public policy design: our code can be extended to test the effects that trading strategies, cognitive models, future climate scenarios, and market designs have on the evolution of trader beliefs.
+This project is a computational test-bed for public policy design: our code can be extended to test the effects of trading strategies, cognitive models, future climate scenarios, and market designs on the evolution of trader beliefs.
 

### 2016-06-29 prediction_market_wsc.Rnw
Some paper edits.

@@
 We simulate a prediction market where traders exchange securities related to climate outcomes to explore whether, and under what social and climate conditions, prediction markets may be useful for increasing convergence of climate beliefs.
-Our work can be extended as part of a computational design process for effective climate prediction markets.
+Our work can be extended as part of a computational design process for effective climate prediction markets and we release all our code along with this paper.
 
@@
 We conclude with details on model dynamics and an overview of the model parameters we experimentally vary.
-The model code and a full \href{https://github.com/JohnNay/predMarket/blob/master/odd.pdf}{Overview, Design concepts and Details (ODD) specification} can be found on \href{https://github.com/JohnNay/predMarket}{the project website}.
+A full Overview, Design concepts and Details (ODD) specification can be found on \href{https://github.com/JohnNay/predMarket}{the project website}.
 
@@
 Most scientific models of the earth's climate include many forcings, including carbon dioxide, other greenhouse gases, aerosols, total solar irradiance, and more.
-In these models, the changing $\ce{CO2}$ concentration is, by a large margin, the strongest single forcing \shortcite[p.~14]{IPCC_WG1_AR5_2013}, so for simplicity we consider only \ce{CO2}.
-Choosing only one forcing term for each of the competing models simplifies comparison because each model has the same number of adjustable parameters.
+In these models, the changing $\ce{CO2}$ concentration is, by a large margin, the strongest single forcing \shortcite[p.~14]{IPCC_WG1_AR5_2013}.
+Choosing only one forcing term for each of the competing models simplifies comparison because each model has the same number of adjustable parameters, therefore we consider only \ce{CO2}.
 

### 2016-06-29 prediction_market_wsc.Rnw
Two minor changes in wording.

@@
 Ideally, we would like to compare belief convergence with and without a prediction market but because the only source of climate belief in our current model is market interactions we cannot make this comparison.
-However, we do use actual temperatures for our model under historical 20th century conditions and observe convergence to the \ce{CO2} model, while in the real world there is no convergence of beliefs.
+However, we do use actual temperatures for our model under historical 20th century conditions and observe convergence to the \ce{CO2} model, whereas in the real world there is no convergence of beliefs.
 
@@
 We thank Yevgeniy Vorobeychik for feedback on this project and Manuel Naumovich Velasco for sharing the data from his Total Solar Irradiance prediction.
-U.S. National Science Foundation grants EAR-1416964, EAR-1204685, and 1526860 partially supported this research.
+U.S. National Science Foundation grants EAR-1416964, EAR-1204685, and IIS-1526860 partially supported this research.
 % The funder had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.

### 2016-06-29 arxiv.Rnw
pasted in the body of the new text to the arXiv.Rnw and compiled that

@@
 We conclude with details on model dynamics and an overview of the model parameters we experimentally vary.
+A full Overview, Design concepts and Details (ODD) specification can be found on \href{https://github.com/JohnNay/predMarket}{the project website}.
 
@@
 Warming coefficients for each model were determined by linear regression of historical temperatures from 1880--2014 against the historical values for each model's forcing term.
-The noise model was determined by fitting an $\ARMA(p,q)$ model to the residuals from the regression, using the \href{http://mc-stan.org}{Stan} probabilistic modeling language and the \href{https://cran.r-project.org/web/packages/rstan/}{\texttt{rstan}} package
-\shortcite{Stan2016,RStan2016}.
+The noise model was determined by fitting an $\ARMA(p,q)$ model to the residuals from the regression, using the \href{http://mc-stan.org}{Stan} probabilistic modeling language and the \href{https://cran.r-project.org/web/packages/rstan/}{\texttt{rstan}} package \shortcite{Stan2016}.
 Stan proved more numerically stable than the \Rstat\ \href{http://CRAN.R-project.org/package=nlme}{\texttt{nlme}} package
-\shortcite{nlme_2016}
+%\shortcite{nlme_2016}
 for fitting ARMA noise models.
-We identified the optimal model for the auto-correlated noise term by performing the regression analysis for all combinations of $p, q \in [0,1,2]$ and using the Widely Applicable Information Criterion to select the optimal noise model \shortcite{Watanabe_2013,Gelman_2014}.
+We identified the optimal model for the auto-correlated noise term by performing the regression analysis for all combinations of $p, q \in \lbrace 0,1,2 \rbrace$ and using the Widely Applicable Information Criterion to select the optimal noise model \shortcite{Watanabe_2013,Gelman_2014}.
 In both cases (TSI and $\log\ce{CO2}$), the optimal noise model was $\AR(1)$.
@@
 Market participation causes traders to converge toward believing the ``true'' climate model under a variety of model parameterizations in a relatively short time: In markets with low segmentation, the number of traders believing the true model rises from 50\% to 75\% in roughly 12 years, and even in highly segmented markets, 75\% convergence is achieved in 18--24 years.
-This suggests that a climate prediction market could be useful for producing broad agreement about the causes of climate change, and could have persuasive power for people who are not persuaded by the overwhelming consensus among climate researchers that greenhouse gases (and especially \ce{CO2}) are responsible for the majority of observed climate change \shortcite{IPCC_WG1_AR5_2013}.
-However, market segmentation has a large effect on the speed of convergence, so transparency and effective communication about the performance of traders with different beliefs will be important.
+Ideally, we would like to compare belief convergence with and without a prediction market but because the only source of climate belief in our current model is market interactions we cannot make this comparison.
+However, we do use actual temperatures for our model under historical 20th century conditions and observe convergence to the \ce{CO2} model, whereas in the real world there is no convergence of beliefs.
+
+Both the historical and future simulation results suggest that a climate prediction market could be useful for producing broad agreement about the causes of climate change, and could have persuasive power for people who are not persuaded by the overwhelming consensus among climate researchers that greenhouse gases (and especially \ce{CO2}) are responsible for the majority of observed climate change \shortcite{IPCC_WG1_AR5_2013}.
+We also find that market segmentation has a large effect on the speed of convergence, so transparency and effective communication about the performance of traders with different beliefs will be important.
 
@@
 
-We thank Yevgeniy Vorobeychik for feedback on this project.
-We thank Manuel Naumovich Velasco for sharing the data from his Total Solar Irradiance prediction.
+We thank Yevgeniy Vorobeychik for feedback on this project and Manuel Naumovich Velasco for sharing the data from his Total Solar Irradiance prediction.
+U.S. National Science Foundation grants EAR-1416964, EAR-1204685, and IIS-1526860 partially supported this research.
+% The funder had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+All views expressed are the authors' and not the funder's.
 %We also thank the authors of \href{http://www.r-project.org}{\textsf{R}} \shortcite{r_core}

### 2016-06-29 arXiv_2/arxiv.tex
All the arxiv and final WSC submission files.

--- 2016-06-29 arxiv.Rnw
@@
 
-\documentclass{article}
+\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
 
-<<knitr_setup, echo=F, results='hide', cache=F>>=
-library(knitr)
-opts_chunk$set(fig.align = "center")
-@
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
 
-<<setup, echo=F, results='hide', cache=F>>=
-library(jgplot2) # devtools::install_github('jonathan-g/jgplot2', ref='jgplot2')
-library(pander)
-source("utils/multi_plot.R")
-theme_set(theme_classic())
-@
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 
@@
 \newcommand{\AR}{\text{AR}}
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 % AUTHOR: Enter the authors of the article, see end of the example document for further examples
-\author{John J. Nay, Martin Van der Linden,
+\author{John J. Nay\thanks{Contact: john.j.nay@gmail.com. For updated information: \href{http://johnjnay.com/}{johnjnay.com}.}, Martin Van der Linden,
 Jonathan M. Gilligan
@@
 
-<<climateplot, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Historical measurements of temperature and a realization of possible future temperatures under two alternate models of climate physics.", fig.width=4, fig.height=3>>=
-source("utils/climate_plot.R")
-climate_plot()
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
+
+{\centering \includegraphics[width=\maxwidth]{figure/climateplot-1}
+
+}
+
+\caption[Historical measurements of temperature and a realization of possible future temperatures under two alternate models of climate physics]{Historical measurements of temperature and a realization of possible future temperatures under two alternate models of climate physics.}\label{fig:climateplot}
+\end{figure}
+
+\end{knitrout}
 
@@
 Consistent with this, our simulated historical trading sequence (Fig.~\ref{fig:time}) shows convergence to belief in the TSI model until the early 1970s, after which traders converge toward believing the \ce{CO2} model.
-<<time, eval=TRUE, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="Convergence over trading sequences for different degrees of social-network connection and segmentation.", fig.width=10.5, fig.height=8.5, fig.pos='t'>>=
-source("utils/time_plot.R")
-p1 <- history_box()
-p2 <- future_box()
-multi_plot(p1, p2, cols=2)
-@
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[t]
+
+{\centering \includegraphics[width=\maxwidth]{figure/time-1}
+
+}
+
+\caption[Convergence over trading sequences for different degrees of social-network connection and segmentation]{Convergence over trading sequences for different degrees of social-network connection and segmentation.}\label{fig:time}
+\end{figure}
+
+\end{knitrout}
 
@@
 
-<<sa_n, eval=TRUE, echo=FALSE, message=FALSE>>=
-load("output/src_future.Rda")
-n_sims <- nrow(src_future@input_set)
-@
-
 Our primary focus is on simulating the market in future scenarios.
 This is more relevant for policy design and more interesting theoretically due to the increasing divergence of global temperature values under the two future scenarios.
-The sensitivity analysis is based on a Latin hypercube sampling of \Sexpr{n_sims} parameter sets from the following distributions \shortcite{beachkofski_improved_2002,carnell_lhs_2012}.
+The sensitivity analysis is based on a Latin hypercube sampling of 500 parameter sets from the following distributions \shortcite{beachkofski_improved_2002,carnell_lhs_2012}.
 
@@
 
-We used the model to perform 10 full simulations for each of the \Sexpr{n_sims} input parameter sets and average the 10 convergence scores.
+We used the model to perform 10 full simulations for each of the 500 input parameter sets and average the 10 convergence scores.
 We conducted multiple simulations for the same parameter set because there is stochasticity in the tempertature time series, the social network structure, and the agent decision models.
@@
 
-<<sa, eval=TRUE, echo=FALSE, message=FALSE, fig.cap="Estimated effects of model parameters on convergence of beliefs in future scenario.
+\begin{knitrout}
+\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}[t]
+
+{\centering \includegraphics[width=\maxwidth]{figure/sa-1}
+
+}
+
+\caption[Estimated effects of model parameters on convergence of beliefs in future scenario]{Estimated effects of model parameters on convergence of beliefs in future scenario.
 Positive numbers accelerate convergence and negative numbers retard it.
-True model refers to the true climate model being \\ce{CO2}, so traders converge to the true model more quickly when carbon dioxide drives climate change than when total solar irradiance does.
-Bars represent 95\\% confidence intervals.", fig.pos='t', fig.width=3, fig.height=3.5>>=
-source("utils/sa_plot.R")
-sa_plot()
-@
+True model refers to the true climate model being \ce{CO2}, so traders converge to the true model more quickly when carbon dioxide drives climate change than when total solar irradiance does.
+Bars represent 95\% confidence intervals.}\label{fig:sa}
+\end{figure}
+
+\end{knitrout}
 

### 2016-06-29 wsc_final_submit/prediction_market_wsc.tex
All the arxiv and final WSC submission files.

--- 2016-06-29 prediction_market_wsc.Rnw
@@
 
-\documentclass{wscpaperproc}
+\documentclass{wscpaperproc}\usepackage[]{graphicx}\usepackage[]{color}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
 
-<<knitr_setup, echo=F, results='hide', cache=F>>=
-library(knitr)
-opts_chunk$set(fig.align = "center", cache = FALSE)
-@
+\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
+\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
+\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
+\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
+\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
+\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
+\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
+\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
+\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
+\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
 
-<<setup, echo=F, results='hide', cache=F>>=
-# replace ggplot2 with jgplot2 that has my enhancement for outlier.alpha on geom_boxplot.
-# library(ggplot2)
-library(jgplot2) # devtools::install_github('jonathan-g/jgplot2', ref='jgplot2')
-library(pander)
-source("utils/multi_plot.R")
-theme_set(theme_classic())
-@
+\usepackage{framed}
+\makeatletter
+\newenvironment{kframe}{%
+ \def\at@end@of@kframe{}%
+ \ifinner\ifhmode%
+  \def\at@end@of@kframe{\end{minipage}}%
+  \begin{minipage}{\columnwidth}%
+ \fi\fi%
+ \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
+ \colorbox{shadecolor}{##1}\hskip-\fboxsep
+     % There is no \\@totalrightmargin, so:
+     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
+ \MakeFramed {\advance\hsize-\width
+   \@totalleftmargin\z@ \linewidth\hsize
+   \@setminipage}}%
+ {\par\unskip\endMakeFramed%
+ \at@end@of@kframe}
+\makeatother
+
+\definecolor{shadecolor}{rgb}{.97, .97, .97}
+\definecolor{messagecolor}{rgb}{0, 0, 0}
+\definecolor{warningcolor}{rgb}{1, 0, 1}
+\definecolor{errorcolor}{rgb}{1, 0, 0}
+\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX
+
+\usepackage{alltt}
 

### 2016-07-11 wsc_final_submit/prediction_market_wsc.tex
Updated references to be all capital for first letters in the titles. Updated the header to be correct list of proceedings editor names.

@@
 Candidate at the Vanderbilt University School of Engineering and a Research Fellow at Vanderbilt Law School's Program on Law \& Innovation.
-He develops machine learning and computer simulation approaches to forecast dynamic systems and improve law and policy.
-His research involves collaboration with lawyers, social scientists, environmental scientists, and engineers.
+He develops computational approaches to forecasting dynamic systems and improving law and policy, e.g.
+\href{http://predictgov.com/}{predictgov.com}.
 His email address is \email{john.j.nay@gmail.com}, and website: \href{http://johnjnay.com/}{http://johnjnay.com/}.\\

### 2016-07-11 README.Rmd
Updated readme with instructions

@@
 theme_set(theme_classic())
-p1 <- history_box() + theme(text=element_text(family='Palatino'))
-p2 <- future_box() + theme(text=element_text(family='Palatino'))
+p1 <- history_box()
+p2 <- future_box()
 multi_plot(p1, p2, cols=2)
 ```
+
+## To run the simulations
+
+In R, install packages and load them to ensure they installed.
+
+```
+install.packages("devtools")
+install.packages("igraph")
+install.packages("ggplot2")
+install.packages("rstan")
+install.packages("shinystan")
+install.packages("dplyr")
+install.packages("tidyr")
+install.packages("magrittr")
+install.packages("purrr")
+install.packages("stringr")
+install.packages("tseries")
+install.packages("lhs")
+
+devtools::install_github("JohnNay/sa")
+library(igraph)
+library(ggplot2)
+library(rstan)
+library(shinystan)
+library(dplyr)
+library(tidyr)
+library(magrittr)
+library(purrr)
+library(stringr)
+library(tseries)
+library(lhs)
+```
+
+Then in R source the main simulation model file and test it out for a run with randomly chosen parameters.
+
+```
+source("main.R")
+main(c(runif(3, min = 0.0001, max = 0.9999), 1, runif(2, min = 0.0001, max = 0.9999)), visu =T)
+```
+
+In the terminal run the following to conduct a sensitivity analysis simulation experiment with 8 cores. The only argument to this script is the number of cores to use.
+
+```
+Rscript master.R 8 &
+```
+
+In the terminal run the following to conduct the time series experiment with 8 cores. The only argument to this script is the number of cores to use.
+
+```
+Rscript master_timeseries_experiments.R 8 &
+```

### 2016-07-11 master.R

@@ -1,5 +1,5 @@
 # master script for conducting ABM experiments
-rm(list=ls()) # make sure to always run this line of code and see that the next two 
+# rm(list=ls()) # make sure to always run this line of code and see that the next two 
 # lines of code work without error, ensuring that the working of the model is not 
 # dependent on anything in your global workspace, if it is, then you need to create 
 # whatever is in your global workpace in the code
@@ -59,7 +59,7 @@ if(estimate_replicates){
 ##############################################################################
 
 future <- TRUE
-past <- FALSE
+past <- TRUE
 sample_count <- 80
 
 if(past){

### 2016-07-11 master_timeseries_experiments.R

@@ -1,10 +1,10 @@
-rm(list=ls())
+# rm(list=ls())
 args <- commandArgs(trailingOnly = TRUE)
 numcores <- as.integer(args[1])
 message(paste("Number of cores", numcores))
 
 source("main.R")
-library(eat)
+library(sa)
 
 ##############################################################################
 ## Set up parameters for climate_model to control level of tracing to screen
@@ -136,7 +136,7 @@ set[[4]] <- list(n.edg = 0.05, seg = 0.05)
 ##############################################################################
 doParallel::registerDoParallel(cores = numcores)
 sample_count <- numcores*6 # numcores*7 takes 8.5 hours to run for past
-past <- FALSE
+past <- TRUE
 future <- TRUE
 true_past <- TRUE
 

### 2016-07-11 assortativity_coefficient.R
Delete assortativity_coefficient.R

@@ -1,24 +1,0 @@
-###
-# From http://www.babelgraph.org/wp/?p=351 for more details
-###
-# calculate the assortativity coefficient for a mixing matrix of a graph
-# ref: MEJ Newman, 'Mixing patterns in networks', Phys Rev E 67, 026126 (2003)
-#
-# define assortativity coefficient as
-#        trace (m) - sum (m^2)
-# ac = -------------------------
-#           1 - sum (m^2)
-# 
-# From Newman " This formula gives 
-#                                 r = 0 when there is no assortative mixing,
-#                                 r = 1 when there is perfect assortative mixing
-#                                 some negative value between -1 and 0 when the
-#                                 is perfectly disassortiative, i.e., every edge
-#                                 connects two vertices of different types
-
-assortcoeff <- function(m) {
-        tr <- 0
-        for (k in 1:nrow(m)) tr <- tr + m[k,k]
-        sumsq <- sum (rowSums(m)*colSums(m))
-        (tr - sumsq) / (1 - sumsq)
-}
